### 2020-05-31 tex/draft.tex
Draft paper in tex.

%
% File acl2020.tex
%
%% Based on the style files for ACL 2020, which were
%% Based on the style files for ACL 2018, NAACL 2018/19, which were
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
%% EACL-2009, IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2020}
\usepackage{times}
\usepackage{latexsym}
\usepackage{amsmath,accents}
\usepackage{hang}
\renewcommand{\UrlFont}{\ttfamily\small}
\usepackage{microtype}

\input{syntax.tex}
\usepackage{wasysym,graphicx}

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.

%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B\textsc{ib}\TeX}

\title{Contstructive Epistemic Semantics in Guarded String Models}

\author{First Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  \texttt{email@domain} \\\And
  Second Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  \texttt{email@domain} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
  Constructive and computable multi-agent epistemic possible worlds
  models are defined, where possible worlds models are guarded string
  models in an epistemic extension of Kleene Algebra with Tests.  The
  account is framed as a formal language Epik (Epistemic KAT) for
  defining such models.  The language is interpreted by translation
  into the finite state calculus, and alternatively by modeling
  propositions as lazy lists. The syntax-semantics interface for
  a fragment of English is defined by a categorial grammar.
\end{abstract}

\section{Introduction and Related Work}
Linguistic semantics in the Montague tradition proceeds by assigning
propositional {\em semantic values} to disambiguated sentences of a
natural language.  A proposition is a set or class of {\em possible
  worlds}.  These worlds are often assumed to be things with the same
nature and complexity as the world we occupy (Lewis 1986).  But
alternatively, one can work with small idealized models, in order to
illustrate and test ideas.  To build such models, spaces of worlds and
individuals are stipulated as small finite sets, and semantic values
of lexical items are constructed as functions or relations from these
small sets.  Such toy or idealized models are useful in research and
in teaching, in that it is possible to represent propositions finitely
and explicitly, and to calculate with them. The point of this paper is
to scale up toy or idealized models to countable sets of worlds, and
to constructive and computable modeling of epistemic alternatives for
agents.  We describe a certain systematic way of defining such models,
and illustrate how to apply them in natural language semantics.
The focus on epistemic semantics and clausal embedding. The fundamental move is to
identify possible worlds with strings primitive events, so that
propositions are sets of strings. An advantage in this is that it
allows for a mathematical description of an algebra of a propositions,
coupled with a computational representation using either lazy lists of
strings, or finite state machines that describe sets of strings.

The approach taken here synthesizes five antecedents in a certain way.
John McCarthy's {\em Situation Calculus} is the source of the idea of
constructing possible worlds as event sequences (McCarthy 1963, Reiter
2001).  The algebraic theory of {\em Kleene Algebra with Tests}
characterizes algebras with elements corresponding to propositions and
event types in our application (Kozen 2003).  The models we propose
are an epistemic extension of guarded string models for KAT, where a
unary operation interpreted as an existential epistemic modality is
included for each agent.  {\em Action models} in dynamic epistemic
semantics introduced the technique of constructing epistemic models
from primitive alternative relations on events, in order to capture
the epistemic consequences of perceptual events (Baltag, Moss, and
Solecki 1998). This is the basis for our construction of epistemic
alternative relations.  Literature on finite state methods in
linguistic semantics has used event strings and sets of event strings
to theorize about tense and aspect in natural language semantics
(Fernando 2003, Carlson 2009) and to express propositions (Fernando
2017).  Literature on finite state intensiononal semantics has 
shown how to do the semantics of intensional complementation
including indirect questions in a setting where propositions are
represented using finite state machines, and compositional semantics
is expressed in a finite state calculus (Rooth 2017, Collard 2018).
We adopt this in our syntax-semantics interface for
English.

We begin with examples of event-sequence models.  {\em The
  Elevator}. An elevator moves up and down in a four-story building,
with floors numbered in the European fashion as 0,1,2,3.  There are
primitive events $u$ (the elevator going up one floor), and $d$ (the
elevator going down on floor).  In worlds $v_1$ and $v_2$, the events
shown in \ex{elevator1} transpire. The truth values for English
sentences shown in \ex{elevator2} are observed.

\begin{example}  \label{elevator1}
\hspace{-4mm} $v_1$
\begin{tabular}[t]{ll}
  $u$ & it goes up from 0 to 1 \\
  $u$ & it goes up from 1 to 2 \\
  $d$ & it goes down from 2 to 1 \\
  $u$ & it goes up from 1 to 2 \\
\end{tabular}

\vspace{2mm}
\hspace{-4mm} $v_2$
\begin{tabular}[t]{ll}
  $u$ & it goes up from 0 to 1 \\
  $u$ & it goes up from 1 to 2 \\
  $u$ & it goes up from 2 to 3 \\
\end{tabular}
\end{example}

\begin{example}  \label{elevator2}
\hspace{-4mm} \begin{tabular}[t]{ccl}
  $v_1$ & $v_2$ & Sentence \\
   false & true & It's on floor 3. \\
   true & true & It has gone up. \\
   true & false & It has gone down. \\
   true & false & It could go up. \\
\end{tabular}
\end{example}

{\em The Concealed Coin}.  Amy and Bob are seated at a table.  There is a coin on the
table under a cup, heads up (H).  The coin could be H (heads) or T (tails), and
neither agent knows which it is.  This initial situation is possible world $w_1$.
Two additional worlds $w_2$ and $w_2$ are
defined by sequencing events after the initial state, with events interpreted as in
\ex{coin1}. The truth values for English
sentences shown in \ex{coin3} are observed.

\begin{example}  \label{coin1}
\hspace{-4mm} \begin{tabular}[t]{ll}
  $a_1$  & Amy peeks at H, by tipping the cup.
  \\ & Bob sees she's peeking, but not what \\
  & she sees. \\
  $b_1$  & Bob peeks at H. \\ 
\end{tabular}
\end{example}

\begin{example} \label{coin2}
\hspace{-4mm} \begin{tabular}[t]{ll}
   $w_1$ \\
   $w_2 = w_1 \hspace{2pt} a_1 $ \\
   $w_2 = w_1 \hspace{2pt} a_1 \hspace{2pt} b_1 $\\ 
 \end{tabular}
\end{example}

\begin{example} \label{coin3}
\hspace{-9mm} \begin{tabular}[t]{cccl}
  $w_1$ & $w_2$  & $w_3$ & Sentence\\
  false & true & true& Amy knows it's H. \\
  false & false & true& Bob knows it's H. \\
  false & false & true& Bob knows Amy \\ &&& knows it's H. \\
  false & true & true& Bob knows Amy knows \\ &&& whether it's H or T. \\
\end{tabular}
\end{example}

The events in the examples come with pre-conditions. The elevator can
not go up if it is already on floor 3, so $u$ has the pre-condition of
the elevator being of floor 0, 1, or 2.  Similarly $d$ has the
precondition that the elevator is on floor 1, 2 or 3.  Amy can peek at
heads only if the coin is heads up, so $a_1$ has the precondition of
the coin being heads up.  Let $h$ be the Boolean proposition that the
coin is heads up.  In the other example, let $q$ be the proposition
that the elevator is on a high floor (2 or 3), and $p$ be the
proposition that it is on an odd floor (1 or 3). Then preconditions
can be described by Boolean formulas, with $h$ being the precondition
of $a_1$, and $!(pq)$ being the precondition of $u$. Juxtaposition is
used for Boolean conjunction, and the exclamation point for for
Boolean negation. Events come as well with a relation between prior
and following state, for instance with $u$ incrementing the floor.
This is expressed using an operator ``$:$'' (read ``and next'') that
pairs Boolean formulas. The first line in \ex{event1} describes $a_1$
(Amy looking at heads) as happening only in an $h$ state, and as not
changing the state.  Symmetrically, $a_0$ (Amy looking at tails) can
happen only in a not-$h$ state, and does not change the state.  The
third line says that $u$ increments the floor, and can happen only on
floors 0, 1, and 2.  The fourth line describes $d$ in similar terms.
Plus is disjunction.

\newcommand{\ev}[2]{#1 \!\! : \!\! #2}
\newcommand{\evr}[2]{#2 \!\! : \!\! #1}

\begin{example}   \label{event1}
\hspace{-5mm}\( \begin{array}[t]{cl}
a_1 & h:h \\
a_0 & \ev{(!h)}{(!h)} \\
% u & (!q!p) \! : \!(!qp) + \\
u & \ev{(!q!p)}{(!qp)} + \ev{(!qp)}{(q!p)} + \ev{(q!p)}{(qp)} \\
d & \evr{(!q!p)}{(!qp)} + \evr{(!qp)}{(q!p)} + \evr{(q!p)}{(qp)} \\ 
\end{array} \)
\end{example}

\section{Epistemic guarded string models}
Figure 1 shows an Epik program that describes a possible worlds model for two agents
with information about one coin, and events of the agents semi-privately
looking at the coin. The line beginning with {\tt state} lists the basic
stative propositions. To illustrate syntax, a second proposition $t$ (tails) is
included.  The line beginning with {\tt constraint} defines compatibilities among the
propositions: the coin is heads or tails and not both.  The lines beginning with {\tt
  event} declare events, their preconditions, and their effect on state, following
the format in \ex{event1}.  Finally the lines beginning with {\tt agent} define {\em
  event alternative} relations for agents.  Each clause with an arrow has a single
event symbol on the left, and a disjunction of alternative events on the right of the
arrow. The interpretation of Amy's alternatives for $b_1$ (Bill peeks at heads), is
that when $b_1$ happens, for Amy either $b_1$ or $b_0$ (Bill peeks at tails) could
be happening.

Kleene Algebra with Tests is an algebraic theory that is defined by equations and
inequalities, which a has model classes including guarded string models, relational
models, finite models, and matrix models.  This paper focuses on defining a family of
concrete guarded string algebras, the elements of which are sets of guarded strings.
Definitions and notation mostly follow Kozen (2003).  Additional syntax and semantics
is included to model multi-agent epistemic semantics.  Guarded strings over a finite
alphabet $E$ are like ordinary strings, but with truth assignments to a set $T$ of
primitive propositions (primitive tests) alternating with the symbols from $E$.  In
the algebra described by Figure 1, $E$ is the set of events $\{a_1,a_0,b_1,b_0\}$,
and in the elevator example, $\{u,d\}$.  In the elevator example, $T$ is $\{p,q\}$,
and in the coin example it is $\{h,t\}$.  We write truth assigments to the primitive
propositions that observe the constraint as Boolean vectors. In the coin example, we
get the vectors 01, 10 and in the elevator example, the vectors 00, 01, 10, 11.
Guarded strings are strings of events, alternating with such vectors, and starting
and ending with vectors.  \ex{gs1} gives the encoding as guarded strings of the
worlds in \ex{elevator1} and \ex{coin1}.

\begin{example}   \label{gs1}
\begin{tabular}[t]{ll}
$v_1$ & $00 u 01 u 10 d 01 u 10$ \\
$v_2$ & $00 u 01 u 10 u 11$ \\   
$w_1$ & $01$ \\
$w_2$ & $01 a_0 01$ \\  
$w_3$ & $01 a_0 01 b_0 01$ \\
\end{tabular}
\end{example}

\begin{figure}
\begin{center}
{\tt
\begin{tabular}{l}
state h t \\
constraint h!t + t!h \\
event a1 h:h \\
event a0 t:t \\
event b1 h:h \\
event b0 t:t \\
agent aly \\
\hspace{3mm} a1 -> a1 \\
\hspace{3mm} a0 -> a0 \\
\hspace{3mm} b1 -> b1 + b0\\
\hspace{3mm} b0 -> b1 + b0\\
agent bob \\
\hspace{3mm} b1 -> b1 \\
\hspace{3mm} b0 -> b0 \\
\hspace{3mm} a1 -> a1 + a0\\
\hspace{3mm} a0 -> a1 + a0\\
\end{tabular} }
\end{center}

\caption{Epik program describing a possible-worlds event sequence model for two agents
  with information about one coin, and events of the agents semi-privately looking
at the coin.}
\end{figure}

The discussion of \ex{coin2} mentioned building worlds by incrementing worlds with events.
This is accomplished in guarded string models with fusion product, a partial operation
that combines two guarded strings, subject to the condition that the truth assignment
at the end of the the first argument is identical to the truth assigment at the
start of the second one.  \ex{fusion1} gives some examples.

\begin{example} \label{fusion1}
\hspace{-2mm}\( 00 u 01 u 10 d 01 \cdot 01 u 10 = 00 u 01 u 10 d 01 u 10 \)

\hspace{-2mm}\( 00 u 01 u 10 d 01 \cdot 10 u 11 = \mbox{\em{undefined}} \)

\hspace{-2mm}\(01 \cdot 01 a_0 01 = 01 a_0 01 \)

\hspace{-2mm}\(01 \cdot 10 a_1 10 = \mbox{\em{undefined}} \)
\end{example}

Rather than guarded strings, elements of a guarded string model for KAT are sets of
guarded strings.  In the application, these elements have the interpretation of
propositions (sets of possible worlds) and/or event types.  An event such as $u$ in
the guarded string model corresponds to the set of guarded strings where the bare
event is flanked by compatible truth assigments, as defined in an Epik program by the
event declaration and the constraint declaration. 
These have the status of event types, in that they
can ``happen'' in different possible worlds. Here happening corresponds to
incrementing the world with some compatible element of the event type.
Where $e$ is a bare event symbol, $\hat{e}$ is the
corresponding set of guarded strings,
consisting of the bare event decorated with compatible truth assignments.
See the examples in \ex{events1}.
% u (!q!p):(!qp) + (!qp):(q!p) + (q!p):(qp) 173
% d (!qp):(!q!p) + (q!p):(!qp) + (qp):(q!p)

\begin{example} \label{events1}
\(
\begin{array}[t]{l|c}
  e & \hat{e} \\ \hline  
  u & \{ 00u01, 01u10, 10u11\}\\
  d & \{ 01d00, 10d01, 11d10\}\\
  a_1 & \{10 a_1 10 \}\\
  a_0 & \{01 a_0 01\}\\
  b_1 & \{10 b_1 10 \}\\
  b_0 & \{01 b_0 01 \}\\
\end{array}
\)
\end{example}

Kleene Algebra with Tests has the algebraic signature
$\tuple{K,+,\cdot,*,\bar{},0,1}$.  To this we add an unary operation $\Diamond_a$ for
each agent, and a unary complent operation $\cdot^{c}$ on propositions.  This results
in the signature $\tuple{K,+,\cdot,*,\bar{},0,1,{}^c,\Diamond_a,\Diamond_b}$ for
epistemic KAT with two agents.  In a guarded string model in our construction, the
elements of $K$ are sets of guarded strings.  The operation $+$ is set union. The
constant $0$ is the empty set.  The constant 1 is the set of all truth assignments
obeying the constraint on truth assignments, i.e. $\{10,10\}$ in the coin example.
The operation $\cdot$ is fusion product raised to sets: $x \cdot y$ is the set of all
defined fusion products of an element of $x$ with and element of $y$.  The operation
$*$ is Kleene star, with $x* = \cup_{i\geq 0}x^i$, where $x^i$ is the $i$-times
product of $x$ with itself and $x^0=1$.  Subsets of $1$ are also elements of $K$, and
these form the Boolean algebra of tests as a set Boolean algebra.  The overbar
operation is complement in this Boolean algebra.  The complement operation ${}^c$ is
complement at the level of sets of guarded strings, with $x^c = 0^c - x$, where the
operation on the right is set difference.  The epistemic modality $\Diamond_a$ is
interpreted using Kripke semantics, as pre-image relative to a fixed relation $R_a$
between guarded strings, $\Diamond_a x = \setabs{u}{\exists v.\element{v}{x} \wedge u
  R_a v}$.  Here $u$ and $v$ are guarded strings, while $x$ is an element of $K$.

\newlength{\dhatheight}
\newcommand{\doublehat}[1]{%
    \settoheight{\dhatheight}{\ensuremath{\hat{#1}}}%
    \addtolength{\dhatheight}{-0.35ex}%
    \hat{\vphantom{\rule{1pt}{\dhatheight}}%
      \smash{\hat{#1}}}}

It remains to define the Kripke relation on guarded strings from an agent specification
as in Figure 1. An an agent specification pairs each bare event with a set of bare events,
and so determines a relation between bare events, call it relation $R_a$ for an
agent $e$.  This determines
a relation $\hat{R_a}$ between decorated events, see \ex{Rhat}.
It is generalized to a relation $\doublehat{R}$ on arbitrary guarded strings by
relational Kleene star, as defined in \ex{Rhathat}. The operation $\cdot$ in the
definition is KAT product, which enforces matching of tests.
The number of terms $n$ is construed as satisfying $n\geq 1$.
{\em What is desired for $n=0$?}

\begin{example}   \label{Rhat}
\hspace{-2mm} \(
\hat{R} =_{\mbox{\footnotesize{def.}}} 
\{\tuple{u,v} | \exists c \exists d. cRd \wedge \element{u}{\hat{c}}  \wedge \element{v}{\hat{d}} \, \} \)
\end{example}

\begin{example}   \label{Rhathat}
\hspace{-2mm}\(
\doublehat{R} =_{\mbox{\footnotesize{def.}}} 
\hspace{-2mm}\begin{array}[t]{l}
\{ \tuple{x,y}| \exists u_1 ... \exists u_n \exists v_1 ... \exists v_n. \\
\hspace{6mm} u_1\hat{R}v_1 \wedge ... \wedge u_n\hat{R}v_n \wedge\\
\hspace{6mm} x = u_1 \cdot ... \cdot u_n \hspace{1mm} \wedge \\
\hspace{6mm} y = v_1 \cdot ... \cdot v_n \hspace{2mm} \}
\end{array}
\)
\end{example}

This defines an epistemic alternative to world $x$ to be a world of
the same length, where each component event in the alternative is an
event-alternative to the event in corresponding position in the base
world. Fusion product enforces preconditions and a correspondence
between pre-states and post-states of events on both sides of the
epistemic alternative relation.  This provides for finitely
specifiable construction of epistemic models that reflect intuitions
about information exchange and epistemic consequences of perceptual
events.  See Section 6 for linguistic examples.  Since an epistemic
alternative has the same lenght as its base world, it follows from the
construction that agents know how many events have transpired in their
base worlds.

Universal box modalites are defined as duals, $\Box_a x = (\Diamond_a
x^c)^c$.  For instance, Aly is certain that the coin is heads if and
only if she does not consider it possible that it is not heads.

Summing up, given an Epik program with $n$ agents, we construct a concrete guarded
string model $\tuple{K,+,\cdot,*,\bar{},0,1,{}^c,\Diamond_1,...,\Diamond_n}$.  The
elements of $K$ are interpreted as propositions.  $0^c$ is the set of worlds, and it
may be countably infinite.  $\Diamond_i$ is an epistemic modality for the $i$th
agent.  Or referring to the Kripke relations $\doublehat{R}_i$, the construction
defines a multi-agent Kripke frame $\tuple{0^c,\doublehat{R}_1,...,\doublehat{R}_n}$
(usually a countable one) from an Epik specification. The frame consists of a set of
worlds, and an epistemic-alternative relation for each agent.  These models are used
as a target for natural-language interpretation in Section 5 and Section 6, where we
obtain semantic values such as \sem{Amy knows that it's heads and Bob knows that Amy
  knows whether it is heads or tails, and does not know that it's heads} as elements
of $K$. Concretely the proposositions are sets of guarded strings (usually countable
ones), construed as sets of worlds as they figure in possible worlds semantics for
natural language.

% $\tuple{K,+,\cdot,*,\bar{},0,1,{}^c,\Diamond_1,...,\Diamond_n}$

\section{Translation into the finite state calculus}
The finite state calculus is an algebra of regular sets of strings and regular
relations between strings that was designed for use in computationonal phonology and
morphographemics (cite Karttunen etc.)  Current implementations allow for the
definition of functions with the status of defined operators on regular sets and
relations.  Such definitions are used to define an embedding of epistemic KAT in a
string algebra. The methodology follows Section 2 closely.  Let ${\cal K}$ be an
epistemic algebra as described in Section 2.  A given element of ${\cal K}$ is
represented in the the string algebra by the very same set of strings, i.e. by a set
of strings that have the form of a sequence of bare event symbols, with interleaved
Boolean vectors. Product in the KAT can not be modeled as concatenation in the string
algebra, because this would not enforce identity of states, and would result in
lengthening Boolean vectors at the concatenation point.  Instead, KAT product and KAT
Kleene star are defined operations in the string algebra, see Figure 2.  The
operations concatenate in the string algebra, delete strings with non-matching tests,
and then delete the second of two tests create a well-formed guarded string.

\begin{figure}
\begin{hangingpar}
{\tt St} \hspace{2mm} Tests such as 0 1 1 0. The length is the number of generators.
\end{hangingpar}

\begin{hangingpar}
{\tt UnequalStPair} \hspace{2mm} Sequence of two unequal tests such as 0 1 1 0 0 1 1 1,
    differing in one or more positions.
\end{hangingpar}

\begin{hangingpar}
\begin{verbatim}
define Wf0 ~[$ UnequalStPair];
\end{verbatim}
String that doesn't contain a non-matching test pair.
\end{hangingpar}

\begin{hangingpar}
\begin{verbatim}
define Squash St -> 0 || St _;
\end{verbatim}
Rewrite relation deleting the second of two tests.
\end{hangingpar}

\begin{hangingpar}
\begin{verbatim}
define Cn(X,Y) 
 [[[X Y] & Wf0] .o. Squash].l;
\end{verbatim}
KAT product in Fst, where {\tt \&} is intersection, {\tt .o.} is relation composition,
and {\tt .l} is relation image.
\end{hangingpar}

\begin{hangingpar}
\begin{verbatim}
define Kpl(X) 
 [[[X+] & Wf0] .o. Squash].l;
\end{verbatim}
\end{hangingpar}

\begin{hangingpar}
\begin{verbatim}
define Kst(X) St |  Kpl(X);
\end{verbatim}
KAT Kleene plus and Kleene star in Fst. The Fst operation {\tt |} is union.
\end{hangingpar}

\caption{Translation into Fst of KAT product and KAT Kleene star.}
\end{figure}

A given bare event such as $a_1$ (Aly looks at heads) is in the KAT algebra a set of
bare events decorated with compatible tests on each side, semantically $\{ 10 a_1 10
\}$ in this case.  This is a unit set rather than a guarded string, because elements
of the KAT algebra are sets. Worlds in the KAT algebra are defined by sequencing
events using {\em Kst}. The operation enforces compatibility of states, so that
$(a_1 + a_0)(b_1 + b_0)$ contains two worlds rather than four.  The program in
Figure 1 as interpreted in FST defines a countably infinite set of possible worlds by
KAT Kleene closure as $\mbox{\em Kst}(a_1 + a_0 +b_1 + b_0)$, and an algebra of
propositions as regular sets of strings drawn from this space of worlds.

It remains to define an epistemic alternative relation on worlds for each agent.  The
relevant information in Figure 1 is a relation between bare events for each
agent. This determines a relation in the guarded string algebra a relation between
bare events decorated with compatible tests.  For agent Aly, this is the relation
described in \ex{alyrel1} as a set of ordered pairs.

\vspace{-3mm}
\begin{example} \label{alyrel1}
 \(
 \left\{
 \begin{array}{l}
   \pair{10a_110}{10a_110}, \\  \pair{01a_001}{01a_001}, \\
   \pair{10b_110}{10b_110}, \\ \pair{10b_110}{01b_001}, \\
   \pair{01b_001}{10b_110}, \\ \pair{01b_001}{01b_001} \\
 \end{array}
 \right\}
 \) 
\end{example}

% Fixes junky vertical spacing of the set above.
% Possibly remove it in the final version.
\raggedbottom

The relation on decorated events needs to be generalized to a relation of worlds.
The principle for this is that an epistemic alternative to a world of the form $we$
is a world of the form $vd$, where $v$ is a world-alternative to $w$, $d$ is an
event-alternative to $e$, and $vd$ is defined (i.e.  the world alternative $v$
satisfies the pre-conditions of the event alternative $d$).  This principle is found
in earlier literature (Moore 198x, Baltag, Moss and Solecki 20xx).  In the
construction in Fst, the definition of world alternatives takes a simple form. Where
$R_a$ is the relation on decorated events for agent $a$, the the corresponding
relation on worlds in is the Kleene closure of $R_a$.  Where $R$ and $S$ are
relations, the concatenation product of $R$ and $S$ is the set of pairs of the form
$\pair{x_1x_2}{y_1y2}$, where $\pair{x_1}{y_1}$ is in relation $R$, and
$\pair{x_2}{y_2}$ is in relation $S$.  The Kleene closure of relation $R$ is
\(\cup_{n \geq 0}R^n\), where $R^n$ is the $n$-times concatenation product of $R$
with itself. This is an operation in the finite state calculus.  Figure
\ref{relclosure} defines the corresponding operation in the guarded string algebra.
The epistemic alternative relation on worlds for an agent is then defined
as the concatenation closure of the event alternative relation for the agent.

%(Note that in both the string algebra and the guarded string algebra, the operation
%on relations is concatenation closure, not reflexive transitive closure.)

\begin{figure} \label{relclosure}
{\raggedright
\begin{hangingpar}
{\tt{define RelKpl(R)}}
$\underbrace{\mbox{\tt{Squash.i\hspace{1pt}.o.}}}_c$
$\underbrace{\mbox{\tt Wf0\hspace{1pt}.o.}}_b$
$\underbrace{\mbox{\tt{[R+]}}}_a$
$\underbrace{\mbox{\tt .o.\hspace{1pt}Wf0}}_b$
$\underbrace{\mbox{\tt{.o.\hspace{1pt}Squash}}}_c$
\end{hangingpar}}

\begin{tabular}{ll}
  $a$ & Relational Kleene plus in the string algebra \\
  $b$ & Constrain domain and co-domain to contain \\
      & no unmatched tests. \\
  $c$ & Reduce doubled tests to a single \\
      & test in the domain and co-domain. \\
      & {\tt{Squash.i}} is the inverse \\
      & of {\tt{Squash}}.
\end{tabular}

\begin{hangingpar}
{\tt{define Kst(R) [St\hspace{1pt}.x.\hspace{1pt}St]\hspace{1pt}|\hspace{1pt}Kpl(X);}}
The Fst operation {\tt{.x.}} is Cartesian product.
\end{hangingpar}

\caption{Definition in Fst of the Kleene concatenation closure of
a relation between guarded strings.}
\end{figure}

Other operations in the guarded string algebra as defined in FST are simpler.
Union is union in the string algebra. The complement of a proposition
is complement relative to the set of worlds, as defined by
set difference in the string algebra.

\section{Interpretation using lazy lists of strings}
%\pagebreak
%Hi Mom!

%\pagebreak

%Hi Sis!
%\pagebreak

\section{Syntax-semantics interface}
An architecture of interpretation by transtlation is employed, where English
sentences are mapped to terms in a logical language via an interpreted grammar, and
these terms are intreted in a guarded string model as propositions. For the latter,
there are options of translation into the finite state calculus, and representation
in Haskell via lazy lists of guarded strings.

The grammar is a semantically interpreted multi-modal categorial grammar, consisting
of a lexicon of words, their categorial types, and interpretations in a logical
lambda language.  \ex{phenomena1} lists and exemplifies phenomena that are covered.

\begin{example}   \label{phenomena1}
{\em Phenomena}

\vspace{1mm}
\hspace{-10mm}\begin{tabular}[t]{l|l}
  Basic statives & It's heads. \\
  & It's tails.\\ \hline
That-complement & Amy knows \\ & \hspace{2mm} that it's heads. \\ \hline
Wh-complement & Amy knows \\ & \hspace{2mm} whether its heads. \\ \hline
Negation & Bob doesn't know \\ & \hspace{2mm} that it isn't heads. \\ \hline 
Tensed and base &  Bob knows \\
verbal forms & \hspace{2mm} that it's heads.      \\ 
\hspace{1mm} & Bob doesn't know \\ & \hspace{2mm} that it's heads.\\ \hline
Sentence conjunction & It's heads and Bob \\ & \hspace{2mm} doesn't know \\ & \hspace{2mm} that it's heads \\ \hline
Predicate conjunction & Bob knows that \\ & Amy knows whether \\ & it's heads and
\\ & doesn't know that \\ & Amy knows that \\ & it's heads.
\end{tabular}    
\end{example}

As illustrated towards the end, there is recursion through conjuction and verbal complementation,
so that the language is infinite, and includes talk of beliefs about beliefs, or in general, talk
of arbitrarily iterated belief.

\ex{lexicon1} gives illustrative lexical entries. The grammar and semantics are in
certain way optimized for the simple fragment of English concerned with clausal
complementation.  The agent names {\em Amy} and {\em Bob} contribute the epistemic
alternative relations for those agents, rather than individuals.  This is possible
because the agents are never arguments of extensional predicates, so what matters
about the agents is their epistemic alternative relations.  The root verb {\em know}
contributes existential modal force.  The complementizers {\em that} and {\em
  whether} are the heads of their dominating clauses, and assemble an alternative
relation, modal force, and proposition contributed by the complement, in part by
introducing the dual of the existential modal force to obtain universal modal force.
These moves are bizarre, but are offered here as a way of constructing a compact
interpreted grammar. They can easily be reformulated with more complicated
constructions that are better motivated in more comprehensive interpreted grammar of
English.

\begin{example}   \label{lexicon1}
{\em Lexicon (partial)}

\hspace{-10mm}\begin{tabular}[t]{l|c|c}
Amy & $e$ & $\doublehat{R}_a$\\
Bob & $e$ & $\doublehat{R}_b$\\ 
heads & $d \backslash_D t$ & $W{\cdot}h$ \\
it's & $t\slash(d \backslash_D t)$ & $\lambda p.p$ \\
\end{tabular}
\end{example}

\section{Examples and discussion}

Page breakdown

\begin{tabular}{lll}
  1,2 & 3.75 \\
  3   & 1.25 & FST translation\\
  4   & 1.25 & Haskell Epik \\
  5   & 0.75 & English CG fragment \\
  6   & 1.0  & Examples and discussion\\
\end{tabular}

\end{document}

\section{Introduction}

The following instructions are directed to authors of papers submitted to ACL 2020 or accepted for publication in its proceedings.
All authors are required to adhere to these specifications.
Authors are required to provide a Portable Document Format (PDF) version of their papers.
\textbf{The proceedings are designed for printing on A4 paper.}

\section{Electronically-available resources}

ACL provides this description and accompanying style files at
\begin{quote}
\url{http://acl2020.org/downloads/acl2020-templates.zip}
\end{quote}
We strongly recommend the use of these style files, which have been appropriately tailored for the ACL 2020 proceedings.

\paragraph{\LaTeX-specific details:}
The templates include the \LaTeX2e{} source (\texttt{\small acl2020.tex}),
the \LaTeX2e{} style file used to format it (\texttt{\small acl2020.sty}),
an ACL bibliography style (\texttt{\small acl\_natbib.bst}),
an example bibliography (\texttt{\small acl2020.bib}),
and the bibliography for the ACL Anthology (\texttt{\small anthology.bib}).

\section{Length of Submission}
\label{sec:length}

The conference accepts submissions of long papers and short papers.
Long papers may consist of up to eight (8) pages of content plus unlimited pages for references.
Upon acceptance, final versions of long papers will be given one additional page -- up to nine (9) pages of content plus unlimited pages for references -- so that reviewers' comments can be taken into account.
Short papers may consist of up to four (4) pages of content, plus unlimited pages for references.
Upon acceptance, short papers will be given five (5) pages in the proceedings and unlimited pages for references. 
For both long and short papers, all illustrations and tables that are part of the main text must be accommodated within these page limits, observing the formatting instructions given in the present document.
Papers that do not conform to the specified length and formatting requirements are subject to be rejected without review.

The conference encourages the submission of additional material that is relevant to the reviewers but not an integral part of the paper.
There are two such types of material: appendices, which can be read, and non-readable supplementary materials, often data or code.
Additional material must be submitted as separate files, and must adhere to the same anonymity guidelines as the main paper.
The paper must be self-contained: it is optional for reviewers to look at the supplementary material.
Papers should not refer, for further detail, to documents, code or data resources that are not available to the reviewers.
Refer to Appendices~\ref{sec:appendix} and \ref{sec:supplemental} for further information.

Workshop chairs may have different rules for allowed length and whether supplemental material is welcome.
As always, the respective call for papers is the authoritative source.

\section{Anonymity}
As reviewing will be double-blind, papers submitted for review should not include any author information (such as names or affiliations). Furthermore, self-references that reveal the author's identity, \emph{e.g.},
\begin{quote}
We previously showed \citep{Gusfield:97} \ldots
\end{quote}
should be avoided. Instead, use citations such as 
\begin{quote}
\citet{Gusfield:97} previously showed\ldots
\end{quote}
Please do not use anonymous citations and do not include acknowledgements.
\textbf{Papers that do not conform to these requirements may be rejected without review.}

Any preliminary non-archival versions of submitted papers should be listed in the submission form but not in the review version of the paper.
Reviewers are generally aware that authors may present preliminary versions of their work in other venues, but will not be provided the list of previous presentations from the submission form.

Once a paper has been accepted to the conference, the camera-ready version of the paper should include the author's names and affiliations, and is allowed to use self-references.

\paragraph{\LaTeX-specific details:}
For an anonymized submission, ensure that {\small\verb|\aclfinalcopy|} at the top of this document is commented out, and that you have filled in the paper ID number (assigned during the submission process on softconf) where {\small\verb|***|} appears in the {\small\verb|\def\aclpaperid{***}|} definition at the top of this document.
For a camera-ready submission, ensure that {\small\verb|\aclfinalcopy|} at the top of this document is not commented out.

\section{Multiple Submission Policy}
Papers that have been or will be submitted to other meetings or publications must indicate this at submission time in the START submission form, and must be withdrawn from the other venues if accepted by ACL 2020. Authors of papers accepted for presentation at ACL 2020 must notify the program chairs by the camera-ready deadline as to whether the paper will be presented. We will not accept for publication or presentation the papers that overlap significantly in content or results with papers that will be (or have been) published elsewhere.

Authors submitting more than one paper to ACL 2020 must ensure that submissions do not overlap significantly (>25\%) with each other in content or results.

\section{Formatting Instructions}

Manuscripts must be in two-column format.
Exceptions to the two-column format include the title, authors' names and complete addresses, which must be centered at the top of the first page, and any full-width figures or tables (see the guidelines in Section~\ref{ssec:title-authors}).
\textbf{Type single-spaced.}
Start all pages directly under the top margin.
The manuscript should be printed single-sided and its length should not exceed the maximum page limit described in Section~\ref{sec:length}.
Pages should be numbered in the version submitted for review, but \textbf{pages should not be numbered in the camera-ready version}.

\paragraph{\LaTeX-specific details:}
The style files will generate page numbers when {\small\verb|\aclfinalcopy|} is commented out, and remove them otherwise.

\subsection{File Format}
\label{sect:pdf}

For the production of the electronic manuscript you must use Adobe's Portable Document Format (PDF).
Please make sure that your PDF file includes all the necessary fonts (especially tree diagrams, symbols, and fonts with Asian characters).
When you print or create the PDF file, there is usually an option in your printer setup to include none, all or just non-standard fonts.
Please make sure that you select the option of including ALL the fonts.
\textbf{Before sending it, test your PDF by printing it from a computer different from the one where it was created.}
Moreover, some word processors may generate very large PDF files, where each page is rendered as an image.
Such images may reproduce poorly.
In this case, try alternative ways to obtain the PDF.
One way on some systems is to install a driver for a postscript printer, send your document to the printer specifying ``Output to a file'', then convert the file to PDF.

It is of utmost importance to specify the \textbf{A4 format} (21 cm x 29.7 cm) when formatting the paper.
Print-outs of the PDF file on A4 paper should be identical to the hardcopy version.
If you cannot meet the above requirements about the production of your electronic submission, please contact the publication chairs as soon as possible.

\paragraph{\LaTeX-specific details:}
PDF files are usually produced from \LaTeX{} using the \texttt{\small pdflatex} command.
If your version of \LaTeX{} produces Postscript files, \texttt{\small ps2pdf} or \texttt{\small dvipdf} can convert these to PDF.
To ensure A4 format in \LaTeX, use the command {\small\verb|\special{papersize=210mm,297mm}|}
in the \LaTeX{} preamble (below the {\small\verb|\usepackage|} commands) and use \texttt{\small dvipdf} and/or \texttt{\small pdflatex}; or specify \texttt{\small -t a4} when working with \texttt{\small dvips}.

\subsection{Layout}
\label{ssec:layout}

Format manuscripts two columns to a page, in the manner these
instructions are formatted.
The exact dimensions for a page on A4 paper are:

\begin{itemize}
\item Left and right margins: 2.5 cm
\item Top margin: 2.5 cm
\item Bottom margin: 2.5 cm
\item Column width: 7.7 cm
\item Column height: 24.7 cm
\item Gap between columns: 0.6 cm
\end{itemize}

\noindent Papers should not be submitted on any other paper size.
If you cannot meet the above requirements about the production of your electronic submission, please contact the publication chairs above as soon as possible.

\subsection{Fonts}

For reasons of uniformity, Adobe's \textbf{Times Roman} font should be used.
If Times Roman is unavailable, you may use Times New Roman or \textbf{Computer Modern Roman}.

Table~\ref{font-table} specifies what font sizes and styles must be used for each type of text in the manuscript.

\begin{table}
\centering
\begin{tabular}{lrl}
\hline \textbf{Type of Text} & \textbf{Font Size} & \textbf{Style} \\ \hline
paper title & 15 pt & bold \\
author names & 12 pt & bold \\
author affiliation & 12 pt & \\
the word ``Abstract'' & 12 pt & bold \\
section titles & 12 pt & bold \\
subsection titles & 11 pt & bold \\
document text & 11 pt  &\\
captions & 10 pt & \\
abstract text & 10 pt & \\
bibliography & 10 pt & \\
footnotes & 9 pt & \\
\hline
\end{tabular}
\caption{\label{font-table} Font guide. }
\end{table}

\paragraph{\LaTeX-specific details:}
To use Times Roman in \LaTeX2e{}, put the following in the preamble:
\begin{quote}
\small
\begin{verbatim}
\usepackage{times}
\usepackage{latexsym}
\end{verbatim}
\end{quote}

\subsection{Ruler}
A printed ruler (line numbers in the left and right margins of the article) should be presented in the version submitted for review, so that reviewers may comment on particular lines in the paper without circumlocution.
The presence or absence of the ruler should not change the appearance of any other content on the page.
The camera ready copy should not contain a ruler.

\paragraph{Reviewers:}
note that the ruler measurements may not align well with lines in the paper -- this turns out to be very difficult to do well when the paper contains many figures and equations, and, when done, looks ugly.
In most cases one would expect that the approximate location will be adequate, although you can also use fractional references (\emph{e.g.}, this line ends at mark $295.5$).

\paragraph{\LaTeX-specific details:}
The style files will generate the ruler when {\small\verb|\aclfinalcopy|} is commented out, and remove it otherwise.

\subsection{Title and Authors}
\label{ssec:title-authors}

Center the title, author's name(s) and affiliation(s) across both columns.
Do not use footnotes for affiliations.
Place the title centered at the top of the first page, in a 15-point bold font.
Long titles should be typed on two lines without a blank line intervening.
Put the title 2.5 cm from the top of the page, followed by a blank line, then the author's names(s), and the affiliation on the following line.
Do not use only initials for given names (middle initials are allowed).
Do not format surnames in all capitals (\emph{e.g.}, use ``Mitchell'' not ``MITCHELL'').
Do not format title and section headings in all capitals except for proper names (such as ``BLEU'') that are
conventionally in all capitals.
The affiliation should contain the author's complete address, and if possible, an electronic mail address.

The title, author names and addresses should be completely identical to those entered to the electronical paper submission website in order to maintain the consistency of author information among all publications of the conference.
If they are different, the publication chairs may resolve the difference without consulting with you; so it is in your own interest to double-check that the information is consistent.

Start the body of the first page 7.5 cm from the top of the page.
\textbf{Even in the anonymous version of the paper, you should maintain space for names and addresses so that they will fit in the final (accepted) version.}

\subsection{Abstract}
Use two-column format when you begin the abstract.
Type the abstract at the beginning of the first column.
The width of the abstract text should be smaller than the
width of the columns for the text in the body of the paper by 0.6 cm on each side.
Center the word \textbf{Abstract} in a 12 point bold font above the body of the abstract.
The abstract should be a concise summary of the general thesis and conclusions of the paper.
It should be no longer than 200 words.
The abstract text should be in 10 point font.

\subsection{Text}
Begin typing the main body of the text immediately after the abstract, observing the two-column format as shown in the present document.

Indent 0.4 cm when starting a new paragraph.

\subsection{Sections}

Format section and subsection headings in the style shown on the present document.
Use numbered sections (Arabic numerals) to facilitate cross references.
Number subsections with the section number and the subsection number separated by a dot, in Arabic numerals.

\subsection{Footnotes}
Put footnotes at the bottom of the page and use 9 point font.
They may be numbered or referred to by asterisks or other symbols.\footnote{This is how a footnote should appear.}
Footnotes should be separated from the text by a line.\footnote{Note the line separating the footnotes from the text.}

\subsection{Graphics}

Place figures, tables, and photographs in the paper near where they are first discussed, rather than at the end, if possible.
Wide illustrations may run across both columns.
Color is allowed, but adhere to Section~\ref{ssec:accessibility}'s guidelines on accessibility.

\paragraph{Captions:}
Provide a caption for every illustration; number each one sequentially in the form:
``Figure 1. Caption of the Figure.''
``Table 1. Caption of the Table.''
Type the captions of the figures and tables below the body, using 10 point text.
Captions should be placed below illustrations.
Captions that are one line are centered (see Table~\ref{font-table}).
Captions longer than one line are left-aligned (see Table~\ref{tab:accents}).

\begin{table}
\centering
\begin{tabular}{lc}
\hline
\textbf{Command} & \textbf{Output}\\
\hline
\verb|{\"a}| & {\"a} \\
\verb|{\^e}| & {\^e} \\
\verb|{\`i}| & {\`i} \\ 
\verb|{\.I}| & {\.I} \\ 
\verb|{\o}| & {\o} \\
\verb|{\'u}| & {\'u}  \\ 
\verb|{\aa}| & {\aa}  \\\hline
\end{tabular}
\begin{tabular}{lc}
\hline
\textbf{Command} & \textbf{Output}\\
\hline
\verb|{\c c}| & {\c c} \\ 
\verb|{\u g}| & {\u g} \\ 
\verb|{\l}| & {\l} \\ 
\verb|{\~n}| & {\~n} \\ 
\verb|{\H o}| & {\H o} \\ 
\verb|{\v r}| & {\v r} \\ 
\verb|{\ss}| & {\ss} \\
\hline
\end{tabular}
\caption{Example commands for accented characters, to be used in, \emph{e.g.}, \BibTeX\ names.}\label{tab:accents}
\end{table}

\paragraph{\LaTeX-specific details:}
The style files are compatible with the caption and subcaption packages; do not add optional arguments.
\textbf{Do not override the default caption sizes.}

\subsection{Hyperlinks}
Within-document and external hyperlinks are indicated with Dark Blue text, Color Hex \#000099.

\subsection{Citations}
Citations within the text appear in parentheses as~\citep{Gusfield:97} or, if the author's name appears in the text itself, as \citet{Gusfield:97}.
Append lowercase letters to the year in cases of ambiguities.  
Treat double authors as in~\citep{Aho:72}, but write as in~\citep{Chandra:81} when more than two authors are involved. Collapse multiple citations as in~\citep{Gusfield:97,Aho:72}.

Refrain from using full citations as sentence constituents.
Instead of
\begin{quote}
  ``\citep{Gusfield:97} showed that ...''
\end{quote}
write
\begin{quote}
``\citet{Gusfield:97} showed that ...''
\end{quote}

\begin{table*}
\centering
\begin{tabular}{lll}
\hline
\textbf{Output} & \textbf{natbib command} & \textbf{Old ACL-style command}\\
\hline
\citep{Gusfield:97} & \small\verb|\citep| & \small\verb|\cite| \\
\citealp{Gusfield:97} & \small\verb|\citealp| & no equivalent \\
\citet{Gusfield:97} & \small\verb|\citet| & \small\verb|\newcite| \\
\citeyearpar{Gusfield:97} & \small\verb|\citeyearpar| & \small\verb|\shortcite| \\
\hline
\end{tabular}
\caption{\label{citation-guide}
Citation commands supported by the style file.
The style is based on the natbib package and supports all natbib citation commands.
It also supports commands defined in previous ACL style files for compatibility.
}
\end{table*}

\paragraph{\LaTeX-specific details:}
Table~\ref{citation-guide} shows the syntax supported by the style files.
We encourage you to use the natbib styles.
You can use the command {\small\verb|\citet|} (cite in text) to get ``author (year)'' citations as in \citet{Gusfield:97}.
You can use the command {\small\verb|\citep|} (cite in parentheses) to get ``(author, year)'' citations as in \citep{Gusfield:97}.
You can use the command {\small\verb|\citealp|} (alternative cite without  parentheses) to get ``author year'' citations (which is useful for  using citations within parentheses, as in \citealp{Gusfield:97}).

\subsection{References}
Gather the full set of references together under the heading \textbf{References}; place the section before any Appendices. 
Arrange the references alphabetically by first author, rather than by order of occurrence in the text.

Provide as complete a citation as possible, using a consistent format, such as the one for \emph{Computational Linguistics\/} or the one in the  \emph{Publication Manual of the American 
Psychological Association\/}~\citep{APA:83}.
Use full names for authors, not just initials.

Submissions should accurately reference prior and related work, including code and data.
If a piece of prior work appeared in multiple venues, the version that appeared in a refereed, archival venue should be referenced.
If multiple versions of a piece of prior work exist, the one used by the authors should be referenced.
Authors should not rely on automated citation indices to provide accurate references for prior and related work.

The following text cites various types of articles so that the references section of the present document will include them.
\begin{itemize}
\item Example article in journal: \citep{Ando2005}.
\item Example article in proceedings, with location: \citep{borschinger-johnson-2011-particle}.
\item Example article in proceedings, without location: \citep{andrew2007scalable}.
\item Example arxiv paper: \citep{rasooli-tetrault-2015}. 
\end{itemize}

\paragraph{\LaTeX-specific details:}
The \LaTeX{} and Bib\TeX{} style files provided roughly follow the American Psychological Association format.
If your own bib file is named \texttt{\small acl2020.bib}, then placing the following before any appendices in your \LaTeX{}  file will generate the references section for you:
\begin{quote}\small
\verb|\bibliographystyle{acl_natbib}|\\
\verb|\bibliography{acl2020}|
\end{quote}

You can obtain the complete ACL Anthology as a Bib\TeX\ file from \url{https://aclweb.org/anthology/anthology.bib.gz}.
To include both the anthology and your own bib file, use the following instead of the above.
\begin{quote}\small
\verb|\bibliographystyle{acl_natbib}|\\
\verb|\bibliography{anthology,acl2020}|
\end{quote}

\subsection{Digital Object Identifiers}
As part of our work to make ACL materials more widely used and cited outside of our discipline, ACL has registered as a CrossRef member, as a registrant of Digital Object Identifiers (DOIs), the standard for registering permanent URNs for referencing scholarly materials.

All camera-ready references are required to contain the appropriate DOIs (or as a second resort, the hyperlinked ACL Anthology Identifier) to all cited works.
Appropriate records should be found for most materials in the current ACL Anthology at \url{http://aclanthology.info/}.
As examples, we cite \citep{goodman-etal-2016-noise} to show you how papers with a DOI will appear in the bibliography.
We cite \citep{harper-2014-learning} to show how papers without a DOI but with an ACL Anthology Identifier will appear in the bibliography.

\paragraph{\LaTeX-specific details:}
Please ensure that you use Bib\TeX\ records that contain DOI or URLs for any of the ACL materials that you reference.
If the Bib\TeX{} file contains DOI fields, the paper title in the references section will appear as a hyperlink to the DOI, using the hyperref \LaTeX{} package.

\subsection{Appendices}
Appendices, if any, directly follow the text and the
references (but only in the camera-ready; see Appendix~\ref{sec:appendix}).
Letter them in sequence and provide an informative title:
\textbf{Appendix A. Title of Appendix}.

\section{Accessibility}
\label{ssec:accessibility}

In an effort to accommodate people who are color-blind (as well as those printing to paper), grayscale readability is strongly encouraged.
Color is not forbidden, but authors should ensure that tables and figures do not rely solely on color to convey critical distinctions.
A simple criterion:
All curves and points in your figures should be clearly distinguishable without color.

\section{Translation of non-English Terms}

It is also advised to supplement non-English characters and terms with appropriate transliterations and/or translations since not all readers understand all such characters and terms.
Inline transliteration or translation can be represented in the order of:
\begin{center}
\begin{tabular}{c}
original-form \\
transliteration \\
``translation''
\end{tabular}
\end{center}

\section{\LaTeX{} Compilation Issues}
You may encounter the following error during compilation: 
\begin{quote}
{\small\verb|\pdfendlink|} ended up in different nesting level than {\small\verb|\pdfstartlink|}.
\end{quote}
This happens when \texttt{\small pdflatex} is used and a citation splits across a page boundary.
To fix this, the style file contains a patch consisting of two lines:
(1) {\small\verb|\RequirePackage{etoolbox}|} (line 455 in \texttt{\small acl2020.sty}), and
(2) A long line below (line 456 in \texttt{\small acl2020.sty}).

If you still encounter compilation issues even with the patch enabled, disable the patch by commenting the two lines, and then disable the \texttt{\small hyperref} package by loading the style file with the \texttt{\small nohyperref} option:

\noindent
{\small\verb|\usepackage[nohyperref]{acl2020}|}

\noindent
Then recompile, find the problematic citation, and rewrite the sentence containing the citation. (See, {\em e.g.}, \url{http://tug.org/errors.html})

\section*{Acknowledgments}

The acknowledgments should go immediately before the references. Do not number the acknowledgments section.
Do not include this section when submitting your paper for review.

\bibliography{anthology,acl2020}
\bibliographystyle{acl_natbib}

\appendix

\section{Appendices}
\label{sec:appendix}
Appendices are material that can be read, and include lemmas, formulas, proofs, and tables that are not critical to the reading and understanding of the paper. 
Appendices should be \textbf{uploaded as supplementary material} when submitting the paper for review.
Upon acceptance, the appendices come after the references, as shown here.

\paragraph{\LaTeX-specific details:}
Use {\small\verb|\appendix|} before any appendix section to switch the section numbering over to letters.

\section{Supplemental Material}
\label{sec:supplemental}
Submissions may include non-readable supplementary material used in the work and described in the paper.
Any accompanying software and/or data should include licenses and documentation of research review as appropriate.
Supplementary material may report preprocessing decisions, model parameters, and other details necessary for the replication of the experiments reported in the paper.
Seemingly small preprocessing decisions can sometimes make a large difference in performance, so it is crucial to record such decisions to precisely characterize state-of-the-art methods.

Nonetheless, supplementary material should be supplementary (rather than central) to the paper.
\textbf{Submissions that misuse the supplementary material may be rejected without review.}
Supplementary material may include explanations or details of proofs or derivations that do not fit into the paper, lists of
features or feature templates, sample inputs and outputs for a system, pseudo-code or source code, and data.
(Source code and data should be separate uploads, rather than part of the paper).

The paper should not rely on the supplementary material: while the paper may refer to and cite the supplementary material and the supplementary material will be available to the reviewers, they will not be asked to review the supplementary material.

\end{document}

### 2020-06-15 tex/draft.tex
Section 5 in draft almost finished.

@@
 
-\title{Contstructive Epistemic Semantics in Guarded String Models}
+% This used to say Constructive ..., which is a better title.
+% Reserve that for a journal paper if we get to that.
+\title{Epistemic Semantics in Guarded String Models}
 
@@
 
+\pagebreak
 \section{Syntax-semantics interface}
 An architecture of interpretation by transtlation is employed, where English
-sentences are mapped to terms in a logical language via an interpreted grammar, and
-these terms are intreted in a guarded string model as propositions. For the latter,
-there are options of translation into the finite state calculus, and representation
-in Haskell via lazy lists of guarded strings.
-
-The grammar is a semantically interpreted multi-modal categorial grammar, consisting
-of a lexicon of words, their categorial types, and interpretations in a logical
-lambda language.  \ex{phenomena1} lists and exemplifies phenomena that are covered.
+sentences are mapped to terms in the logical language (--) via an interpreted
+grammar, and these terms are in turn interpreted as propositions (sets of possible
+worlds). For the latter, there are options of translation into the finite state
+calculus in order to represent propositions as finite state machines (Section 3), and
+representation in Haskell via lazy lists of guarded strings (Section 4).  The grammar
+is a semantically interpreted multi-modal categorial grammar, consisting of a lexicon
+of words, their categorial types, and interpretations in a logical lambda language.
+\ex{phenomena1} lists that are covered.
 
@@
 \ex{lexicon1} gives illustrative lexical entries. The grammar and semantics are in
-certain way optimized for the simple fragment of English concerned with clausal
+certain way optimized for a simple fragment of English concerned with clausal
 complementation.  The agent names {\em Amy} and {\em Bob} contribute the epistemic
@@
   whether} are the heads of their dominating clauses, and assemble an alternative
-relation, modal force, and proposition contributed by the complement, in part by
-introducing the dual of the existential modal force to obtain universal modal force.
-These moves are bizarre, but are offered here as a way of constructing a compact
-interpreted grammar. They can easily be reformulated with more complicated
-constructions that are better motivated in more comprehensive interpreted grammar of
-English.
+relation, modal force, and proposition contributed by the complement.  These
+complementizers introduce the dual via two negations, in order to arrive universal
+modal force.  These moves are offered here as a way of constructing a compact
+interpreted grammar. They can easily be reformulated in a more comprehensive
+interpreted grammar of English.
 
@@
 \hspace{-10mm}\begin{tabular}[t]{l|c|c}
-Amy & $e$ & $\doublehat{R}_a$\\
-Bob & $e$ & $\doublehat{R}_b$\\ 
-heads & $d \backslash_D t$ & $W{\cdot}h$ \\
-it's & $t\slash(d \backslash_D t)$ & $\lambda p.p$ \\
-\end{tabular}
-\end{example}
+Amy & $e$ & ${R}_a$\\
+Bob & $e$ & ${R}_b$\\
+it & $d$ & $d$ \\
+heads & $d \backslash_D t$ & $\lambda x.0^c{\cdot}h$ \\
+tails & $d \backslash_D t$ & $\lambda x.0^c{\cdot}!h$ \\
+is & $(d \backslash t)\slash(d \backslash_D t)$ & $\lambda P.\lambda x.P x$ \\
+knows & $(e \backslash t)\slash_M t$ &  $\lambda p.\lambda R. {\Diamond}R p$\\
+that  & \multicolumn{2}{l}{$(((e \backslash t)\slash_M t) \backslash (e \backslash t)) \slash t$} \\
+& \multicolumn{2}{l}{$\lambda p.\lambda m.\lambda R.{\sim}(m ({\sim}p) R)$} \\
+whether  & \multicolumn{2}{l}{$(((e \backslash t)\slash_M t) \backslash (e \backslash t)) \slash t$} \\
+& \multicolumn{2}{l}{$\lambda p.\lambda m.\lambda R.{\sim}(m ({\sim}p) R)$} \\
+& \multicolumn{2}{r}{$+ {\sim}(m p R)$} \\
+\end{tabular}
+\end{example}
+
+Multimodal categories such as $\backslash_D$ and $\backslash_M$ are used to control
+the derivation.  For instance the category of {\em heads} $d \backslash_D t$.  The
+dummy expletive subject {\em it} has category $d$, but the phrase {\em it heads} of
+category $t$ can not be formed, because $\backslash_D$ is not syntactically active as
+a function.  Instead {\em it is heads} can be formed with a predicator {\em is} of
+category $(d \backslash t)\slash(d \backslash_D t)$.  (This uses Lambek/Bar-Hillel
+notation for slashes, so that $(d \backslash t)\slash(d \backslash_D t)$ combines
+with $d \backslash_D t$ on the right to give a value that combines with $d$ on the
+left to give $t$.)  Similarly {\em knows} has a category with the top-level slash
+$\slash_M$, and combines to form a sentence as an argument of {\em that} or {\em
+  whether}, which has a category that looks for the category of {\em know} on the
+left, after combining with a complement sentence on the right.
+
+The semantic translations in the third column of \ex{lexicon1} use the Epik term
+language, incremented with lambda.  The body of $\lambda x.0^c{\cdot}h$, which is the
+semantic lexical entry for {\em heads}, is a term denoting the set of all worlds
+where the coin is heads, represented as the set of all guarded strings that end with
+a Boolean valuation where the primitive proposition $h$ (it's heads) is true.  There
+is $\lambda x$ at the front because of a correspondence the grammar formalism uses a
+correspondence between syntactic and semantic types. However, it does not bind
+anything, because sentences such as {\em it isn't heads} have an expletive subject.
+The body ${\Diamond}R p$ of $\lambda p.\lambda R. {\Diamond}R p$, which is the
+semantic lexical entry of {\em knows}, is an Epik term denoting the pre-image of the
+set of worlds $p$ according to the relation $\doublehat{R}$ between guarded strings
+that is determined by the event-level relation $R$.  This is not the right semantics
+for {\em Amy knows that it's heads}, because it is an existential modality ${\Diamond_R}p$,
+rather than an universal modality ${\Box_R}p$.  This is corrected by the complementizer
+{\em that} or {\em whether}, which introduces the dual.
+
+Sentences are parsed with a chart parser for categorial grammar.  The semantics for
+complex phrases are obtained by syntactic application of semantic translations,
+accompanied by beta reduction.  Semantic terms in the parsing formalism are
+expressions of untyped lambda calculus.  The grammar is set up so that lambda is
+eliminated by beta reduction in the semantic term corresponding to a sentence. In
+consequence, the semantic term translating a sentence is a term of the Epik language
+(--).  Such a term designates a set of possible words (guarded strings) in the
+possible worlds model deterimened by an Epik specification such as the one in Figure
+1.  \exi{three}{a} is an English sentence with predicate conjunction and three levels
+of clausal embedding.  Using the grammar and parser, the sentence is mapped to the
+Epik term \exi{b}.  Using the result from Section 3, this term can be mapped in an
+implementation of the finite state calculus to a finite state machine that represents
+a countably infinite set of possible worlds, represented as guarded strings.  Using
+the result from Section 4, it can be mapped to an infinite lazy list of guarded
+strings, representing the same set of possible worlds.  Either of these is a concrete
+computational representation of the propositional semantic value \semval{Amy knows
+  that Bob knows that Amy knows whether it is heads and knows that Bob does not know
+  that Amy knows that it is tails}, in the familiar sense of Montague semantics for
+natural language.
+
+\begin{examples}  \label{three}
+\itema  Amy knows that Bob knows that Amy knows whether it is heads and
+knows that Bob does not know that Amy knows that it is tails.
+\itemb
+\end{examples}
 

### 2020-06-16 tex/draft.tex
Figure with term syntax in draft.tex

@@
 
+This scheme provides for an interpretation of the language of Epik
+terms that is defined in Figure 4.  An Epik specifications such as
+Figure 1 is translated to a straight-line program of the finite state
+calculus that defines constants and functions, including the ones from
+Figures 2 and 3.  In Xfst or Foma, which are interpreters for the
+finite state calculus, the program is read, and then propositional
+terms can be mapped to finite state machines that represent sets of
+guarded strings, interpreted as propositions.
+
 \section{Interpretation using lazy lists of strings}
@@
 
-\pagebreak
 \section{Syntax-semantics interface}
+
+\begin{figure}
+\begin{tabular}{lll}
+agents & $a,b...$ & $a {:}{:}{=} \, \tuple{\mbox{atomic agents}}$ 
+\end{tabular}
+
+\begin{tabular}{lll}
+tests & $T,S...$ & $T{:}{:}{=} \, \begin{array}[t]{l}\tuple{\mbox{atomic tests}}\,|\,0\,|\,1\,|\\ T{+}S\;|\;T\,{;}\,S\;|\;!T\end{array}$
+\end{tabular}
+
+\begin{tabular}{lll}
+propositions & $p,q...$ & \(p{:}{:}{=} \,
+\begin{array}[t]{l}\tuple{\mbox{atomic events}}\,| \\
+T \,|\, 
+p{+}q \,|\,
+p\mbox{\&}q \,|\, \\
+p{;}q \;|\;
+p{*} \;|\; {\sim}p \;|\\
+{\diamond} a p
+\end{array}\)
+\end{tabular}
+
+\caption{The language of Epik terms. Propositional terms are interpreted
+as sets of guarded strings.}
+\end{figure}
+
 An architecture of interpretation by transtlation is employed, where English
@@
 
-\begin{example}   \label{phenomena1}
-{\em Phenomena}
-
-\vspace{1mm}
-\hspace{-10mm}\begin{tabular}[t]{l|l}
+\begin{figure}  
+\begin{tabular}[t]{l|l}
   Basic statives & It's heads. \\
@@
 \end{tabular}
-\end{example}
+\caption{Phenomena covered in the English grammar fragment.}
+\end{figure}
 

### 2020-06-16 tex/draft.tex
Latex macros syntax.tex, and bib file fs.bib

@@
 
-%\aclfinalcopy % Uncomment this line for the final submission
+% Uncomment this line for the final submission
+% Or comment it for the submission version with line numbers.
+\aclfinalcopy 
 %\def\aclpaperid{***} %  Enter the acl Paper ID here
+
+% CHECK THE BIB STYLE FOR THE STARSEM SUBMISSION
 
@@
 
-\author{First Author \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  \texttt{email@domain} \\\And
-  Second Author \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  \texttt{email@domain} \\}
+\author{Eric Campbell \\
+  Cornell University\\
+  \texttt{ehc86@cornell.edu} \\\And
+  Mats Rooth \\
+  Cornell University\\
+  \texttt{mr249@cornell.edu} \\}
 
@@
   into the finite state calculus, and alternatively by modeling
-  propositions as lazy lists. The syntax-semantics interface for
+  propositions as lazy lists in Haskell. The syntax-semantics interface for
   a fragment of English is defined by a categorial grammar.
 \end{abstract}
+
+\nocite{beesley2003finite,hulden2009foma}
+\nocite{karttunen2010update}
 
@@
 
-\begin{example}   \label{lexicon1}
-{\em Lexicon (partial)}
-
-\hspace{-10mm}\begin{tabular}[t]{l|c|c}
+\begin{figure}  
+\begin{tabular}[t]{l|c|c}
 Amy & $e$ & ${R}_a$\\
@@
 \end{tabular}
-\end{example}
+\caption{Partial categorial grammar lexicon. The first column has a word form.
+  the second column a categorial type, and third column a semantic translation in
+a notation that extends the Epik term language with lambds.}
+\end{figure}
 
@@
 
+\pagebreak
+
+\bibliographystyle{chicago}
+\bibliography{fs}
+
 \end{document}

### 2020-06-16 tex/draft.tex
Started on bibliography.

@@
   worlds}.  These worlds are often assumed to be things with the same
-nature and complexity as the world we occupy (Lewis 1986).  But
+nature and complexity as the world we occupy \citep{lewis1986plurality}.  But
 alternatively, one can work with small idealized models, in order to
@@
   the second column a categorial type, and third column a semantic translation in
-a notation that extends the Epik term language with lambds.}
+a logical language that extends the Epik term language with lambda.}
 \end{figure}

### 2020-06-16 tex/draft.tex
Most references are in. Labels for figures.

@@
 %\def\aclpaperid{***} %  Enter the acl Paper ID here
-
+% Make the references come out black. TAKE THIS OUT for submission.
+\definecolor{darkblue}{rgb}{0, 0, 0}
 % CHECK THE BIB STYLE FOR THE STARSEM SUBMISSION
@@
 John McCarthy's {\em Situation Calculus} is the source of the idea of
-constructing possible worlds as event sequences (McCarthy 1963, Reiter
-2001).  The algebraic theory of {\em Kleene Algebra with Tests}
-characterizes algebras with elements corresponding to propositions and
-event types in our application (Kozen 2003).  The models we propose
-are an epistemic extension of guarded string models for KAT, where a
-unary operation interpreted as an existential epistemic modality is
-included for each agent.  {\em Action models} in dynamic epistemic
-semantics introduced the technique of constructing epistemic models
-from primitive alternative relations on events, in order to capture
-the epistemic consequences of perceptual events (Baltag, Moss, and
-Solecki 1998). This is the basis for our construction of epistemic
-alternative relations.  Literature on finite state methods in
-linguistic semantics has used event strings and sets of event strings
-to theorize about tense and aspect in natural language semantics
-(Fernando 2003, Carlson 2009) and to express propositions (Fernando
-2017).  Literature on finite state intensiononal semantics has 
-shown how to do the semantics of intensional complementation
-including indirect questions in a setting where propositions are
-represented using finite state machines, and compositional semantics
-is expressed in a finite state calculus (Rooth 2017, Collard 2018).
-We adopt this in our syntax-semantics interface for
-English.
+constructing possible worlds as event sequences
+\citep{mccarthy1963situations,reiter2001knowledge}.  The algebraic
+theory of {\em Kleene Algebra with Tests} characterizes algebras with
+elements corresponding to propositions and event types in our
+application \citep{kozen2001automata}.\marginpar{2001 or 2003?}  The
+models we propose are an epistemic extension of guarded string models
+for KAT, where a unary operation interpreted as an existential
+epistemic modality is included for each agent.  {\em Action models} in
+dynamic epistemic semantics introduced the technique of constructing
+epistemic models from primitive alternative relations on events, in
+order to capture the epistemic consequences of perceptual events
+\citep{baltag1999logic}. This is the basis for our construction of
+epistemic alternative relations.  Literature on finite state methods
+in linguistic semantics has used event strings and sets of event
+strings to theorize about tense and aspect in natural language
+semantics \citep{fernando2004finite,fernando2007observing,carlson2009tense} and to
+express intensionality \cite{fernando2017intensions}.  Literature on
+finite state intensiononal semantics has discussed how to do the
+semantics of intensional complementation including indirect questions
+in a setting where compositional semantics is expressed in a finite
+state calculus \citep{rooth2017finite,collard2018finite}.  We adopt
+this in our syntax-semantics interface for English.
 
@@
 \section{Epistemic guarded string models}
-Figure 1 shows an Epik program that describes a possible worlds model for two agents
+Figure \ref{epikfigure} shows an Epik program that describes a possible worlds model for two agents
 with information about one coin, and events of the agents semi-privately
@@
 
-Kleene Algebra with Tests is an algebraic theory that is defined by equations and
-inequalities, which a has model classes including guarded string models, relational
-models, finite models, and matrix models.  This paper focuses on defining a family of
-concrete guarded string algebras, the elements of which are sets of guarded strings.
-Definitions and notation mostly follow Kozen (2003).  Additional syntax and semantics
-is included to model multi-agent epistemic semantics.  Guarded strings over a finite
-alphabet $E$ are like ordinary strings, but with truth assignments to a set $T$ of
-primitive propositions (primitive tests) alternating with the symbols from $E$.  In
-the algebra described by Figure 1, $E$ is the set of events $\{a_1,a_0,b_1,b_0\}$,
-and in the elevator example, $\{u,d\}$.  In the elevator example, $T$ is $\{p,q\}$,
-and in the coin example it is $\{h,t\}$.  We write truth assigments to the primitive
-propositions that observe the constraint as Boolean vectors. In the coin example, we
-get the vectors 01, 10 and in the elevator example, the vectors 00, 01, 10, 11.
-Guarded strings are strings of events, alternating with such vectors, and starting
-and ending with vectors.  \ex{gs1} gives the encoding as guarded strings of the
-worlds in \ex{elevator1} and \ex{coin1}.
+Kleene Algebra with Tests is an algebraic theory that is defined by
+equations and inequalities, which a has model classes including
+guarded string models, relational models, finite models, and matrix
+models.  This paper focuses on defining a family of concrete guarded
+string algebras, the elements of which are sets of guarded strings.
+Definitions and notation mostly follow \cite{kozen2001automata}.
+Additional syntax and semantics is included to model multi-agent
+epistemic semantics.  Guarded strings over a finite alphabet $E$ are
+like ordinary strings, but with truth assignments to a set $T$ of
+primitive propositions (primitive tests) alternating with the symbols
+from $E$.  In the algebra described by Figure \ref{epikfigure}, $E$ is the set of
+events $\{a_1,a_0,b_1,b_0\}$, and in the elevator example, $\{u,d\}$.
+In the elevator example, $T$ is $\{p,q\}$, and in the coin example it
+is $\{h,t\}$.  We write truth assigments to the primitive propositions
+that observe the constraint as Boolean vectors. In the coin example,
+we get the vectors 01, 10 and in the elevator example, the vectors 00,
+01, 10, 11.  Guarded strings are strings of events, alternating with
+such vectors, and starting and ending with vectors.  \ex{gs1} gives
+the encoding as guarded strings of the worlds in \ex{elevator1} and
+\ex{coin1}.
 
@@
 at the coin.}
+\label{epikfigure}
 \end{figure}
@@
 It remains to define the Kripke relation on guarded strings from an agent specification
-as in Figure 1. An an agent specification pairs each bare event with a set of bare events,
+as in Figure \ref{epikfigure}. An an agent specification pairs each bare event with a set of bare events,
 and so determines a relation between bare events, call it relation $R_a$ for an
@@
 relations between strings that was designed for use in computationonal phonology and
-morphographemics (cite Karttunen etc.)  Current implementations allow for the
+morphographemics 
+\citep{kaplan1994regular,beesley2003finite}.
+Current implementations allow for the
 definition of functions with the status of defined operators on regular sets and
-relations.  Such definitions are used to define an embedding of epistemic KAT in a
+relations
+\citep{hulden2009foma,karttunen2010update}.
+Such definitions are used to define an embedding of epistemic KAT in a
 string algebra. The methodology follows Section 2 closely.  Let ${\cal K}$ be an
@@
 lengthening Boolean vectors at the concatenation point.  Instead, KAT product and KAT
-Kleene star are defined operations in the string algebra, see Figure 2.  The
+Kleene star are defined operations in the string algebra, see Figure \ref{fstfigure1}.  The
 operations concatenate in the string algebra, delete strings with non-matching tests,
@@
 \caption{Translation into Fst of KAT product and KAT Kleene star.}
+\label{fstfigure1}
 \end{figure}
@@
 $(a_1 + a_0)(b_1 + b_0)$ contains two worlds rather than four.  The program in
-Figure 1 as interpreted in FST defines a countably infinite set of possible worlds by
+Figure \ref{epikfigure} as interpreted in FST defines a countably infinite set of possible worlds by
 KAT Kleene closure as $\mbox{\em Kst}(a_1 + a_0 +b_1 + b_0)$, and an algebra of
@@
 It remains to define an epistemic alternative relation on worlds for each agent.  The
-relevant information in Figure 1 is a relation between bare events for each
+relevant information in Figure \ref{epikfigure} is a relation between bare events for each
 agent. This determines a relation in the guarded string algebra a relation between
@@
   a relation between guarded strings.}
+\label{fstfigure2}
 \end{figure}
@@
 This scheme provides for an interpretation of the language of Epik
-terms that is defined in Figure 4.  An Epik specifications such as
+terms that is defined in Figure \ref{termfigure}.  An Epik specifications such as
 Figure 1 is translated to a straight-line program of the finite state
@@
   as sets of guarded strings.}
+\label{termfigure}
 \end{figure}
@@
 of words, their categorial types, and interpretations in a logical lambda language.
-\ex{phenomena1} lists that are covered.
+Figure \ref{phenomenafigure} lists phenomena that are covered.
 
@@
 \caption{Phenomena covered in the English grammar fragment.}
+\label{phenomenafigure}
 \end{figure}
@@
 
-\ex{lexicon1} gives illustrative lexical entries. The grammar and semantics are in
+Figure \ref{lexiconfigure} gives illustrative lexical entries. The grammar and semantics are in
 certain way optimized for a simple fragment of English concerned with clausal
@@
 a logical language that extends the Epik term language with lambda.}
+\label{lexiconfigure}
 \end{figure}
@@
 
-The semantic translations in the third column of \ex{lexicon1} use the Epik term
+The semantic translations in the third column of Figure \ref{lexiconfigure} use the Epik term
 language, incremented with lambda.  The body of $\lambda x.0^c{\cdot}h$, which is the

### 2020-06-16 tex/draft.tex
add section 4 text

--- 2020-06-16 tex/draft.tex
@@
 \usepackage{latexsym}
-\usepackage{amsmath,accents}
+\usepackage{amssymb,amsmath,accents}
+\usepackage{stmaryrd}
+\usepackage{array}
 \usepackage{hang}
@@
 
+\newcommand{\dom}{\ensuremath{\mathrm{dom}}}
 \input{syntax.tex}
@@
 
-% Uncomment this line for the final submission
-% Or comment it for the submission version with line numbers.
-\aclfinalcopy 
+%\aclfinalcopy % Uncomment this line for the final submission
 %\def\aclpaperid{***} %  Enter the acl Paper ID here
-
-% CHECK THE BIB STYLE FOR THE STARSEM SUBMISSION
 
@@
 
-\author{Eric Campbell \\
-  Cornell University\\
-  \texttt{ehc86@cornell.edu} \\\And
-  Mats Rooth \\
-  Cornell University\\
-  \texttt{mr249@cornell.edu} \\}
+\author{First Author \\
+  Affiliation / Address line 1 \\
+  Affiliation / Address line 2 \\
+  Affiliation / Address line 3 \\
+  \texttt{email@domain} \\\And
+  Second Author \\
+  Affiliation / Address line 1 \\
+  Affiliation / Address line 2 \\
+  Affiliation / Address line 3 \\
+  \texttt{email@domain} \\}
 
@@
   into the finite state calculus, and alternatively by modeling
-  propositions as lazy lists in Haskell. The syntax-semantics interface for
+  propositions as lazy lists. The syntax-semantics interface for
   a fragment of English is defined by a categorial grammar.
 \end{abstract}
-
-\nocite{beesley2003finite,hulden2009foma}
-\nocite{karttunen2010update}
 
@@
   at the coin.}
+\label{fig:epik-example}
 \end{figure}
@@
 
-\section{Interpretation using lazy lists of strings}
-%\pagebreak
-%Hi Mom!
-
-%\pagebreak
-
-%Hi Sis!
-%\pagebreak
+\section{Interpretation}
+
+To write down an Epik program, we need to describe the set of worlds
+and then write down terms that can be evaluated in that world. First,
+\texttt{state} declarations describe the domain of the worlds as a
+list of atomic test variables. For instance in
+Figure~\ref{fig:epik-example}, there are two variables in the world,
+\texttt{h} and \texttt{t}, which will correspond to when a the coin is
+\texttt{heads} or \texttt{tails}. Then, we can write down
+\texttt{constraint}s on the allowable set of worlds described by a
+boolean expression. This constraint is denoted by the variable
+$\phi$. In this example, the expression $h!t + t!h$ says that $h$ and
+$t$ are mutually exclusive. Third, we describe the collection of
+\texttt{event}s and their decorations. The set of input-output pairs
+denoted by the \texttt{event} declarations is event pairs, is denoted
+$P(e)$, and the set of all events as $\dom(P)$. For instance, in
+Figure~\ref{fig:epik-example}, the event \texttt{a1}, which
+corresponds to Aly semi-privately sensing the value of the coin, is
+decorated with input test \texttt{h} and output test \texttt{h}. The
+final piece we need to write down is the collection of alternative
+relations for each agent $a$, which we write down as sequence of $ e
+\to e_0 + \cdots + e_n$ where $e$ is an event in the base world and
+each $e_i$ is an alternative for agent $a$ to event $e$. We write each
+agent $a$'s alternative relation, as indicated by these descriptions,
+as $R_a$.
+
+Finally, we can write down terms in the Epik language using the
+grammar described in Figure~\ref{fig:epik-grammar}. We include the
+basic KAT operators and constatns atomic tests, events, $1$, $0$ and
+operators $+$, $;$, $*$, and $!$), which are interpreted in the
+standard way (bearing in mind $P$ and $\phi$) as sets of guarded
+strings in \ref{fig:interp}.  We also have the complement operator
+$\sim p$ which is interpreted as the set of worlds less the set of
+worlds denoted by $p$. The the alternative operator $\langle a
+\rangle~p$, is interpreted as applying the preimage of $a$'s
+alternative relation denoted by $p$. The intersection operator $p\&q$,
+is simply denotes the intersection of the denoted sets, and the
+wild-card pattern $\_$ denotes the set of all events.  These are
+interpreted as sets of guarded strings in Figure~\ref{fig:interp}.
+
+TODO MORE DETAIl.
+
+\begin{figure}
+  \[\begin{array}{l >{\triangleq}c l}
+  \llbracket e \rrbracket^{\phi,P}&& \{\alpha \cdot e \cdot \beta \mid
+  (\alpha,\beta) \in P(e), \alpha, \beta \Rightarrow \phi \}
+  \\ \llbracket \_ \rrbracket^{\phi,P}&& \bigcup_{e \in
+    \dom(P)}\llbracket e \rrbracket^{\phi,P}\\
+  \llbracket t \rrbracket^{\phi,P} && \{\alpha \mid \alpha \Rightarrow \phi\}\\
+  \llbracket{p+q}\rrbracket^{\phi,P} && \llbracket{p}\rrbracket^{\phi,P} \cup \llbracket q \rrbracket^{\phi,P}\\
+  \llbracket{p \& q}\rrbracket^{\phi,P} && \llbracket{p}\rrbracket^{\phi,P} \cap \llbracket q \rrbracket^{\phi,P}\\
+  \llbracket{p;q}\rrbracket^{\phi,P} && \llbracket p \rrbracket^{\phi,P} \cdot \llbracket{q}\rrbracket^{\phi,P}\\
+  \llbracket p^*\rrbracket^{\phi,P} && \displaystyle\bigcup_n {\underbrace{\llbracket p \rrbracket^{\phi,P}\cdot \ldots \cdot \llbracket p \rrbracket^{\phi,P}}_{n}} \\
+  \llbracket \sim p \rrbracket^{\phi,P} && \llbracket \_^*\rrbracket^{\phi,P} \setminus \llbracket p \rrbracket^{\phi,P} \\
+  \llbracket \langle a \rangle~p \rrbracket && \doublehat{R_a}\llbracket p \rrbracket
+
+  \end{array}
+  \]
+  \caption{Interpretation of Epik terms as sets of guarded strings}
+  \label{fig:interp}
+\end{figure}
+
+TODO WRITE ABOUT FINITE INTERPRETATION
 
@@
 propositions & $p,q...$ & \(p{:}{:}{=} \,
-\begin{array}[t]{l}\tuple{\mbox{atomic events}}\,| \\
+\begin{array}[t]{l}\tuple{\mbox{atomic events}}\,|\,\_\, |\\
 T \,|\, 
@@
 p{*} \;|\; {\sim}p \;|\\
-{\diamond} a p
+{\langle a \rangle}~p
 \end{array}\)
@@
   as sets of guarded strings.}
+\label{fig:epik-grammar}
 \end{figure}
@@
 
-\begin{figure}  
-\begin{tabular}[t]{l|c|c}
+\begin{example}   \label{lexicon1}
+{\em Lexicon (partial)}
+
+\hspace{-10mm}\begin{tabular}[t]{l|c|c}
 Amy & $e$ & ${R}_a$\\
@@
 \end{tabular}
-\caption{Partial categorial grammar lexicon. The first column has a word form.
-  the second column a categorial type, and third column a semantic translation in
-a notation that extends the Epik term language with lambds.}
-\end{figure}
+\end{example}
 
@@
 
-\pagebreak
-
-\bibliographystyle{chicago}
-\bibliography{fs}
-
 \end{document}

### 2020-06-16 tex/draft.tex

--- 2020-06-16 tex/draft.tex
@@
   at the coin.}
+\label{fig:epik-example}
 \label{epikfigure}
@@
   as sets of guarded strings.}
+\label{fig:epik-grammar}
 \label{termfigure}

### 2020-06-17 tex/draft.tex
add formal detail to section 2

--- 2020-06-16 tex/draft.tex
@@
 
-Rather than guarded strings, elements of a guarded string model for KAT are sets of
-guarded strings.  In the application, these elements have the interpretation of
-propositions (sets of possible worlds) and/or event types.  An event such as $u$ in
-the guarded string model corresponds to the set of guarded strings where the bare
-event is flanked by compatible truth assigments, as defined in an Epik program by the
-event declaration and the constraint declaration. 
-These have the status of event types, in that they
-can ``happen'' in different possible worlds. Here happening corresponds to
-incrementing the world with some compatible element of the event type.
-Where $e$ is a bare event symbol, $\hat{e}$ is the
-corresponding set of guarded strings,
-consisting of the bare event decorated with compatible truth assignments.
-See the examples in \ex{events1}.
+Rather than guarded strings, elements of a guarded string model for
+KAT are sets of guarded strings.  In the application, these elements
+have the interpretation of propositions (sets of possible worlds)
+and/or event types.  An event such as $u$ in the guarded string model
+corresponds to the set of guarded strings where the bare event is
+flanked by compatible truth assigments, as defined in an Epik program
+by the event declaration and the constraint declaration.  These have
+the status of event types, in that they can ``happen'' in different
+possible worlds. Here happening corresponds to incrementing the world
+with some compatible element of the event type.  Where $e$ is a bare
+event symbol, $\llbracket e \rrbracket^{\phi,P}$ is the corresponding
+set of guarded strings, consisting of the bare event decorated with
+compatible truth assignments denoted by $P$ that are admitted by the
+world-constraints $\phi$.  See this formally in
+Figure~\ref{fig:interp}, and the examples in \ex{events1}.
 % u (!q!p):(!qp) + (!qp):(q!p) + (q!p):(qp) 173
@@
 \begin{array}[t]{l|c}
-  e & \hat{e} \\ \hline  
+  e & \llbracket e \rrbracket^{\phi, P} \\ \hline  
   u & \{ 00u01, 01u10, 10u11\}\\
@@
 Kleene Algebra with Tests has the algebraic signature
-$\tuple{K,+,\cdot,*,\bar{},0,1}$.  To this we add an unary operation $\Diamond_a$ for
-each agent, and a unary complent operation $\cdot^{c}$ on propositions.  This results
-in the signature $\tuple{K,+,\cdot,*,\bar{},0,1,{}^c,\Diamond_a,\Diamond_b}$ for
-epistemic KAT with two agents.  In a guarded string model in our construction, the
-elements of $K$ are sets of guarded strings.  The operation $+$ is set union. The
-constant $0$ is the empty set.  The constant 1 is the set of all truth assignments
-obeying the constraint on truth assignments, i.e. $\{10,10\}$ in the coin example.
-The operation $\cdot$ is fusion product raised to sets: $x \cdot y$ is the set of all
-defined fusion products of an element of $x$ with and element of $y$.  The operation
-$*$ is Kleene star, with $x* = \cup_{i\geq 0}x^i$, where $x^i$ is the $i$-times
-product of $x$ with itself and $x^0=1$.  Subsets of $1$ are also elements of $K$, and
-these form the Boolean algebra of tests as a set Boolean algebra.  The overbar
-operation is complement in this Boolean algebra.  The complement operation ${}^c$ is
-complement at the level of sets of guarded strings, with $x^c = 0^c - x$, where the
-operation on the right is set difference.  The epistemic modality $\Diamond_a$ is
-interpreted using Kripke semantics, as pre-image relative to a fixed relation $R_a$
-between guarded strings, $\Diamond_a x = \setabs{u}{\exists v.\element{v}{x} \wedge u
-  R_a v}$.  Here $u$ and $v$ are guarded strings, while $x$ is an element of $K$.
+$\tuple{K,+,\cdot,*,\bar{},0,1}$.  To this we add an unary operation
+$\Diamond_a$ for each agent, and a unary complent operation
+$\cdot^{c}$ on propositions.  This results in the signature
+$\tuple{K,+,\cdot,*,\bar{},0,1,{}^c,\Diamond_a,\Diamond_b}$ for
+epistemic KAT with two agents.
+
+We interpret each term in as a set of guarded strings as indicated in
+Figure~\ref{fig:interp}. A bare event $e$ is interpreted as the set of
+guarded strings allowed by $\phi$ and $P$. The operation $p + q$ is
+set union. The constant $0$ is the empty set.  The constant 1 is the
+set of all truth assignments obeying the constraint on truth
+assignments, i.e. $\{10,01\}$ in the coin example.  The operation
+$\cdot$ is fusion product raised to sets: $x \cdot y$ is the set of
+all defined fusion products of an element of $x$ with and element of
+$y$.  The operation $*$ is Kleene star, with $x* = \cup_{i\geq 0}x^i$,
+where $x^i$ is the $i$-times product of $x$ with itself and $x^0=1$.
+Subsets of $1$ indicated by $t$ are also elements of $K$, and these
+form the Boolean algebra of tests, denoting all atoms admitted by the
+constraint $\phi$ and the boolean test $t$.  We ``overload'' $\cdot$
+to further indicate boolean conjunction, $+$ to further indicate
+boolean disjunction, and $0,1$ to indicate the zero and one elements
+of the boolean algebra. The overbar operation denotes complement in
+this Boolean algebra: $\overline t$ denotes all atoms $\alpha$ such
+that $\alpha \Rightarrow \neg t$.  Other boolean operators (such as
+implication $\Rightarrow$) are encoded as syntactic sugar in the
+standard way. The complement operation ${p}^c$ is complement at the
+level of sets of guarded strings, defined to every set the universe
+$\_^*$ except those denoted by $p$. Finally, the epistemic modality
+$\Diamond_a$ is interpreted using Kripke semantics, as pre-image
+relative to a fixed relation $R_a$ between guarded strings,
+$\Diamond_a x = \setabs{u}{\exists v.\element{v}{x} \wedge u R_a v}$.
+Here $u$ and $v$ are guarded strings, while $x$ is an element of $K$.
 
@@
 
-It remains to define the Kripke relation on guarded strings from an agent specification
-as in Figure \ref{epikfigure}. An an agent specification pairs each bare event with a set of bare events,
-and so determines a relation between bare events, call it relation $R_a$ for an
-agent $e$.  This determines
-a relation $\hat{R_a}$ between decorated events, see \ex{Rhat}.
-It is generalized to a relation $\doublehat{R}$ on arbitrary guarded strings by
-relational Kleene star, as defined in \ex{Rhathat}. The operation $\cdot$ in the
-definition is KAT product, which enforces matching of tests.
-The number of terms $n$ is construed as satisfying $n\geq 1$.
-{\em What is desired for $n=0$?}
+It remains to define the Kripke relation on guarded strings from an
+agent specification as in Figure \ref{epikfigure}. An an agent
+specification pairs each bare event with a set of bare events, and so
+determines a relation between bare events, call it relation $R_a$ for
+an agent $a$.  This determines a relation $\hat{R_a}$ between
+decorated events, see \ex{Rhat}.  It is generalized to a relation
+$\doublehat{R}$ on arbitrary guarded strings by relational Kleene
+star, as defined in \ex{Rhathat}. The operation $\cdot$ in the
+definition is KAT product, which enforces matching of tests.  The
+number of terms $n$ is construed as satisfying $n\geq 1$.  {\em What
+  is desired for $n=0$?}
 
@@
 base worlds.
+
+\begin{figure}
+  \[\begin{array}{l >{\triangleq}c l}
+  \llbracket 0 \rrbracket^{\phi,P} && \{\} \\  
+  \llbracket 1 \rrbracket^{\phi,P} && \{\alpha \mid \alpha \Rightarrow \phi\}\\
+  \llbracket e \rrbracket^{\phi,P}&& \{\alpha \cdot e \cdot \beta \mid
+  (\alpha,\beta) \in P(e), \alpha, \beta \Rightarrow \phi \}
+  \\ \llbracket \_ \rrbracket^{\phi,P}&& \bigcup_{e \in
+    \dom(P)}\llbracket e \rrbracket^{\phi,P}\\
+  \llbracket t \rrbracket^{\phi,P} &&  \{\alpha \mid \alpha \Rightarrow \phi \wedge t\}\\
+  \llbracket{p+q}\rrbracket^{\phi,P} && \llbracket{p}\rrbracket^{\phi,P} \cup \llbracket q \rrbracket^{\phi,P}\\
+  \llbracket{p \& q}\rrbracket^{\phi,P} && \llbracket{p}\rrbracket^{\phi,P} \cap \llbracket q \rrbracket^{\phi,P}\\
+  \llbracket{p;q}\rrbracket^{\phi,P} && \llbracket p \rrbracket^{\phi,P} \cdot \llbracket{q}\rrbracket^{\phi,P}\\
+  \llbracket p^*\rrbracket^{\phi,P} && \displaystyle\bigcup_n {\underbrace{\llbracket p \rrbracket^{\phi,P}\cdot \ldots \cdot \llbracket p \rrbracket^{\phi,P}}_{n}} \\
+  \llbracket \sim p \rrbracket^{\phi,P} && \llbracket \_^*\rrbracket^{\phi,P} \setminus \llbracket p \rrbracket^{\phi,P} \\
+  \llbracket \langle a \rangle~p \rrbracket && \doublehat{R_a}\llbracket p \rrbracket
+
+  \end{array}
+  \]
+  \caption{Interpretation of Epik terms as sets of guarded strings}
+  \label{fig:interp}
+\end{figure}
 
@@
 
-\section{Interpretation using lazy lists of strings}
-%\pagebreak
-%Hi Mom!
-
-%\pagebreak
-
-%Hi Sis!
-%\pagebreak
+\section{Direct Bounded Implementation}
 

### 2020-06-17 tex/draft.tex
sketch section 4

@@
 with some compatible element of the event type.  Where $e$ is a bare
-event symbol, $\llbracket e \rrbracket^{\phi,P}$ is the corresponding
+event symbol, $\llbracket e \rrbracket^{\varphi,P}$ is the corresponding
 set of guarded strings, consisting of the bare event decorated with
 compatible truth assignments denoted by $P$ that are admitted by the
-world-constraints $\phi$.  See this formally in
+world-constraints $\varphi$.  See this formally in
 Figure~\ref{fig:interp}, and the examples in \ex{events1}.
@@
 \begin{array}[t]{l|c}
-  e & \llbracket e \rrbracket^{\phi, P} \\ \hline  
+  e & \llbracket e \rrbracket^{\varphi, P} \\ \hline  
   u & \{ 00u01, 01u10, 10u11\}\\
@@
 Figure~\ref{fig:interp}. A bare event $e$ is interpreted as the set of
-guarded strings allowed by $\phi$ and $P$. The operation $p + q$ is
+guarded strings allowed by $\varphi$ and $P$. The operation $p + q$ is
 set union. The constant $0$ is the empty set.  The constant 1 is the
@@
 form the Boolean algebra of tests, denoting all atoms admitted by the
-constraint $\phi$ and the boolean test $t$.  We ``overload'' $\cdot$
+constraint $\varphi$ and the boolean test $t$.  We ``overload'' $\cdot$
 to further indicate boolean conjunction, $+$ to further indicate
@@
   \[\begin{array}{l >{\triangleq}c l}
-  \llbracket 0 \rrbracket^{\phi,P} && \{\} \\  
-  \llbracket 1 \rrbracket^{\phi,P} && \{\alpha \mid \alpha \Rightarrow \phi\}\\
-  \llbracket e \rrbracket^{\phi,P}&& \{\alpha \cdot e \cdot \beta \mid
-  (\alpha,\beta) \in P(e), \alpha, \beta \Rightarrow \phi \}
-  \\ \llbracket \_ \rrbracket^{\phi,P}&& \bigcup_{e \in
-    \dom(P)}\llbracket e \rrbracket^{\phi,P}\\
-  \llbracket t \rrbracket^{\phi,P} &&  \{\alpha \mid \alpha \Rightarrow \phi \wedge t\}\\
-  \llbracket{p+q}\rrbracket^{\phi,P} && \llbracket{p}\rrbracket^{\phi,P} \cup \llbracket q \rrbracket^{\phi,P}\\
-  \llbracket{p \& q}\rrbracket^{\phi,P} && \llbracket{p}\rrbracket^{\phi,P} \cap \llbracket q \rrbracket^{\phi,P}\\
-  \llbracket{p;q}\rrbracket^{\phi,P} && \llbracket p \rrbracket^{\phi,P} \cdot \llbracket{q}\rrbracket^{\phi,P}\\
-  \llbracket p^*\rrbracket^{\phi,P} && \displaystyle\bigcup_n {\underbrace{\llbracket p \rrbracket^{\phi,P}\cdot \ldots \cdot \llbracket p \rrbracket^{\phi,P}}_{n}} \\
-  \llbracket \sim p \rrbracket^{\phi,P} && \llbracket \_^*\rrbracket^{\phi,P} \setminus \llbracket p \rrbracket^{\phi,P} \\
+  \llbracket 0 \rrbracket^{\varphi,P} && \{\} \\  
+  \llbracket 1 \rrbracket^{\varphi,P} && \{\alpha \mid \alpha \Rightarrow \varphi\}\\
+  \llbracket e \rrbracket^{\varphi,P}&& \{\alpha \cdot e \cdot \beta \mid
+  (\alpha,\beta) \in P(e), \alpha, \beta \Rightarrow \varphi \}
+  \\ \llbracket \_ \rrbracket^{\varphi,P}&& \bigcup_{e \in
+    \dom(P)}\llbracket e \rrbracket^{\varphi,P}\\
+  \llbracket t \rrbracket^{\varphi,P} &&  \{\alpha \mid \alpha \Rightarrow \varphi \wedge t\}\\
+  \llbracket{p+q}\rrbracket^{\varphi,P} && \llbracket{p}\rrbracket^{\varphi,P} \cup \llbracket q \rrbracket^{\varphi,P}\\
+  \llbracket{p \& q}\rrbracket^{\varphi,P} && \llbracket{p}\rrbracket^{\varphi,P} \cap \llbracket q \rrbracket^{\varphi,P}\\
+  \llbracket{p;q}\rrbracket^{\varphi,P} && \llbracket p \rrbracket^{\varphi,P} \cdot \llbracket{q}\rrbracket^{\varphi,P}\\
+  \llbracket p^*\rrbracket^{\varphi,P} && \displaystyle\bigcup_n {\underbrace{\llbracket p \rrbracket^{\varphi,P}\cdot \ldots \cdot \llbracket p \rrbracket^{\varphi,P}}_{n}} \\
+  \llbracket \sim p \rrbracket^{\varphi,P} && \llbracket \_^*\rrbracket^{\varphi,P} \setminus \llbracket p \rrbracket^{\varphi,P} \\
   \llbracket \langle a \rangle~p \rrbracket && \doublehat{R_a}\llbracket p \rrbracket
@@
 
-\section{Direct Bounded Implementation}
+\section{Direct Bounded Interpretation}
+
+We also implement the semantics of Epik terms using lazy lists~\cite{LazyLists}
+in Haskell, rather than the direct interpretation as sets. Unfortunately,
+regular expressions, and hence also Epik, can denote infinite sets of strings,
+for example the term $\_^*$. Normally regular languages are represented using a
+finite coalgebra~\cite{}. However the non-distributivity of $\diamond$ accross
+$;$ complicates this construction~\footnote{Axiomatic and Coalgebraic models for
+  Epik are open questions}. To sidestep this, we parameterize the interpretation
+function on a positive integer $n$ and only produce guarded strings of length
+$n$ or less.  The length of a guarded string $g$, written $|g|$ is the number of
+events in the string, so $\alpha$ has length $0$, $\alpha e\beta$ has length
+$1$, and $\alpha e \beta e' \gamma$ has length $2$.
+
+We translate the Epik terms into a Haskell algebraic datatype that represents
+terms with the same signature as described in Section 2. We then paramaterize
+the interpretation function, $\llparenthesis p \rrparenthesis^{\varphi,P}_n$ on an
+integer $n$ that describes the maximum length of a string we will produce. The
+bounded interpretation into lists of strings is very similar to the unbounded
+interpretation into sets of strings, except for the bounds checking. The full
+details are shown in Figure~\ref{fig:list-interp}: we use list-comprehension
+notation, which is analogous to set-builder notation, except that it produces a
+list (potentially with duplicated elements) instead of a set, and so is written
+with square brackets $[]$. We use $+$ to indicate list concatenation, and $\sum
+l$ to indicate the concatenation of a list of lists $l$. We overload the set
+membership symbol $\in$ to also indicate membership in a list. We also lift the
+fusion operator $\cdot$ to lists. For a list of guarded strings $l$, we write
+$l|_n$ as shorthand for the set $[ g~|~g \in l , |g| \leq n]$.
+
+First of all, whenever $n = 0$, the set of guarded strings is empty. TODO ADD
+MORE TEXT.
+
+The final step is to convert the list of guarded strings into a set of guarded
+strings. Since a list is just an ordered multiset, we forget the order, and
+remove all duplicates. Haskell's libraries have a built-in function for this,
+\texttt{Data.Set.fromList}, which converts a list into a set.
+
+\begin{figure}
+  \[\begin{array}{r >{\triangleq}c l}
+  \llparenthesis p \rrparenthesis^{\varphi,P}_0 && [] \\
+  \llparenthesis 0 \rrparenthesis^{\varphi,P}_n && [] \\
+  \llparenthesis 1 \rrparenthesis^{\varphi,P}_n && [\alpha \mid \alpha \leq \varphi] \\
+  \llparenthesis e \rrparenthesis^{\varphi,P}_n && [\alpha \cdot e \cdot \beta \mid (\alpha, \beta) \in P(e), \alpha\vee \beta \Rightarrow \varphi] \\
+  \llparenthesis \_ \rrparenthesis^{\varphi,P}_n && \sum [\llparenthesis e \rrparenthesis^{\varphi,P} \mid e \in \dom(P) ] \\
+  \llparenthesis t \rrparenthesis^{\varphi,P}_n && [\alpha\mid \alpha \Rightarrow \varphi \wedge t] \\
+  \llparenthesis p + q \rrparenthesis^{\varphi,P}_n && \llparenthesis p \rrparenthesis^{\varphi,P}_n + \llparenthesis q \rrparenthesis^{\varphi,P}_n \\
+  \llparenthesis p \& q \rrparenthesis^{\varphi,P}_n &&
+  [g~|~g \in \llparenthesis p \rrparenthesis_n^{\varphi,P}, g \in \llparenthesis q \rrparenthesis_n^{\varphi,P}] \\
+  \llparenthesis p;q \rrparenthesis^{\varphi,P}_n &&
+  \left(\llparenthesis p \rrparenthesis^{\varphi,P}_n \cdot \llparenthesis q\rrparenthesis^{\varphi,P}_n\right)|_n\\
+  \llparenthesis p^* \rrparenthesis^{\varphi,P}_n && [] + \left(\llparenthesis p\rrparenthesis^{\varphi,P}_n \cdot  \llparenthesis p^* \rrparenthesis^{\varphi,P}_{n-i}\right)|_n \\
+  \multicolumn 2 r {} & \text{where}~i = \max\{1,\min\{|g|\mid g \in \llbracket p \rrbracket^{\varphi,P}\}\}\\
+  \llparenthesis \sim p \rrparenthesis^{\varphi,P}_n && [g \mid g \in \llbracket \_^* \rrbracket^{\varphi,P}_n, g \not\in \llbracket p \rrbracket^{\varphi,P}_n] \\
+  \llparenthesis \langle a \rangle p \rrparenthesis^{\varphi,P}_n &&
+  [g' \mid g \in g'~\doublehat{R_a}~g, g \in \llparenthesis p \rrparenthesis^{\varphi,P}]
+  \end{array}
+  \]
+\end{figure}
+
+TODO Describe advantages of Lazy Lists
 

### 2020-06-17 tex/draft.tex
update atoms

@@
 \section{Epistemic guarded string models}
-Figure \ref{epikfigure} shows an Epik program that describes a possible worlds model for two agents
-with information about one coin, and events of the agents semi-privately
-looking at the coin. The line beginning with {\tt state} lists the basic
-stative propositions. To illustrate syntax, a second proposition $t$ (tails) is
-included.  The line beginning with {\tt constraint} defines compatibilities among the
-propositions: the coin is heads or tails and not both.  The lines beginning with {\tt
-  event} declare events, their preconditions, and their effect on state, following
-the format in \ex{event1}.  Finally the lines beginning with {\tt agent} define {\em
-  event alternative} relations for agents.  Each clause with an arrow has a single
-event symbol on the left, and a disjunction of alternative events on the right of the
-arrow. The interpretation of Amy's alternatives for $b_1$ (Bill peeks at heads), is
-that when $b_1$ happens, for Amy either $b_1$ or $b_0$ (Bill peeks at tails) could
-be happening.
+Figure \ref{epikfigure} shows an Epik program that describes a
+possible worlds model for two agents with information about one coin,
+and events of the agents semi-privately looking at the coin. The line
+beginning with {\tt state} lists the basic stative propositions. To
+illustrate syntax, a second proposition $t$ (tails) is included.  The
+line beginning with {\tt constraint} defines compatibilities among the
+propositions: the coin is heads or tails and not both.  The lines
+beginning with {\tt event} declare events, their preconditions, and
+their effect on state, following the format in \ex{event1}.  Finally
+the lines beginning with {\tt agent} define {\em event alternative}
+relations for agents.  Each clause with an arrow has a single event
+symbol on the left, and a disjunction of alternative events on the
+right of the arrow. The interpretation of Amy's alternatives for $b_1$
+(Bill peeks at heads), is that when $b_1$ happens, for Amy either
+$b_1$ or $b_0$ (Bill peeks at tails) could be happening.
 
@@
 primitive propositions (primitive tests) alternating with the symbols
-from $E$.  In the algebra described by Figure \ref{epikfigure}, $E$ is the set of
+from $E$. Following \cite{kozen2001automata}, we call these truth
+assignments \emph{atoms}; the set of atoms is notated $\mathcal A$.
+In the algebra described by Figure \ref{epikfigure}, $E$ is the set of
 events $\{a_1,a_0,b_1,b_0\}$, and in the elevator example, $\{u,d\}$.
 In the elevator example, $T$ is $\{p,q\}$, and in the coin example it
-is $\{h,t\}$.  We write truth assigments to the primitive propositions
-that observe the constraint as Boolean vectors. In the coin example,
-we get the vectors 01, 10 and in the elevator example, the vectors 00,
-01, 10, 11.  Guarded strings are strings of events, alternating with
-such vectors, and starting and ending with vectors.  \ex{gs1} gives
-the encoding as guarded strings of the worlds in \ex{elevator1} and
-\ex{coin1}.
+is $\{h,t\}$.  We write atoms as the boolean formulae that uniquely
+specify the truth assignment. In the elevator example, we have atoms
+the vectors $\bar q \bar p$, $\bar q p$, $q \bar p$, $q p$.  In the
+coin example, we get $h\bar t$ and $\bar h t$, as well as the bogus
+atoms $\bar h \bar t$ and $ht$!  To preclude bogosity, the
+\texttt{constraint} declaration in Figure~\ref{epikfigure} describes
+the set of \emph{feasible atoms}, which we parse into a formula
+$\varphi$ (if there are multiple \texttt{constraint declarations},
+$\varphi$ is the conjunction of the parsed formulae). For a formula
+$\phi$, the set of feasible atoms is $\mathcal A_\phi$. In the coin
+example, the constraint $\varphi$ is $h\bar t + \bar h t$. Since $\bar
+h t \models \varphi$ and $h \bar t \models \varphi$, they are
+permitted feasible atoms, and since $\bar h \bar t$ and $h t$ do not
+satisfy $\phi$ they are not permitted.
+
+Guarded strings are strings of events, alternating with such atoms,
+and starting and ending with atoms.  \ex{gs1} gives the encoding as
+guarded strings of the worlds in \ex{elevator1} and \ex{coin1}.
 
@@
 \begin{tabular}[t]{ll}
-$v_1$ & $00 u 01 u 10 d 01 u 10$ \\
-$v_2$ & $00 u 01 u 10 u 11$ \\   
-$w_1$ & $01$ \\
-$w_2$ & $01 a_0 01$ \\  
-$w_3$ & $01 a_0 01 b_0 01$ \\
+$v_1$ & $\bar p \bar q ~u~ \bar p q ~u~ q \bar p ~d~ \bar p q ~u~ q \bar p$ \\
+$v_2$ & $\bar p \bar q ~u~ \bar p q ~u~ p\bar q ~u~ pq$ \\   
+$w_1$ & $\bar h t$ \\
+$w_2$ & $\bar h t ~a_0~ \bar h t$ \\  
+$w_3$ & $\bar h t ~a_0~ \bar h t ~b_0~ \bar h t$ \\
 \end{tabular}
@@
 \begin{example} \label{fusion1}
-\hspace{-2mm}\( 00 u 01 u 10 d 01 \cdot 01 u 10 = 00 u 01 u 10 d 01 u 10 \)
-
-\hspace{-2mm}\( 00 u 01 u 10 d 01 \cdot 10 u 11 = \mbox{\em{undefined}} \)
-
-\hspace{-2mm}\(01 \cdot 01 a_0 01 = 01 a_0 01 \)
-
-\hspace{-2mm}\(01 \cdot 10 a_1 10 = \mbox{\em{undefined}} \)
+\hspace{-2mm}\( \bar p q ~u~ \bar p q ~u~ p \bar q ~d~ \bar p q \cdot \bar p q ~u~ p \bar q = \bar p q ~u~ \bar p q ~u~ p \bar q ~d~ \bar p q ~u~ p \bar q \)
+
+\hspace{-2mm}\( \bar p \bar q ~u~ \bar p q ~u~ p \bar q ~d~ \bar p q \cdot p \bar q ~u~ pq = \mbox{\em{undefined}} \)
+
+\hspace{-2mm}\(\bar h t \cdot \bar h t ~a_0~ \bar h t = \bar h t ~a_0~ \bar h t \)
+
+\hspace{-2mm}\(\bar h t \cdot h \bar t ~a_1~ h \bar t = \mbox{\em{undefined}} \)
 \end{example}
 
-Rather than guarded strings, elements of a guarded string model for
-KAT are sets of guarded strings.  In the application, these elements
-have the interpretation of propositions (sets of possible worlds)
-and/or event types.  An event such as $u$ in the guarded string model
-corresponds to the set of guarded strings where the bare event is
-flanked by compatible truth assigments, as defined in an Epik program
-by the event declaration and the constraint declaration.  These have
-the status of event types, in that they can ``happen'' in different
-possible worlds. Here happening corresponds to incrementing the world
-with some compatible element of the event type.  Where $e$ is a bare
-event symbol, $\llbracket e \rrbracket^{\varphi,P}$ is the corresponding
-set of guarded strings, consisting of the bare event decorated with
-compatible truth assignments denoted by $P$ that are admitted by the
-world-constraints $\varphi$.  See this formally in
+Rather than individual guarded strings, elements of a guarded string
+model for KAT are sets of guarded strings.  In the application, these
+elements have the interpretation of propositions (sets of possible
+worlds) and/or event types.  An event such as $u$ in the guarded
+string model corresponds to the set of guarded strings where the bare
+event is flanked by compatible atoms. In standard KATs, this is  as defined in an Epik
+program by the event declaration and the constraint declaration.
+These have the status of event types, in that they can ``happen'' in
+different possible worlds. Here happening corresponds to incrementing
+the world with some compatible element of the event type.  Where $e$
+is a bare event symbol, $\llbracket e \rrbracket^{\varphi,P}$ is the
+corresponding set of guarded strings, consisting of the bare event
+decorated with compatible truth assignments denoted by $P$ that are
+admitted by the world-constraints $\varphi$.  See this formally in
 Figure~\ref{fig:interp}, and the examples in \ex{events1}.
@@
   e & \llbracket e \rrbracket^{\varphi, P} \\ \hline  
-  u & \{ 00u01, 01u10, 10u11\}\\
-  d & \{ 01d00, 10d01, 11d10\}\\
-  a_1 & \{10 a_1 10 \}\\
-  a_0 & \{01 a_0 01\}\\
-  b_1 & \{10 b_1 10 \}\\
-  b_0 & \{01 b_0 01 \}\\
+  u & \{ \bar p \bar q ~u~ \bar p q, \bar p q ~u ~p \bar q,  p \bar q ~u~ pq\}\\
+  d & \{ \bar p q ~d~ \bar p \bar q, p \bar q~d~\bar p q, p q~d~ p \bar q\}\\
+  a_1 & \{h \bar t ~a_1~ h \bar t \}\\
+  a_0 & \{t \bar h ~a_0~ t \bar h\}\\
+  b_1 & \{h\bar t ~b_1~ h \bar t \}\\
+  b_0 & \{t \bar h ~b_0~ t \bar h \}\\
 \end{array}

### 2020-06-17 tex/draft.tex
polish atom description

@@
 specify the truth assignment. In the elevator example, we have atoms
-the vectors $\bar q \bar p$, $\bar q p$, $q \bar p$, $q p$.  In the
-coin example, we get $h\bar t$ and $\bar h t$, as well as the bogus
-atoms $\bar h \bar t$ and $ht$!  To preclude bogosity, the
-\texttt{constraint} declaration in Figure~\ref{epikfigure} describes
-the set of \emph{feasible atoms}, which we parse into a formula
-$\varphi$ (if there are multiple \texttt{constraint declarations},
-$\varphi$ is the conjunction of the parsed formulae). For a formula
-$\phi$, the set of feasible atoms is $\mathcal A_\phi$. In the coin
-example, the constraint $\varphi$ is $h\bar t + \bar h t$. Since $\bar
-h t \models \varphi$ and $h \bar t \models \varphi$, they are
-permitted feasible atoms, and since $\bar h \bar t$ and $h t$ do not
-satisfy $\phi$ they are not permitted.
+$\bar q \bar p$, $\bar q p$, $q \bar p$, $q p$, for notational
+simplicity we will write these as $\hat0$, $\hat1$, $\hat2$, $\hat3$
+respectivelyto indicate the floor each atom represents. In the coin
+example, we get $h\bar t$ and $\bar h t$ (for which we use the
+shorthand $H$ and $T$), as well as the bogus atoms $\bar h \bar t$ and
+$ht$!  To preclude bogosity, the \texttt{constraint} declaration in
+Figure~\ref{epikfigure} describes the set of \emph{feasible atoms},
+which we parse into a formula $\varphi$ (if there are multiple
+\texttt{constraint} declarations, $\varphi$ is the conjunction of the
+parsed formulae in each). For a formula $\phi$, the set of feasible atoms is
+$\mathcal A_\phi$. In the coin example, the constraint $\varphi$ is
+$h\bar t + \bar h t$. Since $\bar h t \models \varphi$ and $h \bar t
+\models \varphi$, they are permitted feasible atoms, and since $\bar h
+\bar t$ and $h t$ do not satisfy $\phi$ they are not permitted.
 
@@
 
+%% \begin{example}   \label{gs1}
+%% \begin{tabular}[t]{ll}
+%% $v_1$ & $\bar p \bar q ~u~ \bar p q ~u~ q \bar p ~d~ \bar p q ~u~ q \bar p$ \\
+%% $v_2$ & $\bar p \bar q ~u~ \bar p q ~u~ p\bar q ~u~ pq$ \\   
+%% $w_1$ & $\bar h t$ \\
+%% $w_2$ & $\bar h t ~a_0~ \bar h t$ \\  
+%% $w_3$ & $\bar h t ~a_0~ \bar h t ~b_0~ \bar h t$ \\
+%% \end{tabular}
+%% \end{example}
 \begin{example}   \label{gs1}
 \begin{tabular}[t]{ll}
-$v_1$ & $\bar p \bar q ~u~ \bar p q ~u~ q \bar p ~d~ \bar p q ~u~ q \bar p$ \\
-$v_2$ & $\bar p \bar q ~u~ \bar p q ~u~ p\bar q ~u~ pq$ \\   
-$w_1$ & $\bar h t$ \\
-$w_2$ & $\bar h t ~a_0~ \bar h t$ \\  
-$w_3$ & $\bar h t ~a_0~ \bar h t ~b_0~ \bar h t$ \\
+$v_1$ & $\hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 ~u~ \hat 2$ \\
+$v_2$ & $\hat 0 ~u~ \hat 1 ~u~ \hat 2 ~u~ \hat 3$ \\   
+$w_1$ & $H$ \\
+$w_2$ & $H ~a_0~ H$ \\  
+$w_3$ & $T~a_0~ T ~b_0~ T$ \\
 \end{tabular}
@@
 \begin{example} \label{fusion1}
-\hspace{-2mm}\( \bar p q ~u~ \bar p q ~u~ p \bar q ~d~ \bar p q \cdot \bar p q ~u~ p \bar q = \bar p q ~u~ \bar p q ~u~ p \bar q ~d~ \bar p q ~u~ p \bar q \)
-
-\hspace{-2mm}\( \bar p \bar q ~u~ \bar p q ~u~ p \bar q ~d~ \bar p q \cdot p \bar q ~u~ pq = \mbox{\em{undefined}} \)
-
-\hspace{-2mm}\(\bar h t \cdot \bar h t ~a_0~ \bar h t = \bar h t ~a_0~ \bar h t \)
-
-\hspace{-2mm}\(\bar h t \cdot h \bar t ~a_1~ h \bar t = \mbox{\em{undefined}} \)
+\hspace{-2mm}\(\hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 \cdot \hat 1 ~u~ \hat 2 = \hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 ~u~ \hat 2 \)
+
+\hspace{-2mm}\( \hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 \cdot \hat 2 ~u~ \hat 3 = \mbox{\em{undefined}} \)
+
+\hspace{-2mm}\(H \cdot H ~a_0~ H = H ~a_0~ H \)
+
+\hspace{-2mm}\(T \cdot T ~a_1~ T = \mbox{\em{undefined}} \)
 \end{example}
@@
   e & \llbracket e \rrbracket^{\varphi, P} \\ \hline  
-  u & \{ \bar p \bar q ~u~ \bar p q, \bar p q ~u ~p \bar q,  p \bar q ~u~ pq\}\\
-  d & \{ \bar p q ~d~ \bar p \bar q, p \bar q~d~\bar p q, p q~d~ p \bar q\}\\
-  a_1 & \{h \bar t ~a_1~ h \bar t \}\\
-  a_0 & \{t \bar h ~a_0~ t \bar h\}\\
-  b_1 & \{h\bar t ~b_1~ h \bar t \}\\
-  b_0 & \{t \bar h ~b_0~ t \bar h \}\\
+  u & \{ \hat 0 ~u~ \hat 1, \hat 1 ~u~ \hat 2,  \hat 2 ~u~ \hat 3\}\\
+  d & \{ \hat 1 ~d~ \hat 0, \hat 2~d~\hat 1, \hat 3 ~d~ \hat 2\}\\
+  a_1 & \{H ~a_1~ H\}\\
+  a_0 & \{T ~a_0~ T\}\\
+  b_1 & \{H ~b_1~ H \}\\
+  b_0 & \{T ~b_0~ T \}
 \end{array}

### 2020-06-17 tex/draft.tex
add description of P

@@
 string model corresponds to the set of guarded strings where the bare
-event is flanked by compatible atoms. In standard KATs, this is  as defined in an Epik
-program by the event declaration and the constraint declaration.
-These have the status of event types, in that they can ``happen'' in
-different possible worlds. Here happening corresponds to incrementing
-the world with some compatible element of the event type.  Where $e$
-is a bare event symbol, $\llbracket e \rrbracket^{\varphi,P}$ is the
+event is flanked by compatible atoms. In standard KATs, every possible
+atom is allowed to surround every event, which leads to nonsensical
+strings like $\hat 1 ~u~\hat 0$ in the denotation of $u$, indicating
+that the elevator travelled up from floor 1 to floor 0 (a physical
+impossibility).
+
+Fortunately, Epik provides a syntax to specify the input-output
+behavior via the \texttt{event} declarations. From these we construct
+a function $P$ from agents to relations on atoms that specify the
+effects each action can have. If there are multiple event declarations
+for the same event, only the latter one is incorporated. To construct
+the effect relation $P(e)$ for event $e$, we add a pair $\alpha,
+\beta$ to the relation $P(e)$ for every disjunct $\phi:\phi'$ such
+that $\alpha \models \phi$ and $\beta \models \phi'$.  These have the
+status of event types, in that they can ``happen'' in different
+possible worlds. Here happening corresponds to incrementing the world
+with some compatible element of the event type.  Where $e$ is a bare
+event symbol, $\llbracket e \rrbracket^{\varphi,P}$ is the
 corresponding set of guarded strings, consisting of the bare event
@@
 \begin{example} \label{events1}
+  \renewcommand{\arraystretch}{1.1}
 \(
@@
 \)
+  \renewcommand{\arraystretch}{1}
+
 \end{example}

### 2020-06-17 tex/draft.tex
update algebraic stuff in section 2

@@
 
-Kleene Algebra with Tests has the algebraic signature
-$\tuple{K,+,\cdot,*,\bar{},0,1}$.  To this we add an unary operation
-$\Diamond_a$ for each agent, and a unary complent operation
-$\cdot^{c}$ on propositions.  This results in the signature
-$\tuple{K,+,\cdot,*,\bar{},0,1,{}^c,\Diamond_a,\Diamond_b}$ for
-epistemic KAT with two agents.
-
-We interpret each term in as a set of guarded strings as indicated in
-Figure~\ref{fig:interp}. A bare event $e$ is interpreted as the set of
-guarded strings allowed by $\varphi$ and $P$. The operation $p + q$ is
-set union. The constant $0$ is the empty set.  The constant 1 is the
-set of all truth assignments obeying the constraint on truth
-assignments, i.e. $\{10,01\}$ in the coin example.  The operation
-$\cdot$ is fusion product raised to sets: $x \cdot y$ is the set of
-all defined fusion products of an element of $x$ with and element of
-$y$.  The operation $*$ is Kleene star, with $x* = \cup_{i\geq 0}x^i$,
-where $x^i$ is the $i$-times product of $x$ with itself and $x^0=1$.
-Subsets of $1$ indicated by $t$ are also elements of $K$, and these
-form the Boolean algebra of tests, denoting all atoms admitted by the
-constraint $\varphi$ and the boolean test $t$.  We ``overload'' $\cdot$
-to further indicate boolean conjunction, $+$ to further indicate
-boolean disjunction, and $0,1$ to indicate the zero and one elements
-of the boolean algebra. The overbar operation denotes complement in
-this Boolean algebra: $\overline t$ denotes all atoms $\alpha$ such
-that $\alpha \Rightarrow \neg t$.  Other boolean operators (such as
+These operators combine to form the algebraic structure Kleene Algebra
+with Tests (\textsc{Kat}), which has the algebraic signature
+$\tuple{K,+,\cdot,*,\bar{},0,1}$.  We interpret each term in as a set
+of guarded strings as indicated in Figure~\ref{fig:interp}. A bare
+event $e$ is interpreted as the set of guarded strings allowed by
+$\varphi$ and $P$. The operation $p + q$ is set union. The constant
+$0$ is the empty set.  The constant 1 is the set of all truth
+assignments obeying the constraint on truth assignments,
+i.e. $\{10,01\}$ in the coin example.  The operation $\cdot$ is fusion
+product raised to sets: $x \cdot y$ is the set of all defined fusion
+products of an element of $x$ with and element of $y$.  The operation
+$*$ is Kleene star, with $x* = \cup_{i\geq 0}x^i$, where $x^i$ is the
+$i$-times product of $x$ with itself and $x^0=1$.  Subsets of $1$
+indicated by $t$ are also elements of $K$, and these form the Boolean
+algebra of tests, denoting all atoms admitted by the constraint
+$\varphi$ and the boolean test $t$.  We ``overload'' $\cdot$ to
+further indicate boolean conjunction, $+$ to further indicate boolean
+disjunction, and $0,1$ to indicate the zero and one elements of the
+boolean algebra. The overbar operation denotes complement in this
+Boolean algebra: $\overline t$ denotes all atoms $\alpha$ such that
+$\alpha \Rightarrow \neg t$.  Other boolean operators (such as
 implication $\Rightarrow$) are encoded as syntactic sugar in the
-standard way. The complement operation ${p}^c$ is complement at the
-level of sets of guarded strings, defined to every set the universe
-$\_^*$ except those denoted by $p$. Finally, the epistemic modality
-$\Diamond_a$ is interpreted using Kripke semantics, as pre-image
+standard way.
+
+To this we add an unary operation $\Diamond_a$ for each agent $a$, and a
+unary complent operation $\cdot^{c}$ on terms.  This results in the
+signature $\tuple{K,+,\cdot,*,\bar{},0,1,{}^c,\Diamond_a,\Diamond_b}$
+for epistemic KAT with two agents.  We interpret the epistemic
+epistemic modality $\Diamond_a$ using Kripke semantics, as pre-image
 relative to a fixed relation $R_a$ between guarded strings,
-$\Diamond_a x = \setabs{u}{\exists v.\element{v}{x} \wedge u R_a v}$.
+$\Diamond_a~x = \{u \mid \exists v. v \in x \wedge u \mathrel{R_a} v\}$.
 Here $u$ and $v$ are guarded strings, while $x$ is an element of $K$.
@@
 It remains to define the Kripke relation on guarded strings from an
-agent specification as in Figure \ref{epikfigure}. An an agent
-specification pairs each bare event with a set of bare events, and so
-determines a relation between bare events, call it relation $R_a$ for
-an agent $a$.  This determines a relation $\hat{R_a}$ between
-decorated events, see \ex{Rhat}.  It is generalized to a relation
-$\doublehat{R}$ on arbitrary guarded strings by relational Kleene
-star, as defined in \ex{Rhathat}. The operation $\cdot$ in the
-definition is KAT product, which enforces matching of tests.  The
-number of terms $n$ is construed as satisfying $n\geq 1$.  {\em What
-  is desired for $n=0$?}
-
-\begin{example}   \label{Rhat}
-\hspace{-2mm} \(
-\hat{R} =_{\mbox{\footnotesize{def.}}} 
-\{\tuple{u,v} | \exists c \exists d. cRd \wedge \element{u}{\hat{c}}  \wedge \element{v}{\hat{d}} \, \} \)
-\end{example}
+agent specification as in Figure \ref{epikfigure}. In Epik, an agent
+specification pairs each bare event with a set of ($+$-separated) bare
+events, and so determines a relation between bare events, call it
+relation $R_a$ for an agent $a$.  We can lift this to a relation
+$\hat{R_a}$ between decorated events for a given $\varphi$ and $P$,
+see \ex{Rhat}, and again to generalized to a relation $\doublehat{R}$
+on arbitrary guarded strings by relational Kleene star (as opposed to
+the Kleene star from aboce) , as defined in \ex{Rhathat}. The
+operation $\cdot$ in the definition is KAT product, which enforces
+matching of tests.
+
+%% \begin{example}   \label{Rhat}
+%% \hspace{-2mm} \(
+%% \hat{R} \triangleq
+%% \{\tuple{u,v} \mid \exists c \exists d. c\mathrel R d \wedge u \in \hat{c}  \wedge v \in \hat{d} \, \} \)
+%% \end{example}
 
@@
 \hspace{-2mm}\(
-\doublehat{R} =_{\mbox{\footnotesize{def.}}} 
+\doublehat{R} \triangleq
 \hspace{-2mm}\begin{array}[t]{l}
-\{ \tuple{x,y}| \exists u_1 ... \exists u_n \exists v_1 ... \exists v_n. \\
-\hspace{6mm} u_1\hat{R}v_1 \wedge ... \wedge u_n\hat{R}v_n \wedge\\
-\hspace{6mm} x = u_1 \cdot ... \cdot u_n \hspace{1mm} \wedge \\
-\hspace{6mm} y = v_1 \cdot ... \cdot v_n \hspace{2mm} \}
+\{ \tuple{x,y} \mid \exists u_1 ... \exists u_n \exists v_1 ... \exists v_n. \\
+\hspace{6mm} u_1\mathrel{R}v_1 \wedge ... \wedge u_n\mathrel{R}v_n \wedge\\
+\hspace{6mm} x \in \llbracket u_1 \cdot ... \cdot u_n \rrbracket^{\varphi,P}\hspace{1mm} \wedge \\
+\hspace{6mm} y \in \llbracket v_1 \cdot ... \cdot v_n \rrbracket^{\varphi,P} \hspace{2mm} \}
 \end{array}
@@
 
-Universal box modalites are defined as duals, $\Box_a x = (\Diamond_a
-x^c)^c$.  For instance, Aly is certain that the coin is heads if and
-only if she does not consider it possible that it is not heads.
-
-Summing up, given an Epik program with $n$ agents, we construct a concrete guarded
-string model $\tuple{K,+,\cdot,*,\bar{},0,1,{}^c,\Diamond_1,...,\Diamond_n}$.  The
-elements of $K$ are interpreted as propositions.  $0^c$ is the set of worlds, and it
-may be countably infinite.  $\Diamond_i$ is an epistemic modality for the $i$th
-agent.  Or referring to the Kripke relations $\doublehat{R}_i$, the construction
-defines a multi-agent Kripke frame $\tuple{0^c,\doublehat{R}_1,...,\doublehat{R}_n}$
-(usually a countable one) from an Epik specification. The frame consists of a set of
-worlds, and an epistemic-alternative relation for each agent.  These models are used
-as a target for natural-language interpretation in Section 5 and Section 6, where we
-obtain semantic values such as \sem{Amy knows that it's heads and Bob knows that Amy
-  knows whether it is heads or tails, and does not know that it's heads} as elements
-of $K$. Concretely the proposositions are sets of guarded strings (usually countable
-ones), construed as sets of worlds as they figure in possible worlds semantics for
-natural language.
+The complement operation ${p}^c$ is complement at the level of sets of
+guarded strings, defined to every set the universe $\_^*$ except those
+denoted by $p$. Note that $\llbracket 0^c \rrbracket^{\varphi,P} =
+\llbracket\_^*\rrbracket^{\varphi,P}$. We use the complement to define
+universal box modalites as the dual of the diamond modality, $\Box_a x
+= (\Diamond_a x^c)^c$.  For instance, Aly is certain that the coin is
+heads if and only if she does not consider it possible that it is not
+heads.
+
+Summing up, given an Epik program with $n$ agents, we construct a
+concrete guarded string model
+$\tuple{K,+,\cdot,*,\bar{},0,1,{}^c,\Diamond_{a_1},...,\Diamond_{a_n}}$.
+The elements of $K$ are interpreted as propositions.  $0^c$ is the set
+of worlds, and it may be countably infinite.  $\Diamond_{a_i}$ is an
+epistemic modality for the $i$th agent.  Or referring to the Kripke
+relations $\doublehat{R}_i$, the construction defines a multi-agent
+Kripke frame $\tuple{0^c,\doublehat{R}_1,...,\doublehat{R}_n}$
+(usually a countable one) from an Epik specification. The frame
+consists of a set of worlds, and an epistemic-alternative relation for
+each agent.  These models are used as a target for natural-language
+interpretation in Section 5 and Section 6, where we obtain semantic
+values such as \sem{Amy knows that it's heads and Bob knows that Amy
+  knows whether it is heads or tails, and does not know that it's
+  heads} as elements of $K$. Concretely the proposositions are sets of
+guarded strings (usually countable ones), construed as sets of worlds
+as they figure in possible worlds semantics for natural language.
 

### 2020-06-18 tex/draft.tex
Edits at the end of Section 1.

@@
 can be described by Boolean formulas, with $h$ being the precondition
-of $a_1$, and $!(pq)$ being the precondition of $u$. Juxtaposition is
-used for Boolean conjunction, and the exclamation point for for
+of $a_1$, and $\overline{pq}$ being the precondition of $u$. Juxtaposition is
+used for Boolean conjunction, and the overbar for
 Boolean negation. Events come as well with a relation between prior
@@
 a_1 & h:h \\
-a_0 & \ev{(!h)}{(!h)} \\
-% u & (!q!p) \! : \!(!qp) + \\
-u & \ev{(!q!p)}{(!qp)} + \ev{(!qp)}{(q!p)} + \ev{(q!p)}{(qp)} \\
-d & \evr{(!q!p)}{(!qp)} + \evr{(!qp)}{(q!p)} + \evr{(q!p)}{(qp)} \\ 
+a_0 & \ev{\bar{h}}{\bar{h}} \\
+% u & (\bar{q}\bar{p}) \! : \!(\bar{q}p) + \\
+u & \ev{(\bar{q}\bar{p})}{(\bar{q}p)} + \ev{(\bar{q}p)}{(q\bar{p})} + \ev{(q\bar{p})}{(qp)} \\
+d & \evr{(\bar{q}\bar{p})}{(\bar{q}p)} + \evr{(\bar{q}p)}{(q\bar{p})} + \evr{(q\bar{p})}{(qp)} \\ 
 \end{array} \)
 \end{example}
+
+In this discussion, a sequence such as $\bar{q}{p}$ can be viewed as a
+valuation of the primitive propositions $q$ and $p$, which is used to
+describe world state. The primitives are listed in fixed order, and
+left unmarked (indicating true) or marked with the overbar (indicating
+false).  Suppose that in the coin example, we had an additional
+primitive stative proposition (or atomic test) $t$, interpreted as
+tails.  Since a coin is heads or tails but not both, we want to allow
+the valuations $h\bar{t}$ and $\bar{h}t$, and disallow $ht$ and
+$\bar{h}\bar{t}$.  This is enforced by a {\em state formula},
+which is a simply Boolean formula, in this case \exi{effect}{a}.
+Where
+{\sf{B}} is a set of atomic tests and $\rho$ is a state constraint
+over {\sf{B}}, ${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\rho}$ is the set of valuations of
+{\sf{B}} that make formula $\rho$ true.
+
+Formulas like the ones in \ex{event1} that describe pre- and
+post-conditions are {\em effect formulas}. They are interpreted as
+defining relations between valuations, as defined in Figure
+\ref{effectfigure}.  The valuations they relate are constrained by the
+state formula as well.  For the heads-tails example, let the state
+formula and the effect formula for $a_1$ (Aly peeking at heads) be as
+specified in \ex{effect}. Then the set of valuations determined by
+$\rho$ and the relation on valuations for the event $u$ are \sem{rho}
+and $\sem{\zeta}^{\sem{rho}}$ as given at the bottom in \ex{6}.  Given
+a state formula $\rho$, the the relation characterizing pre- and
+post-conditions of an event with effect formula $\zeta$ is defined to
+be $\sem{\zeta}^{\sem{\rho}}$.
+
+\begin{example}   \label{effect}
+\begin{tabular}[t]{l}
+State formula ($\rho$)  $h\bar{t} + \bar{h}t$ \\
+Effect formula for $a_1$ ($\zeta$)  $\ev{h}{h}$ \\
+\( \sem{\rho} = \{h\bar{t},\bar{h}t\} \)  \\
+\( \sem{\zeta}^{\sem{\rho}} = \{\pair{h\bar{t}}{h\bar{t}}\} \)  \\
+\end{tabular}
+\end{example}
+
+\begin{figure}
+\hspace{-2mm}\begin{tabular}{lll}
+  state formulas & $\rho,\sigma...$ & $\rho{:}{:}{=} \, \begin{array}[t]{l}\tuple{\mbox{atomic tests}}\,|\\
+    0\,|\,1\,| \; \rho{+}\sigma\;| \\ \rho\,{\&}\,\sigma\;|\;\bar{\rho}\end{array}$ \\
+  effect formulas & $\zeta,\eta...$ & $\zeta{:}{:}{=} \, \begin{array}[t]{l} \rho\,{:}\,\sigma \,|\, \zeta{+}\eta \\
+    \zeta\,{\&}\,\eta\;|\;\bar{\zeta}\end{array}$
+\end{tabular}
+\vspace{1mm}
+
+\(
+\begin{array}{cll}
+  \sem{\rho} & = & \mbox{Set of Boolean valuations of} \\
+            && \mbox{the atomic tests that verify $\rho$.} \\
+\sem{\rho\,{:}\,\sigma}^V & = & (\sem{\rho}\cap V)\times(\sem{\sigma}\cap V)\\
+\sem{\zeta{+}\eta}^V & = & \sem{\zeta}^V\cup\sem{\eta}^V\\
+\sem{\zeta{\&}\eta}^V & = & \sem{\zeta}^V\cap\sem{\eta}^V\\
+\sem{\bar{\zeta}}^V & = & V \times V \setminus \sem{\zeta}^V\\
+\end{array}  
+\)
+
+\caption{Syntax and semantics of state formulas and effect formulas. Effect formulas
+  are evaluated relative to as set Boolean valuations $V$, and denote relations between
+  Boolean valuations.}
+\label{effectfigure}
+\end{figure}
 
@@
 Additional syntax and semantics is included to model multi-agent
-epistemic semantics.  Guarded strings over a finite alphabet $E$ are
+epistemic semantics.  Guarded strings over a finite alphabet {\sf{P}} are
 like ordinary strings, but with truth assignments to a set $T$ of
@@
 Figure~\ref{fig:interp}, and the examples in \ex{events1}.
-% u (!q!p):(!qp) + (!qp):(q!p) + (q!p):(qp) 173
-% d (!qp):(!q!p) + (q!p):(!qp) + (qp):(q!p)
+% u (\bar{q}\bar{p}):(\bar{q}p) + (\bar{q}p):(q\bar{p}) + (q\bar{p}):(qp) 173
+% d (\bar{q}p):(\bar{q}\bar{p}) + (q\bar{p}):(\bar{q}p) + (qp):(q\bar{p})
 

### 2020-06-18 tex/draft.tex
add separate language section

@@
 specified in \ex{effect}. Then the set of valuations determined by
-$\rho$ and the relation on valuations for the event $u$ are \sem{rho}
-and $\sem{\zeta}^{\sem{rho}}$ as given at the bottom in \ex{6}.  Given
-a state formula $\rho$, the the relation characterizing pre- and
-post-conditions of an event with effect formula $\zeta$ is defined to
-be $\sem{\zeta}^{\sem{\rho}}$.
+$\rho$ and the relation on valuations for the event $u$ are
+$\sem{\rho}$ and $\sem{\zeta}^{\sem{\rho}}$ as given at the bottom in
+\ex{6}.  Given a state formula $\rho$, the the relation characterizing
+pre- and post-conditions of an event with effect formula $\zeta$ is
+defined to be $\sem{\zeta}^{\sem{\rho}}$.
 
@@
   \sem{\rho} & = & \mbox{Set of Boolean valuations of} \\
-            && \mbox{the atomic tests that verify $\rho$.} \\
+            && \mbox{the atomic tests that satisfy $\rho$.} \\
 \sem{\rho\,{:}\,\sigma}^V & = & (\sem{\rho}\cap V)\times(\sem{\sigma}\cap V)\\
@@
 Additional syntax and semantics is included to model multi-agent
-epistemic semantics.  Guarded strings over a finite alphabet {\sf{P}} are
-like ordinary strings, but with truth assignments to a set $T$ of
+epistemic semantics.  Guarded strings over a finite alphabet {\sf{P}}
+are like ordinary strings, but with truth assignments to a set $T$ of
 primitive propositions (primitive tests) alternating with the symbols
@@
 \texttt{constraint} declarations, $\varphi$ is the conjunction of the
-parsed formulae in each). For a formula $\phi$, the set of feasible atoms is
-$\mathcal A_\phi$. In the coin example, the constraint $\varphi$ is
-$h\bar t + \bar h t$. Since $\bar h t \models \varphi$ and $h \bar t
-\models \varphi$, they are permitted feasible atoms, and since $\bar h
-\bar t$ and $h t$ do not satisfy $\phi$ they are not permitted.
+parsed formulae in each). For a formula $\varphi$, the set of atoms
+that satisfy $\varphi$ is $\mathcal A_\varphi$ -- these are the
+feasible atoms. In the coin example, the constraint $\varphi$ is
+$h\bar t + \bar h t$. Since $\bar h t$ satisfies $\varphi$ and $h \bar
+t$ satisfies $\varphi$, they are permitted feasible atoms, and since
+$\bar h \bar t$ and $h t$ do not satisfy $\varphi$ they are not
+permitted. We omit boolean satisfiability since its definition is
+standard.
 
@@
 \begin{example} \label{fusion1}
-\hspace{-2mm}\(\hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 \cdot \hat 1 ~u~ \hat 2 = \hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 ~u~ \hat 2 \)
-
-\hspace{-2mm}\( \hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 \cdot \hat 2 ~u~ \hat 3 = \mbox{\em{undefined}} \)
-
-\hspace{-2mm}\(H \cdot H ~a_0~ H = H ~a_0~ H \)
-
-\hspace{-2mm}\(T \cdot T ~a_1~ T = \mbox{\em{undefined}} \)
+\hspace{-2mm}\(\hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 \diamond \hat 1 ~u~ \hat 2 = \hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 ~u~ \hat 2 \)
+
+\hspace{-2mm}\( \hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 \diamond \hat 2 ~u~ \hat 3 = \mbox{\em{undefined}} \)
+
+\hspace{-2mm}\(H \diamond H ~a_0~ H = H ~a_0~ H \)
+
+\hspace{-2mm}\(T \diamond T ~a_1~ T = \mbox{\em{undefined}} \)
 \end{example}
@@
 the effect relation $P(e)$ for event $e$, we add a pair $\alpha,
-\beta$ to the relation $P(e)$ for every disjunct $\phi:\phi'$ such
-that $\alpha \models \phi$ and $\beta \models \phi'$.  These have the
+\beta$ to the relation $P(e)$ for every disjunct $\varphi:\varphi'$ such
+that $\alpha$ satisfies $\phi$ and $\beta$ satisfies $\varphi'$.  These have the
 status of event types, in that they can ``happen'' in different
@@
 
-These operators combine to form the algebraic structure Kleene Algebra
-with Tests (\textsc{Kat}), which has the algebraic signature
-$\tuple{K,+,\cdot,*,\bar{},0,1}$.  We interpret each term in as a set
-of guarded strings as indicated in Figure~\ref{fig:interp}. A bare
-event $e$ is interpreted as the set of guarded strings allowed by
-$\varphi$ and $P$. The operation $p + q$ is set union. The constant
-$0$ is the empty set.  The constant 1 is the set of all truth
-assignments obeying the constraint on truth assignments,
-i.e. $\{10,01\}$ in the coin example.  The operation $\cdot$ is fusion
-product raised to sets: $x \cdot y$ is the set of all defined fusion
-products of an element of $x$ with and element of $y$.  The operation
-$*$ is Kleene star, with $x* = \cup_{i\geq 0}x^i$, where $x^i$ is the
-$i$-times product of $x$ with itself and $x^0=1$.  Subsets of $1$
-indicated by $t$ are also elements of $K$, and these form the Boolean
-algebra of tests, denoting all atoms admitted by the constraint
-$\varphi$ and the boolean test $t$.  We ``overload'' $\cdot$ to
-further indicate boolean conjunction, $+$ to further indicate boolean
-disjunction, and $0,1$ to indicate the zero and one elements of the
-boolean algebra. The overbar operation denotes complement in this
-Boolean algebra: $\overline t$ denotes all atoms $\alpha$ such that
-$\alpha \Rightarrow \neg t$.  Other boolean operators (such as
-implication $\Rightarrow$) are encoded as syntactic sugar in the
-standard way.
-
-To this we add an unary operation $\Diamond_a$ for each agent $a$, and a
-unary complent operation $\cdot^{c}$ on terms.  This results in the
-signature $\tuple{K,+,\cdot,*,\bar{},0,1,{}^c,\Diamond_a,\Diamond_b}$
-for epistemic KAT with two agents.  We interpret the epistemic
-epistemic modality $\Diamond_a$ using Kripke semantics, as pre-image
-relative to a fixed relation $R_a$ between guarded strings,
-$\Diamond_a~x = \{u \mid \exists v. v \in x \wedge u \mathrel{R_a} v\}$.
-Here $u$ and $v$ are guarded strings, while $x$ is an element of $K$.
+\section{A Language for Epistemic Sets of Guarded Strings}
+
+To reason about epistemic alternatives, guarded strings are used to
+represent different possible worlds, so agents' knowledge/beliefs are
+represented by sets of guarded strings. The standard algebra for
+manipulating guarded strings is Kleene Algebra with Tests
+(\textsc{KAT}), which has the algebraic signature
+$\tuple{K,+,;,*,\bar{},0,1}$ \cite{kozen2001automata}. However, the
+KATs don't permit native reasoning about modal operators, so Epik's
+signature needs to subsume the KAT operators.
+
+Figure~\ref{fig:epik-grammar} depicts the syntax of Epik's formal
+term language. As in KAT, $0$ is the null element, $1$ is the unit
+element, $+$ indicates disjuction, $;$ indicates event sequencing or
+boolean conjunction, $!$ indicates boolean negation, and $*$ indicates
+iteration. We add some additional operations to reason about epistemic
+alternatives: the term $\Diamond_a~p$ represents epistemic possibility
+for agent $a$ in worlds $p$, and complement $\neg p$ represents the
+worlds not described by $p$. We can encode $\Box_a~p$, i.e. epistemic
+certainty for agent $a$ in worlds $p$, by $\Box_a~p =
+(\Diamond_a~p^c)^c$.
+
+\begin{figure}
+
+  \[\begin{array}{l>{\hspace{-.5em}}c>{\hspace{-0.75em}}l}
+  \multicolumn 2 l {\textrm{tests}}& a~\textrm{is an atomic test} \\
+  t,s & ::=& a~|~0~|~1~|~t+t~|~t;t~|~!t\\
+  \multicolumn 2 l {\textrm{terms}} & e~\textrm{is an atomic event} \\
+  p,q & ::=& e~|~0~|~1~|~t~|~p+p~|~p;p~|~p^*~|\neg p~|\Diamond_a~p
+  \end{array}\]
+
+  Derived Operators\\
+  \mbox{}\qquad \(\Box_a p \triangleq \neg\Diamond_a\neg p \qquad p\&q \triangleq \neg(\neg p + \neg q) \)
+
+\caption{The language of Epik terms and key derived
+  operators}
+\label{fig:epik-grammar}
+\label{termfigure}
+\end{figure}
+
+We interpret each term in as a set of guarded strings as indicated in
+Figure~\ref{fig:interp}. The function $\llbracket p
+\rrbracket^{\varphi,P}$ interprets a term $p$ as a set of guarded
+strings consistent with state constraints $\varphi$ and effect
+relation $P$. A bare event $e$ is interpreted as the set of guarded
+strings allowed by $\varphi$ and $P$. The constant $0$ is interpreted
+as the empty set.  The constant $1$ is interpreted as the set of all
+atoms obeying $\varphi$, i.e. $\{h\bar t, \bar h t\}$ in the coin
+example. The operation $p + q$ is interpreted as the union of the
+interpretations of $p$ and $q$. The operation $p;q$ interpreted as the
+fusion product of the interpretations of $p$ and $q$ raised to sets:
+$X \diamond Y = \{ x \diamond y \mid x \in X, y \in Y, x\diamond
+y~\textit{defined}\}$.  The operation $p*$ is Kleene star, which is
+interpreted to be least upper bound of the iterated fusion product of
+the denotation of $p$ with itself. Subsets of $1$ indicated by $t$ are
+also elements of $K$, and these form the Boolean algebra of tests,
+denoting all atoms admitted by the constraint $\varphi$ and the
+boolean test $t$.  We ``overload'' $;$ to further indicate boolean
+conjunction, $+$ to further indicate boolean disjunction, and $0,1$ to
+indicate the zero and one elements of the boolean algebra. The overbar
+operation denotes complement in this Boolean algebra: $\overline t$
+denotes all atoms $\alpha$ such that $\alpha \Rightarrow \neg t$.
+Other boolean operators (such as implication $\Rightarrow$) are
+encoded as syntactic sugar in the standard way. A boolean test $t$
+denotes the set of atoms that satisfy $t;\varphi$, that is, they
+satisfy both $t$ and $\varphi$.
+
+The unary epistemic alternative operation $\Diamond_a$ for each agent
+$a$, is interpreted using using Kripke semantics, as pre-image
+relative to a fixed relation $\hat{R_a}$ between guarded strings,
+$\Diamond_a~x = \{u \mid \exists v. v \in x \wedge u \mathrel{R_a}
+v\}$.  Here $u$ and $v$ are guarded strings, while $x$ is an element
+of $K$.
 
@@
 relation $R_a$ for an agent $a$.  We can lift this to a relation
-$\hat{R_a}$ between decorated events for a given $\varphi$ and $P$,
-see \ex{Rhat}, and again to generalized to a relation $\doublehat{R}$
-on arbitrary guarded strings by relational Kleene star (as opposed to
-the Kleene star from aboce) , as defined in \ex{Rhathat}. The
-operation $\cdot$ in the definition is KAT product, which enforces
+$\hat{R_a}$ between guarded strings, as defined in \ex{Rhathat}. The
+operation $(\mathop;)$ in the definition is KAT product, which enforces
 matching of tests.
@@
 \hspace{-2mm}\(
-\doublehat{R} \triangleq
+\hat{R} \triangleq
 \hspace{-2mm}\begin{array}[t]{l}
@@
 \hspace{6mm} u_1\mathrel{R}v_1 \wedge ... \wedge u_n\mathrel{R}v_n \wedge\\
-\hspace{6mm} x \in \llbracket u_1 \cdot ... \cdot u_n \rrbracket^{\varphi,P}\hspace{1mm} \wedge \\
-\hspace{6mm} y \in \llbracket v_1 \cdot ... \cdot v_n \rrbracket^{\varphi,P} \hspace{2mm} \}
+\hspace{6mm} x \in \llbracket u_1 ; ... ; u_n \rrbracket^{\varphi,P}\hspace{1mm} \wedge \\
+\hspace{6mm} y \in \llbracket v_1 ; ... ; v_n \rrbracket^{\varphi,P} \hspace{2mm} \}
 \end{array}
@@
   \llbracket 0 \rrbracket^{\varphi,P} && \{\} \\  
-  \llbracket 1 \rrbracket^{\varphi,P} && \{\alpha \mid \alpha \Rightarrow \varphi\}\\
-  \llbracket e \rrbracket^{\varphi,P}&& \{\alpha \cdot e \cdot \beta \mid
+  \llbracket 1 \rrbracket^{\varphi,P} && \{\alpha \mid \alpha \in \mathcal A_\varphi\}\\
+  \llbracket e \rrbracket^{\varphi,P}&& \{\alpha~e~\beta \mid
   (\alpha,\beta) \in P(e), \alpha, \beta \Rightarrow \varphi \}
@@
     \dom(P)}\llbracket e \rrbracket^{\varphi,P}\\
-  \llbracket t \rrbracket^{\varphi,P} &&  \{\alpha \mid \alpha \Rightarrow \varphi \wedge t\}\\
+  \llbracket t \rrbracket^{\varphi,P} &&  \{\alpha \mid \alpha \in \mathcal A_{\varphi;t}\}\\
   \llbracket{p+q}\rrbracket^{\varphi,P} && \llbracket{p}\rrbracket^{\varphi,P} \cup \llbracket q \rrbracket^{\varphi,P}\\
-  \llbracket{p \& q}\rrbracket^{\varphi,P} && \llbracket{p}\rrbracket^{\varphi,P} \cap \llbracket q \rrbracket^{\varphi,P}\\
-  \llbracket{p;q}\rrbracket^{\varphi,P} && \llbracket p \rrbracket^{\varphi,P} \cdot \llbracket{q}\rrbracket^{\varphi,P}\\
-  \llbracket p^*\rrbracket^{\varphi,P} && \displaystyle\bigcup_n {\underbrace{\llbracket p \rrbracket^{\varphi,P}\cdot \ldots \cdot \llbracket p \rrbracket^{\varphi,P}}_{n}} \\
-  \llbracket \sim p \rrbracket^{\varphi,P} && \llbracket \_^*\rrbracket^{\varphi,P} \setminus \llbracket p \rrbracket^{\varphi,P} \\
-  \llbracket \langle a \rangle~p \rrbracket && \doublehat{R_a}\llbracket p \rrbracket
+  \llbracket{p;q}\rrbracket^{\varphi,P} && \llbracket p \rrbracket^{\varphi,P} \diamond \llbracket{q}\rrbracket^{\varphi,P}\\
+  \llbracket p^*\rrbracket^{\varphi,P} && \displaystyle\bigcup_n {\underbrace{\llbracket p \rrbracket^{\varphi,P}\diamond \ldots \diamond \llbracket p \rrbracket^{\varphi,P}}_{n}} \\
+  \llbracket \neg p \rrbracket^{\varphi,P} && \llbracket \_^*\rrbracket^{\varphi,P} \setminus \llbracket p \rrbracket^{\varphi,P} \\
+  \llbracket \langle a \rangle~p \rrbracket && \{x \mid x~\hat{R_a}~y, y \in \llbracket p \rrbracket\}
 
@@
 
-Summing up, given an Epik program with $n$ agents, we construct a
-concrete guarded string model
-$\tuple{K,+,\cdot,*,\bar{},0,1,{}^c,\Diamond_{a_1},...,\Diamond_{a_n}}$.
-The elements of $K$ are interpreted as propositions.  $0^c$ is the set
-of worlds, and it may be countably infinite.  $\Diamond_{a_i}$ is an
-epistemic modality for the $i$th agent.  Or referring to the Kripke
-relations $\doublehat{R}_i$, the construction defines a multi-agent
-Kripke frame $\tuple{0^c,\doublehat{R}_1,...,\doublehat{R}_n}$
-(usually a countable one) from an Epik specification. The frame
-consists of a set of worlds, and an epistemic-alternative relation for
-each agent.  These models are used as a target for natural-language
-interpretation in Section 5 and Section 6, where we obtain semantic
-values such as \sem{Amy knows that it's heads and Bob knows that Amy
-  knows whether it is heads or tails, and does not know that it's
-  heads} as elements of $K$. Concretely the proposositions are sets of
-guarded strings (usually countable ones), construed as sets of worlds
-as they figure in possible worlds semantics for natural language.
+Summing up, given an Epik program with $n$ agents, we construct state
+constraints $\varphi$, effect relations $P$, and a concrete guarded
+string model for each term using $\llbracket -
+\rrbracket^{\varphi,P}$.
+%% $\tuple{K,+,;,*,\bar{},0,1,{}^c,\Diamond_{a_1},...,\Diamond_{a_n}}$.
+ $0^c$ is the set of worlds, and it may be countably infinite.
+$\Diamond_{a_i}$ is an epistemic modality for the $i$th agent.  Or
+referring to the Kripke relations $\hat{R}_i$, the construction
+defines a multi-agent Kripke frame
+$\tuple{0^c,\hat{R}_1,...,\hat{R}_n}$ (usually a countable one) from
+an Epik specification. The frame consists of a set of worlds, and an
+epistemic-alternative relation for each agent.  These models are used
+as a target for natural-language interpretation in Section 5 and
+Section 6, where we obtain semantic values such as \sem{Amy knows that
+  it's heads and Bob knows that Amy knows whether it is heads or
+  tails, and does not know that it's heads} as elements of
+$K$. Concretely the proposositions are sets of guarded strings
+(usually countable ones), construed as sets of worlds as they figure
+in possible worlds semantics for natural language.
 
@@
 
-\begin{figure}
-\begin{tabular}{lll}
-agents & $a,b...$ & $a {:}{:}{=} \, \tuple{\mbox{atomic agents}}$
-\end{tabular}
-
-\begin{tabular}{lll}
-tests & $T,S...$ & $T{:}{:}{=} \, \begin{array}[t]{l}\tuple{\mbox{atomic tests}}\,|\,0\,|\,1\,|\\ T{+}S\;|\;T\,{;}\,S\;|\;!T\end{array}$
-\end{tabular}
-
-\begin{tabular}{lll}
-propositions & $p,q...$ & \(p{:}{:}{=} \,
-\begin{array}[t]{l}\tuple{\mbox{atomic events}}\,|\,\_\, |\\
-T \,|\, 
-p{+}q \,|\,
-p\mbox{\&}q \,|\, \\
-p{;}q \;|\;
-p{*} \;|\; {\sim}p \;|\\
-{\langle a \rangle}~p
-\end{array}\)
-\end{tabular}
-
-\caption{The language of Epik terms. Propositional terms are interpreted
-  as sets of guarded strings.}
-\label{fig:epik-grammar}
-\label{termfigure}
-\end{figure}
-
 An architecture of interpretation by transtlation is employed, where English

### 2020-06-18 tex/draft.tex
restructuring 1-3

@@
 can be described by Boolean formulas, with $h$ being the precondition
-of $a_1$, and $\overline{pq}$ being the precondition of $u$. Juxtaposition is
-used for Boolean conjunction, and the overbar for
-Boolean negation. Events come as well with a relation between prior
-and following state, for instance with $u$ incrementing the floor.
-This is expressed using an operator ``$:$'' (read ``and next'') that
-pairs Boolean formulas. The first line in \ex{event1} describes $a_1$
-(Amy looking at heads) as happening only in an $h$ state, and as not
-changing the state.  Symmetrically, $a_0$ (Amy looking at tails) can
-happen only in a not-$h$ state, and does not change the state.  The
-third line says that $u$ increments the floor, and can happen only on
-floors 0, 1, and 2.  The fourth line describes $d$ in similar terms.
-Plus is disjunction.
+of $a_1$, and $\overline{pq}$ being the precondition of
+$u$. Juxtaposition is used for Boolean conjunction, and the overbar
+for Boolean negation. Events come as well with a relation between
+prior and following state, for instance with $u$ incrementing the
+floor.  This is expressed using an operator ``$:$'' (read ``and
+next'') that pairs Boolean formulas. The first line in \ex{event1}
+describes $a_1$ (Amy looking at heads) as happening only in an $h$
+state, and as not changing the state.  Symmetrically, $a_0$ (Amy
+looking at tails) can happen only in a not-$h$ state, and does not
+change the state.  The third line says that $u$ increments the floor,
+and can happen only on floors 0, 1, and 2.  The fourth line describes
+$d$ in similar terms.  Plus is disjunction.
 
@@
 the valuations $h\bar{t}$ and $\bar{h}t$, and disallow $ht$ and
-$\bar{h}\bar{t}$.  This is enforced by a {\em state formula},
-which is a simply Boolean formula, in this case \exi{effect}{a}.
-Where
+$\bar{h}\bar{t}$.  This is enforced by a {\em state formula}, which is
+a simply Boolean formula, in this case \exi{effect}{a}.  Where
 {\sf{B}} is a set of atomic tests and $\rho$ is a state constraint
-over {\sf{B}}, ${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\rho}$ is the set of valuations of
-{\sf{B}} that make formula $\rho$ true.
+over {\sf{B}}, ${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\rho}$ is the set of
+valuations of {\sf{B}} that make formula $\rho$ true. Following
+\cite{kozen2001automata}, we call the valuations
+\emph{atoms}.
 
@@
 specified in \ex{effect}. Then the set of valuations determined by
-$\rho$ and the relation on valuations for the event $u$ are
-$\sem{\rho}$ and $\sem{\zeta}^{\sem{\rho}}$ as given at the bottom in
-\ex{6}.  Given a state formula $\rho$, the the relation characterizing
-pre- and post-conditions of an event with effect formula $\zeta$ is
-defined to be $\sem{\zeta}^{\sem{\rho}}$.
+$\varphi$ and the relation on valuations for the event $u$ are
+$\mathcal A_{\mathsf B}^\varphi$ and
+$\llbracket\zeta\rrbracket^\varphi$ as given at the bottom in \ex{6}.
+Given a state formula $\rho$, the the relation characterizing pre- and
+post-conditions of an event with effect formula $\zeta$ is defined to
+be $\sem{\zeta}^{\sem{\rho}}$.
 
@@
 \begin{tabular}[t]{l}
-State formula ($\rho$)  $h\bar{t} + \bar{h}t$ \\
-Effect formula for $a_1$ ($\zeta$)  $\ev{h}{h}$ \\
-\( \sem{\rho} = \{h\bar{t},\bar{h}t\} \)  \\
-\( \sem{\zeta}^{\sem{\rho}} = \{\pair{h\bar{t}}{h\bar{t}}\} \)  \\
+State formula: $\varphi = h\bar{t} + \bar{h}t$ \\
+Effect formula for $a_1$: $\zeta = h \mathbin{:} h$ \\
+\( \mathcal{A}^\varphi_{\mathsf B} = \{h\bar{t},\bar{h}t\} \)  \\
+\( \sem{\zeta}^\varphi = \{\pair{h\bar{t}}{h\bar{t}}\} \)  \\
 \end{tabular}
@@
 \begin{figure}
-\hspace{-2mm}\begin{tabular}{lll}
-  state formulas & $\rho,\sigma...$ & $\rho{:}{:}{=} \, \begin{array}[t]{l}\tuple{\mbox{atomic tests}}\,|\\
-    0\,|\,1\,| \; \rho{+}\sigma\;| \\ \rho\,{\&}\,\sigma\;|\;\bar{\rho}\end{array}$ \\
-  effect formulas & $\zeta,\eta...$ & $\zeta{:}{:}{=} \, \begin{array}[t]{l} \rho\,{:}\,\sigma \,|\, \zeta{+}\eta \\
-    \zeta\,{\&}\,\eta\;|\;\bar{\zeta}\end{array}$
-\end{tabular}
-\vspace{1mm}
-
-\(
+  \[\begin{array}{lcl}
+  \multicolumn 3 l {\textrm{state formulas}\hfill (a \in \mathsf{B}) \hfill \mbox{}}\\
+  \varphi,\psi &::=& a \mid0 \mid 1 \mid\varphi{+}\psi \mid \varphi~\psi \mid \bar{\varphi} \\
+  \multicolumn {3} l {\textrm{effect formulas}}\\
+  \zeta,\eta & ::= & \varphi \mathbin{:} \sigma \mid \zeta{+}\eta
+    \mid \zeta \mathbin{\&} \eta \mid \bar{\zeta}
+  \end{array}\]
+
+\[
 \begin{array}{cll}
-  \sem{\rho} & = & \mbox{Set of Boolean valuations of} \\
-            && \mbox{the atomic tests that satisfy $\rho$.} \\
-\sem{\rho\,{:}\,\sigma}^V & = & (\sem{\rho}\cap V)\times(\sem{\sigma}\cap V)\\
-\sem{\zeta{+}\eta}^V & = & \sem{\zeta}^V\cup\sem{\eta}^V\\
-\sem{\zeta{\&}\eta}^V & = & \sem{\zeta}^V\cap\sem{\eta}^V\\
-\sem{\bar{\zeta}}^V & = & V \times V \setminus \sem{\zeta}^V\\
+  \llbracket\psi_1\mathbin{:}\psi_2\rrbracket^\varphi & = &
+  \mathcal A^{(\varphi~\psi_1)}_{\mathsf{B}}\times\mathcal A^{(\varphi~\psi_2)}_{\mathsf{B}}\\[.1em]
+\llbracket\zeta{+}\eta\rrbracket^\varphi & = & \llbracket\zeta\rrbracket^\varphi\cup\llbracket\eta\rrbracket^\varphi\\
+\llbracket\zeta{\&}\eta\rrbracket^\varphi & = & \llbracket\zeta\rrbracket^\varphi\cap\llbracket\eta\rrbracket^\varphi\\
+\llbracket\bar{\zeta}\rrbracket^\varphi & = & V \times V \setminus \llbracket\zeta\rrbracket^\varphi\\
 \end{array}  
-\)
+\]
 
@@
 primitive propositions (primitive tests) alternating with the symbols
-from $E$. Following \cite{kozen2001automata}, we call these truth
-assignments \emph{atoms}; the set of atoms is notated $\mathcal A$.
-In the algebra described by Figure \ref{epikfigure}, $E$ is the set of
-events $\{a_1,a_0,b_1,b_0\}$, and in the elevator example, $\{u,d\}$.
-In the elevator example, $T$ is $\{p,q\}$, and in the coin example it
-is $\{h,t\}$.  We write atoms as the boolean formulae that uniquely
-specify the truth assignment. In the elevator example, we have atoms
-$\bar q \bar p$, $\bar q p$, $q \bar p$, $q p$, for notational
-simplicity we will write these as $\hat0$, $\hat1$, $\hat2$, $\hat3$
-respectivelyto indicate the floor each atom represents. In the coin
-example, we get $h\bar t$ and $\bar h t$ (for which we use the
-shorthand $H$ and $T$), as well as the bogus atoms $\bar h \bar t$ and
-$ht$!  To preclude bogosity, the \texttt{constraint} declaration in
-Figure~\ref{epikfigure} describes the set of \emph{feasible atoms},
-which we parse into a formula $\varphi$ (if there are multiple
-\texttt{constraint} declarations, $\varphi$ is the conjunction of the
-parsed formulae in each). For a formula $\varphi$, the set of atoms
-that satisfy $\varphi$ is $\mathcal A_\varphi$ -- these are the
-feasible atoms. In the coin example, the constraint $\varphi$ is
-$h\bar t + \bar h t$. Since $\bar h t$ satisfies $\varphi$ and $h \bar
-t$ satisfies $\varphi$, they are permitted feasible atoms, and since
-$\bar h \bar t$ and $h t$ do not satisfy $\varphi$ they are not
-permitted. We omit boolean satisfiability since its definition is
-standard.
+from $E$.  In the algebra described by Figure \ref{epikfigure}, $E$ is
+the set of events $\{a_1,a_0,b_1,b_0\}$, and in the elevator example,
+$\{u,d\}$.  In the elevator example, $T$ is $\{p,q\}$, and in the coin
+example it is $\{h,t\}$.  We write atoms as the boolean formulae that
+uniquely specify the truth assignment. In the elevator example, we
+have atoms $\bar q \bar p$, $\bar q p$, $q \bar p$, $q p$, for
+notational simplicity we will write these as $\hat0$, $\hat1$,
+$\hat2$, $\hat3$ respectivelyto indicate the floor each atom
+represents. In the coin example, we get $h\bar t$ and $\bar h t$ (for
+which we use the shorthand $H$ and $T$), as well as the bogus atoms
+$\bar h \bar t$ and $ht$!  To preclude bogosity, the
+\texttt{constraint} declaration in Figure~\ref{epikfigure} describes a
+state formula $\varphi$ which denotes the set of \emph{feasible atoms}
+$\mathcal A_{\mathsf B}^\varphi$.
 
@@
 behavior via the \texttt{event} declarations. From these we construct
-a function $P$ from agents to relations on atoms that specify the
-effects each action can have. If there are multiple event declarations
-for the same event, only the latter one is incorporated. To construct
-the effect relation $P(e)$ for event $e$, we add a pair $\alpha,
-\beta$ to the relation $P(e)$ for every disjunct $\varphi:\varphi'$ such
-that $\alpha$ satisfies $\phi$ and $\beta$ satisfies $\varphi'$.  These have the
-status of event types, in that they can ``happen'' in different
-possible worlds. Here happening corresponds to incrementing the world
-with some compatible element of the event type.  Where $e$ is a bare
-event symbol, $\llbracket e \rrbracket^{\varphi,P}$ is the
-corresponding set of guarded strings, consisting of the bare event
-decorated with compatible truth assignments denoted by $P$ that are
-admitted by the world-constraints $\varphi$.  See this formally in
-Figure~\ref{fig:interp}, and the examples in \ex{events1}.
+a function $\mathcal E$ from agents to relations on atoms that specify
+the effects each action can have. Notably for every declaration
+$\texttt{event}~e~\zeta$, we have $\mathcal E^\varphi(e) =
+\llbracket\zeta\rrbracket^\varphi$. We can lift $\mathcal E^\varphi$
+to $\hat{\mathcal E}$ which decorates the events with their permitted
+atoms, that is $\hat{\mathcal E}^\varphi(e) = \{\alpha~e~\beta \mid
+\alpha, \beta \in \mathcal E^\varphi (e) \}$. See the examples in
+\ex{events1}.
 % u (\bar{q}\bar{p}):(\bar{q}p) + (\bar{q}p):(q\bar{p}) + (q\bar{p}):(qp) 173
@@
 \begin{array}[t]{l|c}
-  e & \llbracket e \rrbracket^{\varphi, P} \\ \hline
+  e &  \hat{\mathcal E}^\varphi(e) \\ \hline
   u & \{ \hat 0 ~u~ \hat 1, \hat 1 ~u~ \hat 2,  \hat 2 ~u~ \hat 3\}\\
@@
 
-\section{A Language for Epistemic Sets of Guarded Strings}
-
-To reason about epistemic alternatives, guarded strings are used to
-represent different possible worlds, so agents' knowledge/beliefs are
-represented by sets of guarded strings. The standard algebra for
-manipulating guarded strings is Kleene Algebra with Tests
-(\textsc{KAT}), which has the algebraic signature
-$\tuple{K,+,;,*,\bar{},0,1}$ \cite{kozen2001automata}. However, the
-KATs don't permit native reasoning about modal operators, so Epik's
-signature needs to subsume the KAT operators.
-
-Figure~\ref{fig:epik-grammar} depicts the syntax of Epik's formal
-term language. As in KAT, $0$ is the null element, $1$ is the unit
-element, $+$ indicates disjuction, $;$ indicates event sequencing or
-boolean conjunction, $!$ indicates boolean negation, and $*$ indicates
-iteration. We add some additional operations to reason about epistemic
-alternatives: the term $\Diamond_a~p$ represents epistemic possibility
-for agent $a$ in worlds $p$, and complement $\neg p$ represents the
-worlds not described by $p$. We can encode $\Box_a~p$, i.e. epistemic
-certainty for agent $a$ in worlds $p$, by $\Box_a~p =
-(\Diamond_a~p^c)^c$.
-
-\begin{figure}
-
-  \[\begin{array}{l>{\hspace{-.5em}}c>{\hspace{-0.75em}}l}
-  \multicolumn 2 l {\textrm{tests}}& a~\textrm{is an atomic test} \\
-  t,s & ::=& a~|~0~|~1~|~t+t~|~t;t~|~!t\\
-  \multicolumn 2 l {\textrm{terms}} & e~\textrm{is an atomic event} \\
-  p,q & ::=& e~|~0~|~1~|~t~|~p+p~|~p;p~|~p^*~|\neg p~|\Diamond_a~p
-  \end{array}\]
-
-  Derived Operators\\
-  \mbox{}\qquad \(\Box_a p \triangleq \neg\Diamond_a\neg p \qquad p\&q \triangleq \neg(\neg p + \neg q) \)
-
-\caption{The language of Epik terms and key derived
-  operators}
-\label{fig:epik-grammar}
-\label{termfigure}
-\end{figure}
-
-We interpret each term in as a set of guarded strings as indicated in
-Figure~\ref{fig:interp}. The function $\llbracket p
-\rrbracket^{\varphi,P}$ interprets a term $p$ as a set of guarded
-strings consistent with state constraints $\varphi$ and effect
-relation $P$. A bare event $e$ is interpreted as the set of guarded
-strings allowed by $\varphi$ and $P$. The constant $0$ is interpreted
-as the empty set.  The constant $1$ is interpreted as the set of all
-atoms obeying $\varphi$, i.e. $\{h\bar t, \bar h t\}$ in the coin
-example. The operation $p + q$ is interpreted as the union of the
-interpretations of $p$ and $q$. The operation $p;q$ interpreted as the
-fusion product of the interpretations of $p$ and $q$ raised to sets:
-$X \diamond Y = \{ x \diamond y \mid x \in X, y \in Y, x\diamond
-y~\textit{defined}\}$.  The operation $p*$ is Kleene star, which is
-interpreted to be least upper bound of the iterated fusion product of
-the denotation of $p$ with itself. Subsets of $1$ indicated by $t$ are
-also elements of $K$, and these form the Boolean algebra of tests,
-denoting all atoms admitted by the constraint $\varphi$ and the
-boolean test $t$.  We ``overload'' $;$ to further indicate boolean
-conjunction, $+$ to further indicate boolean disjunction, and $0,1$ to
-indicate the zero and one elements of the boolean algebra. The overbar
-operation denotes complement in this Boolean algebra: $\overline t$
-denotes all atoms $\alpha$ such that $\alpha \Rightarrow \neg t$.
-Other boolean operators (such as implication $\Rightarrow$) are
-encoded as syntactic sugar in the standard way. A boolean test $t$
-denotes the set of atoms that satisfy $t;\varphi$, that is, they
-satisfy both $t$ and $\varphi$.
-
-The unary epistemic alternative operation $\Diamond_a$ for each agent
-$a$, is interpreted using using Kripke semantics, as pre-image
-relative to a fixed relation $\hat{R_a}$ between guarded strings,
-$\Diamond_a~x = \{u \mid \exists v. v \in x \wedge u \mathrel{R_a}
-v\}$.  Here $u$ and $v$ are guarded strings, while $x$ is an element
-of $K$.
-
-\newlength{\dhatheight}
-\newcommand{\doublehat}[1]{%
-    \settoheight{\dhatheight}{\ensuremath{\hat{#1}}}%
-    \addtolength{\dhatheight}{-0.35ex}%
-    \hat{\vphantom{\rule{1pt}{\dhatheight}}%
-      \smash{\hat{#1}}}}
-
 It remains to define the Kripke relation on guarded strings from an
@@
 \hspace{6mm} u_1\mathrel{R}v_1 \wedge ... \wedge u_n\mathrel{R}v_n \wedge\\
-\hspace{6mm} x \in \llbracket u_1 ; ... ; u_n \rrbracket^{\varphi,P}\hspace{1mm} \wedge \\
-\hspace{6mm} y \in \llbracket v_1 ; ... ; v_n \rrbracket^{\varphi,P} \hspace{2mm} \}
+\hspace{6mm} x \in \llbracket u_1 ; ... ; u_n \rrbracket^{\varphi,{\mathcal{E}}}\hspace{1mm} \wedge \\
+\hspace{6mm} y \in \llbracket v_1 ; ... ; v_n \rrbracket^{\varphi,{\mathcal{E}}} \hspace{2mm} \}
 \end{array}
@@
 
+\section{A Language for Epistemic Sets of Guarded Strings}
+
+To reason about epistemic alternatives, guarded strings are used to
+represent different possible worlds, so agents' knowledge/beliefs are
+represented by sets of guarded strings. The standard algebra for
+manipulating guarded strings is Kleene Algebra with Tests
+(\textsc{KAT}), which has the algebraic signature
+$\tuple{K,+,;,*,\bar{},0,1}$ \cite{kozen2001automata}. However, the
+KATs don't permit native reasoning about modal operators, so Epik's
+signature needs to subsume the KAT operators.
+
+Figure~\ref{fig:epik-grammar} depicts the syntax of Epik's formal term
+language. As in KAT, $0$ is the null element, $1$ is the unit element,
+$p + q$ indicates disjuction, $p;q$ (or simply $pq$) indicates
+sequence, the boolean operators for tests $\varphi$ are the standard
+ones (as in Figure~\ref{fig:effect-syntax}), and ${p}^*$ indicates
+iteration of $p$. We add some additional operations to reason about epistemic
+alternatives: the term $\Diamond_a~p$ represents epistemic possibility
+for agent $a$ in worlds $p$, and complement $\neg p$ represents the
+worlds not described by $p$. We can encode $\Box_a~p$, i.e. epistemic
+certainty for agent $a$ in worlds $p$, by $\Box_a~p =
+(\Diamond_a~p^c)^c$.
+
+\begin{figure}
+
+  \[\begin{array}{l>{\hspace{-.5em}}c>{\hspace{-0.75em}}l}
+  \multicolumn 2 l {\textrm{terms}} & e \in \mathsf{P} \\
+  p,q & ::=& e~|~0~|~1~|~\phi~|~p+p~|~p;p~|~p^*~|\neg p~|\Diamond_a~p
+  \end{array}\]
+
+  Derived Operators\\[.1em]
+  \(\Box_a p \triangleq \neg\Diamond_a\neg p\) \hfill \(p\&q \triangleq \neg(\neg p + \neg q)\) \hfill \(\_ \triangleq \sum\mathsf P\)
+
+\caption{The language of Epik terms and key derived
+  operators}
+\label{fig:epik-grammar}
+\label{termfigure}
+\end{figure}
+
+We interpret each term in as a set of guarded strings as indicated in
+Figure~\ref{fig:interp}. The function $\llbracket p
+\rrbracket^{\varphi,{\mathcal{E}}}$ interprets a term $p$ as a set of
+guarded strings consistent with state constraints $\varphi$ and effect
+relations ${\mathcal{E}^\varphi}$. A bare event $e$ is interpreted as
+the set $\hat{\mathcal E}^\varphi(e)$. The constant $0$ is interpreted
+as the empty set.  The constant $1$ is interpreted as the set
+$\mathcal A^\varphi_{\mathsf B}$, i.e. $\{h\bar t, \bar h t\}$ in the
+coin example. The operation $p + q$ is interpreted as the union of the
+interpretations of $p$ and $q$. The operation $p;q$ interpreted as the
+fusion product of the interpretations of $p$ and $q$ raised to sets:
+$X \diamond Y = \{ x \diamond y \mid x \in X, y \in Y, x\diamond
+y~\textit{defined}\}$.  The operation $p*$ is Kleene star, which is
+interpreted to be least upper bound of the iterated fusion product of
+the denotation of $p$ with itself. Subsets of $1$ indicated by
+propositions $\psi$ are also elements of $K$, and these form the
+Boolean algebra of tests, denoting all atoms admitted by the
+constraint $\varphi$ and the boolean test $\psi$.
+
+The unary epistemic alternative operation $\Diamond_a$ for each agent
+$a$, is interpreted using using Kripke semantics, as pre-image
+relative to a fixed relation $\hat{R_a}$ between guarded strings,
+$\Diamond_a~p = \{u \mid \exists v. v \in \llbracket
+p\rrbracket^{\phi, \mathcal E} \wedge u \mathrel{\hat{R_a}} v\}$.  Here $u$
+and $v$ are guarded strings, while $x$ is an element of $K$.
+
+\newlength{\dhatheight}
+\newcommand{\doublehat}[1]{%
+    \settoheight{\dhatheight}{\ensuremath{\hat{#1}}}%
+    \addtolength{\dhatheight}{-0.35ex}%
+    \hat{\vphantom{\rule{1pt}{\dhatheight}}%
+      \smash{\hat{#1}}}}
+
 \begin{figure}
   \[\begin{array}{l >{\triangleq}c l}
-  \llbracket 0 \rrbracket^{\varphi,P} && \{\} \\  
-  \llbracket 1 \rrbracket^{\varphi,P} && \{\alpha \mid \alpha \in \mathcal A_\varphi\}\\
-  \llbracket e \rrbracket^{\varphi,P}&& \{\alpha~e~\beta \mid
-  (\alpha,\beta) \in P(e), \alpha, \beta \Rightarrow \varphi \}
-  \\ \llbracket \_ \rrbracket^{\varphi,P}&& \bigcup_{e \in
-    \dom(P)}\llbracket e \rrbracket^{\varphi,P}\\
-  \llbracket t \rrbracket^{\varphi,P} &&  \{\alpha \mid \alpha \in \mathcal A_{\varphi;t}\}\\
-  \llbracket{p+q}\rrbracket^{\varphi,P} && \llbracket{p}\rrbracket^{\varphi,P} \cup \llbracket q \rrbracket^{\varphi,P}\\
-  \llbracket{p;q}\rrbracket^{\varphi,P} && \llbracket p \rrbracket^{\varphi,P} \diamond \llbracket{q}\rrbracket^{\varphi,P}\\
-  \llbracket p^*\rrbracket^{\varphi,P} && \displaystyle\bigcup_n {\underbrace{\llbracket p \rrbracket^{\varphi,P}\diamond \ldots \diamond \llbracket p \rrbracket^{\varphi,P}}_{n}} \\
-  \llbracket \neg p \rrbracket^{\varphi,P} && \llbracket \_^*\rrbracket^{\varphi,P} \setminus \llbracket p \rrbracket^{\varphi,P} \\
-  \llbracket \langle a \rangle~p \rrbracket && \{x \mid x~\hat{R_a}~y, y \in \llbracket p \rrbracket\}
+  \llbracket 0 \rrbracket^{\varphi,{\mathcal{E}}} && \{\} \\  
+  \llbracket 1 \rrbracket^{\varphi,{\mathcal{E}}} && \mathcal A^\varphi_{\mathsf B}\\
+  \llbracket e \rrbracket^{\varphi,{\mathcal{E}}}&& \hat{\mathcal E}^\varphi(e) \\
+  \llbracket \psi \rrbracket^{\varphi,{\mathcal{E}}} &&  \mathcal A^{\varphi~\psi}\\
+  \llbracket{p+q}\rrbracket^{\varphi,{\mathcal{E}}} && \llbracket{p}\rrbracket^{\varphi,{\mathcal{E}}} \cup \llbracket q \rrbracket^{\varphi,{\mathcal{E}}}\\
+  \llbracket{p;q}\rrbracket^{\varphi,{\mathcal{E}}} && \llbracket p \rrbracket^{\varphi,{\mathcal{E}}} \diamond \llbracket{q}\rrbracket^{\varphi,{\mathcal{E}}}\\
+  \llbracket p^*\rrbracket^{\varphi,{\mathcal{E}}} && \displaystyle\bigcup_n {\underbrace{\llbracket p \rrbracket^{\varphi,{\mathcal{E}}}\diamond \ldots \diamond \llbracket p \rrbracket^{\varphi,{\mathcal{E}}}}_{n}} \\
+  \llbracket \neg p \rrbracket^{\varphi,{\mathcal{E}}} && \llbracket \_^*\rrbracket^{\varphi,{\mathcal{E}}} \setminus \llbracket p \rrbracket^{\varphi,{\mathcal{E}}} \\
+  \llbracket \Diamond_a~p \rrbracket && \{x \mid x \mathbin{\hat{R_a}} y, y \in \llbracket p \rrbracket^{\varphi,\mathcal E}\}
 
@@
 guarded strings, defined to every set the universe $\_^*$ except those
-denoted by $p$. Note that $\llbracket 0^c \rrbracket^{\varphi,P} =
-\llbracket\_^*\rrbracket^{\varphi,P}$. We use the complement to define
-universal box modalites as the dual of the diamond modality, $\Box_a x
-= (\Diamond_a x^c)^c$.  For instance, Aly is certain that the coin is
-heads if and only if she does not consider it possible that it is not
-heads.
+denoted by $p$. Note that $\llbracket 0^c
+\rrbracket^{\varphi,{\mathcal{E}}} =
+\llbracket\_^*\rrbracket^{\varphi,{\mathcal{E}}}$.
 
 Summing up, given an Epik program with $n$ agents, we construct state
-constraints $\varphi$, effect relations $P$, and a concrete guarded
+constraints $\varphi$, effect relations ${\mathcal{E}}$, and a concrete guarded
 string model for each term using $\llbracket -
-\rrbracket^{\varphi,P}$.
+\rrbracket^{\varphi,{\mathcal{E}}}$.
 %% $\tuple{K,+,;,*,\bar{},0,1,{}^c,\Diamond_{a_1},...,\Diamond_{a_n}}$.
@@
   tails, and does not know that it's heads} as elements of
-$K$. Concretely the proposositions are sets of guarded strings
-(usually countable ones), construed as sets of worlds as they figure
-in possible worlds semantics for natural language.
+$K$. Concretely the propositions are sets of guarded strings (usually
+countable ones), construed as sets of worlds as they figure in
+possible worlds semantics for natural language.
 
@@
 
-The relation on decorated events needs to be generalized to a relation of worlds.
-The principle for this is that an epistemic alternative to a world of the form $we$
-is a world of the form $vd$, where $v$ is a world-alternative to $w$, $d$ is an
-event-alternative to $e$, and $vd$ is defined (i.e.  the world alternative $v$
-satisfies the pre-conditions of the event alternative $d$).  This principle is found
-in earlier literature (Moore 198x, Baltag, Moss and Solecki 20xx).  In the
-construction in Fst, the definition of world alternatives takes a simple form. Where
-$R_a$ is the relation on decorated events for agent $a$, the the corresponding
-relation on worlds in is the Kleene closure of $R_a$.  Where $R$ and $S$ are
-relations, the concatenation product of $R$ and $S$ is the set of pairs of the form
-$\pair{x_1x_2}{y_1y2}$, where $\pair{x_1}{y_1}$ is in relation $R$, and
-$\pair{x_2}{y_2}$ is in relation $S$.  The Kleene closure of relation $R$ is
-\(\cup_{n \geq 0}R^n\), where $R^n$ is the $n$-times concatenation product of $R$
-with itself. This is an operation in the finite state calculus.  Figure
-\ref{relclosure} defines the corresponding operation in the guarded string algebra.
-The epistemic alternative relation on worlds for an agent is then defined
-as the concatenation closure of the event alternative relation for the agent.
+The relation on decorated events needs to be generalized to a relation
+of worlds.  The principle for this is that an epistemic alternative to
+a world of the form $we$ is a world of the form $vd$, where $v$ is a
+world-alternative to $w$, $d$ is an event-alternative to $e$, and $vd$
+is defined (i.e.  the world alternative $v$ satisfies the
+pre-conditions of the event alternative $d$).  This principle is found
+in earlier literature (Moore 198x, Baltag, Moss and Solecki 20xx).  In
+the construction in Fst, the definition of world alternatives takes a
+simple form. Where $R_a$ is the relation on decorated events for agent
+$a$, the the corresponding relation on worlds in is the Kleene closure
+of $R_a$.  Where $R$ and $S$ are relations, the concatenation product
+of $R$ and $S$ is the set of pairs of the form $\pair{x_1x_2}{y_1y2}$,
+where $\pair{x_1}{y_1}$ is in relation $R$, and $\pair{x_2}{y_2}$ is
+in relation $S$.  The Kleene closure of relation $R$ is \(\cup_{n \geq
+  0}R^n\), where $R^n$ is the $n$-times concatenation product of $R$
+with itself (the $0$-times concatenation product is $\llbracket 1
+\rrbracket^{\varphi,\mathcal E}$). This is an operation in the finite
+state calculus.  Figure \ref{relclosure} defines the corresponding
+operation in the guarded string algebra.  The epistemic alternative
+relation on worlds for an agent is then defined as the concatenation
+closure of the event alternative relation for the agent.
 
@@
 terms with the same signature as described in Section 2. We then paramaterize
-the interpretation function, $\llparenthesis p \rrparenthesis^{\varphi,P}_n$ on an
+the interpretation function, $\llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n$ on an
 integer $n$ that describes the maximum length of a string we will produce. The
@@
   \[\begin{array}{r >{\triangleq}c l}
-  \llparenthesis p \rrparenthesis^{\varphi,P}_0 && [] \\
-  \llparenthesis 0 \rrparenthesis^{\varphi,P}_n && [] \\
-  \llparenthesis 1 \rrparenthesis^{\varphi,P}_n && [\alpha \mid \alpha \leq \varphi] \\
-  \llparenthesis e \rrparenthesis^{\varphi,P}_n && [\alpha \cdot e \cdot \beta \mid (\alpha, \beta) \in P(e), \alpha\vee \beta \Rightarrow \varphi] \\
-  \llparenthesis \_ \rrparenthesis^{\varphi,P}_n && \sum [\llparenthesis e \rrparenthesis^{\varphi,P} \mid e \in \dom(P) ] \\
-  \llparenthesis t \rrparenthesis^{\varphi,P}_n && [\alpha\mid \alpha \Rightarrow \varphi \wedge t] \\
-  \llparenthesis p + q \rrparenthesis^{\varphi,P}_n && \llparenthesis p \rrparenthesis^{\varphi,P}_n + \llparenthesis q \rrparenthesis^{\varphi,P}_n \\
-  \llparenthesis p \& q \rrparenthesis^{\varphi,P}_n &&
-  [g~|~g \in \llparenthesis p \rrparenthesis_n^{\varphi,P}, g \in \llparenthesis q \rrparenthesis_n^{\varphi,P}] \\
-  \llparenthesis p;q \rrparenthesis^{\varphi,P}_n &&
-  \left(\llparenthesis p \rrparenthesis^{\varphi,P}_n \cdot \llparenthesis q\rrparenthesis^{\varphi,P}_n\right)|_n\\
-  \llparenthesis p^* \rrparenthesis^{\varphi,P}_n && [] + \left(\llparenthesis p\rrparenthesis^{\varphi,P}_n \cdot  \llparenthesis p^* \rrparenthesis^{\varphi,P}_{n-i}\right)|_n \\
-  \multicolumn 2 r {} & \text{where}~i = \max\{1,\min\{|g|\mid g \in \llbracket p \rrbracket^{\varphi,P}\}\}\\
-  \llparenthesis \sim p \rrparenthesis^{\varphi,P}_n && [g \mid g \in \llbracket \_^* \rrbracket^{\varphi,P}_n, g \not\in \llbracket p \rrbracket^{\varphi,P}_n] \\
-  \llparenthesis \langle a \rangle p \rrparenthesis^{\varphi,P}_n &&
-  [g' \mid g \in g'~\doublehat{R_a}~g, g \in \llparenthesis p \rrparenthesis^{\varphi,P}]
+  \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_0 && [] \\
+  \llparenthesis 0 \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [] \\
+  \llparenthesis 1 \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [\alpha \mid \alpha \leq \varphi] \\
+  \llparenthesis e \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [\alpha \cdot e \cdot \beta \mid (\alpha, \beta) \in {\mathcal{E}}(e), \alpha\vee \beta \Rightarrow \varphi] \\
+  \llparenthesis \_ \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \sum [\llparenthesis e \rrparenthesis^{\varphi,{\mathcal{E}}} \mid e \in \dom({\mathcal{E}}) ] \\
+  \llparenthesis t \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [\alpha\mid \alpha \Rightarrow \varphi \wedge t] \\
+  \llparenthesis p + q \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n + \llparenthesis q \rrparenthesis^{\varphi,{\mathcal{E}}}_n \\
+  \llparenthesis p \& q \rrparenthesis^{\varphi,{\mathcal{E}}}_n &&
+  [g~|~g \in \llparenthesis p \rrparenthesis_n^{\varphi,{\mathcal{E}}}, g \in \llparenthesis q \rrparenthesis_n^{\varphi,{\mathcal{E}}}] \\
+  \llparenthesis p;q \rrparenthesis^{\varphi,{\mathcal{E}}}_n &&
+  \left(\llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n \cdot \llparenthesis q\rrparenthesis^{\varphi,{\mathcal{E}}}_n\right)|_n\\
+  \llparenthesis p^* \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [] + \left(\llparenthesis p\rrparenthesis^{\varphi,{\mathcal{E}}}_n \cdot  \llparenthesis p^* \rrparenthesis^{\varphi,{\mathcal{E}}}_{n-i}\right)|_n \\
+  \multicolumn 2 r {} & \text{where}~i = \max\{1,\min\{|g|\mid g \in \llbracket p \rrbracket^{\varphi,{\mathcal{E}}}\}\}\\
+  \llparenthesis \sim p \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [g \mid g \in \llbracket \_^* \rrbracket^{\varphi,{\mathcal{E}}}_n, g \not\in \llbracket p \rrbracket^{\varphi,{\mathcal{E}}}_n] \\
+  \llparenthesis \langle a \rangle p \rrparenthesis^{\varphi,{\mathcal{E}}}_n &&
+  [g' \mid g \in g'~\doublehat{R_a}~g, g \in \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}]
   \end{array}

### 2020-06-18 tex/draft.tex
update guarded string length

@@
 Guarded strings are strings of events, alternating with such atoms,
-and starting and ending with atoms.  \ex{gs1} gives the encoding as
-guarded strings of the worlds in \ex{elevator1} and \ex{coin1}.
+and starting and ending with atoms. The length of a guarded string
+$g$, written $|g|$ is the number of events in $g$, ignoring the atoms.
+\ex{gs1} gives the encoding as guarded strings of the worlds in
+\ex{elevator1} and \ex{coin1}.
 
@@
 \begin{example}   \label{gs1}
-\begin{tabular}[t]{ll}
-$v_1$ & $\hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 ~u~ \hat 2$ \\
-$v_2$ & $\hat 0 ~u~ \hat 1 ~u~ \hat 2 ~u~ \hat 3$ \\   
-$w_1$ & $H$ \\
-$w_2$ & $H ~a_0~ H$ \\  
-$w_3$ & $T~a_0~ T ~b_0~ T$ \\
-\end{tabular}
+  \renewcommand{\arraystretch}{1.2}
+  \begin{tabular}[t]{llc}
+     \multicolumn 2 c {\textit{Guarded String}} & \textit{Length}\\ \hline
+$v_1$ & $\hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 ~u~ \hat 2$ & $4$ \\
+$v_2$ & $\hat 0 ~u~ \hat 1 ~u~ \hat 2 ~u~ \hat 3$ & $3$ \\   
+$w_1$ & $H$ & $0$\\
+$w_2$ & $H ~a_0~ H$ & $1$ \\  
+$w_3$ & $T~a_0~ T ~b_0~ T$ & $3$\\
+  \end{tabular}
+    \renewcommand{\arraystretch}{1}
 \end{example}
@@
 
-We also implement the semantics of Epik terms using lazy lists~\cite{LazyLists}
-in Haskell, rather than the direct interpretation as sets. Unfortunately,
-regular expressions, and hence also Epik, can denote infinite sets of strings,
-for example the term $\_^*$. Normally regular languages are represented using a
-finite coalgebra~\cite{}. However the non-distributivity of $\diamond$ accross
-$;$ complicates this construction~\footnote{Axiomatic and Coalgebraic models for
-  Epik are open questions}. To sidestep this, we parameterize the interpretation
-function on a positive integer $n$ and only produce guarded strings of length
-$n$ or less.  The length of a guarded string $g$, written $|g|$ is the number of
-events in the string, so $\alpha$ has length $0$, $\alpha e\beta$ has length
-$1$, and $\alpha e \beta e' \gamma$ has length $2$.
-
-We translate the Epik terms into a Haskell algebraic datatype that represents
-terms with the same signature as described in Section 2. We then paramaterize
-the interpretation function, $\llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n$ on an
-integer $n$ that describes the maximum length of a string we will produce. The
-bounded interpretation into lists of strings is very similar to the unbounded
-interpretation into sets of strings, except for the bounds checking. The full
-details are shown in Figure~\ref{fig:list-interp}: we use list-comprehension
-notation, which is analogous to set-builder notation, except that it produces a
-list (potentially with duplicated elements) instead of a set, and so is written
-with square brackets $[]$. We use $+$ to indicate list concatenation, and $\sum
-l$ to indicate the concatenation of a list of lists $l$. We overload the set
-membership symbol $\in$ to also indicate membership in a list. We also lift the
-fusion operator $\cdot$ to lists. For a list of guarded strings $l$, we write
-$l|_n$ as shorthand for the set $[ g~|~g \in l , |g| \leq n]$.
+We also implement the semantics of Epik terms using lazy
+lists~\cite{LazyLists} in Haskell, rather than the direct
+interpretation as sets. Unfortunately, regular expressions, and hence
+also Epik, can denote infinite sets of strings, for example the term
+$\_^*$. Normally regular languages are represented using a finite
+coalgebra~\cite{}. However the non-distributivity of $\diamond$
+accross $;$ complicates this construction\footnote{Axiomatic and
+  Coalgebraic models for Epik are open questions}. To sidestep this,
+we parameterize the interpretation function on a positive integer $n$
+and only produce guarded strings of length $n$ or less.
+
+We translate the Epik terms into a Haskell algebraic datatype that
+represents terms with the same signature as described in Section 2. We
+then paramaterize the interpretation function, $\llparenthesis p
+\rrparenthesis^{\varphi,{\mathcal{E}}}_n$ on an integer $n$ that
+describes the maximum length of a string we will produce. The bounded
+interpretation into lists of strings is very similar to the unbounded
+interpretation into sets of strings, except for the bounds
+checking. The full details are shown in Figure~\ref{fig:list-interp}:
+we use list-comprehension notation, which is analogous to set-builder
+notation, except that it produces a list (potentially with duplicated
+elements) instead of a set, and so is written with square brackets
+$[]$. We use $+$ to indicate list concatenation, and $\sum l$ to
+indicate the concatenation of a list of lists $l$. We overload the set
+membership symbol $\in$ to also indicate membership in a list. We also
+lift the fusion operator $\diamond$ to lists. For a list of guarded
+strings $l$, we write $l|_n$ as shorthand for the set $[ g~|~g \in l ,
+  |g| \leq n]$. Finally, we write $\lfloor X \rfloor$ to convert a set
+$X$ into a list containing same elements with an arbitrary order.
 
@@
 
-The final step is to convert the list of guarded strings into a set of guarded
-strings. Since a list is just an ordered multiset, we forget the order, and
-remove all duplicates. Haskell's libraries have a built-in function for this,
-\texttt{Data.Set.fromList}, which converts a list into a set.
+To convert from a list of guarded strings $l$ to a set of guarded
+strings, we simply write $\lceil l\rceil$. Note that for any $p$,
+$\varphi$, $\mathcal E$, and $n$, $\left\lceil\llparenthesis p
+\rrparenthesis^{\varphi,\mathcal E}_n\right\rceil = \{g \mid g \in
+\llbracket p \rrbracket^{\varphi,\mathcal E}, |g| \leq n\}$.
 
@@
   \llparenthesis 0 \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [] \\
-  \llparenthesis 1 \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [\alpha \mid \alpha \leq \varphi] \\
-  \llparenthesis e \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [\alpha \cdot e \cdot \beta \mid (\alpha, \beta) \in {\mathcal{E}}(e), \alpha\vee \beta \Rightarrow \varphi] \\
-  \llparenthesis \_ \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \sum [\llparenthesis e \rrparenthesis^{\varphi,{\mathcal{E}}} \mid e \in \dom({\mathcal{E}}) ] \\
-  \llparenthesis t \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [\alpha\mid \alpha \Rightarrow \varphi \wedge t] \\
+  \llparenthesis 1 \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \lfloor \mathcal A_{\mathsf B}^\varphi \rfloor \\
+  \llparenthesis e \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \lfloor \hat{\mathcal E}^\varphi(e)\rfloor \\
+  \llparenthesis \psi \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \lfloor \mathcal A_{\mathsf B}^{\varphi~\psi} \rfloor \\
   \llparenthesis p + q \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n + \llparenthesis q \rrparenthesis^{\varphi,{\mathcal{E}}}_n \\
-  \llparenthesis p \& q \rrparenthesis^{\varphi,{\mathcal{E}}}_n &&
-  [g~|~g \in \llparenthesis p \rrparenthesis_n^{\varphi,{\mathcal{E}}}, g \in \llparenthesis q \rrparenthesis_n^{\varphi,{\mathcal{E}}}] \\
   \llparenthesis p;q \rrparenthesis^{\varphi,{\mathcal{E}}}_n &&
   \left(\llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n \cdot \llparenthesis q\rrparenthesis^{\varphi,{\mathcal{E}}}_n\right)|_n\\
-  \llparenthesis p^* \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [] + \left(\llparenthesis p\rrparenthesis^{\varphi,{\mathcal{E}}}_n \cdot  \llparenthesis p^* \rrparenthesis^{\varphi,{\mathcal{E}}}_{n-i}\right)|_n \\
+  \llparenthesis p^* \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [] + \left(\llparenthesis p\rrparenthesis^{\varphi,{\mathcal{E}}}_n \diamond  \llparenthesis p^* \rrparenthesis^{\varphi,{\mathcal{E}}}_{n-i}\right)|_n \\
   \multicolumn 2 r {} & \text{where}~i = \max\{1,\min\{|g|\mid g \in \llbracket p \rrbracket^{\varphi,{\mathcal{E}}}\}\}\\
-  \llparenthesis \sim p \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [g \mid g \in \llbracket \_^* \rrbracket^{\varphi,{\mathcal{E}}}_n, g \not\in \llbracket p \rrbracket^{\varphi,{\mathcal{E}}}_n] \\
-  \llparenthesis \langle a \rangle p \rrparenthesis^{\varphi,{\mathcal{E}}}_n &&
-  [g' \mid g \in g'~\doublehat{R_a}~g, g \in \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}]
+  \llparenthesis \neg p \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [g \mid g \in \llbracket \_^* \rrbracket^{\varphi,{\mathcal{E}}}_n, g \not\in \llbracket p \rrbracket^{\varphi,{\mathcal{E}}}_n] \\
+  \llparenthesis \Diamond_a~p \rrparenthesis^{\varphi,{\mathcal{E}}}_n &&
+  [g' \mid g \in g'~\hat{R_a}~g, g \in \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}]
   \end{array}
   \]
+  \caption{Bounded interpretation using lazy lists}
 \end{figure}

### 2020-06-18 tex/draft.tex
polish 5

@@
 
-\section{Direct Bounded Interpretation}
+\section{Bounded Lazy Interpretation}
 
@@
 \rrparenthesis^{\varphi,{\mathcal{E}}}_n$ on an integer $n$ that
-describes the maximum length of a string we will produce. The bounded
-interpretation into lists of strings is very similar to the unbounded
-interpretation into sets of strings, except for the bounds
-checking. The full details are shown in Figure~\ref{fig:list-interp}:
-we use list-comprehension notation, which is analogous to set-builder
-notation, except that it produces a list (potentially with duplicated
-elements) instead of a set, and so is written with square brackets
-$[]$. We use $+$ to indicate list concatenation, and $\sum l$ to
-indicate the concatenation of a list of lists $l$. We overload the set
-membership symbol $\in$ to also indicate membership in a list. We also
-lift the fusion operator $\diamond$ to lists. For a list of guarded
-strings $l$, we write $l|_n$ as shorthand for the set $[ g~|~g \in l ,
-  |g| \leq n]$. Finally, we write $\lfloor X \rfloor$ to convert a set
-$X$ into a list containing same elements with an arbitrary order.
-
-First of all, whenever $n = 0$, the set of guarded strings is empty. TODO ADD
-MORE TEXT.
+describes the maximum length of a string we will produce.
+
+The bounded interpretation into lists of strings is very similar to
+the unbounded interpretation into sets of strings, except for the
+bounds checking. The full details are shown in
+Figure~\ref{fig:list-interp}. First note that when $n=0$, we simply
+return the empty list, denoted $[]$. Terms of the form $0$, $1$, $e$,
+and $\psi$ have the same denotation as before, translated into a list
+(for a set $X$, $\lfloor X \rfloor$ is a list with the same elements
+as $X$ arbitrarily ordered). The term $p+q$ denotes the list
+concatenation (written $\texttt{+\!+}$) of the denotation of $p$ and of
+$q$. The term $p;q$ denotes the fusion product (lifted to lists) of
+the denotations of $p$ and in $q$ restricted to only those strings
+shorter than $n$ (for a list $l$, $l|_n$ filters out elements longer
+than $n$). The denotation of $p^*$ uses the fact that $p^*$ and $1 +
+p;p^*$ are equivalent, and decrements the size threshold on the
+recursive denotation of $p^*$ by $i$, where $i$ is the length of the
+longest (nonzero) string in the denotation of $p$, making sure to
+filter out guarded strings that are too long. The denotation of $\neg
+p$ is the strings that occur in $\_^*$ and not in $p$ (the $\setminus$
+operator on lists is analogous to the $\setminus$ operator on
+sets). The denotation of $\Diamond_a~p$ is analogous to the set
+denotation, depicted in Figure~\ref{fig:list-interp}-comprehension
+notation\footnote{List comprehension notation is analogous to set
+  builder notation, except that it is written using square
+  brackets. The order preservation is indicated by the keyword
+  \texttt{for}}.
 
@@
   \llparenthesis \psi \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \lfloor \mathcal A_{\mathsf B}^{\varphi~\psi} \rfloor \\
-  \llparenthesis p + q \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n + \llparenthesis q \rrparenthesis^{\varphi,{\mathcal{E}}}_n \\
+  \llparenthesis p + q \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n \mathbin{\texttt{+\!+}} \llparenthesis q \rrparenthesis^{\varphi,{\mathcal{E}}}_n \\
   \llparenthesis p;q \rrparenthesis^{\varphi,{\mathcal{E}}}_n &&
-  \left(\llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n \cdot \llparenthesis q\rrparenthesis^{\varphi,{\mathcal{E}}}_n\right)|_n\\
-  \llparenthesis p^* \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [] + \left(\llparenthesis p\rrparenthesis^{\varphi,{\mathcal{E}}}_n \diamond  \llparenthesis p^* \rrparenthesis^{\varphi,{\mathcal{E}}}_{n-i}\right)|_n \\
-  \multicolumn 2 r {} & \text{where}~i = \max\{1,\min\{|g|\mid g \in \llbracket p \rrbracket^{\varphi,{\mathcal{E}}}\}\}\\
-  \llparenthesis \neg p \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [g \mid g \in \llbracket \_^* \rrbracket^{\varphi,{\mathcal{E}}}_n, g \not\in \llbracket p \rrbracket^{\varphi,{\mathcal{E}}}_n] \\
+  \left(\llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n \diamond \llparenthesis q\rrparenthesis^{\varphi,{\mathcal{E}}}_n\right)|_n\\
+  \llparenthesis p^* \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [] + \left(\llparenthesis p\rrparenthesis^{\varphi,{\mathcal{E}}}_n \diamond  \llparenthesis p^* \rrparenthesis^{\varphi,{\mathcal{E}}}_{m}\right)|_n \\
+  \multicolumn 3 r {\text{where}~
+    \begin{array}[t]{r >{\hspace{-.7em}}l}
+      i &= \max\{1,\min\{|g|\mid g \in \llparenthesis p \rrparenthesis_n^{\varphi,{\mathcal{E}}}\}\}\\
+      m &= \max\{0, n - i\}
+    \end{array}}\\
+  \llparenthesis \neg p \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \llparenthesis \_^* \rrparenthesis^{\varphi,{\mathcal{E}}}_n \setminus \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n \\
   \llparenthesis \Diamond_a~p \rrparenthesis^{\varphi,{\mathcal{E}}}_n &&
-  [g' \mid g \in g'~\hat{R_a}~g, g \in \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}]
+  [g' \mid g'~\hat{R_a}~g,~\texttt{for}~g~\texttt{in}~\llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n]
   \end{array}
@@
   \caption{Bounded interpretation using lazy lists}
+  \label{fig:bound-interp}
 \end{figure}
 
-TODO Describe advantages of Lazy Lists
+The lists we use here are \emph{lazy} (as opposed to \emph{strict}),
+which broadly means that computation is delayed until the value is
+needed. This allows us to avoid computing large, unnecessary
+iterations. In the elevator example, this means that $\hat 0;u;u;u^*$
+will unroll the $u^*$ 0 times, once, and then discover that it is
+impossible to unroll it a second time, because $\llparenthesis\hat
+3;u\rrparenthesis^{\varphi,\mathcal E}_n = []$, for any $n$. So
+$\llparenthesis\hat 3;u\rrparenthesis^{\varphi,\mathcal E}_n \diamond
+\llparenthesis u^*\rrparenthesis^{\varphi,\mathcal E}_{n-1}$ will
+return $[]$ without having to compute the full fixpoint of
+$\llparenthesis u\rrparenthesis^{\varphi,\mathcal E}$. Similar
+behavior occurs once $n = 0$.
+
+Conversely, if we used sets (as in the math) instead of lists we would
+need to constantly verify the set invariant (that every element is
+unique) which means processing every element in the set. Verifying
+that every element in $\llparenthesis
+u^*\rrparenthesis^{\varphi,\mathcal E}$ is unique would require the full
+computation of the fixpoint.
 

### 2020-06-21 tex/draft.tex
Edits in Sections 1 an 2. 2 is not finished.

@@
 natural language.  A proposition is a set or class of {\em possible
-  worlds}.  These worlds are often assumed to be things with the same
+  worlds}.  These are often assumed to be things with the same
 nature and complexity as the world we occupy \citep{lewis1986plurality}.  But
@@
 and illustrate how to apply them in natural language semantics.
-The focus on epistemic semantics and clausal embedding. The fundamental move is to
-identify possible worlds with strings primitive events, so that
+The focus is on epistemic semantics and clausal embedding. The fundamental move is to
+identify possible worlds with strings of primitive events, so that
 propositions are sets of strings. An advantage in this is that it
@@
 strings to theorize about tense and aspect in natural language
-semantics \citep{fernando2004finite,fernando2007observing,carlson2009tense} and to
-express intensionality \cite{fernando2017intensions}.  Literature on
-finite state intensiononal semantics has discussed how to do the
-semantics of intensional complementation including indirect questions
-in a setting where compositional semantics is expressed in a finite
-state calculus \citep{rooth2017finite,collard2018finite}.  We adopt
-this in our syntax-semantics interface for English.
+semantics
+\citep{fernando2004finite,fernando2007observing,carlson2009tense} and
+to express intensionality \cite{fernando2017intensions}.  Literature
+on finite state intensiononal semantics has investigated how to do the
+semantics of intensional complementation, including indirect
+questions, in a setting where compositional semantics is expressed in
+a finite state calculus \citep{rooth2017finite,collard2018finite}.  We
+adopt this in our syntax-semantics interface for English.
 
@@
 {\em The Concealed Coin}.  Amy and Bob are seated at a table.  There is a coin on the
-table under a cup, heads up (H).  The coin could be H (heads) or T (tails), and
+table under a cup, heads up.  The coin could be heads or tails, and
 neither agent knows which it is.  This initial situation is possible world $w_1$.
@@
 \hspace{-4mm} \begin{tabular}[t]{ll}
-  $a_1$  & Amy peeks at H, by tipping the cup.
-  \\ & Bob sees she's peeking, but not what \\
-  & she sees. \\
-  $b_1$  & Bob peeks at H. \\ 
+  $a_1$  & Amy peeks at heads, by tipping the
+  \\ & cup. Bob sees she's peeking, but not\\
+  & what she sees. \\
+  $b_1$  & Bob peeks at heads. \\ 
 \end{tabular}
@@
   $w_1$ & $w_2$  & $w_3$ & Sentence\\
-  false & true & true& Amy knows it's H. \\
-  false & false & true& Bob knows it's H. \\
-  false & false & true& Bob knows Amy \\ &&& knows it's H. \\
-  false & true & true& Bob knows Amy knows \\ &&& whether it's H or T. \\
+  false & true & true& Amy knows that it's \\ &&& heads. \\
+  false & false & true& Bob knows thats it's \\&&&heads. \\
+  false & false & true& Bob knows Amy \\ &&& knows it's heads. \\
+  false & true & true& Bob knows Amy \\ &&& knows  whether it's \\ &&& heads or tails. \\
 \end{tabular}
@@
 
+\renewcommand{\arraystretch}{1.0}
+
 \begin{example}   \label{event1}
@@
 
-In this discussion, a sequence such as $\bar{q}{p}$ can be viewed as a
-valuation of the primitive propositions $q$ and $p$, which is used to
-describe world state. The primitives are listed in fixed order, and
-left unmarked (indicating true) or marked with the overbar (indicating
-false).  Suppose that in the coin example, we had an additional
-primitive stative proposition (or atomic test) $t$, interpreted as
-tails.  Since a coin is heads or tails but not both, we want to allow
-the valuations $h\bar{t}$ and $\bar{h}t$, and disallow $ht$ and
-$\bar{h}\bar{t}$.  This is enforced by a {\em state formula}, which is
-a simply Boolean formula, in this case \exi{effect}{a}.  Where
-{\sf{B}} is a set of atomic tests and $\rho$ is a state constraint
-over {\sf{B}}, ${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\rho}$ is the set of
-valuations of {\sf{B}} that make formula $\rho$ true. Following
-\cite{kozen2001automata}, we call the valuations
-\emph{atoms}.
+\section{Epistemic guarded string models}
+
+In the discussion at the end of the last section, a sequence such as
+$\bar{q}{p}$ can be viewed as a valuation of primitive test
+propositions, which is used to describe world state. The primitives
+are listed in fixed order, and left unmarked (indicating true) or
+marked with the overbar (indicating false).  Suppose that in the coin
+example, we have an additional primitive stative proposition (or
+atomic test) $t$, interpreted as tails.  Since a coin is heads or
+tails but not both, we want to allow the valuations $h\bar{t}$ and
+$\bar{h}t$, and disallow $ht$ and $\bar{h}\bar{t}$.  This is enforced
+by a {\em state formula}, which is a simply Boolean formula, in this
+case \exi{effect}{a}.  Where {\sf{B}} is a set of atomic tests and
+$\phi$ is a state constraint over {\sf{B}},
+${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\phi}$ is the set of valuations of
+{\sf{B}} that make formula $\phi$ true. The valuations are called
+atoms, because they correspond to the atoms of Boolean algebra of
+tests \citep{kozen2001automata}.
 
@@
 post-conditions are {\em effect formulas}. They are interpreted as
-defining relations between valuations, as defined in Figure
-\ref{effectfigure}.  The valuations they relate are constrained by the
+defining relations between atoms, as defined in Figure
+\ref{effectfigure}.  The atoms they relate are constrained by the
 state formula as well.  For the heads-tails example, let the state
 formula and the effect formula for $a_1$ (Aly peeking at heads) be as
-specified in \ex{effect}. Then the set of valuations determined by
-$\varphi$ and the relation on valuations for the event $u$ are
-$\mathcal A_{\mathsf B}^\varphi$ and
-$\llbracket\zeta\rrbracket^\varphi$ as given at the bottom in \ex{6}.
-Given a state formula $\rho$, the the relation characterizing pre- and
-post-conditions of an event with effect formula $\zeta$ is defined to
-be $\sem{\zeta}^{\sem{\rho}}$.
+specified in \ex{effect}. Then ${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\varphi}$
+and the relation on atoms for the event $a_1$ are as given at
+the bottom in \ex{effect}.
 
 \begin{example}   \label{effect}
-\begin{tabular}[t]{l}
-State formula: $\varphi = h\bar{t} + \bar{h}t$ \\
-Effect formula for $a_1$: $\zeta = h \mathbin{:} h$ \\
-\( \mathcal{A}^\varphi_{\mathsf B} = \{h\bar{t},\bar{h}t\} \)  \\
-\( \sem{\zeta}^\varphi = \{\pair{h\bar{t}}{h\bar{t}}\} \)  \\
-\end{tabular}
+\(\begin{array}[t]{ll}
+{\sf{B}} & \{h,t\}\\
+\mbox{state formula $\varphi$} & h\bar{t} + \bar{h}t \\
+\mbox{effect formula $\zeta$ for $a_1$} &  h \mathbin{:} h \\
+ \mathcal{A}^\varphi_{\mathsf B} & \{h\bar{t},\bar{h}t\}   \\
+\sem{\zeta}^\varphi & \{\pair{h\bar{t}}{h\bar{t}}\}   \\
+\end{array}\)
 \end{example}
@@
   \multicolumn 3 l {\textrm{state formulas}\hfill (a \in \mathsf{B}) \hfill \mbox{}}\\
-  \varphi,\psi &::=& a \mid0 \mid 1 \mid\varphi{+}\psi \mid \varphi~\psi \mid \bar{\varphi} \\
+  \rho,\sigma,\varphi &::=& a \mid0 \mid 1 \mid\rho+\sigma \mid \rho\,\sigma \mid \bar{\rho} \\ \\
   \multicolumn {3} l {\textrm{effect formulas}}\\
-  \zeta,\eta & ::= & \varphi \mathbin{:} \sigma \mid \zeta{+}\eta
+  \zeta,\eta & ::= & \rho \mathbin{:} \sigma \mid \zeta+\eta
     \mid \zeta \mathbin{\&} \eta \mid \bar{\zeta}
@@
 \[
-\begin{array}{cll}
-  \llbracket\psi_1\mathbin{:}\psi_2\rrbracket^\varphi & = &
-  \mathcal A^{(\varphi~\psi_1)}_{\mathsf{B}}\times\mathcal A^{(\varphi~\psi_2)}_{\mathsf{B}}\\[.1em]
-\llbracket\zeta{+}\eta\rrbracket^\varphi & = & \llbracket\zeta\rrbracket^\varphi\cup\llbracket\eta\rrbracket^\varphi\\
-\llbracket\zeta{\&}\eta\rrbracket^\varphi & = & \llbracket\zeta\rrbracket^\varphi\cap\llbracket\eta\rrbracket^\varphi\\
-\llbracket\bar{\zeta}\rrbracket^\varphi & = & V \times V \setminus \llbracket\zeta\rrbracket^\varphi\\
+\begin{array}{lll}
+  \sem{\rho\mathbin{:}\sigma}^\varphi & = &
+  {\mathcal A}^{\rho\varphi}_{\mathsf{B}}\times{\mathcal A}^{\sigma\varphi}_{\mathsf{B}}\\[.1em]
+\sem{\zeta +\eta}^\varphi & = & \sem{\zeta}^\varphi\cup\sem{\eta}^\varphi\\
+\sem{\zeta \mathbin{\&}  \eta}^\varphi & = & \sem{\zeta}^\varphi\cap\sem{\eta}^\varphi\\
+\sem{\bar{\zeta}}^\varphi & = & {\mathcal A}^{\varphi}_{\mathsf{B}} \times {\mathcal A}^{\varphi}_{\mathsf{B}} \setminus \sem{\zeta}^\varphi\\
 \end{array}  
@@
 
-\caption{Syntax and semantics of state formulas and effect formulas. Effect formulas
-  are evaluated relative to as set Boolean valuations $V$, and denote relations between
-  Boolean valuations.}
+\vspace{-2mm}
+
+\caption{Syntax of state formulas and syntax and semantics of effect formulas. Effect formulas
+  denote relations between atoms. In a state formula, juxtaposition $\rho\,\sigma$ is conjunction.}
 \label{effectfigure}
@@
 
-\section{Epistemic guarded string models}
+\vspace{1mm}
+
 Figure \ref{epikfigure} shows an Epik program that describes a
@@
 and events of the agents semi-privately looking at the coin. The line
-beginning with {\tt state} lists the basic stative propositions. To
-illustrate syntax, a second proposition $t$ (tails) is included.  The
-line beginning with {\tt constraint} defines compatibilities among the
-propositions: the coin is heads or tails and not both.  The lines
-beginning with {\tt event} declare events, their preconditions, and
-their effect on state, following the format in \ex{event1}.  Finally
+beginning with {\tt state} enumerates ${\sf{B}}$. The
+line beginning with {\tt constraint} gives the state formula.
+The lines
+beginning with {\tt event} declare events and their effect formulas.
+Finally
 the lines beginning with {\tt agent} define {\em event alternative}
@@
 $b_1$ or $b_0$ (Bill peeks at tails) could be happening.
-
-Kleene Algebra with Tests is an algebraic theory that is defined by
-equations and inequalities, which a has model classes including
-guarded string models, relational models, finite models, and matrix
-models.  This paper focuses on defining a family of concrete guarded
-string algebras, the elements of which are sets of guarded strings.
-Definitions and notation mostly follow \cite{kozen2001automata}.
-Additional syntax and semantics is included to model multi-agent
-epistemic semantics.  Guarded strings over a finite alphabet {\sf{P}}
-are like ordinary strings, but with truth assignments to a set $T$ of
-primitive propositions (primitive tests) alternating with the symbols
-from $E$.  In the algebra described by Figure \ref{epikfigure}, $E$ is
-the set of events $\{a_1,a_0,b_1,b_0\}$, and in the elevator example,
-$\{u,d\}$.  In the elevator example, $T$ is $\{p,q\}$, and in the coin
-example it is $\{h,t\}$.  We write atoms as the boolean formulae that
-uniquely specify the truth assignment. In the elevator example, we
-have atoms $\bar q \bar p$, $\bar q p$, $q \bar p$, $q p$, for
-notational simplicity we will write these as $\hat0$, $\hat1$,
-$\hat2$, $\hat3$ respectivelyto indicate the floor each atom
-represents. In the coin example, we get $h\bar t$ and $\bar h t$ (for
-which we use the shorthand $H$ and $T$), as well as the bogus atoms
-$\bar h \bar t$ and $ht$!  To preclude bogosity, the
-\texttt{constraint} declaration in Figure~\ref{epikfigure} describes a
-state formula $\varphi$ which denotes the set of \emph{feasible atoms}
-$\mathcal A_{\mathsf B}^\varphi$.
-
-Guarded strings are strings of events, alternating with such atoms,
-and starting and ending with atoms. The length of a guarded string
-$g$, written $|g|$ is the number of events in $g$, ignoring the atoms.
-\ex{gs1} gives the encoding as guarded strings of the worlds in
-\ex{elevator1} and \ex{coin1}.
-
-%% \begin{example}   \label{gs1}
-%% \begin{tabular}[t]{ll}
-%% $v_1$ & $\bar p \bar q ~u~ \bar p q ~u~ q \bar p ~d~ \bar p q ~u~ q \bar p$ \\
-%% $v_2$ & $\bar p \bar q ~u~ \bar p q ~u~ p\bar q ~u~ pq$ \\   
-%% $w_1$ & $\bar h t$ \\
-%% $w_2$ & $\bar h t ~a_0~ \bar h t$ \\  
-%% $w_3$ & $\bar h t ~a_0~ \bar h t ~b_0~ \bar h t$ \\
-%% \end{tabular}
-%% \end{example}
-\begin{example}   \label{gs1}
-  \renewcommand{\arraystretch}{1.2}
-  \begin{tabular}[t]{llc}
-     \multicolumn 2 c {\textit{Guarded String}} & \textit{Length}\\ \hline
-$v_1$ & $\hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 ~u~ \hat 2$ & $4$ \\
-$v_2$ & $\hat 0 ~u~ \hat 1 ~u~ \hat 2 ~u~ \hat 3$ & $3$ \\   
-$w_1$ & $H$ & $0$\\
-$w_2$ & $H ~a_0~ H$ & $1$ \\  
-$w_3$ & $T~a_0~ T ~b_0~ T$ & $3$\\
-  \end{tabular}
-    \renewcommand{\arraystretch}{1}
-\end{example}
 
@@
 
-The discussion of \ex{coin2} mentioned building worlds by incrementing worlds with events.
-This is accomplished in guarded string models with fusion product, a partial operation
-that combines two guarded strings, subject to the condition that the truth assignment
-at the end of the the first argument is identical to the truth assigment at the
-start of the second one.  \ex{fusion1} gives some examples.
+Kleene Algebra with Tests is an algebraic theory that is defined by
+equations and inequalities, which has model classes including
+guarded string models, relational models, finite models, and matrix
+models.  This paper focuses on defining a a concrete guarded
+string algebra, the elements of which are sets of guarded strings, from
+an Epik specification.
+Definitions and notation mostly follow \cite{kozen2001automata}.
+Additional syntax and semantics is included to cover multi-agent
+epistemic semantics.  Guarded strings over a finite alphabet ${\sf{P}}$
+are like ordinary strings, but with atoms 
+alternating with the symbols
+from ${\sf{P}}$.  In the algebra described by Figure \ref{epikfigure}, 
+is the set of events $\{a_1,a_0,b_1,b_0\}$, and in the elevator example,
+$\sf{P}$ is $\{u,d\}$.  In the elevator example, $\sf{B}$ is $\{p,q\}$, and in the coin
+example it is $\{h,t\}$.
+Assuming a trivially true state formula $\rho$,
+${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\rho}$
+in the elevator example is
+$\{\bar q \bar p$, $\bar q p$, $q \bar p$, $q p\}$,
+which we write $\{\hat0$, $\hat1$,
+$\hat2$, $\hat3\}$.
+In the coin example, as we already saw in
+\ex{effect},
+$\mathcal{A}^\varphi_{\mathsf B}$ is $\{h\bar{t},\bar{h}t\}$,
+for which we use the shorthand $\{H,T\}$.
+Guarded strings are strings of events, alternating with atoms,
+and beginning and ending with atoms.
+\ex{gs1} gives the encoding as guarded strings of the worlds in
+\ex{elevator1} and \ex{coin1}.
+The length of a guarded string
+$p$, written $|p|$ is the number of events in $p$.
+An atom such as $H$ is a guarded string of length $0$.
+
+\begin{example}   \label{gs1}
+  \renewcommand{\arraystretch}{1.2}
+  \begin{tabular}[t]{clc}
+{\textit{World}} & {\textit{Guarded string}} & \textit{Length}\\ \hline
+$v_1$ & $\hat 0 \,u\, \hat 1 \,u\, \hat 2 \,d\, \hat 1 \,u\, \hat 2$ & $4$ \\
+$v_2$ & $\hat 0 \,u\, \hat 1 \,u\, \hat 2 \,u\, \hat 3$ & $3$ \\   
+$w_1$ & $H$ & $0$\\
+$w_2$ & $H \,a_0\, H$ & $1$ \\  
+$w_3$ & $T\,a_0\, T \,b_0\, T$ & $2$\\
+  \end{tabular}
+    \renewcommand{\arraystretch}{1}
+\end{example}
+
+The discussion of \ex{coin2} mentioned building worlds by incrementing
+worlds with events.  This is accomplished in guarded string models
+with fusion product $\diamond$, a partial operation that combines two
+guarded strings, subject to the condition that the atom at the end of
+the the first argument is identical to the atom at the start of the
+second one.  \ex{fusion1} gives some examples.
 
 \begin{example} \label{fusion1}
-\hspace{-2mm}\(\hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 \diamond \hat 1 ~u~ \hat 2 = \hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 ~u~ \hat 2 \)
-
-\hspace{-2mm}\( \hat 0 ~u~ \hat 1 ~u~ \hat 2 ~d~ \hat 1 \diamond \hat 2 ~u~ \hat 3 = \mbox{\em{undefined}} \)
-
-\hspace{-2mm}\(H \diamond H ~a_0~ H = H ~a_0~ H \)
-
-\hspace{-2mm}\(T \diamond T ~a_1~ T = \mbox{\em{undefined}} \)
+\hspace{-2mm}\(\hat 0 \,u\, \hat 1 \,u\, \hat 2 \,d\, \hat 1 \; \diamond \; \hat 1 \,u\, \hat 2 = \hat 0 \,u\, \hat 1 \,u\, \hat 2 \,d\, \hat 1 \,u\, \hat 2 \)
+
+\hspace{-2mm}\( \hat 0 \,u\, \hat 1 \,u\, \hat 2 \,d\, \hat 1 \; \diamond \; \hat 2 \,u\, \hat 3 = \textit{undefined} \)
+
+\hspace{-2mm}\(H \; \diamond \; H \,a_1\, H = H \,a_0\, H \)
+
+\hspace{-2mm}\(H \; \diamond \; T \,a_1\, T = \textit{undefined} \)
 \end{example}
@@
 elements have the interpretation of propositions (sets of possible
-worlds) and/or event types.  An event such as $u$ in the guarded
-string model corresponds to the set of guarded strings where the bare
-event is flanked by compatible atoms. In standard KATs, every possible
-atom is allowed to surround every event, which leads to nonsensical
-strings like $\hat 1 ~u~\hat 0$ in the denotation of $u$, indicating
-that the elevator travelled up from floor 1 to floor 0 (a physical
-impossibility).
-
-Fortunately, Epik provides a syntax to specify the input-output
-behavior via the \texttt{event} declarations. From these we construct
-a function $\mathcal E$ from agents to relations on atoms that specify
-the effects each action can have. Notably for every declaration
-$\texttt{event}~e~\zeta$, we have $\mathcal E^\varphi(e) =
-\llbracket\zeta\rrbracket^\varphi$. We can lift $\mathcal E^\varphi$
-to $\hat{\mathcal E}$ which decorates the events with their permitted
-atoms, that is $\hat{\mathcal E}^\varphi(e) = \{\alpha~e~\beta \mid
-\alpha, \beta \in \mathcal E^\varphi (e) \}$. See the examples in
-\ex{events1}.
-% u (\bar{q}\bar{p}):(\bar{q}p) + (\bar{q}p):(q\bar{p}) + (q\bar{p}):(qp) 173
-% d (\bar{q}p):(\bar{q}\bar{p}) + (q\bar{p}):(\bar{q}p) + (qp):(q\bar{p})
+worlds).  In a free guarded string model for KAT, any event can be
+adjacent to any atom in a guarded string that is an element of the
+underlying set for the algebra.  We instead impose the constraints
+coming from the state and effect formulas.  \ex{wf1} defines the
+well-formed guarded strings determined by and Epic specification.  (i)
+says that each atom is consistent with the state constraint, and (ii)
+says that each constituent token event $\alpha_ie_i\alpha_{i+1}$ is
+consistent with the effect constraint on $e_i$.\footnote{
+An alternative is to define equations such as
+$\bar{\phi}=0$ (from the state formula $\phi$)
+and $a_1\,{=}\,ha_{1}h$ (from the effect formula $h\mathbin{:}h$ for event $a_1$),
+and construct a quotient algebra from the equivalence relation generated
+by these equations.  This results in equating sets of guarded strings in the free algebra that
+differ by guarded strings that are ill-formed according to the state and
+effect formulas.  In the development in the text, we instead use
+a set of guarded strings that are well-formed according to the state
+and effect formulas as the representative of the equivalence class.
+}
+
+\begin{example}   \label{wf1}
+Given $\sf{P}$, $\sf{B}$, a
+state formula $\varphi$, and an effect formula $\zeta_e$ for each
+event $e$ in $\sf{P}$,    
+$\alpha_0e_0 ... e_n\alpha_{n+1}$ is well-formed iff
+
+\vspace{1mm}
+
+\hspace{2mm}(i) $\element{\alpha_i}{{\mathcal A}^{\varphi}_{\mathsf{B}}}$ ($0{\leq}i{\leq}n$), and
+
+\hspace{2mm}(ii) $\element{\tuple{\alpha_i,\alpha_{i+1}}}{\sem{\zeta_{ei}}^{\varphi}}$, ($0{\leq} i {\le} n$).
+\end{example}
+
+Well-formed guarded strings have the interpretation of worlds in the
+application to natural-language semantics.  The set of possible worlds
+in the Kripke frame determined by an Epik specification is the set of
+well-formed guarded strings.  At this point, we could say in the way
+familiar from the type theory of possible worlds semantics that any
+set of worlds is a proposition, so that the set of propositions is the
+power set of the set of worlds \citep{montague,gallin}.  We will
+instead define a more restrictive set of propositions corresponding to the regular sets
+of strings. This is deferred to the next section.
+
+In the natural language application, sets of well-formed guarded strings 
+have the addional interpretation of event types.
+An event-type is something that can ``happen'' is different worlds.
+The event of the elevator going up is modeled not as the bare event
+symbol $u$ or its unit set $\{u\}$, but as the set of guarded strings
+\( \hat 0 \,u\, \hat 1, \hat 1 \,u\, \hat 2,  \hat 2 \,u\, \hat 3\).
+The event of the elevator going up in a given world $w$ corresponds
+to $w$ being incremented to form $w \diamond x$, where $x$ is an element
+of the event type.  \ex{events1} gives the event types for
+the examples.
 
@@
 \begin{array}[t]{l|c}
-  e &  \hat{\mathcal E}^\varphi(e) \\ \hline
-  u & \{ \hat 0 ~u~ \hat 1, \hat 1 ~u~ \hat 2,  \hat 2 ~u~ \hat 3\}\\
-  d & \{ \hat 1 ~d~ \hat 0, \hat 2~d~\hat 1, \hat 3 ~d~ \hat 2\}\\
-  a_1 & \{H ~a_1~ H\}\\
-  a_0 & \{T ~a_0~ T\}\\
-  b_1 & \{H ~b_1~ H \}\\
-  b_0 & \{T ~b_0~ T \}
+   &  {\mbox{Event type}}\\ \hline
+  u & \{ \hat 0 \,u\, \hat 1, \hat 1 \,u\, \hat 2,  \hat 2 \,u\, \hat 3\}\\
+  d & \{ \hat 1 \,d\, \hat 0, \hat 2\,d\,\hat 1, \hat 3 \,d\, \hat 2\}\\
+  a_1 & \{H \,a_1\, H\}\\
+  a_0 & \{T \,a_0\, T\}\\
+  b_1 & \{H \,b_1\, H \}\\
+  b_0 & \{T \,b_0\, T \}
 \end{array}
@@
   \renewcommand{\arraystretch}{1}
-
 \end{example}
 
-It remains to define the Kripke relation on guarded strings from an
-agent specification as in Figure \ref{epikfigure}. In Epik, an agent
-specification pairs each bare event with a set of ($+$-separated) bare
-events, and so determines a relation between bare events, call it
-relation $R_a$ for an agent $a$.  We can lift this to a relation
-$\hat{R_a}$ between guarded strings, as defined in \ex{Rhathat}. The
-operation $(\mathop;)$ in the definition is KAT product, which enforces
-matching of tests.
+The construction so far defines a set of worlds $W$ from an Epik specification.
+Normally and in our examples, $W$ is countably infinite, though some choices of effect
+formulas can result in a finite set of worlds.
+The next step is to define an alternative relation $R_a$ on worlds for each
+agent $a$. This will result in a Kripke frame $\tuple{W,R_a,...}$ in the 
+standard sense, consisting of a set of worlds, and a world-alternative relation
+for each agent \citep{kripke, hintikka}.
+An Epik specification defines an alternative relation on bare events for each
+agent $a$, which we notate as $\hat{R}_{a}$.  This should be lifted to a relation
+$R_a$ on worlds.  The rough idea is that when a world $w$ is incremented with
+an event $e$, in the resulting world $w \diamond e$, epistemic alternatives
+for agent $a$ are of the form $w' \diamond e'$, where $w'$ is an alternative to
+for $a$ in $w$, and $e'$ is and event-alternative for $a$ (Moore --, Baltag --).\footnote{
+ In this it is important that the event-alternative relation for an agent
+ is constant across worlds. We anticipate that the definition given here produces
+ results equivalent to what is found in literature on event alternatives in
+ dynamic epistemic semantics, though we have not verified this.  That literature primarily focuses on mapping
+ an epistemic model for a single time and situation to another, and uses general first-order models, rather than
+ guarded string models. See ---, and articles in ---.  We picked up the idea from papers by Moss and his colleagues, together
+ with Moore () and subsequent literature in situation theory.  This previous literature is motivated by epistemic
+ logic and AI planning, rather than natural language semantics as in our application.
+} This need to be implemented in a way that takes account of pre- and post-conditions for
+events. For this, our approach is to refer the definition of well-formed guarded strings.
+\ex{raise1} defines a relation on worlds from a relation on bare events.
+
+\begin{example}  \label{raise1}
+Let a set of guarded strings $W$ over events P and primitive tests B, and a relation $\hat{R}$ on P be given.
+The corresponding relation $R$ on $W$ holds between a guarded string
+\[
+\alpha_0e_0 ... e_n\alpha_{n+1} 
+\]
+in $W$ and a guarded string $q$ in $W$ iff $q$ is of the form
+
+\[
+\alpha_0'e_0' ... e_n'\alpha_{n+1}', 
+\]
+
+where for $0 \leq n$, $\element{\tuple{e_i,e'_i}}{R}$.
+
+\end{example}
+
+This requires that in the alternative world $q$, each constituent event $e'_i$ is an alternative
+to the event $e_i$ in the base world according to $R$.  Compatibilities between events in the alternative
+world are enforced by the general definition of worlds. Consider a scenario like the one from Figure 1, with
+an additional agent Cem. The base world $T b_0 c_0$ is one where the coin is tails, and first Bob looks at tails,
+and them Cem looks at tails. The first event $b_0$ has the alternatives $b_0$ and $b_1$ for Aly,
+and the second event $c_0$ has the alternatives $c_0$ and $c_1$ for Aly.
+This results in the combinations $b_0c_0$, $b_0c_1$, $b_1c_0$, $b_1c_1$. But these
+are filtered by post- and pre-conditions of events in the alternative world, so that
+the set of alernatives for Aly in $T b_0 c_0$ is  $\{T b_0 c_0,H b_1 c_1\}$,
+with two world-alternatives instead of four.
 
@@
 \hspace{6mm} u_1\mathrel{R}v_1 \wedge ... \wedge u_n\mathrel{R}v_n \wedge\\
-\hspace{6mm} x \in \llbracket u_1 ; ... ; u_n \rrbracket^{\varphi,{\mathcal{E}}}\hspace{1mm} \wedge \\
-\hspace{6mm} y \in \llbracket v_1 ; ... ; v_n \rrbracket^{\varphi,{\mathcal{E}}} \hspace{2mm} \}
+\hspace{6mm} x \in \sem{ u_1 ; ... ; u_n }^{\varphi,{\mathcal{E}}}\hspace{1mm} \wedge \\
+\hspace{6mm} y \in \sem{ v_1 ; ... ; v_n }^{\varphi,{\mathcal{E}}} \hspace{2mm} \}
 \end{array}
@@
 
+Fortunately, Epik provides a syntax to specify the input-output
+behavior via the \texttt{event} declarations. From these we construct
+a function $\mathcal E$ from agents to relations on atoms that specify
+the effects each action can have. Notably for every declaration
+$\texttt{event}~e~\zeta$, we have $\mathcal E^\varphi(e) =
+\sem{\zeta}^\varphi$. We can lift $\mathcal E^\varphi$
+to $\hat{\mathcal E}$ which decorates the events with their permitted
+atoms, that is $\hat{\mathcal E}^\varphi(e) = \{\alpha~e~\beta \mid
+\alpha, \beta \in \mathcal E^\varphi (e) \}$. See the examples in
+\ex{events1}.
+% u (\bar{q}\bar{p}):(\bar{q}p) + (\bar{q}p):(q\bar{p}) + (q\bar{p}):(qp) 173
+% d (\bar{q}p):(\bar{q}\bar{p}) + (q\bar{p}):(\bar{q}p) + (qp):(q\bar{p})
+
 \section{A Language for Epistemic Sets of Guarded Strings}
@@
 We interpret each term in as a set of guarded strings as indicated in
-Figure~\ref{fig:interp}. The function $\llbracket p
-\rrbracket^{\varphi,{\mathcal{E}}}$ interprets a term $p$ as a set of
+Figure~\ref{fig:interp}. The function $\sem{ p
+}^{\varphi,{\mathcal{E}}}$ interprets a term $p$ as a set of
 guarded strings consistent with state constraints $\varphi$ and effect
@@
 relative to a fixed relation $\hat{R_a}$ between guarded strings,
-$\Diamond_a~p = \{u \mid \exists v. v \in \llbracket
-p\rrbracket^{\phi, \mathcal E} \wedge u \mathrel{\hat{R_a}} v\}$.  Here $u$
+$\Diamond_a~p = \{u \mid \exists v. v \in \sem{
+p}^{\phi, \mathcal E} \wedge u \mathrel{\hat{R_a}} v\}$.  Here $u$
 and $v$ are guarded strings, while $x$ is an element of $K$.
@@
   \[\begin{array}{l >{\triangleq}c l}
-  \llbracket 0 \rrbracket^{\varphi,{\mathcal{E}}} && \{\} \\  
-  \llbracket 1 \rrbracket^{\varphi,{\mathcal{E}}} && \mathcal A^\varphi_{\mathsf B}\\
-  \llbracket e \rrbracket^{\varphi,{\mathcal{E}}}&& \hat{\mathcal E}^\varphi(e) \\
-  \llbracket \psi \rrbracket^{\varphi,{\mathcal{E}}} &&  \mathcal A^{\varphi~\psi}\\
-  \llbracket{p+q}\rrbracket^{\varphi,{\mathcal{E}}} && \llbracket{p}\rrbracket^{\varphi,{\mathcal{E}}} \cup \llbracket q \rrbracket^{\varphi,{\mathcal{E}}}\\
-  \llbracket{p;q}\rrbracket^{\varphi,{\mathcal{E}}} && \llbracket p \rrbracket^{\varphi,{\mathcal{E}}} \diamond \llbracket{q}\rrbracket^{\varphi,{\mathcal{E}}}\\
-  \llbracket p^*\rrbracket^{\varphi,{\mathcal{E}}} && \displaystyle\bigcup_n {\underbrace{\llbracket p \rrbracket^{\varphi,{\mathcal{E}}}\diamond \ldots \diamond \llbracket p \rrbracket^{\varphi,{\mathcal{E}}}}_{n}} \\
-  \llbracket \neg p \rrbracket^{\varphi,{\mathcal{E}}} && \llbracket \_^*\rrbracket^{\varphi,{\mathcal{E}}} \setminus \llbracket p \rrbracket^{\varphi,{\mathcal{E}}} \\
-  \llbracket \Diamond_a~p \rrbracket && \{x \mid x \mathbin{\hat{R_a}} y, y \in \llbracket p \rrbracket^{\varphi,\mathcal E}\}
+  \sem{ 0 }^{\varphi,{\mathcal{E}}} && \{\} \\  
+  \sem{ 1 }^{\varphi,{\mathcal{E}}} && \mathcal A^\varphi_{\mathsf B}\\
+  \sem{ e }^{\varphi,{\mathcal{E}}}&& \hat{\mathcal E}^\varphi(e) \\
+  \sem{ \psi }^{\varphi,{\mathcal{E}}} &&  \mathcal A^{\varphi~\psi}\\
+  \sem{{p+q}}^{\varphi,{\mathcal{E}}} && \sem{{p}}^{\varphi,{\mathcal{E}}} \cup \sem{ q }^{\varphi,{\mathcal{E}}}\\
+  \sem{{p;q}}^{\varphi,{\mathcal{E}}} && \sem{ p }^{\varphi,{\mathcal{E}}} \diamond \sem{{q}}^{\varphi,{\mathcal{E}}}\\
+  \sem{ p^*}^{\varphi,{\mathcal{E}}} && \displaystyle\bigcup_n {\underbrace{\sem{ p }^{\varphi,{\mathcal{E}}}\diamond \ldots \diamond \sem{ p }^{\varphi,{\mathcal{E}}}}_{n}} \\
+  \sem{ \neg p }^{\varphi,{\mathcal{E}}} && \sem{ \_^*}^{\varphi,{\mathcal{E}}} \setminus \sem{ p }^{\varphi,{\mathcal{E}}} \\
+  \sem{ \Diamond_a~p } && \{x \mid x \mathbin{\hat{R_a}} y, y \in \sem{ p }^{\varphi,\mathcal E}\}
 
@@
 guarded strings, defined to every set the universe $\_^*$ except those
-denoted by $p$. Note that $\llbracket 0^c
-\rrbracket^{\varphi,{\mathcal{E}}} =
-\llbracket\_^*\rrbracket^{\varphi,{\mathcal{E}}}$.
+denoted by $p$. Note that $\sem{ 0^c
+}^{\varphi,{\mathcal{E}}} =
+\sem{\_^*}^{\varphi,{\mathcal{E}}}$.
 
@@
 constraints $\varphi$, effect relations ${\mathcal{E}}$, and a concrete guarded
-string model for each term using $\llbracket -
-\rrbracket^{\varphi,{\mathcal{E}}}$.
+string model for each term using $\sem{ -
+}^{\varphi,{\mathcal{E}}}$.
 %% $\tuple{K,+,;,*,\bar{},0,1,{}^c,\Diamond_{a_1},...,\Diamond_{a_n}}$.
@@
   0}R^n\), where $R^n$ is the $n$-times concatenation product of $R$
-with itself (the $0$-times concatenation product is $\llbracket 1
-\rrbracket^{\varphi,\mathcal E}$). This is an operation in the finite
+with itself (the $0$-times concatenation product is $\sem{ 1
+}^{\varphi,\mathcal E}$). This is an operation in the finite
 state calculus.  Figure \ref{relclosure} defines the corresponding
@@
 \rrparenthesis^{\varphi,\mathcal E}_n\right\rceil = \{g \mid g \in
-\llbracket p \rrbracket^{\varphi,\mathcal E}, |g| \leq n\}$.
+\sem{ p }^{\varphi,\mathcal E}, |g| \leq n\}$.
 

### 2020-06-22 tex/draft.tex
Edits through Section 2.

@@
 Let a set of guarded strings $W$ over events P and primitive tests B, and a relation $\hat{R}$ on P be given.
-The corresponding relation $R$ on $W$ holds between a guarded string
+The corresponding relation $R$ holds between a guarded string
 \[
@@
 \]
-in $W$ and a guarded string $q$ in $W$ iff $q$ is of the form
+in $W$ and a guarded string  $q$ iff $q$ is an element of $W$ and is of the form
 
@@
 
-where for $0 \leq n$, $\element{\tuple{e_i,e'_i}}{R}$.
+where for $0 \leq n$, $\element{\tuple{e_i,e'_i}}{\hat{R}}$.
 
@@
 
-This requires that in the alternative world $q$, each constituent event $e'_i$ is an alternative
-to the event $e_i$ in the base world according to $R$.  Compatibilities between events in the alternative
-world are enforced by the general definition of worlds. Consider a scenario like the one from Figure 1, with
-an additional agent Cem. The base world $T b_0 c_0$ is one where the coin is tails, and first Bob looks at tails,
-and them Cem looks at tails. The first event $b_0$ has the alternatives $b_0$ and $b_1$ for Aly,
-and the second event $c_0$ has the alternatives $c_0$ and $c_1$ for Aly.
-This results in the combinations $b_0c_0$, $b_0c_1$, $b_1c_0$, $b_1c_1$. But these
-are filtered by post- and pre-conditions of events in the alternative world, so that
-the set of alernatives for Aly in $T b_0 c_0$ is  $\{T b_0 c_0,H b_1 c_1\}$,
-with two world-alternatives instead of four.
-
-%% \begin{example}   \label{Rhat}
-%% \hspace{-2mm} \(
-%% \hat{R} \triangleq
-%% \{\tuple{u,v} \mid \exists c \exists d. c\mathrel R d \wedge u \in \hat{c}  \wedge v \in \hat{d} \, \} \)
-%% \end{example}
-
-\begin{example}   \label{Rhathat}
-\hspace{-2mm}\(
-\hat{R} \triangleq
-\hspace{-2mm}\begin{array}[t]{l}
-\{ \tuple{x,y} \mid \exists u_1 ... \exists u_n \exists v_1 ... \exists v_n. \\
-\hspace{6mm} u_1\mathrel{R}v_1 \wedge ... \wedge u_n\mathrel{R}v_n \wedge\\
-\hspace{6mm} x \in \sem{ u_1 ; ... ; u_n }^{\varphi,{\mathcal{E}}}\hspace{1mm} \wedge \\
-\hspace{6mm} y \in \sem{ v_1 ; ... ; v_n }^{\varphi,{\mathcal{E}}} \hspace{2mm} \}
-\end{array}
-\)
+This requires that in the alternative world $q$, each constituent event $e'_i$ is an
+alternative to the event $e_i$ in the base world according to $\hat{R}$.
+Compatibilities between events in the alternative world are enforced by the
+requirement that the alternative $q$ is an element of $W$, so that 
+state and effect formulas are enforced. Consider a scenario like the one from
+Figure 1, with an additional agent Cem. The base world $T b_0 c_0$ is one where the
+coin is tails, and first Bob looks at tails, and them Cem looks at tails. The first
+event $b_0$ has the alternatives $b_0$ and $b_1$ for Aly, and the second event $c_0$
+has the alternatives $c_0$ and $c_1$ for Aly.  This results in the combinations
+$b_0c_0$, $b_0c_1$, $b_1c_0$, and $b_1c_1$. But these are filtered by post- and
+pre-conditions of events in the alternative world, so that the set of alernatives for
+Aly in $T b_0 c_0$ is $\{T b_0 c_0,H b_1 c_1\}$, with two world-alternatives instead
+of four.
+
+\section{The language of Epistemic KAT}
+In addition to a set of possible worlds and an algebra of propositions, the
+application in natural language requires a logical language for naming propositions.
+The standard language for Kleene algebra with tests has the signature
+$\tuple{K,+,\cdot,*,\bar{},0,1}$ \citep{kozen2001automata}.  In a guarded string
+model for KAT, $K$ is a set of guarded strings, $+$ is set union, $\cdot$ is fusion
+product raised to sets, $*$ is Kleene star, $\bar{}$ is complement for tests, $0$ the
+empty set (functioning as an identity for $0$, and as False for tests) and 1 is the
+set of atoms (functioning as the identity element for $\cdot$, and as True for
+tests).  To this signature we add a unary modal operation $\Diamond_a$ for each
+agent, and a unary complement operation $\not$ on elements of $K$. The latter is
+included because natural languages have sentence negation, and it should be possible
+to represent the propositional complement of a formula that captures the semantics of
+a natural language sentence.  In addition, universal box modalities are defined as
+duals of existential diamond modalities.
+
+With modalities and propositional negation added, the signature of $n$-agent
+epistemic KAT is $\tuple{K,+,\cdot,*,\bar{},0,1,\neg,\Diamond_1
+  ... \Diamond_n}$. Figure \ref{termfigure} defines the syntax of the language.
+\ex{interpretation1} illustrates the application of the language with some natural
+language sentences, and their representations in the logical language.  While there
+is no constant for the set of worlds, the term $\neg 0$ (the complement of the empty
+set) can be used as one.  Where $h$ is the primitive test ``it' heads'' from the coin
+example, $(\neg 0)h$ designates the set of worlds where it is heads.  The reason is
+that $h$ corresponds to the set of atoms that evaluate $h$ to true. Then the product
+term $(\neg 0)h$ restricts the set of worlds to the set of guarded strings that end
+with and atom that evaluate $h$ to True.
+
+\begin{example}   \label{interpretation1}
+\begin{tabular}[t]{ll}
+  It's heads. & It's tails. \\
+  $(\neg 0)h$ & $(\neg 0)t$ \\ \hline
+\end{tabular}
 \end{example}  
-
-This defines an epistemic alternative to world $x$ to be a world of
-the same length, where each component event in the alternative is an
-event-alternative to the event in corresponding position in the base
-world. Fusion product enforces preconditions and a correspondence
-between pre-states and post-states of events on both sides of the
-epistemic alternative relation.  This provides for finitely
-specifiable construction of epistemic models that reflect intuitions
-about information exchange and epistemic consequences of perceptual
-events.  See Section 6 for linguistic examples.  Since an epistemic
-alternative has the same lenght as its base world, it follows from the
-construction that agents know how many events have transpired in their
-base worlds.
-
-Fortunately, Epik provides a syntax to specify the input-output
-behavior via the \texttt{event} declarations. From these we construct
-a function $\mathcal E$ from agents to relations on atoms that specify
-the effects each action can have. Notably for every declaration
-$\texttt{event}~e~\zeta$, we have $\mathcal E^\varphi(e) =
-\sem{\zeta}^\varphi$. We can lift $\mathcal E^\varphi$
-to $\hat{\mathcal E}$ which decorates the events with their permitted
-atoms, that is $\hat{\mathcal E}^\varphi(e) = \{\alpha~e~\beta \mid
-\alpha, \beta \in \mathcal E^\varphi (e) \}$. See the examples in
-\ex{events1}.
-% u (\bar{q}\bar{p}):(\bar{q}p) + (\bar{q}p):(q\bar{p}) + (q\bar{p}):(qp) 173
-% d (\bar{q}p):(\bar{q}\bar{p}) + (q\bar{p}):(\bar{q}p) + (qp):(q\bar{p})
-
-\section{A Language for Epistemic Sets of Guarded Strings}
-
-To reason about epistemic alternatives, guarded strings are used to
-represent different possible worlds, so agents' knowledge/beliefs are
-represented by sets of guarded strings. The standard algebra for
-manipulating guarded strings is Kleene Algebra with Tests
-(\textsc{KAT}), which has the algebraic signature
-$\tuple{K,+,;,*,\bar{},0,1}$ \cite{kozen2001automata}. However, the
-KATs don't permit native reasoning about modal operators, so Epik's
-signature needs to subsume the KAT operators.
-
-Figure~\ref{fig:epik-grammar} depicts the syntax of Epik's formal term
-language. As in KAT, $0$ is the null element, $1$ is the unit element,
-$p + q$ indicates disjuction, $p;q$ (or simply $pq$) indicates
-sequence, the boolean operators for tests $\varphi$ are the standard
-ones (as in Figure~\ref{fig:effect-syntax}), and ${p}^*$ indicates
-iteration of $p$. We add some additional operations to reason about epistemic
-alternatives: the term $\Diamond_a~p$ represents epistemic possibility
-for agent $a$ in worlds $p$, and complement $\neg p$ represents the
-worlds not described by $p$. We can encode $\Box_a~p$, i.e. epistemic
-certainty for agent $a$ in worlds $p$, by $\Box_a~p =
-(\Diamond_a~p^c)^c$.
+\vspace{-3mm}
+\hspace{2mm}\begin{tabular}{l}
+Amy doesn't know that it's heads.  \\
+$\neg {\Box}_a (\neg 0)h$\\ \hline
+Amy knows whether it's tails. \\
+$\Box_a(\neg 0)t + \Box_a(\neg 0)\neg t$ \\ \hline
+It's heads and Bob knows that it's tails. \\
+$(\neg 0)h + \Box_a(\neg 0)\neg t$ \\  \hline
+Bob knows that Amy knows whether it's tails. \\
+$\Box_b(\Box_a(\neg 0)t + \Box_a(\neg 0)\neg t)$ \\ \hline
+\end{tabular}
+%% (\neg 0)t
 
 \begin{figure}
+
+  {\em Consider adding the syntax of tests including $\bar{}$. Consider making product implicit,
+    in place of $\mathbin{;}$.}
 
@@
 
+%% \begin{example}   \label{Rhat}
+%% \hspace{-2mm} \(
+%% \hat{R} \triangleq
+%% \{\tuple{u,v} \mid \exists c \exists d. c\mathrel R d \wedge u \in \hat{c}  \wedge v \in \hat{d} \, \} \)
+%% \end{example}
+
+{\em This stuff was at the end of Section 2, it became redundant. Also it refers to the interpretation function, which is introduced in Section 3.}
+
+\begin{example}   \label{Rhathat}
+\hspace{-2mm}\(
+\hat{R} \triangleq
+\hspace{-2mm}\begin{array}[t]{l}
+\{ \tuple{x,y} \mid \exists u_1 ... \exists u_n \exists v_1 ... \exists v_n. \\
+\hspace{6mm} u_1\mathrel{R}v_1 \wedge ... \wedge u_n\mathrel{R}v_n \wedge\\
+\hspace{6mm} x \in \sem{ u_1 ; ... ; u_n }^{\varphi,{\mathcal{E}}}\hspace{1mm} \wedge \\
+\hspace{6mm} y \in \sem{ v_1 ; ... ; v_n }^{\varphi,{\mathcal{E}}} \hspace{2mm} \}
+\end{array}
+\)
+\end{example}
+
+This defines an epistemic alternative to world $x$ to be a world of
+the same length, where each component event in the alternative is an
+event-alternative to the event in corresponding position in the base
+world. Fusion product enforces preconditions and a correspondence
+between pre-states and post-states of events on both sides of the
+epistemic alternative relation.  This provides for finitely
+specifiable construction of epistemic models that reflect intuitions
+about information exchange and epistemic consequences of perceptual
+events.  See Section 6 for linguistic examples.  Since an epistemic
+alternative has the same lenght as its base world, it follows from the
+construction that agents know how many events have transpired in their
+base worlds.
+
+Fortunately, Epik provides a syntax to specify the input-output
+behavior via the \texttt{event} declarations. From these we construct
+a function $\mathcal E$ from agents to relations on atoms that specify
+the effects each action can have. Notably for every declaration
+$\texttt{event}~e~\zeta$, we have $\mathcal E^\varphi(e) =
+\sem{\zeta}^\varphi$. We can lift $\mathcal E^\varphi$
+to $\hat{\mathcal E}$ which decorates the events with their permitted
+atoms, that is $\hat{\mathcal E}^\varphi(e) = \{\alpha~e~\beta \mid
+\alpha, \beta \in \mathcal E^\varphi (e) \}$. See the examples in
+\ex{events1}.
+% u (\bar{q}\bar{p}):(\bar{q}p) + (\bar{q}p):(q\bar{p}) + (q\bar{p}):(qp) 173
+% d (\bar{q}p):(\bar{q}\bar{p}) + (q\bar{p}):(\bar{q}p) + (qp):(q\bar{p})
+
 \pagebreak

### 2020-06-23 tex/draft.tex
Edits at start of Section 3

@@
 \usepackage{wasysym,graphicx}
+
+\newcommand{\event}{\sf{P}}
+\newcommand{\test}{\sf{B}}
 
@@
 epistemic KAT is $\tuple{K,+,\cdot,*,\bar{},0,1,\neg,\Diamond_1
-  ... \Diamond_n}$. Figure \ref{termfigure} defines the syntax of the language.
-\ex{interpretation1} illustrates the application of the language with some natural
-language sentences, and their representations in the logical language.  While there
-is no constant for the set of worlds, the term $\neg 0$ (the complement of the empty
-set) can be used as one.  Where $h$ is the primitive test ``it' heads'' from the coin
-example, $(\neg 0)h$ designates the set of worlds where it is heads.  The reason is
-that $h$ corresponds to the set of atoms that evaluate $h$ to true. Then the product
-term $(\neg 0)h$ restricts the set of worlds to the set of guarded strings that end
-with and atom that evaluate $h$ to True.
+  ... \Diamond_n}$. Figure \ref{termfigure} defines the syntax of the language, but
+using juxtaposition in place of the product symbol $\cdot$.
+As illustrated in \ex{interpretation1}, terms in this language are used to
+represent the proposositional semantic values of English sentences.\footnote{See Sections
+6 and 7 for explanation and further examples.  }
 
@@
 \begin{tabular}[t]{ll}
-  It's heads. & It's tails. \\
-  $(\neg 0)h$ & $(\neg 0)t$ \\ \hline
+Bob knows that Amy knows whether it's tails.  \\
+$\Box_b(\Box_a(\neg 0)t + \Box_a(\neg 0)\neg t)$
 \end{tabular}
 \end{example}
-\vspace{-3mm}
-\hspace{2mm}\begin{tabular}{l}
-Amy doesn't know that it's heads.  \\
-$\neg {\Box}_a (\neg 0)h$\\ \hline
-Amy knows whether it's tails. \\
-$\Box_a(\neg 0)t + \Box_a(\neg 0)\neg t$ \\ \hline
-It's heads and Bob knows that it's tails. \\
-$(\neg 0)h + \Box_a(\neg 0)\neg t$ \\  \hline
-Bob knows that Amy knows whether it's tails. \\
-$\Box_b(\Box_a(\neg 0)t + \Box_a(\neg 0)\neg t)$ \\ \hline
-\end{tabular}
-%% (\neg 0)t
 
@@
 
-  {\em Consider adding the syntax of tests including $\bar{}$. Consider making product implicit,
-    in place of $\mathbin{;}$.}
-
   \[\begin{array}{l>{\hspace{-.5em}}c>{\hspace{-0.75em}}l}
-  \multicolumn 2 l {\textrm{terms}} & e \in \mathsf{P} \\
-  p,q & ::=& e~|~0~|~1~|~\phi~|~p+p~|~p;p~|~p^*~|\neg p~|\Diamond_a~p
+  \multicolumn{3}{l}{\mbox{state terms $\sigma$ as in Figure 1}} \\
+  \multicolumn 2 l {\textrm{events}} & e \in \mathsf{P} \\
+  p,q & ::=& e~|~\sigma~|~p+q~|~pq~|~p^*~|~ \neg p ~|~ \Diamond_a\,p
   \end{array}\]
@@
 
-We interpret each term in as a set of guarded strings as indicated in
-Figure~\ref{fig:interp}. The function $\sem{ p
-}^{\varphi,{\mathcal{E}}}$ interprets a term $p$ as a set of
+Each term in the language is interpreted as a set of guarded strings.  Where $p$ is
+a term, $\sem{p}^{\test,\event,\varphi,\zeta}$ is the interpretation of the term.
+The superscript captures dependence on an Epik specification.
+
+interprets a term $p$ as a set of
 guarded strings consistent with state constraints $\varphi$ and effect
@@
       \smash{\hat{#1}}}}
+
+\newcommand{\epikspec}{\test,\event,\varphi,\zeta}
+
+\begin{figure}
+  \[\begin{array}{l >{\triangleq}c l}
+  \sem{ 0 }^{\epikspec} && \emptyset \\  
+  \sem{ 1 }^{\varphi,{\mathcal{E}}} && \mathcal A^\varphi_{\mathsf B}\\
+  \sem{ e }^{\varphi,{\mathcal{E}}}&& \hat{\mathcal E}^\varphi(e) \\
+  \sem{ \psi }^{\varphi,{\mathcal{E}}} &&  \mathcal A^{\varphi~\psi}\\
+  \sem{{p+q}}^{\varphi,{\mathcal{E}}} && \sem{{p}}^{\varphi,{\mathcal{E}}} \cup \sem{ q }^{\varphi,{\mathcal{E}}}\\
+  \sem{{p;q}}^{\varphi,{\mathcal{E}}} && \sem{ p }^{\varphi,{\mathcal{E}}} \diamond \sem{{q}}^{\varphi,{\mathcal{E}}}\\
+  \sem{ p^*}^{\varphi,{\mathcal{E}}} && \displaystyle\bigcup_n {\underbrace{\sem{ p }^{\varphi,{\mathcal{E}}}\diamond \ldots \diamond \sem{ p }^{\varphi,{\mathcal{E}}}}_{n}} \\
+  \sem{ \neg p }^{\varphi,{\mathcal{E}}} && \sem{ \_^*}^{\varphi,{\mathcal{E}}} \setminus \sem{ p }^{\varphi,{\mathcal{E}}} \\
+  \sem{ \Diamond_a~p } && \{x \mid x \mathbin{\hat{R_a}} y, y \in \sem{ p }^{\varphi,\mathcal E}\}
+
+  \end{array}
+  \]
+  \caption{Interpretation of Epik terms as sets of guarded strings}
+  \label{fig:interp2}
+\end{figure}
 

### 2020-07-18 tex/draft.tex
Different versions of Figure 4.
Edits up to fig. 4 in draft.text

@@
 \newcommand{\test}{\sf{B}}
+
+\pagestyle{plain}
 
@@
 propositions are sets of strings. An advantage in this is that it
-allows for a mathematical description of an algebra of a propositions,
+allows for a mathematical description of an algebra of propositions,
 coupled with a computational representation using either lazy lists of
@@
 epistemic modality is included for each agent.  {\em Action models} in
-dynamic epistemic semantics introduced the technique of constructing
+dynamic epistemic semantics introduce the technique of constructing
 epistemic models from primitive alternative relations on events, in
-order to capture the epistemic consequences of perceptual events
-\citep{baltag1999logic}. This is the basis for our construction of
-epistemic alternative relations.  Literature on finite state methods
-in linguistic semantics has used event strings and sets of event
-strings to theorize about tense and aspect in natural language
-semantics
+order to capture the epistemic consequences of perceptual and
+communicative events \citep{baltag1999logic}. This is the basis for
+our construction of epistemic alternative relations.  Literature on
+finite state methods in linguistic semantics has used event strings
+and sets of event strings to theorize about tense and aspect in
+natural language semantics
 \citep{fernando2004finite,fernando2007observing,carlson2009tense} and
-to express intensionality \cite{fernando2017intensions}.  Literature
-on finite state intensiononal semantics has investigated how to do the
+to express intensionality \cite{fernando2017intensions}.  Work on
+finite state intensiononal semantics investigates how to do the
 semantics of intensional complementation, including indirect
@@
   $b_1$  & Bob peeks at heads. \\
+  $a_0$  & Amy peeks at tails. \\
+  $b_0$  & Bob peeks at tails. \\     
 \end{tabular}
@@
 $\bar{h}t$, and disallow $ht$ and $\bar{h}\bar{t}$.  This is enforced
-by a {\em state formula}, which is a simply Boolean formula, in this
-case \exi{effect}{a}.  Where {\sf{B}} is a set of atomic tests and
-$\phi$ is a state constraint over {\sf{B}},
+by a {\em state formula}, which is a Boolean formula, in this case the
+one given on the second line of \ex{effect}.  Where {\sf{B}} is a set
+of atomic tests and $\phi$ is a state constraint over {\sf{B}},
 ${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\phi}$ is the set of valuations of
-{\sf{B}} that make formula $\phi$ true. The valuations are called
-atoms, because they correspond to the atoms of Boolean algebra of
-tests \citep{kozen2001automata}.
+{\sf{B}} that make formula $\phi$ true. Valuations are called atoms,
+because they correspond to the atoms of a Boolean algebra of tests
+\citep{kozen2001automata}.
 
@@
 
-Figure \ref{epikfigure} shows an Epik program that describes a
-possible worlds model for two agents with information about one coin,
-and events of the agents semi-privately looking at the coin. The line
-beginning with {\tt state} enumerates ${\sf{B}}$. The
-line beginning with {\tt constraint} gives the state formula.
-The lines
-beginning with {\tt event} declare events and their effect formulas.
-Finally
-the lines beginning with {\tt agent} define {\em event alternative}
-relations for agents.  Each clause with an arrow has a single event
-symbol on the left, and a disjunction of alternative events on the
-right of the arrow. The interpretation of Amy's alternatives for $b_1$
-(Bill peeks at heads), is that when $b_1$ happens, for Amy either
-$b_1$ or $b_0$ (Bill peeks at tails) could be happening.
+Epik is a specification language for possible worlds models that
+includes declarations of events and states, state formulas, effect
+formulas, and additional information.  Figure \ref{epikfigure} shows
+an Epik program that describes a possible worlds model for two agents
+with information about one coin, and events of the agents
+semi-privately looking at the coin. The line beginning with {\tt
+  state} enumerates ${\sf{B}}$. The line beginning with {\tt
+  constraint} gives the state formula.  The lines beginning with {\tt
+  event} declare events and their effect formulas.  Finally the lines
+beginning with {\tt agent} define {\em event alternative} relations
+for agents.  Each clause with an arrow has a single event symbol on
+the left, and a disjunction of alternative events on the right of the
+arrow. The interpretation of Amy's alternatives for $b_1$ (Bill peeks
+at heads), is that when $b_1$ happens, for Amy either $b_1$ or $b_0$
+(Bill peeks at tails) could be happening.
 
@@
 
-Kleene Algebra with Tests is an algebraic theory that is defined by
-equations and inequalities, which has model classes including
-guarded string models, relational models, finite models, and matrix
-models.  This paper focuses on defining a a concrete guarded
-string algebra, the elements of which are sets of guarded strings, from
-an Epik specification.
-Definitions and notation mostly follow \cite{kozen2001automata}.
-Additional syntax and semantics is included to cover multi-agent
-epistemic semantics.  Guarded strings over a finite alphabet ${\sf{P}}$
-are like ordinary strings, but with atoms 
+This paper focuses on defining a concrete possible worlds model from
+an Epik specification.  The models are an extension of guarded-string
+models for Kleene Algebra with Tests (KAT). This is an algebraic
+theory that has model classes including guarded string models,
+relational models, finite models, and matrix models.  Our definitions
+and notation follow \cite{kozen2001automata}.  We add syntax and
+semantics is included to cover multi-agent epistemic semantics.
+
+Guarded strings over a finite alphabet ${\sf{P}}$
+are like ordinary strings, but with atoms over a set ${\sf{B}}$  
 alternating with the symbols
 from ${\sf{P}}$.  In the algebra described by Figure \ref{epikfigure}, 
-is the set of events $\{a_1,a_0,b_1,b_0\}$, and in the elevator example,
-$\sf{P}$ is $\{u,d\}$.  In the elevator example, $\sf{B}$ is $\{p,q\}$, and in the coin
-example it is $\{h,t\}$.
-Assuming a trivially true state formula $\rho$,
-${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\rho}$
-in the elevator example is
-$\{\bar q \bar p$, $\bar q p$, $q \bar p$, $q p\}$,
-which we write $\{\hat0$, $\hat1$,
-$\hat2$, $\hat3\}$.
-In the coin example, as we already saw in
-\ex{effect},
-$\mathcal{A}^\varphi_{\mathsf B}$ is $\{h\bar{t},\bar{h}t\}$,
-for which we use the shorthand $\{H,T\}$.
-Guarded strings are strings of events, alternating with atoms,
-and beginning and ending with atoms.
-\ex{gs1} gives the encoding as guarded strings of the worlds in
-\ex{elevator1} and \ex{coin1}.
-The length of a guarded string
-$p$, written $|p|$ is the number of events in $p$.
-An atom such as $H$ is a guarded string of length $0$.
+${\sf{P}}$ is the set of events $\{a_1,a_0,b_1,b_0\}$, and
+${\sf{B}}$ is $\{h,t\}$.
+In the elevator example,
+$\sf{P}$ is $\{u,d\}$, and $\sf{B}$ is $\{p,q\}$
+
+Assuming a trivially true state formula $\rho$, the set of atoms
+${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\rho}$ in the elevator example is
+$\{\bar q \bar p$, $\bar q p$, $q \bar p$, $q p\}$, which we write
+$\{\hat0$, $\hat1$, $\hat2$, $\hat3\}$.  In the coin example, as we
+already saw in \ex{effect}, $\mathcal{A}^\varphi_{\mathsf B}$ is
+$\{h\bar{t},\bar{h}t\}$, for which we use the shorthand $\{H,T\}$.  A
+guarded string over ${\sf{P}}$ and ${\sf{B}}$ is a strings of events
+from ${\sf{P}}$, alternating with atoms over ${\sf{B}}$, and beginning
+and ending with atoms.  \ex{gs1} gives the encoding as guarded strings
+of the worlds in \ex{elevator1} and \ex{coin1}.  The length of a
+guarded string $p$, written $|p|$ is the number of events in $p$.  An
+atom such as $H$ is a guarded string of length $0$.
 
@@
 Rather than individual guarded strings, elements of a guarded string
-model for KAT are sets of guarded strings.  In the application, these
-elements have the interpretation of propositions (sets of possible
-worlds).  In a free guarded string model for KAT, any event can be
+model for KAT are sets of guarded strings.  In our application, these
+elements have the interpretation of propositions, which are sets of possible
+worlds.  In a free guarded string model for KAT, any event can be
 adjacent to any atom in a guarded string that is an element of the
@@
 coming from the state and effect formulas.  \ex{wf1} defines the
-well-formed guarded strings determined by and Epic specification.  (i)
-says that each atom is consistent with the state constraint, and (ii)
+well-formed guarded strings determined by and Epik specification.  Condition (i)
+says that each atom is consistent with the state constraint, and condition (ii)
 says that each constituent token event $\alpha_ie_i\alpha_{i+1}$ is
@@
 in the Kripke frame determined by an Epik specification is the set of
-well-formed guarded strings.  At this point, we could say in the way
-familiar from the type theory of possible worlds semantics that any
-set of worlds is a proposition, so that the set of propositions is the
-power set of the set of worlds \citep{montague,gallin}.  We will
-instead define a more restrictive set of propositions corresponding to the regular sets
-of strings. This is deferred to the next section.
+well-formed guarded strings.  At this point, we could say that any set
+of worlds is a proposition, so that the set of propositions is the
+power set of the set of worlds
+\citep{montague1975formal,gallin1975intensional}.  We will instead
+define a more restrictive set of propositions corresponding to the
+regular sets of strings. This is deferred to the next section.
 
@@
 symbol $u$ or its unit set $\{u\}$, but as the set of guarded strings
-\( \hat 0 \,u\, \hat 1, \hat 1 \,u\, \hat 2,  \hat 2 \,u\, \hat 3\).
+\(\{ \hat 0 \,u\, \hat 1, \hat 1 \,u\, \hat 2,  \hat 2 \,u\, \hat 3 \}\).
 The event of the elevator going up in a given world $w$ corresponds
-to $w$ being incremented to form $w \diamond x$, where $x$ is an element
-of the event type.  \ex{events1} gives the event types for
+to $w$ being incremented to form $w \diamond e$, where $e$ is an element
+of the event type.  \ex{events1} gives the event types in
 the examples.
@@
 
-The construction so far defines a set of worlds $W$ from an Epik specification.
-Normally and in our examples, $W$ is countably infinite, though some choices of effect
-formulas can result in a finite set of worlds.
-The next step is to define an alternative relation $R_a$ on worlds for each
-agent $a$. This will result in a Kripke frame $\tuple{W,R_a,...}$ in the 
-standard sense, consisting of a set of worlds, and a world-alternative relation
-for each agent \citep{kripke, hintikka}.
-An Epik specification defines an alternative relation on bare events for each
-agent $a$, which we notate as $\hat{R}_{a}$.  This should be lifted to a relation
-$R_a$ on worlds.  The rough idea is that when a world $w$ is incremented with
-an event $e$, in the resulting world $w \diamond e$, epistemic alternatives
-for agent $a$ are of the form $w' \diamond e'$, where $w'$ is an alternative to
-for $a$ in $w$, and $e'$ is and event-alternative for $a$ (Moore --, Baltag --).\footnote{
- In this it is important that the event-alternative relation for an agent
- is constant across worlds. We anticipate that the definition given here produces
- results equivalent to what is found in literature on event alternatives in
- dynamic epistemic semantics, though we have not verified this.  That literature primarily focuses on mapping
- an epistemic model for a single time and situation to another, and uses general first-order models, rather than
- guarded string models. See ---, and articles in ---.  We picked up the idea from papers by Moss and his colleagues, together
- with Moore () and subsequent literature in situation theory.  This previous literature is motivated by epistemic
- logic and AI planning, rather than natural language semantics as in our application.
-} This need to be implemented in a way that takes account of pre- and post-conditions for
-events. For this, our approach is to refer the definition of well-formed guarded strings.
+The construction so far defines a set of worlds from an Epik specification.  Normally
+and in our examples, $W$ is countably infinite, though some choices of effect
+formulas can result in a finite set of worlds.  The next step is to define an
+alternative relation $R_a$ on worlds for each agent $a$. This will result in a Kripke
+frame $\tuple{W,R_1,...,R_n}$ consisting of a set of worlds, and a world-alternative
+relation for each agent \citep{kripke1963semantical}.  An Epik specification defines
+an alternative relation on bare events for each agent $a$, which we notate as
+${R}_{a}$.  This should be lifted to a relation $\hat{R}_a$ on worlds.  The basic
+idea is that when a world $w$ is incremented with an event $e$, in the resulting
+world $w \diamond e$, epistemic alternatives for agent $a$ are of the form $w'
+\diamond e'$, where $w'$ is an alternative to for $a$ in $w$, and $e'$ is and
+event-alternative to $e$ for $a$.\footnote{In this it is important that the
+  event-alternative relation for an agent is constant across worlds. We anticipate
+  that the definition given here produces results equivalent to what is found in
+  literature on event alternatives in dynamic epistemic semantics, though we have not
+  verified this.  That literature primarily focuses on mapping an epistemic model for
+  a single time and situation to another, and uses general first-order models, rather
+  than guarded string models. See \citet{baltag1999logic}, \citet{van2007dynamic}, and
+  articles in \citet{ditmarsch2015handbook}.  We picked up the idea from papers by
+  Moss and his colleagues, together with \citet{moore1984formal} and subsequent
+  literature in situation theory.  This previous literature is motivated by epistemic
+  logic and AI planning, rather than natural language semantics.  } This needs to be
+implemented in a way that takes account of pre- and post-conditions for events. For
+this, our approach is to refer the definition of well-formed guarded strings.
 \ex{raise1} defines a relation on worlds from a relation on bare events.
@@
 \begin{example}  \label{raise1}
-Let a set of guarded strings $W$ over events P and primitive tests B, and a relation $\hat{R}$ on P be given.
-The corresponding relation $R$ holds between a guarded string
+  Let $W$ be a set of guarded strings over  events P and primitive tests B, a relation ${R}$ be a relation
+  on $P$. The corresponding relation $\hat{R}$ on $W$ holds between a guarded string
 \[
@@
 where for $0 \leq n$, $\element{\tuple{e_i,e'_i}}{\hat{R}}$.
-
 \end{example}
@@
 Compatibilities between events in the alternative world are enforced by the
-requirement that the alternative $q$ is an element of $W$, so that 
-state and effect formulas are enforced. Consider a scenario like the one from
-Figure 1, with an additional agent Cem. The base world $T b_0 c_0$ is one where the
-coin is tails, and first Bob looks at tails, and them Cem looks at tails. The first
-event $b_0$ has the alternatives $b_0$ and $b_1$ for Aly, and the second event $c_0$
-has the alternatives $c_0$ and $c_1$ for Aly.  This results in the combinations
-$b_0c_0$, $b_0c_1$, $b_1c_0$, and $b_1c_1$. But these are filtered by post- and
-pre-conditions of events in the alternative world, so that the set of alernatives for
-Aly in $T b_0 c_0$ is $\{T b_0 c_0,H b_1 c_1\}$, with two world-alternatives instead
-of four.
-
-\section{The language of Epistemic KAT}
+requirement that the alternative $q$ is an element of $W$, so that state and effect
+formulas are enforced.
+
+Consider a scenario like the one from Figure 1, but with an additional agent Cem. The
+base world $T b_0 T c_0 T$ is one where the coin is tails, and first Bob looks at
+tails, and them Cem looks at tails. The first event $b_0$ has the alternatives $b_0$
+and $b_1$ for Amy, and the second event $c_0$ has the alternatives $c_0$ and $c_1$
+for Amy.  This results in four combinations $b_0c_0$, $b_0c_1$, $b_1c_0$, and
+$b_1c_1$. But these are filtered by post- and pre-conditions of events in the
+alternative world, so that the set of alernatives for Amy in $T b_0 T c_0 T$ is $\{T
+b_0 T c_0 T,H b_1 H c_1 H\}$, with two world-alternatives instead of four.
+
+\section{The logical language of Epistemic KAT}
 In addition to a set of possible worlds and an algebra of propositions, the
@@
 $\tuple{K,+,\cdot,*,\bar{},0,1}$ \citep{kozen2001automata}.  In a guarded string
-model for KAT, $K$ is a set of guarded strings, $+$ is set union, $\cdot$ is fusion
-product raised to sets, $*$ is Kleene star, $\bar{}$ is complement for tests, $0$ the
-empty set (functioning as an identity for $0$, and as False for tests) and 1 is the
-set of atoms (functioning as the identity element for $\cdot$, and as True for
-tests).  To this signature we add a unary modal operation $\Diamond_a$ for each
-agent, and a unary complement operation $\not$ on elements of $K$. The latter is
-included because natural languages have sentence negation, and it should be possible
-to represent the propositional complement of a formula that captures the semantics of
-a natural language sentence.  In addition, universal box modalities are defined as
-duals of existential diamond modalities.
+model for KAT, $K$ is a set of sets of guarded strings, $+$ is set union, the
+operation $\cdot$ is fusion product raised to sets, $*$ is Kleene star, the operation
+$\;\bar{}\;$ is complement for tests, $0$ the empty set, and $1$ is the set of
+atoms.\footnote{$0$ has the dual role the identity for $+$ (union), and as False for
+  operations on tests.  $1$ has the dual role of the identity for product (fusion
+  product raised to sets), and True for tests.  }. To this signature we add a unary
+modal operation $\Diamond_a$ for each agent, and a unary complement operation $\neg$
+on elements of $K$.  The intended interpretation of $\Diamond_a p$ is the set of
+worlds where proposition $p$ is epistemically possible for agent $a$.  Propositional
+complement is included because natural languages have sentence negation.  In
+addition, universal box modalities are defined as duals of existential diamond
+modalities.
 
@@
 epistemic KAT is $\tuple{K,+,\cdot,*,\bar{},0,1,\neg,\Diamond_1
-  ... \Diamond_n}$. Figure \ref{termfigure} defines the syntax of the language, but
-using juxtaposition in place of the product symbol $\cdot$.
-As illustrated in \ex{interpretation1}, terms in this language are used to
-represent the proposositional semantic values of English sentences.\footnote{See Sections
-6 and 7 for explanation and further examples.  }
+  ... \Diamond_n}$. Figure \ref{termfigure} defines the syntax of the language.
+Juxtaposition is used in place of the product symbol $\cdot$.  Terms in this language
+are used to represent the proposositional semantic values of English sentences.
+\ex{interpretation1} gives some examples.  To explain the first one $\bullet$ as
+defined in Figure \ref{fig:epik-grammar} is the disjunction of the primitive events.
+Since a world is a well-formed sequence of events, $\bullet *$ is the set of worlds.
+Multiplying by the state symbol $h$ in the term $\bullet \! * \! h$ has the effect of
+conjoining $h$ with the atom at the end of the world.  So $\bullet\!*\!h$ is the set
+of worlds where the coin is heads.\footnote{A mapping from English sentences to logical
+  terms in epistemic KAT is presented in Section 6}.
 
@@
 \begin{tabular}[t]{ll}
-Bob knows that Amy knows whether it's tails.  \\
-$\Box_b(\Box_a(\neg 0)t + \Box_a(\neg 0)\neg t)$
+${\bullet}\!* t$ & It's tails. \\
+$\bullet \!* h$ & It's heads. \\  
+$\Box_a \bullet \!* h$  & Amy knows that it's heads. \\
+\end{tabular}
+\vspace{1mm}
+
+\begin{tabular}[t]{l}
+  $\Box_b(\Box_a\!\bullet\!* t + \Box_a\!\bullet \!* \neg t)$\\
+ Bob knows that Amy knows whether \\ \hspace{3mm} it's tails.  \\
 \end{tabular}
@@
 \begin{figure}
-
-  \[\begin{array}{l>{\hspace{-.5em}}c>{\hspace{-0.75em}}l}
-  \multicolumn{3}{l}{\mbox{state terms $\sigma$ as in Figure 1}} \\
-  \multicolumn 2 l {\textrm{events}} & e \in \mathsf{P} \\
-  p,q & ::=& e~|~\sigma~|~p+q~|~pq~|~p^*~|~ \neg p ~|~ \Diamond_a\,p
-  \end{array}\]
-
-  Derived Operators\\[.1em]
-  \(\Box_a p \triangleq \neg\Diamond_a\neg p\) \hfill \(p\&q \triangleq \neg(\neg p + \neg q)\) \hfill \(\_ \triangleq \sum\mathsf P\)
+\begin{tabular}{ll}
+events & $e \in \mathsf{P}$ \\
+states & $\sigma$ as in Figure 1 \\    
+$p,q$ & $::= e~|~\sigma~|~p+q~|~pq~|~p^*~|~ \neg p ~|~ \Diamond_a\,p$\\
+\end{tabular}
+
+\vspace{1mm}
+\begin{tabular}{cl}
+$\Box_a p$ & $\triangleq \neg\Diamond_a\neg p$ \\
+\(p \wedge q\) &  \(\triangleq \neg(\neg p + \neg q)\) \\
+$\bullet$ & \( \triangleq \sum_{e \in \mathsf P}e\)
+\end{tabular}
 
 \caption{The language of Epik terms and key derived
-  operators}
+  operators.}
 \label{fig:epik-grammar}
@@
 
-Each term in the language is interpreted as a set of guarded strings.  Where $p$ is
-a term, $\sem{p}^{\test,\event,\varphi,\zeta}$ is the interpretation of the term.
-The superscript captures dependence on an Epik specification.
+Each term in the language is interpreted as a set of guarded strings.  Where $p$ is a
+term, $\sem{p}^{\test,\event,\varphi,\zeta}$ is the interpretation of the term.  The
+superscript captures dependence on an Epik specification.
 
@@
 
+\newcommand{\AphiB}{\mathcal A^\varphi_{\mathsf B}}
+
 \begin{figure}
@@
   \sem{ 0 }^{\epikspec} && \emptyset \\  
-  \sem{ 1 }^{\varphi,{\mathcal{E}}} && \mathcal A^\varphi_{\mathsf B}\\
-  \sem{ e }^{\varphi,{\mathcal{E}}}&& \hat{\mathcal E}^\varphi(e) \\
-  \sem{ \psi }^{\varphi,{\mathcal{E}}} &&  \mathcal A^{\varphi~\psi}\\
-  \sem{{p+q}}^{\varphi,{\mathcal{E}}} && \sem{{p}}^{\varphi,{\mathcal{E}}} \cup \sem{ q }^{\varphi,{\mathcal{E}}}\\
-  \sem{{p;q}}^{\varphi,{\mathcal{E}}} && \sem{ p }^{\varphi,{\mathcal{E}}} \diamond \sem{{q}}^{\varphi,{\mathcal{E}}}\\
+  \sem{ 1 }^{\epikspec} && \mathcal A^\varphi_{\mathsf B}\\
+  \sem{ a }^{\epikspec} && \mathcal A^{a\varphi}_{\mathsf B}\\
+  \sem{ \bar{\sigma} }^{\epikspec} && \AphiB \setminus \sem{\sigma}^{\epikspec}\\
+  \sem{ e }^{\epikspec}&&  \setabs{\alpha e \beta}{\alpha \zeta_e \beta} \\
+  \sem{{p+q}}^{\epikspec} && \sem{{p}}^{\epikspec} \cup \sem{ q }^{\epikspec}\\
+  \sem{{pq}}^{\epikspec} && \setabs{x \diamond y}{ \begin{array}{l}x \in \sem{p}^{\epikspec} \\ y \in \sem{q}^{\epikspec}\\\mbox{$x \diamond y$ is defined}\end{array}} \\
   \sem{ p^*}^{\varphi,{\mathcal{E}}} && \displaystyle\bigcup_n {\underbrace{\sem{ p }^{\varphi,{\mathcal{E}}}\diamond \ldots \diamond \sem{ p }^{\varphi,{\mathcal{E}}}}_{n}} \\

### 2020-07-21 tex/draft.tex
Edits in Section 3 of draft.tex

@@
 \begin{example}  \label{raise1}
-  Let $W$ be a set of guarded strings over  events P and primitive tests B, a relation ${R}$ be a relation
+  Let $W$ be a set of guarded strings over  events P and primitive tests B, and ${R}$ be a relation
   on $P$. The corresponding relation $\hat{R}$ on $W$ holds between a guarded string
@@
 
+\newcommand{\epikspec}{\test,\event,\varphi,\zeta}
+\newcommand{\AphiB}{\mathcal A^\varphi_{\mathsf B}}
+
 Each term in the language is interpreted as a set of guarded strings.  Where $p$ is a
-term, $\sem{p}^{\test,\event,\varphi,\zeta}$ is the interpretation of the term.  The
-superscript captures dependence on an Epik specification.
-
-interprets a term $p$ as a set of
-guarded strings consistent with state constraints $\varphi$ and effect
-relations ${\mathcal{E}^\varphi}$. A bare event $e$ is interpreted as
-the set $\hat{\mathcal E}^\varphi(e)$. The constant $0$ is interpreted
-as the empty set.  The constant $1$ is interpreted as the set
-$\mathcal A^\varphi_{\mathsf B}$, i.e. $\{h\bar t, \bar h t\}$ in the
-coin example. The operation $p + q$ is interpreted as the union of the
-interpretations of $p$ and $q$. The operation $p;q$ interpreted as the
-fusion product of the interpretations of $p$ and $q$ raised to sets:
-$X \diamond Y = \{ x \diamond y \mid x \in X, y \in Y, x\diamond
-y~\textit{defined}\}$.  The operation $p*$ is Kleene star, which is
-interpreted to be least upper bound of the iterated fusion product of
-the denotation of $p$ with itself. Subsets of $1$ indicated by
-propositions $\psi$ are also elements of $K$, and these form the
-Boolean algebra of tests, denoting all atoms admitted by the
-constraint $\varphi$ and the boolean test $\psi$.
-
-The unary epistemic alternative operation $\Diamond_a$ for each agent
-$a$, is interpreted using using Kripke semantics, as pre-image
-relative to a fixed relation $\hat{R_a}$ between guarded strings,
-$\Diamond_a~p = \{u \mid \exists v. v \in \sem{
-p}^{\phi, \mathcal E} \wedge u \mathrel{\hat{R_a}} v\}$.  Here $u$
-and $v$ are guarded strings, while $x$ is an element of $K$.
+term, $\sem{p}^{\epikspec}$ is the interpretation of the term, where superscript
+captures dependence on an Epik specification.  Figure \ref{fig:interp2} defines the
+interpretation of terms recursively.  The interpretation $\sem{ 1 }^{\epikspec}$ of
+the multiplicative identity is the set of atoms that satisfy the state constraint
+$\varphi$.  Where $b$ is a primitive Boolean, $\sem{ b }^{\epikspec}$ is the set of
+atoms that satisfy the state constraint and where $b$ is true.  Where $e$ is a
+primitive event, $\sem{ e }^{\epikspec}$ is the set of guarded strings that have the
+form of $e$ flanked by compatible atoms, as determined by the event formula
+$\zeta_e$. The product $pq$ of terms $p$ and $q$ is interpreted with fusion product
+raised to sets of guarded strings.
+Kleene star is interpreted as the union of iterated products. $p^n$ is the $n$-times product
+of $p$ with itself, with $p^0 = 1$.
+Propositional complement is complement relative to the set of worlds.
+The epistemic formula $\Diamond_a p$ is interpreted with Kripke
+semantics for epistemic modality, as the pre-image of the
+complement proposition under the relation
+$\hat{R}_a$.  Here $\hat{R}_a$ is the world-alternative for agent $a$,
+as defined in \ex{raise1}.
 
@@
 
-\newcommand{\epikspec}{\test,\event,\varphi,\zeta}
-
-\newcommand{\AphiB}{\mathcal A^\varphi_{\mathsf B}}
-
 \begin{figure}
-  \[\begin{array}{l >{\triangleq}c l}
+  \(\begin{array}{l >{\triangleq}c l}
   \sem{ 0 }^{\epikspec} && \emptyset \\  
   \sem{ 1 }^{\epikspec} && \mathcal A^\varphi_{\mathsf B}\\
-  \sem{ a }^{\epikspec} && \mathcal A^{a\varphi}_{\mathsf B}\\
+  \sem{ b }^{\epikspec} && \mathcal A^{b\varphi}_{\mathsf B}\\
   \sem{ \bar{\sigma} }^{\epikspec} && \AphiB \setminus \sem{\sigma}^{\epikspec}\\
@@
   \sem{{pq}}^{\epikspec} && \setabs{x \diamond y}{ \begin{array}{l}x \in \sem{p}^{\epikspec} \\ y \in \sem{q}^{\epikspec}\\\mbox{$x \diamond y$ is defined}\end{array}} \\
-  \sem{ p^*}^{\varphi,{\mathcal{E}}} && \displaystyle\bigcup_n {\underbrace{\sem{ p }^{\varphi,{\mathcal{E}}}\diamond \ldots \diamond \sem{ p }^{\varphi,{\mathcal{E}}}}_{n}} \\
-  \sem{ \neg p }^{\varphi,{\mathcal{E}}} && \sem{ \_^*}^{\varphi,{\mathcal{E}}} \setminus \sem{ p }^{\varphi,{\mathcal{E}}} \\
-  \sem{ \Diamond_a~p } && \{x \mid x \mathbin{\hat{R_a}} y, y \in \sem{ p }^{\varphi,\mathcal E}\}
-
+  \sem{p\!*\!}^{\epikspec} && \displaystyle\bigcup_{n \geq 0} {\sem{ p^n }^{\epikspec}} \\
+  \sem{ \neg p }^{\epikspec} && \sem{\bullet\!*}^{\epikspec} \setminus \sem{ p }^{\epikspec} \\
   \end{array}
-  \]
-  \caption{Interpretation of Epik terms as sets of guarded strings}
+  \)
+
+\hspace{2mm}\(\sem{ \Diamond_a\,p } \hspace{4mm}{\triangleq} \hspace{4mm}\{x \mid \exists y . x{\hat{R}_a}y \wedge y \in \sem{ p }^{\epikspec}\} \)
+  \caption{Interpretation of Epik terms as sets of guarded strings (Version M)}
   \label{fig:interp2}
@@
 
-The complement operation ${p}^c$ is complement at the level of sets of
-guarded strings, defined to every set the universe $\_^*$ except those
-denoted by $p$. Note that $\sem{ 0^c
-}^{\varphi,{\mathcal{E}}} =
-\sem{\_^*}^{\varphi,{\mathcal{E}}}$.
-
-Summing up, given an Epik program with $n$ agents, we construct state
-constraints $\varphi$, effect relations ${\mathcal{E}}$, and a concrete guarded
-string model for each term using $\sem{ -
-}^{\varphi,{\mathcal{E}}}$.
-%% $\tuple{K,+,;,*,\bar{},0,1,{}^c,\Diamond_{a_1},...,\Diamond_{a_n}}$.
- $0^c$ is the set of worlds, and it may be countably infinite.
-$\Diamond_{a_i}$ is an epistemic modality for the $i$th agent.  Or
-referring to the Kripke relations $\hat{R}_i$, the construction
-defines a multi-agent Kripke frame
-$\tuple{0^c,\hat{R}_1,...,\hat{R}_n}$ (usually a countable one) from
-an Epik specification. The frame consists of a set of worlds, and an
-epistemic-alternative relation for each agent.  These models are used
-as a target for natural-language interpretation in Section 5 and
-Section 6, where we obtain semantic values such as \sem{Amy knows that
-  it's heads and Bob knows that Amy knows whether it is heads or
-  tails, and does not know that it's heads} as elements of
-$K$. Concretely the propositions are sets of guarded strings (usually
-countable ones), construed as sets of worlds as they figure in
-possible worlds semantics for natural language.
+Summing up, given an Epik specification, a state constraint $\varphi$, and effect
+relation $\zeta_e$ for each primitive are constructed. Given these, each term $p$ (as
+defined syntactically in Figure \ref{termfigure}) is interpreted as a set of guarded
+strings $\sem{p}^{\epikspec}$ over the Boolean primitives and event primitives in the
+specification.  Let $K^{\epikspec}$ be the sets that are interpretations of terms.
+Then
+$\tuple{K^{\epikspec},+,\cdot,*,\bar{},0,1,\neg,\Diamond_{a_1},...,\Diamond_{a_n}}$
+is a concrete guarded string interpretation for the signature of epistemic KAT, with
+operations as in Figure \ref{fig:interp2} (e.g. the binary operation $+$ is union,
+and the unary operation $\Diamond_a$ is pre-image relative to $\hat{R}_a$).  The
+construction can also be viewed as providing a concrete $n$-agent Kripke frame
+$\tuple{\sem{\bullet *}^{\epikspec}, \hat{R}_1,...,\hat{R}_n}$.\footnote{ The domain
+  of the Kripke frame differs from the domain of the guarded string model, because
+  the former is the set of worlds, while the latter is the set of propositions.  }
+The frame consists of a set of worlds, and an epistemic-alternative relation for each
+agent.  These models are used as a target for natural-language interpretation in
+Section 5 and Section 6.
 
@@
 relations between strings that was designed for use in computationonal phonology and
-morphographemics 
-\citep{kaplan1994regular,beesley2003finite}.
-Current implementations allow for the
-definition of functions with the status of defined operators on regular sets and
-relations
-\citep{hulden2009foma,karttunen2010update}.
-Such definitions are used to define an embedding of epistemic KAT in a
+morphographemics \citep{kaplan1994regular,beesley2003finite}.  Current
+implementations allow for the definition of functions that have the status of defined
+operations on regular sets and relations \citep{hulden2009foma,karttunen2010update}.
+Such definitions are used here to construct of a model for epistemic KAT inside the
+finite state calculus.  The space of worlds is a set of ordinary strings.  Bit
+sequences (sequences of 0's and 1's) are used for atoms, and these alternate with
+event symbols in the encoding of a world. \ex{gs1fst} gives the encoding of some of
+the worlds from the examples.  A string is a finite sequence of symbols, and {\tt 0},
+{\tt 1}, {\tt u}, and {\tt d}, are symbols.  {\tt a0} and {\tt b0} are special
+multi-character symbols that are found in implementations of the finite state
+calculus.
+
+\begin{example}   \label{gs1fst}
+  \renewcommand{\arraystretch}{1.2}
+Worlds coded as strings
+\vspace{1mm}
+
+ \hspace{-13mm} \begin{tabular}[t]{cl}
+{\textit{World}} & {\textit{String}} \\
+% $v_1$ & $ 0 \,u\,  1 \,u\,  2 \,d\,  1 \,u\,  2$ \\
+$v_1$ & {\tt 0 0 u 0 1 u 1 0 d  0 1 u  1 0} \\
+$v_2$ & $ 0 \,u\,  1 \,u\,  2 \,u\,  3$ \\   
+$w_1$ & $1 0$ \\
+$w_2$ & $1 0 \,a_0\, 1 0$ \\  
+$w_3$ & $0 1 \,a_0\, 0 1 \,b_0\, 0 1$ \\
+  \end{tabular}
+    \renewcommand{\arraystretch}{1}
+\end{example}
+
+Operations such as KAT product are defined operations in the finite
+state calculus.
+
+of epistemic KAT in a
 string algebra. The methodology follows Section 2 closely.  Let ${\cal K}$ be an

### 2020-09-18 tex/draft.tex
tex edits

@@
 natural language.  A proposition is a set or class of {\em possible
-  worlds}.  These are often assumed to be things with the same
-nature and complexity as the world we occupy \citep{lewis1986plurality}.  But
+  worlds}.  These are often assumed to be things with the same nature
+and complexity as the world we occupy \citep{lewis1986plurality}.  But
 alternatively, one can work with small idealized models, in order to
@@
 of lexical items are constructed as functions or relations from these
-small sets.  Such toy or idealized models are useful in research and
-in teaching, in that it is possible to represent propositions finitely
+sets.  Such toy or idealized models are useful in research and in
+teaching, in that it is possible to represent propositions finitely
 and explicitly, and to calculate with them. The point of this paper is
@@
 agents.  We describe a certain systematic way of defining such models,
-and illustrate how to apply them in natural language semantics.
-The focus is on epistemic semantics and clausal embedding. The fundamental move is to
-identify possible worlds with strings of primitive events, so that
-propositions are sets of strings. An advantage in this is that it
-allows for a mathematical description of an algebra of propositions,
-coupled with a computational representation using either lazy lists of
-strings, or finite state machines that describe sets of strings.
+and illustrate how to apply them in natural language semantics.  The
+focus is on epistemic semantics and clausal embedding. The fundamental
+move is to identify possible worlds with strings of primitive events,
+so that propositions are sets of strings. An advantage in this is that
+it allows for a mathematical description of an algebra of
+propositions, coupled with a computational representation using either
+lazy lists of strings, or finite state machines that describe sets of
+strings.
 
@@
 elements corresponding to propositions and event types in our
-application \citep{kozen2001automata}.\marginpar{2001 or 2003?}  The
-models we propose are an epistemic extension of guarded string models
-for KAT, where a unary operation interpreted as an existential
-epistemic modality is included for each agent.  {\em Action models} in
-dynamic epistemic semantics introduce the technique of constructing
-epistemic models from primitive alternative relations on events, in
-order to capture the epistemic consequences of perceptual and
-communicative events \citep{baltag1999logic}. This is the basis for
-our construction of epistemic alternative relations.  Literature on
-finite state methods in linguistic semantics has used event strings
-and sets of event strings to theorize about tense and aspect in
-natural language semantics
+application \citep{kozen2001automata}. The models we propose are an
+epistemic extension of guarded string models for KAT, where a unary
+operation interpreted as an existential epistemic modality is included
+for each agent.  {\em Action models} in dynamic epistemic semantics
+introduce the technique of constructing epistemic models from
+primitive alternative relations on events, in order to capture the
+epistemic consequences of perceptual and communicative events
+\citep{baltag1999logic}. This is the basis for our construction of
+epistemic alternative relations.  Literature on {\em finite state
+  methods in linguistic semantics} has used event strings and sets of
+event strings to theorize about tense and aspect in natural language
+semantics
 \citep{fernando2004finite,fernando2007observing,carlson2009tense} and
-to express intensionality \cite{fernando2017intensions}.  Work on
-finite state intensiononal semantics investigates how to do the
+to express intensions \cite{fernando2017intensions}.  Work on {\em
+  finite state intensiononal semantics} has investigated how to do the
 semantics of intensional complementation, including indirect
@@
 $w_1$ & $H$ & $0$\\
-$w_2$ & $H \,a_0\, H$ & $1$ \\  
-$w_3$ & $T\,a_0\, T \,b_0\, T$ & $2$\\
+$w_2$ & $H \,a_1\, H$ & $1$ \\  
+$w_3$ & $H\,a_1\, H \,b_1\, H$ & $2$\\
   \end{tabular}
@@
 the worlds from the examples.  A string is a finite sequence of symbols, and {\tt 0},
-{\tt 1}, {\tt u}, and {\tt d}, are symbols.  {\tt a0} and {\tt b0} are special
+{\tt 1}, {\tt u}, and {\tt d}, are symbols.  {\tt a0} and {\tt b0} are
 multi-character symbols that are found in implementations of the finite state
-calculus.
+calculus. The multi-character symbol {\tt a0} is an element of the alphabet that has
+no connection with the element {\tt a}.
 
@@
 $v_1$ & {\tt 0 0 u 0 1 u 1 0 d  0 1 u 1 0} \\
-$v_2$ & $ 0 \,u\,  1 \,u\,  2 \,u\,  3$ \\   
-$w_1$ & $1 0$ \\
-$w_2$ & $1 0 \,a_0\, 1 0$ \\  
-$w_3$ & $0 1 \,a_0\, 0 1 \,b_0\, 0 1$ \\
+$v_2$ & {\tt 0 0 u 0 1 u 1 0 u 1 1} \\
+$w_1$ & {\tt 1 0} \\
+$w_2$ & {\tt 1 0 a1 1 0} \\
+$w_3$ & {\tt 1 0 a1 1 0 b1 10} \\  
   \end{tabular}
@@
 
-Operations such as KAT product are defined operations in the finite
-state calculus.
-
-of epistemic KAT in a
-string algebra. The methodology follows Section 2 closely.  Let ${\cal K}$ be an
-epistemic algebra as described in Section 2.  A given element of ${\cal K}$ is
-represented in the the string algebra by the very same set of strings, i.e. by a set
-of strings that have the form of a sequence of bare event symbols, with interleaved
-Boolean vectors. Product in the KAT can not be modeled as concatenation in the string
-algebra, because this would not enforce identity of states, and would result in
-lengthening Boolean vectors at the concatenation point.  Instead, KAT product and KAT
-Kleene star are defined operations in the string algebra, see Figure \ref{fstfigure1}.  The
-operations concatenate in the string algebra, delete strings with non-matching tests,
-and then delete the second of two tests create a well-formed guarded string.
+Terms in the finite state calculus are interpreted as sets of strings, or for
+relational terms, as relations between strings.  Computationally, the sets and
+relations are represented by finite state acceptors.  As used here, a program in the
+Fst language of the finite state calculus is a straight-line program that defines a
+sequence of constants naming sets, constants naming relations, and functions mapping
+one or more regular relations to a regular set or relation.  The functions are
+defined as macros.  To illustrate, \ex{stex1} is part of an Epik program that
+declares states and a state constraint.  This is equivalent to the Fst program
+\ex{stex2}. The result of the sequence of definitions is that {\tt St} is the set of
+Boolean vectors or length four that are well-formed according to the state formula;
+Nst is the complement operation for tests; H1 is the set of well-formed Boolean
+vectors that have 1 in the first position; and so forth.  Translation from an Epik
+program to an Fst program is accomplished in Haskell.
+
+\begin{example}\label{stex1}
+Part of an Epik program
+\end{example}
+
+\vspace{-3mm} {\small
+\begin{verbatim}  
+   state H1 T1 H2 T2
+   constraint !(H1 T1) & (H1 + T1) 
+              !(H2 T2) & (H2 + T2) 
+\end{verbatim}}
+
+\begin{example}\label{stex2}
+Corresponding Fst program
+\end{example}
+
+\vspace{-3mm}
+{\small 
+\begin{verbatim} 
+   define Bool ["0"|"1"];
+   define St0 [Bool Bool Bool Bool];
+   define H1 [1 Bool Bool Bool];
+   define T1 [Bool 1 Bool Bool];
+   define H2 [Bool Bool 1 Bool];
+   define T2 [Bool Bool Bool 1];
+   define St St0 & ((((St0 - (H1 & T1)) 
+                 & (H1 | T1)) 
+                 & (St0 - (H2 & T2))) 
+                 & (H2 | T2));
+   define Nst(X) [St - X];
+   define H1 St & H1;
+   define T1 St & T1;
+   define H2 St & H2;
+   define T2 St & T2;
+\end{verbatim}}
+
+Event symbols are interpreted as event types, in the way discussed in connection with
+\ref{events1}.  For instance, defines an event type {\tt paT2type} of Amy peeking at the second coin
+in a scenario with two coins, and seeing that it is tails.\footnote{After the definition,
+  {\tt paT2type} is a constant symbol denoting a set of two strings. On the right hand side of the definition,
+  {\tt paT2} is the unit set of the symbol {\tt paT2}.  
+}  The first conjunct
+expresses that this event can only happen if the second coin is tails ({\tt H2}), and that the
+state of the second coin does not change.  The second conjunct expresses that the
+event does not change the state of the first coin -- either it is heads ({\tt H1}) and
+remains heads, or it is tails ({\tt T1}) and remains tails.  Such formulas are mapped from
+effect formulas in Epik programs. The state formula is imposed as well, since it
+figures in the definition of state constants such as {\tt H1}.
+
+\begin{example} \label{peek1}
+{\small 
+\begin{verbatim}
+define paT2type [[[T2 paT2] T2] & 
+  [[[H1 paT2] H1] | [[T1 paT2] T1]]];
+\end{verbatim} }
+\end{example}
+
+Semantically in Fst, {\tt paT2type} is a set of two strings, which can be displayed
+in an interpreter for the language by reading {\tt paT2type} as a regular expression
+and printing its extension, see \ex{printwords1}.
+
+\begin{example}
+\begin{verbatim}
+xfst[1]: regex paT2type;
+3.0 Kb. 16 states, 16 arcs, 2 paths.
+xfst[2]: print words
+0 1 0 1 paT2 0 1 0 1
+1 0 0 1 paT2 1 0 0 1
+\end{verbatim}
+\end{example}
+
+The finite state calculus has a product operation of string concatenation raised to sets.
+Concatenation of strings with atoms (Boolean vectors) at both ends has the effect of
+doubling atoms at the juncture, and does not enforce matching of atoms at the juncture.  Therefore
+KAT product can not be identified with product in the finite state calculus. Instead,
+KAT product and KAT Kleene star are defined operations, see Figure \ref{fstfigure1}.
+The binary product operation {\tt Cn} and the unary Kleene star operation
+{\tt Kst} combine strings in the string algebra, remove strings with
+non-matching atoms, and then delete the second of two tests to create well-formed
+guarded strings.  Matching of atoms is enforced with the set {\tt Wf0}, which is
+the set of strings that do not contain non-matching Boolean vectors.  The containment
+operator (expressed by the dollar sign) and complement (expressed by tilde) are operators of the finite
+state calculus.  The set of non-matching sequences of atoms {\tt UnequalStPair} is defined by a finite disjunction.
+Deletion is accomplished by a re-write rule in the finite state calculus, which is a
+notation for defining regular relations by contextually constrained substitutions
+(REFS).  In this case, is {\tt Squash} us a regular relation that deletes an atom
+(a sequence of 0's and 1's of a certain length) when it follows an atom.\footnote{
+  This is a non-equal length regular relation. The finite state calculus includes such
+  relations, and they can be used with relation composition and relation domain and co-domain.
+  They are restricted in that the complement and set difference for non-equal length relations is not defined.
+The epistemic alterative relations that are defined in Figure -- are equal-length relations.
+  }
 
@@
 \end{verbatim}
-KAT product in Fst, where {\tt \&} is intersection, {\tt .o.} is relation composition,
-and {\tt .l} is relation image.
+KAT product.
 \end{hangingpar}
@@
 \end{verbatim}
-KAT Kleene plus and Kleene star in Fst. The Fst operation {\tt |} is union.
+KAT Kleene plus and Kleene star. The Fst operation {\tt |} is union.
 \end{hangingpar}
 
-\caption{Translation into Fst of KAT product and KAT Kleene star.}
+\caption{\raggedright Definition in Fst of KAT product and KAT Kleene star.
+  Where {\tt X} and {\tt Y} are regular sets and
+  {\tt R} and {\tt S} are regular relatiions,
+  {\tt X{\&}Y} is the intersection of {\tt X} and {\tt Y},
+  {\tt X|Y} is the union of {\tt X} and {\tt Y},
+  {\tt \~{}X} is the complement of {\tt X},
+  {\tt R{.o.}S} is the composition of {\tt R} and {\tt S},
+  {\tt R{.}l} is the co-domain of {\tt R}, and
+  {\tt {\$}X} is the set of strings that have a substring in $X$.   
+}
 \label{fstfigure1}
@@
       & test in the domain and co-domain. \\
-      & {\tt{Squash.i}} is the inverse \\
-      & of {\tt{Squash}}.
 \end{tabular}
@@
 \begin{hangingpar}
-{\tt{define Kst(R) [St\hspace{1pt}.x.\hspace{1pt}St]\hspace{1pt}|\hspace{1pt}Kpl(X);}}
-The Fst operation {\tt{.x.}} is Cartesian product.
+{\tt{define Kst(R)
+    [St\hspace{1pt}.x.\hspace{1pt}St]\hspace{1pt}|\hspace{1pt}Kpl(X);}} The Fst
+operation {\tt{.x.}} is Cartesian product.  {\tt{R.i}} is the inverse of relation
+{\tt{R}}.
 \end{hangingpar}
@@
 
-An architecture of interpretation by transtlation is employed, where English
-sentences are mapped to terms in the logical language (--) via an interpreted
-grammar, and these terms are in turn interpreted as propositions (sets of possible
-worlds). For the latter, there are options of translation into the finite state
-calculus in order to represent propositions as finite state machines (Section 3), and
-representation in Haskell via lazy lists of guarded strings (Section 4).  The grammar
-is a semantically interpreted multi-modal categorial grammar, consisting of a lexicon
-of words, their categorial types, and interpretations in a logical lambda language.
-Figure \ref{phenomenafigure} lists phenomena that are covered.
+An architecture of interpretation by transtlation is employed, where
+English sentences are mapped to terms in the logical language via an
+interpreted grammar, and these terms are in turn interpreted as
+propositions (sets of possible worlds). For the latter, there are
+options of translation into the finite state calculus in order to
+represent propositions as finite state machines (Section 4), and
+representation in Haskell via lazy lists of guarded strings (Section
+5).  The grammar is a semantically interpreted multi-modal categorial
+grammar, consisting of a lexicon of words, their categorial types, and
+interpretations in a logical lambda language.  Figure
+\ref{phenomenafigure} lists phenomena that are covered.  There is
+recursion through conjuction and verbal complementation, so that the
+language is infinite. It includes talk of beliefs about beliefs, or in
+general, talk of arbitrarily iterated belief.
 
@@
 
-As illustrated towards the end, there is recursion through conjuction and verbal complementation,
-so that the language is infinite, and includes talk of beliefs about beliefs, or in general, talk
-of arbitrarily iterated belief.
-
-Figure \ref{lexiconfigure} gives illustrative lexical entries. The grammar and semantics are in
-certain way optimized for a simple fragment of English concerned with clausal
-complementation.  The agent names {\em Amy} and {\em Bob} contribute the epistemic
-alternative relations for those agents, rather than individuals.  This is possible
-because the agents are never arguments of extensional predicates, so what matters
-about the agents is their epistemic alternative relations.  The root verb {\em know}
-contributes existential modal force.  The complementizers {\em that} and {\em
-  whether} are the heads of their dominating clauses, and assemble an alternative
-relation, modal force, and proposition contributed by the complement.  These
-complementizers introduce the dual via two negations, in order to arrive universal
-modal force.  These moves are offered here as a way of constructing a compact
-interpreted grammar. They can easily be reformulated in a more comprehensive
-interpreted grammar of English.
+Figure \ref{lexiconfigure} gives illustrative lexical
+entries.\footnote{Category symbols use Lambek/Bar-Hillel notation for
+  slashes, so that $(d \backslash t)\slash(d \backslash_D t)$ combines
+  with $d \backslash_D t$ on the right to give a value that combines
+  with $d$ on the left to give $t$.} The grammar and semantics are
+optimized for a simple fragment of English concerned with
+clausal complementation.  The agent names {\em Amy} and {\em Bob}
+contribute the epistemic alternative relations for those agents,
+rather than individuals.  This is possible because the agents are
+never arguments of extensional predicates, so what matters about the
+agents is their epistemic alternative relations.  The root verb {\em
+  know} contributes existential modal force.  The complementizers {\em
+  that} and {\em whether} are the heads of their dominating clauses,
+and assemble an alternative relation, modal force, and proposition
+contributed by the complement.  These complementizers introduce the
+dual via two negations, in order to express universal modal force.
+These moves are offered here as a way of constructing a compact
+interpreted grammar. They can easily be reformulated in a more
+comprehensive interpreted grammar of English.
 
@@
 knows & $(e \backslash t)\slash_M t$ &  $\lambda p.\lambda R. {\Diamond}R p$\\
-that  & \multicolumn{2}{l}{$(((e \backslash t)\slash_M t) \backslash (e \backslash t)) \slash t$} \\
+that  & \multicolumn{2}{l}{$(((e \backslash t){\slash_M}t) \backslash (e \backslash t)) \slash t$} \\
 & \multicolumn{2}{l}{$\lambda p.\lambda m.\lambda R.{\sim}(m ({\sim}p) R)$} \\
-whether  & \multicolumn{2}{l}{$(((e \backslash t)\slash_M t) \backslash (e \backslash t)) \slash t$} \\
+whether  & \multicolumn{2}{l}{$(((e \backslash t){\slash_M} t) \backslash (e \backslash t)) \slash t$} \\
 & \multicolumn{2}{l}{$\lambda p.\lambda m.\lambda R.{\sim}(m ({\sim}p) R)$} \\
@@
 
-Multimodal categories such as $\backslash_D$ and $\backslash_M$ are used to control
-the derivation.  For instance the category of {\em heads} $d \backslash_D t$.  The
-dummy expletive subject {\em it} has category $d$, but the phrase {\em it heads} of
-category $t$ can not be formed, because $\backslash_D$ is not syntactically active as
-a function.  Instead {\em it is heads} can be formed with a predicator {\em is} of
-category $(d \backslash t)\slash(d \backslash_D t)$.  (This uses Lambek/Bar-Hillel
-notation for slashes, so that $(d \backslash t)\slash(d \backslash_D t)$ combines
-with $d \backslash_D t$ on the right to give a value that combines with $d$ on the
-left to give $t$.)  Similarly {\em knows} has a category with the top-level slash
-$\slash_M$, and combines to form a sentence as an argument of {\em that} or {\em
-  whether}, which has a category that looks for the category of {\em know} on the
-left, after combining with a complement sentence on the right.
-
-The semantic translations in the third column of Figure \ref{lexiconfigure} use the Epik term
-language, incremented with lambda.  The body of $\lambda x.0^c{\cdot}h$, which is the
-semantic lexical entry for {\em heads}, is a term denoting the set of all worlds
-where the coin is heads, represented as the set of all guarded strings that end with
-a Boolean valuation where the primitive proposition $h$ (it's heads) is true.  There
-is $\lambda x$ at the front because of a correspondence the grammar formalism uses a
-correspondence between syntactic and semantic types. However, it does not bind
-anything, because sentences such as {\em it isn't heads} have an expletive subject.
-The body ${\Diamond}R p$ of $\lambda p.\lambda R. {\Diamond}R p$, which is the
-semantic lexical entry of {\em knows}, is an Epik term denoting the pre-image of the
-set of worlds $p$ according to the relation $\doublehat{R}$ between guarded strings
-that is determined by the event-level relation $R$.  This is not the right semantics
-for {\em Amy knows that it's heads}, because it is an existential modality ${\Diamond_R}p$,
-rather than an universal modality ${\Box_R}p$.  This is corrected by the complementizer
-{\em that} or {\em whether}, which introduces the dual.
-
-Sentences are parsed with a chart parser for categorial grammar.  The semantics for
-complex phrases are obtained by syntactic application of semantic translations,
-accompanied by beta reduction.  Semantic terms in the parsing formalism are
-expressions of untyped lambda calculus.  The grammar is set up so that lambda is
-eliminated by beta reduction in the semantic term corresponding to a sentence. In
-consequence, the semantic term translating a sentence is a term of the Epik language
-(--).  Such a term designates a set of possible words (guarded strings) in the
-possible worlds model deterimened by an Epik specification such as the one in Figure
-1.  \exi{three}{a} is an English sentence with predicate conjunction and three levels
+Multimodal categories such as $\backslash_D$ and $\backslash_M$ are
+used to control the derivation.  For instance the category of {\em
+  heads} is $d \backslash_D t$.  The dummy expletive subject {\em it}
+has category $d$.  However the phrase {\em it heads} of category $t$
+can not be formed, because $\backslash_D$ is not syntactically active
+as a function.  Instead {\em it is heads} is formed with a predicator
+{\em is} of category $(d \backslash t)\slash(d \backslash_D t)$.
+Similarly {\em knows} has a category with the top-level slash
+$\slash\!_M$, and combines to form a sentence as an argument of {\em
+  that} or {\em whether}. These complementizers have a category that
+looks for the category of {\em know} on the left, after combining with
+a complement sentence on the right.
+
+The semantic translations in the third column of Figure
+\ref{lexiconfigure} use the Epik term language, incremented with
+lambda.  The body of $\lambda x.0^c{\cdot}h$, which is the semantic
+lexical entry for {\em heads}, is a term denoting the set of all
+worlds where the coin is heads, expressed as the set of all guarded
+strings that end with a Boolean valuation where the primitive
+proposition $h$ (it's heads) is true.  There is $\lambda x$ at the
+front because the grammar formalism enforces a strict correspondence
+between syntactic and semantic types. However, the lambda does not
+bind anything, because sentences such as {\em it isn't heads} have an
+expletive subject.  The body ${\Diamond}R p$ of $\lambda p.\lambda
+R. {\Diamond}R p$, which is the semantic lexical entry of {\em knows},
+is an Epik term denoting the pre-image of the world-alternative
+relation contributed by the subject.  This is not the right semantics
+for {\em Amy knows that it's heads}, because it is an existential
+modality ${\Diamond_R}p$, rather than an universal modality
+${\Box_R}p$.  This is corrected by the complementizer {\em that} or
+{\em whether}, which introduces the dual.
+
+Sentences are parsed with a chart parser for categorial grammar.  The
+semantics for complex phrases are obtained by syntactic application of
+semantic translations, accompanied by beta reduction.  Semantic terms
+in the parsing formalism are expressions of untyped lambda calculus.
+The grammar is set up so that lambda is eliminated by beta reduction
+in the semantic term corresponding to a sentence. In consequence, the
+semantic term translating a sentence is a term of the Epik term
+language.  Such a term designates a set of possible words (guarded
+strings) in the possible worlds model deterimened by an Epik
+specification.
+By way of example, 
+\exi{three}{a} is an English sentence with predicate conjunction and three levels
 of clausal embedding.  Using the grammar and parser, the sentence is mapped to the
-Epik term \exi{b}.  Using the result from Section 3, this term can be mapped in an
+Epik term \exi{b}, which here takes the form of a Scheme s-expression.
+
+Using the result from Section 3, this term can be mapped in an
 implementation of the finite state calculus to a finite state machine that represents
@@
 
-\begin{examples}  \label{three}
-\itema  Amy knows that Bob knows that Amy knows whether it is heads and
-knows that Bob does not know that Amy knows that it is tails.
-\itemb
+\begin{example}  \label{three}
+Amy knows that Bob knows that Amy knows whether it is heads and
+knows that Bob doesnt know that Amy knows that it is tails.
+\end{example}
+
+\section{Examples and discussion}
+
+\begin{examples}
+\itema Amy knows that it is tails
+\itemb {\tt (Not (Diamond amy (Not tails)))}
 \end{examples}
 
-\section{Examples and discussion}
+\begin{examples}
+\itema Bob doesnt know that Amy knows that it is tails
+\itemb {\tt (Not (Not (Diamond bob (Not (Not (Diamond amy (Not tails)))))))}
+\end{examples}
+
+\begin{examples}
+\itema Bob knows that Amy knows whether it is heads.
+\itemb {\tt (Not (Diamond bob (Not (Or (Not (Diamond amy (Not heads))) (Not (Diamond amy heads))))))}
+\end{examples}
+
+\begin{examples}
+\itema Amy knows that Bob knows that Amy knows whether it is heads and knows that Bob doesnt know that Amy knows that it is tails.
+\itemb {\tt (And (Not (Diamond amy (Not (Not (Diamond bob (Not (Or (Not (Diamond amy (Not heads))) (Not (Diamond amy heads))))))))) (Not (Diamond amy (Not (Not (Not (Diamond bob (Not (Not (Diamond amy (Not tails))))))))))) }
+\end{examples}
 

### 2020-09-19 tex/sdraft.tex
Submission dradt sdraft.

--- 2020-09-18 tex/draft.tex
@@
 alternatively, one can work with small idealized models, in order to
-illustrate and test ideas.  To build such models, spaces of worlds and
-individuals are stipulated as small finite sets, and semantic values
-of lexical items are constructed as functions or relations from these
-sets.  Such toy or idealized models are useful in research and in
-teaching, in that it is possible to represent propositions finitely
-and explicitly, and to calculate with them. The point of this paper is
-to scale up toy or idealized models to countable sets of worlds, and
-to constructive and computable modeling of epistemic alternatives for
-agents.  We describe a certain systematic way of defining such models,
-and illustrate how to apply them in natural language semantics.  The
-focus is on epistemic semantics and clausal embedding. The fundamental
-move is to identify possible worlds with strings of primitive events,
-so that propositions are sets of strings. An advantage in this is that
-it allows for a mathematical description of an algebra of
-propositions, coupled with a computational representation using either
-lazy lists of strings, or finite state machines that describe sets of
-strings.
+illustrate and test ideas.  The point of this paper is to scale up toy
+or idealized models to countable sets of worlds, and to constructive
+and computable modeling of epistemic alternatives for agents.  We
+describe a certain systematic way of defining such models, and
+illustrate how to apply them in natural language semantics.  The focus
+is on epistemic semantics and clausal embedding. The fundamental move
+is to identify possible worlds with strings of primitive events, so
+that propositions are sets of strings. An advantage in this is that it
+allows for a mathematical description of an algebra of propositions,
+coupled with a computational representation using either lazy lists of
+strings, or finite state machines that describe sets of strings.
 
@@
 
-We begin with examples of event-sequence models.  {\em The
-  Elevator}. An elevator moves up and down in a four-story building,
-with floors numbered in the European fashion as 0,1,2,3.  There are
-primitive events $u$ (the elevator going up one floor), and $d$ (the
-elevator going down on floor).  In worlds $v_1$ and $v_2$, the events
-shown in \ex{elevator1} transpire. The truth values for English
-sentences shown in \ex{elevator2} are observed.
-
-\begin{example}  \label{elevator1}
-\hspace{-4mm} $v_1$
-\begin{tabular}[t]{ll}
-  $u$ & it goes up from 0 to 1 \\
-  $u$ & it goes up from 1 to 2 \\
-  $d$ & it goes down from 2 to 1 \\
-  $u$ & it goes up from 1 to 2 \\
-\end{tabular}
-
-\vspace{2mm}
-\hspace{-4mm} $v_2$
-\begin{tabular}[t]{ll}
-  $u$ & it goes up from 0 to 1 \\
-  $u$ & it goes up from 1 to 2 \\
-  $u$ & it goes up from 2 to 3 \\
-\end{tabular}
-\end{example}
-
-\begin{example}  \label{elevator2}
-\hspace{-4mm} \begin{tabular}[t]{ccl}
-  $v_1$ & $v_2$ & Sentence \\
-   false & true & It's on floor 3. \\
-   true & true & It has gone up. \\
-   true & false & It has gone down. \\
-   true & false & It could go up. \\
-\end{tabular}
-\end{example}
-
-{\em The Concealed Coin}.  Amy and Bob are seated at a table.  There is a coin on the
+A running example of an event-sequence model is {\em The Concealed
+  Coin}.  Amy and Bob are seated at a table.  There is a coin on the
 table under a cup, heads up.  The coin could be heads or tails, and
-neither agent knows which it is.  This initial situation is possible world $w_1$.
-Two additional worlds $w_2$ and $w_2$ are
-defined by sequencing events after the initial state, with events interpreted as in
-\ex{coin1}. The truth values for English
-sentences shown in \ex{coin3} are observed.
+neither agent knows which it is.  This initial situation is possible
+world $w_1$.  Two additional worlds $w_2$ and $w_2$ are defined by
+sequencing events after the initial state, with events interpreted as
+in \ex{coin1}. The truth values for English sentences shown in
+\ex{coin3} are observed.
 
@@
 
-The events in the examples come with pre-conditions. The elevator can
-not go up if it is already on floor 3, so $u$ has the pre-condition of
-the elevator being of floor 0, 1, or 2.  Similarly $d$ has the
-precondition that the elevator is on floor 1, 2 or 3.  Amy can peek at
-heads only if the coin is heads up, so $a_1$ has the precondition of
+The events come with pre-conditions.
+Amy can peek at heads only if the coin is heads up, so $a_1$ has the precondition of
 the coin being heads up.  Let $h$ be the Boolean proposition that the
-coin is heads up.  In the other example, let $q$ be the proposition
-that the elevator is on a high floor (2 or 3), and $p$ be the
-proposition that it is on an odd floor (1 or 3). Then preconditions
+coin is heads up.
+Then preconditions
 can be described by Boolean formulas, with $h$ being the precondition
-of $a_1$, and $\overline{pq}$ being the precondition of
-$u$. Juxtaposition is used for Boolean conjunction, and the overbar
-for Boolean negation. Events come as well with a relation between
+of $a_1$
+$u$. Events come as well with a relation between
 prior and following state, for instance with $u$ incrementing the
@@
 looking at tails) can happen only in a not-$h$ state, and does not
-change the state.  The third line says that $u$ increments the floor,
-and can happen only on floors 0, 1, and 2.  The fourth line describes
-$d$ in similar terms.  Plus is disjunction.
+change the state.
 
@@
 % u & (\bar{q}\bar{p}) \! : \!(\bar{q}p) + \\
-u & \ev{(\bar{q}\bar{p})}{(\bar{q}p)} + \ev{(\bar{q}p)}{(q\bar{p})} + \ev{(q\bar{p})}{(qp)} \\
-d & \evr{(\bar{q}\bar{p})}{(\bar{q}p)} + \evr{(\bar{q}p)}{(q\bar{p})} + \evr{(q\bar{p})}{(qp)} \\ 
+% u & \ev{(\bar{q}\bar{p})}{(\bar{q}p)} + \ev{(\bar{q}p)}{(q\bar{p})} + \ev{(q\bar{p})}{(qp)} \\
+% d & \evr{(\bar{q}\bar{p})}{(\bar{q}p)} + \evr{(\bar{q}p)}{(q\bar{p})} + \evr{(q\bar{p})}{(qp)} \\ 
 \end{array} \)
@@
 
-In the discussion at the end of the last section, a sequence such as
-$\bar{q}{p}$ can be viewed as a valuation of primitive test
-propositions, which is used to describe world state. The primitives
-are listed in fixed order, and left unmarked (indicating true) or
-marked with the overbar (indicating false).  Suppose that in the coin
-example, we have an additional primitive stative proposition (or
-atomic test) $t$, interpreted as tails.  Since a coin is heads or
-tails but not both, we want to allow the valuations $h\bar{t}$ and
-$\bar{h}t$, and disallow $ht$ and $\bar{h}\bar{t}$.  This is enforced
-by a {\em state formula}, which is a Boolean formula, in this case the
-one given on the second line of \ex{effect}.  Where {\sf{B}} is a set
-of atomic tests and $\phi$ is a state constraint over {\sf{B}},
-${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\phi}$ is the set of valuations of
-{\sf{B}} that make formula $\phi$ true. Valuations are called atoms,
-because they correspond to the atoms of a Boolean algebra of tests
-\citep{kozen2001automata}.
+Suppose that in the coin example, we have an additional primitive
+stative proposition (or atomic test) $t$, interpreted as tails.  A
+sequence such as $\bar{h}{t}$ can be viewed as a valuation of
+primitive propositions, which is used to describe world state. The
+primitives are listed in fixed order, and left unmarked (indicating
+true) or marked with the overbar (indicating false).  Since a coin is
+heads or tails but not both, we want to allow the valuations
+$h\bar{t}$ and $\bar{h}t$, and disallow $ht$ and $\bar{h}\bar{t}$.
+This is enforced by a {\em state formula}, which is a Boolean formula,
+in this case the one given on the second line of \ex{effect}.  Where
+{\sf{B}} is a set of atomic tests and $\phi$ is a state constraint
+over {\sf{B}}, ${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\phi}$ is the set of
+valuations of {\sf{B}} that make formula $\phi$ true. Valuations are
+called atoms, because they correspond to the atoms of a Boolean
+algebra of tests \citep{kozen2001automata}.
 
@@
 {\textit{World}} & {\textit{Guarded string}} & \textit{Length}\\ \hline
-$v_1$ & $\hat 0 \,u\, \hat 1 \,u\, \hat 2 \,d\, \hat 1 \,u\, \hat 2$ & $4$ \\
-$v_2$ & $\hat 0 \,u\, \hat 1 \,u\, \hat 2 \,u\, \hat 3$ & $3$ \\   
+%$v_1$ & $\hat 0 \,u\, \hat 1 \,u\, \hat 2 \,d\, \hat 1 \,u\, \hat 2$ & $4$ \\
+%$v_2$ & $\hat 0 \,u\, \hat 1 \,u\, \hat 2 \,u\, \hat 3$ & $3$ \\   
 $w_1$ & $H$ & $0$\\
@@
 \begin{example} \label{fusion1}
-\hspace{-2mm}\(\hat 0 \,u\, \hat 1 \,u\, \hat 2 \,d\, \hat 1 \; \diamond \; \hat 1 \,u\, \hat 2 = \hat 0 \,u\, \hat 1 \,u\, \hat 2 \,d\, \hat 1 \,u\, \hat 2 \)
-
-\hspace{-2mm}\( \hat 0 \,u\, \hat 1 \,u\, \hat 2 \,d\, \hat 1 \; \diamond \; \hat 2 \,u\, \hat 3 = \textit{undefined} \)
+%\hspace{-2mm}\(\hat 0 \,u\, \hat 1 \,u\, \hat 2 \,d\, \hat 1 \; \diamond \; \hat 1 \,u\, \hat 2 = \hat 0 \,u\, \hat 1 \,u\, \hat 2 \,d\, \hat 1 \,u\, \hat 2 \)
+
+%\hspace{-2mm}\( \hat 0 \,u\, \hat 1 \,u\, \hat 2 \,d\, \hat 1 \; \diamond \; \hat 2 \,u\, \hat 3 = \textit{undefined} \)
 
@@
 \hspace{2mm}\(\sem{ \Diamond_a\,p } \hspace{4mm}{\triangleq} \hspace{4mm}\{x \mid \exists y . x{\hat{R}_a}y \wedge y \in \sem{ p }^{\epikspec}\} \)
-  \caption{Interpretation of Epik terms as sets of guarded strings (Version M)}
+  \caption{Interpretation of Epik terms as sets of guarded strings}
   \label{fig:interp2}
-\end{figure}
-
-\begin{figure}
-  \[\begin{array}{l >{\triangleq}c l}
-  \sem{ 0 }^{\varphi,{\mathcal{E}}} && \{\} \\  
-  \sem{ 1 }^{\varphi,{\mathcal{E}}} && \mathcal A^\varphi_{\mathsf B}\\
-  \sem{ e }^{\varphi,{\mathcal{E}}}&& \hat{\mathcal E}^\varphi(e) \\
-  \sem{ \psi }^{\varphi,{\mathcal{E}}} &&  \mathcal A^{\varphi~\psi}\\
-  \sem{{p+q}}^{\varphi,{\mathcal{E}}} && \sem{{p}}^{\varphi,{\mathcal{E}}} \cup \sem{ q }^{\varphi,{\mathcal{E}}}\\
-  \sem{{p;q}}^{\varphi,{\mathcal{E}}} && \sem{ p }^{\varphi,{\mathcal{E}}} \diamond \sem{{q}}^{\varphi,{\mathcal{E}}}\\
-  \sem{ p^*}^{\varphi,{\mathcal{E}}} && \displaystyle\bigcup_n {\underbrace{\sem{ p }^{\varphi,{\mathcal{E}}}\diamond \ldots \diamond \sem{ p }^{\varphi,{\mathcal{E}}}}_{n}} \\
-  \sem{ \neg p }^{\varphi,{\mathcal{E}}} && \sem{ \_^*}^{\varphi,{\mathcal{E}}} \setminus \sem{ p }^{\varphi,{\mathcal{E}}} \\
-  \sem{ \Diamond_a~p } && \{x \mid x \mathbin{\hat{R_a}} y, y \in \sem{ p }^{\varphi,\mathcal E}\}
-
-  \end{array}
-  \]
-  \caption{Interpretation of Epik terms as sets of guarded strings}
-  \label{fig:interp}
 \end{figure}
@@
 \end{figure}
-
-Other operations in the guarded string algebra as defined in FST are simpler.
-Union is union in the string algebra. The complement of a proposition
-is complement relative to the set of worlds, as defined by
-set difference in the string algebra.
-
-This scheme provides for an interpretation of the language of Epik
-terms that is defined in Figure \ref{termfigure}.  An Epik specifications such as
-Figure 1 is translated to a straight-line program of the finite state
-calculus that defines constants and functions, including the ones from
-Figures 2 and 3.  In Xfst or Foma, which are interpreters for the
-finite state calculus, the program is read, and then propositional
-terms can be mapped to finite state machines that represent sets of
-guarded strings, interpreted as propositions.
 
@@
 \end{examples}
-
-Page breakdown
-
-\begin{tabular}{lll}
-  1,2 & 3.75 \\
-  3   & 1.25 & FST translation\\
-  4   & 1.25 & Haskell Epik \\
-  5   & 0.75 & English CG fragment \\
-  6   & 1.0  & Examples and discussion\\
-\end{tabular}
-
-%% \begin{example}   \label{Rhat}
-%% \hspace{-2mm} \(
-%% \hat{R} \triangleq
-%% \{\tuple{u,v} \mid \exists c \exists d. c\mathrel R d \wedge u \in \hat{c}  \wedge v \in \hat{d} \, \} \)
-%% \end{example}
-
-{\em This stuff was at the end of Section 2, it became redundant. Also it refers to the interpretation function, which is introduced in Section 3.}
-
-\begin{example}   \label{Rhathat}
-\hspace{-2mm}\(
-\hat{R} \triangleq
-\hspace{-2mm}\begin{array}[t]{l}
-\{ \tuple{x,y} \mid \exists u_1 ... \exists u_n \exists v_1 ... \exists v_n. \\
-\hspace{6mm} u_1\mathrel{R}v_1 \wedge ... \wedge u_n\mathrel{R}v_n \wedge\\
-\hspace{6mm} x \in \sem{ u_1 ; ... ; u_n }^{\varphi,{\mathcal{E}}}\hspace{1mm} \wedge \\
-\hspace{6mm} y \in \sem{ v_1 ; ... ; v_n }^{\varphi,{\mathcal{E}}} \hspace{2mm} \}
-\end{array}
-\)
-\end{example}
-
-This defines an epistemic alternative to world $x$ to be a world of
-the same length, where each component event in the alternative is an
-event-alternative to the event in corresponding position in the base
-world. Fusion product enforces preconditions and a correspondence
-between pre-states and post-states of events on both sides of the
-epistemic alternative relation.  This provides for finitely
-specifiable construction of epistemic models that reflect intuitions
-about information exchange and epistemic consequences of perceptual
-events.  See Section 6 for linguistic examples.  Since an epistemic
-alternative has the same lenght as its base world, it follows from the
-construction that agents know how many events have transpired in their
-base worlds.
-
-Fortunately, Epik provides a syntax to specify the input-output
-behavior via the \texttt{event} declarations. From these we construct
-a function $\mathcal E$ from agents to relations on atoms that specify
-the effects each action can have. Notably for every declaration
-$\texttt{event}~e~\zeta$, we have $\mathcal E^\varphi(e) =
-\sem{\zeta}^\varphi$. We can lift $\mathcal E^\varphi$
-to $\hat{\mathcal E}$ which decorates the events with their permitted
-atoms, that is $\hat{\mathcal E}^\varphi(e) = \{\alpha~e~\beta \mid
-\alpha, \beta \in \mathcal E^\varphi (e) \}$. See the examples in
-\ex{events1}.
-% u (\bar{q}\bar{p}):(\bar{q}p) + (\bar{q}p):(q\bar{p}) + (q\bar{p}):(qp) 173
-% d (\bar{q}p):(\bar{q}\bar{p}) + (q\bar{p}):(\bar{q}p) + (qp):(q\bar{p})
 

### 2020-09-19 tex/sdraft.tex
Cuts in sdraft

@@
 \usepackage{latexsym}
+%\usepackage{caption}
+\renewcommand{\UrlFont}{\ttfamily\small}
+
+%\captionsetup{belowskip=0pt}
+
+%\setlength{\abovecaptionskip}{-10pt}
+\setlength{\belowcaptionskip}{-15pt}
+
+\setlength{\abovecaptionskip}{5pt plus 3pt minus 2pt}
+
+% This is not strictly necessary, and may be commented out,
+% but it will improve the layout of the manuscript,
+% and will typically save some space.
+\usepackage{microtype}
+
+%\aclfinalcopy % Uncomment this line for the final submission
+%\def\aclpaperid{***} %  Enter the acl Paper ID here
+
+%\setlength\titlebox{5cm}
+% You can expand the titlebox if you need extra space
+% to show all the authors. Please do not make the titlebox
+% smaller than 5cm (the original size); we will check this
+% in the camera-ready version and ask you to change it back.
+
+\newcommand\BibTeX{B\textsc{ib}\TeX}
+
+\title{Epistemic Semantics in Guarded String Models}
+
+%\author{Eric Campbell \\
+%  Cornell University\\
+%  \texttt{ehc86@cornell.edu} \\\And
+%  Mats Rooth \\
+%  Cornell University\\
+%  \texttt{mr249@cornell.edu} \\}
+
+\date{}
+
 \usepackage{amssymb,amsmath,accents}
@@
 \newcommand{\test}{\sf{B}}
-
-\pagestyle{plain}
-
-% This is not strictly necessary, and may be commented out,
-% but it will improve the layout of the manuscript,
-% and will typically save some space.
-
-% Uncomment this line for the final submission
-% Or comment it for the submission version with line numbers.
-\aclfinalcopy 
-%\def\aclpaperid{***} %  Enter the acl Paper ID here
-% Make the references come out black. TAKE THIS OUT for submission.
-\definecolor{darkblue}{rgb}{0, 0, 0}
-% CHECK THE BIB STYLE FOR THE STARSEM SUBMISSION
-
-%\setlength\titlebox{5cm}
-% You can expand the titlebox if you need extra space
-% to show all the authors. Please do not make the titlebox
-% smaller than 5cm (the original size); we will check this
-% in the camera-ready version and ask you to change it back.
-
-\newcommand\BibTeX{B\textsc{ib}\TeX}
-
-% This used to say Constructive ..., which is a better title.
-% Reserve that for a journal paper if we get to that.
-\title{Epistemic Semantics in Guarded String Models}
-
-\author{Eric Campbell \\
-  Cornell University\\
-  \texttt{ehc86@cornell.edu} \\\And
-  Mats Rooth \\
-  Cornell University\\
-  \texttt{mr249@cornell.edu} \\}
-
-\date{}
 
@@
   into the finite state calculus, and alternatively by modeling
-  propositions as lazy lists in Haskell. The syntax-semantics interface for
-  a fragment of English is defined by a categorial grammar.
+  propositions as lazy lists in Haskell. The syntax-semantics
+  interface for a fragment of English is defined by a categorial
+  grammar.
 \end{abstract}
@@
 elements corresponding to propositions and event types in our
-application \citep{kozen2001automata}. The models we propose are an
-epistemic extension of guarded string models for KAT, where a unary
-operation interpreted as an existential epistemic modality is included
-for each agent.  {\em Action models} in dynamic epistemic semantics
-introduce the technique of constructing epistemic models from
-primitive alternative relations on events, in order to capture the
-epistemic consequences of perceptual and communicative events
-\citep{baltag1999logic}. This is the basis for our construction of
-epistemic alternative relations.  Literature on {\em finite state
+application \citep{kozen2001automata}.  {\em Action models} in dynamic
+epistemic semantics introduce the technique of constructing epistemic
+models from primitive alternative relations on events, in order to
+capture the epistemic consequences of perceptual and communicative
+events \citep{baltag1999logic}.  Literature on {\em finite state
   methods in linguistic semantics} has used event strings and sets of
@@
   finite state intensiononal semantics} has investigated how to do the
-semantics of intensional complementation, including indirect
-questions, in a setting where compositional semantics is expressed in
-a finite state calculus \citep{rooth2017finite,collard2018finite}.  We
-adopt this in our syntax-semantics interface for English.
+semantics of intensional complementation in a setting where
+compositional semantics is expressed in a finite state calculus
+\citep{rooth2017finite,collard2018finite}.
 
@@
 \begin{example} \label{coin2}
-\hspace{-4mm} \begin{tabular}[t]{ll}
-   $w_1$ \\
-   $w_2 = w_1 \hspace{2pt} a_1 $ \\
-   $w_2 = w_1 \hspace{2pt} a_1 \hspace{2pt} b_1 $\\ 
- \end{tabular}
+   $w_2 = w_1 \hspace{2pt} a_1 $ \hspace{2mm}
+   $w_2 = w_1 \hspace{2pt} a_1 \hspace{2pt} b_1 $ 
 \end{example}
@@
 \end{example}
+
+\newcommand{\ev}[2]{#1 \!\! : \!\! #2}
+\newcommand{\evr}[2]{#2 \!\! : \!\! #1}
 
@@
 floor.  This is expressed using an operator ``$:$'' (read ``and
-next'') that pairs Boolean formulas. The first line in \ex{event1}
+next'') that pairs Boolean formulas. The formula $h:h$
 describes $a_1$ (Amy looking at heads) as happening only in an $h$
@@
 looking at tails) can happen only in a not-$h$ state, and does not
-change the state.
-
-\newcommand{\ev}[2]{#1 \!\! : \!\! #2}
-\newcommand{\evr}[2]{#2 \!\! : \!\! #1}
-
-\renewcommand{\arraystretch}{1.0}
-
-\begin{example}   \label{event1}
-\hspace{-5mm}\( \begin{array}[t]{cl}
-a_1 & h:h \\
-a_0 & \ev{\bar{h}}{\bar{h}} \\
-% u & (\bar{q}\bar{p}) \! : \!(\bar{q}p) + \\
-% u & \ev{(\bar{q}\bar{p})}{(\bar{q}p)} + \ev{(\bar{q}p)}{(q\bar{p})} + \ev{(q\bar{p})}{(qp)} \\
-% d & \evr{(\bar{q}\bar{p})}{(\bar{q}p)} + \evr{(\bar{q}p)}{(q\bar{p})} + \evr{(q\bar{p})}{(qp)} \\ 
-\end{array} \)
-\end{example}
+change the state ($\ev{\bar{h}}{\bar{h}}$).
+
+%\renewcommand{\arraystretch}{1.0}
 
@@
 regular sets of strings. This is deferred to the next section.
-
-In the natural language application, sets of well-formed guarded strings 
+Certai sets of well-formed guarded strings 
 have the addional interpretation of event types.
 An event-type is something that can ``happen'' is different worlds.
-The event of the elevator going up is modeled not as the bare event
-symbol $u$ or its unit set $\{u\}$, but as the set of guarded strings
-\(\{ \hat 0 \,u\, \hat 1, \hat 1 \,u\, \hat 2,  \hat 2 \,u\, \hat 3 \}\).
-The event of the elevator going up in a given world $w$ corresponds
-to $w$ being incremented to form $w \diamond e$, where $e$ is an element
-of the event type.  \ex{events1} gives the event types in
-the examples.
+\ex{events1} gives the event types in
+the example.
 
@@
 \(
-\begin{array}[t]{l|c}
-   &  {\mbox{Event type}}\\ \hline
-  u & \{ \hat 0 \,u\, \hat 1, \hat 1 \,u\, \hat 2,  \hat 2 \,u\, \hat 3\}\\
-  d & \{ \hat 1 \,d\, \hat 0, \hat 2\,d\,\hat 1, \hat 3 \,d\, \hat 2\}\\
-  a_1 & \{H \,a_1\, H\}\\
-  a_0 & \{T \,a_0\, T\}\\
-  b_1 & \{H \,b_1\, H \}\\
-  b_0 & \{T \,b_0\, T \}
+\begin{array}[t]{l|cl|c}
+  &  {\mbox{Event type}} && {\mbox{Event type}} \\ \hline
+  a_1 & \{H \,a_1\, H\} &   b_1 & \{H \,b_1\, H \} \\
+  a_0 & \{T \,a_0\, T\} &   b_0 & \{T \,b_0\, T \} \\
 \end{array}
@@
 The construction so far defines a set of worlds from an Epik specification.  Normally
-and in our examples, $W$ is countably infinite, though some choices of effect
+the set is countably infinite, though some choices of effect
 formulas can result in a finite set of worlds.  The next step is to define an
@@
 an alternative relation on bare events for each agent $a$, which we notate as
-${R}_{a}$.  This should be lifted to a relation $\hat{R}_a$ on worlds.  The basic
+$\hat{R}_{a}$.  This should be lifted to a relation ${R}_a$ on worlds.  The basic
 idea is that when a world $w$ is incremented with an event $e$, in the resulting
@@
   than guarded string models. See \citet{baltag1999logic}, \citet{van2007dynamic}, and
-  articles in \citet{ditmarsch2015handbook}.  We picked up the idea from papers by
-  Moss and his colleagues, together with \citet{moore1984formal} and subsequent
-  literature in situation theory.  This previous literature is motivated by epistemic
-  logic and AI planning, rather than natural language semantics.  } This needs to be
+  articles in \citet{ditmarsch2015handbook}.  Previous literature is motivated by epistemic
+  logic and AI planning, rather computable possible worlds models in natural language semantics.
+} This needs to be
 implemented in a way that takes account of pre- and post-conditions for events. For
@@
 \begin{example}  \label{raise1}
-  Let $W$ be a set of guarded strings over  events P and primitive tests B, and ${R}$ be a relation
-  on $P$. The corresponding relation $\hat{R}$ on $W$ holds between a guarded string
-\[
+  Let $W$ be a set of guarded strings over events P and primitive
+  tests B, and $\hat{R}$ be a relation on $P$. The corresponding
+  relation ${R}$ on $W$ holds between a guarded string
+\(
 \alpha_0e_0 ... e_n\alpha_{n+1} 
-\]
+\)
 in $W$ and a guarded string  $q$ iff $q$ is an element of $W$ and is of the form
-
-\[
+\(
 \alpha_0'e_0' ... e_n'\alpha_{n+1}', 
-\]
-
-where for $0 \leq n$, $\element{\tuple{e_i,e'_i}}{\hat{R}}$.
+\), where for $0 \leq n$, $\element{\tuple{e_i,e'_i}}{\hat{R}}$.
 \end{example}
 
-This requires that in the alternative world $q$, each constituent event $e'_i$ is an
-alternative to the event $e_i$ in the base world according to $\hat{R}$.
+This requires that in an alternative world, each constituent event $e'_i$ is an
+alternative to the event $e_i$ in the base world.
 Compatibilities between events in the alternative world are enforced by the
-requirement that the alternative $q$ is an element of $W$, so that state and effect
+requirement that the alternative world is an element of $W$, so that state and effect
 formulas are enforced.
@@
 
+\begin{figure}
+\begin{tabular}{ll}
+events & $e \in \mathsf{P}$ \\
+states & $\sigma$ as in Figure 1 \\    
+$p,q$ & $::= e~|~\sigma~|~p+q~|~pq~|~p^*~|~ \neg p ~|~ \Diamond_a\,p$\\
+\end{tabular}
+
+\vspace{1mm}
+\begin{tabular}{cl}
+$\Box_a p$ & $\triangleq \neg\Diamond_a\neg p$ \\
+\(p \wedge q\) &  \(\triangleq \neg(\neg p + \neg q)\) \\
+$\bullet$ & \( \triangleq \sum_{e \in \mathsf P}e\)
+\end{tabular}
+
+\caption{The language of Epik terms and key derived
+  operators.}
+\label{fig:epik-grammar}
+\label{termfigure}
+\end{figure}
+
 \section{The logical language of Epistemic KAT}
-In addition to a set of possible worlds and an algebra of propositions, the
-application in natural language requires a logical language for naming propositions.
 The standard language for Kleene algebra with tests has the signature
@@
   operations on tests.  $1$ has the dual role of the identity for product (fusion
-  product raised to sets), and True for tests.  }. To this signature we add a unary
+  product raised to sets), and True for tests.  } To this we add a unary
 modal operation $\Diamond_a$ for each agent, and a unary complement operation $\neg$
@@
   ... \Diamond_n}$. Figure \ref{termfigure} defines the syntax of the language.
-Juxtaposition is used in place of the product symbol $\cdot$.  Terms in this language
+Juxtaposition is used for product.  Terms in this language
 are used to represent the proposositional semantic values of English sentences.
@@
 
-\begin{figure}
-\begin{tabular}{ll}
-events & $e \in \mathsf{P}$ \\
-states & $\sigma$ as in Figure 1 \\    
-$p,q$ & $::= e~|~\sigma~|~p+q~|~pq~|~p^*~|~ \neg p ~|~ \Diamond_a\,p$\\
-\end{tabular}
-
-\vspace{1mm}
-\begin{tabular}{cl}
-$\Box_a p$ & $\triangleq \neg\Diamond_a\neg p$ \\
-\(p \wedge q\) &  \(\triangleq \neg(\neg p + \neg q)\) \\
-$\bullet$ & \( \triangleq \sum_{e \in \mathsf P}e\)
-\end{tabular}
-
-\caption{The language of Epik terms and key derived
-  operators.}
-\label{fig:epik-grammar}
-\label{termfigure}
-\end{figure}
-
 \newcommand{\epikspec}{\test,\event,\varphi,\zeta}
@@
 
-Each term in the language is interpreted as a set of guarded strings.  Where $p$ is a
-term, $\sem{p}^{\epikspec}$ is the interpretation of the term, where superscript
-captures dependence on an Epik specification.  Figure \ref{fig:interp2} defines the
-interpretation of terms recursively.  The interpretation $\sem{ 1 }^{\epikspec}$ of
-the multiplicative identity is the set of atoms that satisfy the state constraint
-$\varphi$.  Where $b$ is a primitive Boolean, $\sem{ b }^{\epikspec}$ is the set of
-atoms that satisfy the state constraint and where $b$ is true.  Where $e$ is a
-primitive event, $\sem{ e }^{\epikspec}$ is the set of guarded strings that have the
-form of $e$ flanked by compatible atoms, as determined by the event formula
-$\zeta_e$. The product $pq$ of terms $p$ and $q$ is interpreted with fusion product
-raised to sets of guarded strings.
-Kleene star is interpreted as the union of iterated products. $p^n$ is the $n$-times product
-of $p$ with itself, with $p^0 = 1$.
+A term $p$ of the logical language is interpreted as a set of guarded
+strings $\sem{p}^{\epikspec}$, where superscript captures dependence
+on an Epik specification.  Figure \ref{fig:interp2} defines the
+interpretation.  The interpretation $\sem{ 1
+}^{\epikspec}$ of the multiplicative identity 1 is the set of atoms that
+satisfy the state constraint $\varphi$.  Where $b$ is a primitive
+Boolean, $\sem{ b }^{\epikspec}$ is the set of atoms that satisfy the
+state constraint and where $b$ is true.  Where $e$ is a primitive
+event, $\sem{ e }^{\epikspec}$ is the set of guarded strings that have
+the form of $e$ flanked by compatible atoms, as determined by the
+event formula $\zeta_e$. The product $pq$ is
+interpreted with fusion product raised to sets of guarded strings.
+Kleene star is interpreted as the union of exponents ($p^n$ is
+the $n$-times product of $p$ with itself, with $p^0 = 1$).
 Propositional complement is complement relative to the set of worlds.
 The epistemic formula $\Diamond_a p$ is interpreted with Kripke
-semantics for epistemic modality, as the pre-image of the
-complement proposition under the relation
-$\hat{R}_a$.  Here $\hat{R}_a$ is the world-alternative for agent $a$,
-as defined in \ex{raise1}.
+semantics for epistemic modality, as the pre-image of the embedded
+proposition under the world-alternative relation ${R}_a$.
 
@@
 
-Summing up, given an Epik specification, a state constraint $\varphi$, and effect
-relation $\zeta_e$ for each primitive are constructed. Given these, each term $p$ (as
-defined syntactically in Figure \ref{termfigure}) is interpreted as a set of guarded
-strings $\sem{p}^{\epikspec}$ over the Boolean primitives and event primitives in the
-specification.  Let $K^{\epikspec}$ be the sets that are interpretations of terms.
-Then
+Summing up, given an Epik specification $\epikspec$, term $p$ (as
+defined syntactically in Figure \ref{termfigure}) is interpreted as a
+set of guarded strings $\sem{p}^{\epikspec}$.  Let $K^{\epikspec}$ be
+the sets that are interpretations of terms.  Then
 $\tuple{K^{\epikspec},+,\cdot,*,\bar{},0,1,\neg,\Diamond_{a_1},...,\Diamond_{a_n}}$
-is a concrete guarded string interpretation for the signature of epistemic KAT, with
-operations as in Figure \ref{fig:interp2} (e.g. the binary operation $+$ is union,
-and the unary operation $\Diamond_a$ is pre-image relative to $\hat{R}_a$).  The
-construction can also be viewed as providing a concrete $n$-agent Kripke frame
-$\tuple{\sem{\bullet *}^{\epikspec}, \hat{R}_1,...,\hat{R}_n}$.\footnote{ The domain
-  of the Kripke frame differs from the domain of the guarded string model, because
-  the former is the set of worlds, while the latter is the set of propositions.  }
-The frame consists of a set of worlds, and an epistemic-alternative relation for each
-agent.  These models are used as a target for natural-language interpretation in
-Section 5 and Section 6.
+is a concrete guarded string interpretation for the signature of
+epistemic KAT, with operations as in Figure \ref{fig:interp2}
+(e.g. the binary operation $+$ is union, and the unary operation
+$\Diamond_a$ is pre-image relative to $\hat{R}_a$).  This provides a
+concrete $n$-agent Kripke frame $\tuple{\sem{\bullet *}^{\epikspec},
+  \hat{R}_1,...,\hat{R}_n}$.\footnote{ The domain of the Kripke frame
+  differs from the domain of the guarded string model, because the
+  former is the set of worlds, while the latter is the set of
+  propositions.  } The frame consists of a set of worlds, and an
+epistemic-alternative relation for each agent.  It is used as a target
+for natural-language interpretation in Section 6.
 
@@
 sequences (sequences of 0's and 1's) are used for atoms, and these alternate with
-event symbols in the encoding of a world. \ex{gs1fst} gives the encoding of some of
-the worlds from the examples.  A string is a finite sequence of symbols, and {\tt 0},
+event symbols in the encoding of a world. \ex{gs1fst} gives the encoding of worlds
+from the example.  A string is a finite sequence of symbols, and {\tt 0},
 {\tt 1}, {\tt u}, and {\tt d}, are symbols.  {\tt a0} and {\tt b0} are
@@
 no connection with the element {\tt a}.
-
-\begin{example}   \label{gs1fst}
-  \renewcommand{\arraystretch}{1.2}
-Worlds coded as strings
-\vspace{1mm}
-
- \hspace{-13mm} \begin{tabular}[t]{cl}
-{\textit{World}} & {\textit{String}} \\
-% $v_1$ & $ 0 \,u\,  1 \,u\,  2 \,d\,  1 \,u\,  2$ \\
-$v_1$ & {\tt 0 0 u 0 1 u 1 0 d  0 1 u 1 0} \\
-$v_2$ & {\tt 0 0 u 0 1 u 1 0 u 1 1} \\
-$w_1$ & {\tt 1 0} \\
-$w_2$ & {\tt 1 0 a1 1 0} \\
-$w_3$ & {\tt 1 0 a1 1 0 b1 10} \\  
-  \end{tabular}
-    \renewcommand{\arraystretch}{1}
-\end{example}
-
-Terms in the finite state calculus are interpreted as sets of strings, or for
-relational terms, as relations between strings.  Computationally, the sets and
-relations are represented by finite state acceptors.  As used here, a program in the
-Fst language of the finite state calculus is a straight-line program that defines a
-sequence of constants naming sets, constants naming relations, and functions mapping
-one or more regular relations to a regular set or relation.  The functions are
-defined as macros.  To illustrate, \ex{stex1} is part of an Epik program that
-declares states and a state constraint.  This is equivalent to the Fst program
-\ex{stex2}. The result of the sequence of definitions is that {\tt St} is the set of
-Boolean vectors or length four that are well-formed according to the state formula;
-Nst is the complement operation for tests; H1 is the set of well-formed Boolean
-vectors that have 1 in the first position; and so forth.  Translation from an Epik
-program to an Fst program is accomplished in Haskell.
-
-\begin{example}\label{stex1}
-Part of an Epik program
-\end{example}
-
-\vspace{-3mm} {\small
-\begin{verbatim}  
-   state H1 T1 H2 T2
-   constraint !(H1 T1) & (H1 + T1) 
-              !(H2 T2) & (H2 + T2) 
-\end{verbatim}}
-
-\begin{example}\label{stex2}
-Corresponding Fst program
-\end{example}
-
-\vspace{-3mm}
-{\small 
-\begin{verbatim} 
-   define Bool ["0"|"1"];
-   define St0 [Bool Bool Bool Bool];
-   define H1 [1 Bool Bool Bool];
-   define T1 [Bool 1 Bool Bool];
-   define H2 [Bool Bool 1 Bool];
-   define T2 [Bool Bool Bool 1];
-   define St St0 & ((((St0 - (H1 & T1)) 
-                 & (H1 | T1)) 
-                 & (St0 - (H2 & T2))) 
-                 & (H2 | T2));
-   define Nst(X) [St - X];
-   define H1 St & H1;
-   define T1 St & T1;
-   define H2 St & H2;
-   define T2 St & T2;
-\end{verbatim}}
-
-Event symbols are interpreted as event types, in the way discussed in connection with
-\ref{events1}.  For instance, defines an event type {\tt paT2type} of Amy peeking at the second coin
-in a scenario with two coins, and seeing that it is tails.\footnote{After the definition,
-  {\tt paT2type} is a constant symbol denoting a set of two strings. On the right hand side of the definition,
-  {\tt paT2} is the unit set of the symbol {\tt paT2}.  
-}  The first conjunct
-expresses that this event can only happen if the second coin is tails ({\tt H2}), and that the
-state of the second coin does not change.  The second conjunct expresses that the
-event does not change the state of the first coin -- either it is heads ({\tt H1}) and
-remains heads, or it is tails ({\tt T1}) and remains tails.  Such formulas are mapped from
-effect formulas in Epik programs. The state formula is imposed as well, since it
-figures in the definition of state constants such as {\tt H1}.
-
-\begin{example} \label{peek1}
-{\small 
-\begin{verbatim}
-define paT2type [[[T2 paT2] T2] & 
-  [[[H1 paT2] H1] | [[T1 paT2] T1]]];
-\end{verbatim} }
-\end{example}
-
-Semantically in Fst, {\tt paT2type} is a set of two strings, which can be displayed
-in an interpreter for the language by reading {\tt paT2type} as a regular expression
-and printing its extension, see \ex{printwords1}.
-
-\begin{example}
-\begin{verbatim}
-xfst[1]: regex paT2type;
-3.0 Kb. 16 states, 16 arcs, 2 paths.
-xfst[2]: print words
-0 1 0 1 paT2 0 1 0 1
-1 0 0 1 paT2 1 0 0 1
-\end{verbatim}
-\end{example}
-
-The finite state calculus has a product operation of string concatenation raised to sets.
-Concatenation of strings with atoms (Boolean vectors) at both ends has the effect of
-doubling atoms at the juncture, and does not enforce matching of atoms at the juncture.  Therefore
-KAT product can not be identified with product in the finite state calculus. Instead,
-KAT product and KAT Kleene star are defined operations, see Figure \ref{fstfigure1}.
-The binary product operation {\tt Cn} and the unary Kleene star operation
-{\tt Kst} combine strings in the string algebra, remove strings with
-non-matching atoms, and then delete the second of two tests to create well-formed
-guarded strings.  Matching of atoms is enforced with the set {\tt Wf0}, which is
-the set of strings that do not contain non-matching Boolean vectors.  The containment
-operator (expressed by the dollar sign) and complement (expressed by tilde) are operators of the finite
-state calculus.  The set of non-matching sequences of atoms {\tt UnequalStPair} is defined by a finite disjunction.
-Deletion is accomplished by a re-write rule in the finite state calculus, which is a
-notation for defining regular relations by contextually constrained substitutions
-(REFS).  In this case, is {\tt Squash} us a regular relation that deletes an atom
-(a sequence of 0's and 1's of a certain length) when it follows an atom.\footnote{
-  This is a non-equal length regular relation. The finite state calculus includes such
-  relations, and they can be used with relation composition and relation domain and co-domain.
-  They are restricted in that the complement and set difference for non-equal length relations is not defined.
-The epistemic alterative relations that are defined in Figure -- are equal-length relations.
-  }
 
@@
 
-A given bare event such as $a_1$ (Aly looks at heads) is in the KAT algebra a set of
-bare events decorated with compatible tests on each side, semantically $\{ 10 a_1 10
+\begin{example}   \label{gs1fst}
+  \renewcommand{\arraystretch}{1.2}
+Worlds coded as strings
+\vspace{1mm}
+
+ \hspace{-13mm} \begin{tabular}[t]{cl}
+{\textit{World}} & {\textit{String}} \\
+% $v_1$ & $ 0 \,u\,  1 \,u\,  2 \,d\,  1 \,u\,  2$ \\
+$w_1$ & {\tt 1 0} \\
+$w_2$ & {\tt 1 0 a1 1 0} \\
+$w_3$ & {\tt 1 0 a1 1 0 b1 1 0} \\  
+  \end{tabular}
+    \renewcommand{\arraystretch}{1}
+\end{example}
+
+Terms in the finite state calculus are interpreted as sets of strings,
+or for relational terms, as relations between strings.
+Computationally, the sets and relations are represented by finite
+state acceptors.  As used here, a program in the Fst language of the
+finite state calculus is a straight-line program that defines a
+sequence of constants naming sets, constants naming relations, and
+functions (defined as macros) mapping one or more regular sets or
+relations to a regular set or relation.  The finite state calculus has
+a product operation of string concatenation raised to sets.
+Concatenation of strings with atoms (Boolean vectors) at both ends has
+the effect of doubling atoms at the juncture, and does not enforce
+matching of atoms at the juncture.  Therefore KAT product can not be
+identified with product in the finite state calculus. Instead, KAT
+product and KAT Kleene star are defined operations, see Figure
+\ref{fstfigure1}.  The binary product operation {\tt Cn} and the unary
+Kleene star operation {\tt Kst} combine strings in the string algebra,
+remove strings with non-matching atoms, and then delete the second of
+two tests to create well-formed guarded strings.  Matching of atoms is
+enforced with the set {\tt Wf0}, which is the set of strings that do
+not contain non-matching Boolean vectors.  The containment operator
+(expressed by the dollar sign) and complement (expressed by tilde) are
+operators of the finite state calculus.  The set of non-matching
+sequences of atoms {\tt UnequalStPair} is defined by a finite
+disjunction.  Deletion is accomplished by a re-write rule in the
+finite state calculus, which is a notation for defining regular
+relations by contextually constrained substitutions.  In this case, is
+{\tt Squash} us a regular relation that deletes an atom (a sequence of
+0's and 1's of a certain length) when it follows an atom.\footnote{
+  This is a non-equal length regular relation. The finite state
+  calculus includes such relations, and they can be used with relation
+  composition and relation domain and co-domain.  They are restricted
+  in that the complement and set difference for non-equal length
+  relations is not defined.  The epistemic alterative relations that
+  are defined in Figure -- are equal-length relations.  }
+
+\begin{figure} \label{relclosure}
+{\raggedright
+\begin{hangingpar}
+{\tt{define RelKpl(R)}}
+$\underbrace{\mbox{\tt{Squash.i\hspace{1pt}.o.}}}_c$
+$\underbrace{\mbox{\tt Wf0\hspace{1pt}.o.}}_b$
+$\underbrace{\mbox{\tt{[R+]}}}_a$
+$\underbrace{\mbox{\tt .o.\hspace{1pt}Wf0}}_b$
+$\underbrace{\mbox{\tt{.o.\hspace{1pt}Squash}}}_c$
+\end{hangingpar}}
+
+\begin{tabular}{ll}
+  $a$ & Relational Kleene plus in the string algebra \\
+  $b$ & Constrain domain and co-domain to contain \\
+      & no unmatched tests. \\
+  $c$ & Reduce doubled tests to a single \\
+      & test in the domain and co-domain. \\
+\end{tabular}
+
+\begin{hangingpar}
+{\tt{define Kst(R)
+    [St\hspace{1pt}.x.\hspace{1pt}St]\hspace{1pt}|\hspace{1pt}Kpl(X);}} The Fst
+operation {\tt{.x.}} is Cartesian product.  {\tt{R.i}} is the inverse of relation
+{\tt{R}}.
+\end{hangingpar}
+
+\caption{Definition in Fst of the Kleene concatenation closure of
+  a relation between guarded strings.}
+\label{fstfigure2}
+\end{figure}
+
+An event symbol as $a_1$ (Aly looks at heads) is in the KAT algebra a set of
+bare events decorated with compatible tests on each side, $\{ 10 a_1 10
 \}$ in this case.  This is a unit set rather than a guarded string, because elements
@@
 Figure \ref{epikfigure} as interpreted in FST defines a countably infinite set of possible worlds by
-KAT Kleene closure as $\mbox{\em Kst}(a_1 + a_0 +b_1 + b_0)$, and an algebra of
-propositions as regular sets of strings drawn from this space of worlds.
+KAT Kleene closure as $\mbox{\em Kst}(a_1 + a_0 +b_1 + b_0)$.
 
@@
 
-\vspace{-3mm}
+\vspace{-2mm}
 \begin{example} \label{alyrel1}
- \(
+ \hspace{-4mm} \(
  \left\{
  \begin{array}{l}
-   \pair{10a_110}{10a_110}, \\  \pair{01a_001}{01a_001}, \\
-   \pair{10b_110}{10b_110}, \\ \pair{10b_110}{01b_001}, \\
-   \pair{01b_001}{10b_110}, \\ \pair{01b_001}{01b_001} \\
+   \pair{10a_110}{10a_110},  \pair{01a_001}{01a_001}, \\
+   \pair{10b_110}{10b_110},  \pair{10b_110}{01b_001}, \\
+   \pair{01b_001}{10b_110},  \pair{01b_001}{01b_001} \\
  \end{array}
@@
 %on relations is concatenation closure, not reflexive transitive closure.)
-
-\begin{figure} \label{relclosure}
-{\raggedright
-\begin{hangingpar}
-{\tt{define RelKpl(R)}}
-$\underbrace{\mbox{\tt{Squash.i\hspace{1pt}.o.}}}_c$
-$\underbrace{\mbox{\tt Wf0\hspace{1pt}.o.}}_b$
-$\underbrace{\mbox{\tt{[R+]}}}_a$
-$\underbrace{\mbox{\tt .o.\hspace{1pt}Wf0}}_b$
-$\underbrace{\mbox{\tt{.o.\hspace{1pt}Squash}}}_c$
-\end{hangingpar}}
-
-\begin{tabular}{ll}
-  $a$ & Relational Kleene plus in the string algebra \\
-  $b$ & Constrain domain and co-domain to contain \\
-      & no unmatched tests. \\
-  $c$ & Reduce doubled tests to a single \\
-      & test in the domain and co-domain. \\
-\end{tabular}
-
-\begin{hangingpar}
-{\tt{define Kst(R)
-    [St\hspace{1pt}.x.\hspace{1pt}St]\hspace{1pt}|\hspace{1pt}Kpl(X);}} The Fst
-operation {\tt{.x.}} is Cartesian product.  {\tt{R.i}} is the inverse of relation
-{\tt{R}}.
-\end{hangingpar}
-
-\caption{Definition in Fst of the Kleene concatenation closure of
-  a relation between guarded strings.}
-\label{fstfigure2}
-\end{figure}
 
@@
 
-An architecture of interpretation by transtlation is employed, where
 English sentences are mapped to terms in the logical language via an
 interpreted grammar, and these terms are in turn interpreted as
-propositions (sets of possible worlds). For the latter, there are
-options of translation into the finite state calculus in order to
-represent propositions as finite state machines (Section 4), and
-representation in Haskell via lazy lists of guarded strings (Section
-5).  The grammar is a semantically interpreted multi-modal categorial
+propositions (sets of possible worlds).
+The grammar is a semantically interpreted multi-modal categorial
 grammar, consisting of a lexicon of words, their categorial types, and
-interpretations in a logical lambda language.  Figure
-\ref{phenomenafigure} lists phenomena that are covered.  There is
-recursion through conjuction and verbal complementation, so that the
-language is infinite. It includes talk of beliefs about beliefs, or in
-general, talk of arbitrarily iterated belief.
-
-\begin{figure}  
-\begin{tabular}[t]{l|l}
-  Basic statives & It's heads. \\
-  & It's tails.\\ \hline
-That-complement & Amy knows \\ & \hspace{2mm} that it's heads. \\ \hline
-Wh-complement & Amy knows \\ & \hspace{2mm} whether its heads. \\ \hline
-Negation & Bob doesn't know \\ & \hspace{2mm} that it isn't heads. \\ \hline 
-Tensed and base &  Bob knows \\
-verbal forms & \hspace{2mm} that it's heads.      \\ 
-\hspace{1mm} & Bob doesn't know \\ & \hspace{2mm} that it's heads.\\ \hline
-Sentence conjunction & It's heads and Bob \\ & \hspace{2mm} doesn't know \\ & \hspace{2mm} that it's heads \\ \hline
-Predicate conjunction & Bob knows that \\ & Amy knows whether \\ & it's heads and
-\\ & doesn't know that \\ & Amy knows that \\ & it's heads.
-\end{tabular}
-\caption{Phenomena covered in the English grammar fragment.}
-\label{phenomenafigure}
-\end{figure}
-
+interpretations in a logical lambda language.
+The grammar covers basic statives (it's heads), that- and whether-complements
+of {\em know}, sentence negation, and predicate and sentence conjunction.
 Figure \ref{lexiconfigure} gives illustrative lexical
@@
   with $d$ on the left to give $t$.} The grammar and semantics are
-optimized for a simple fragment of English concerned with
-clausal complementation.  The agent names {\em Amy} and {\em Bob}
-contribute the epistemic alternative relations for those agents,
-rather than individuals.  This is possible because the agents are
-never arguments of extensional predicates, so what matters about the
-agents is their epistemic alternative relations.  The root verb {\em
-  know} contributes existential modal force.  The complementizers {\em
-  that} and {\em whether} are the heads of their dominating clauses,
-and assemble an alternative relation, modal force, and proposition
-contributed by the complement.  These complementizers introduce the
-dual via two negations, in order to express universal modal force.
-These moves are offered here as a way of constructing a compact
-interpreted grammar. They can easily be reformulated in a more
-comprehensive interpreted grammar of English.
+optimized for a simple fragment of English concerned with clausal
+complementation.  The agent names {\em Amy} and {\em Bob} contribute
+the epistemic alternative relations for those agents, rather than
+individuals.  This is possible because the agents are never arguments
+of extensional predicates.  The root verb {\em know} contributes
+existential modal force.  The complementizers {\em that} and {\em
+  whether} are the heads of their dominating clauses, and assemble an
+alternative relation, modal force, and proposition contributed by the
+complement.  These complementizers introduce the dual via two
+negations, in order to express universal modal force.
 
@@
 Multimodal categories such as $\backslash_D$ and $\backslash_M$ are
-used to control the derivation.  For instance the category of {\em
-  heads} is $d \backslash_D t$.  The dummy expletive subject {\em it}
-has category $d$.  However the phrase {\em it heads} of category $t$
-can not be formed, because $\backslash_D$ is not syntactically active
-as a function.  Instead {\em it is heads} is formed with a predicator
-{\em is} of category $(d \backslash t)\slash(d \backslash_D t)$.
-Similarly {\em knows} has a category with the top-level slash
-$\slash\!_M$, and combines to form a sentence as an argument of {\em
-  that} or {\em whether}. These complementizers have a category that
-looks for the category of {\em know} on the left, after combining with
-a complement sentence on the right.
-
+used to control the derivation.
+%
+%For instance the category of {\em
+%  heads} is $d \backslash_D t$.  The dummy expletive subject {\em it}
+%has category $d$.  However the phrase {\em it heads} of category $t$
+%can not be formed, because $\backslash_D$ is not syntactically active
+%as a function.  Instead {\em it is heads} is formed with a predicator
+%{\em is} of category $(d \backslash t)\slash(d \backslash_D t)$.
+%Similarly {\em knows} has a category with the top-level slash
+%$\slash\!_M$, and combines to form a sentence as an argument of {\em
+%  that} or {\em whether}. These complementizers have a category that
+%looks for the category of {\em know} on the left, after combining with
+%a complement sentence on the right.
 The semantic translations in the third column of Figure
@@
 strings that end with a Boolean valuation where the primitive
-proposition $h$ (it's heads) is true.  There is $\lambda x$ at the
-front because the grammar formalism enforces a strict correspondence
-between syntactic and semantic types. However, the lambda does not
-bind anything, because sentences such as {\em it isn't heads} have an
-expletive subject.  The body ${\Diamond}R p$ of $\lambda p.\lambda
+proposition $h$ (it's heads) is true.
+%There is $\lambda x$ at the
+%front because the grammar formalism enforces a strict correspondence
+%between syntactic and semantic types. However, the lambda does not
+%bind anything, because sentences such as {\em it isn't heads} have an
+%expletive subject.
+The body of $\lambda p.\lambda
 R. {\Diamond}R p$, which is the semantic lexical entry of {\em knows},
@@
 semantics for complex phrases are obtained by syntactic application of
-semantic translations, accompanied by beta reduction.  Semantic terms
-in the parsing formalism are expressions of untyped lambda calculus.
+semantic translations, accompanied by beta reduction.
 The grammar is set up so that lambda is eliminated by beta reduction
 in the semantic term corresponding to a sentence. In consequence, the
-semantic term translating a sentence is a term of the Epik term
+semantic term translating a sentence is a term of the logical language
 language.  Such a term designates a set of possible words (guarded
 strings) in the possible worlds model deterimened by an Epik
-specification.
-By way of example, 
-\exi{three}{a} is an English sentence with predicate conjunction and three levels
-of clausal embedding.  Using the grammar and parser, the sentence is mapped to the
-Epik term \exi{b}, which here takes the form of a Scheme s-expression.
-
-Using the result from Section 3, this term can be mapped in an
-implementation of the finite state calculus to a finite state machine that represents
-a countably infinite set of possible worlds, represented as guarded strings.  Using
-the result from Section 4, it can be mapped to an infinite lazy list of guarded
-strings, representing the same set of possible worlds.  Either of these is a concrete
-computational representation of the propositional semantic value \semval{Amy knows
-  that Bob knows that Amy knows whether it is heads and knows that Bob does not know
-  that Amy knows that it is tails}, in the familiar sense of Montague semantics for
-natural language.
-
-\begin{example}  \label{three}
-Amy knows that Bob knows that Amy knows whether it is heads and
-knows that Bob doesnt know that Amy knows that it is tails.
-\end{example}
-
-\section{Examples and discussion}
-
-\begin{examples}
-\itema Amy knows that it is tails
-\itemb {\tt (Not (Diamond amy (Not tails)))}
-\end{examples}
-
-\begin{examples}
-\itema Bob doesnt know that Amy knows that it is tails
-\itemb {\tt (Not (Not (Diamond bob (Not (Not (Diamond amy (Not tails)))))))}
-\end{examples}
-
-\begin{examples}
-\itema Bob knows that Amy knows whether it is heads.
-\itemb {\tt (Not (Diamond bob (Not (Or (Not (Diamond amy (Not heads))) (Not (Diamond amy heads))))))}
-\end{examples}
-
-\begin{examples}
+specification.  By way of example, \exi{three}{a} is an English
+sentence with predicate conjunction and three levels of clausal
+embedding.  Using the grammar and parser, the sentence is mapped to
+the term \exi{b}, which here takes the form of a Scheme
+S-expression.
+
+\begin{examples} \label{three}
 \itema Amy knows that Bob knows that Amy knows whether it is heads and knows that Bob doesnt know that Amy knows that it is tails.
@@
 
+\exi{three}{b} is compiled in Fst into a finite state machine with --
+nodes and -- edges that accepts a countably infinite set of strings
+(worlds).  This is a concrete computational representation of the
+propositiion denoted by \exi{three}{a}.  This contrasts with standard
+computational approaches to intensional semantics, sentences are
+translated into logical formulas that have a mathematical
+interpretation as a set of worlds, but not a computational one.  The
+finite state representations can be used to check entailment between
+sentences (via a subset check), print random worlds that satisfy the
+sentence, or the like.
+
 \pagebreak
 
-\bibliographystyle{chicago}
+\bibliographystyle{acl_natbib}
+% \bibliographystyle{chicago}
 \bibliography{fs}
@@
 \end{document}
+
+\section{Credits}
+
+This document has been adapted
+by Steven Bethard, Ryan Cotterrell and Rui Yan
+from the instructions for earlier ACL and NAACL proceedings, including those for 
+ACL 2019 by Douwe Kiela and Ivan Vuli\'{c},
+NAACL 2019 by Stephanie Lukin and Alla Roskovskaya, 
+ACL 2018 by Shay Cohen, Kevin Gimpel, and Wei Lu, 
+NAACL 2018 by Margaret Michell and Stephanie Lukin,
+2017/2018 (NA)ACL bibtex suggestions from Jason Eisner,
+ACL 2017 by Dan Gildea and Min-Yen Kan, 
+NAACL 2017 by Margaret Mitchell, 
+ACL 2012 by Maggie Li and Michael White, 
+ACL 2010 by Jing-Shing Chang and Philipp Koehn, 
+ACL 2008 by Johanna D. Moore, Simone Teufel, James Allan, and Sadaoki Furui, 
+ACL 2005 by Hwee Tou Ng and Kemal Oflazer, 
+ACL 2002 by Eugene Charniak and Dekang Lin, 
+and earlier ACL and EACL formats written by several people, including
+John Chen, Henry S. Thompson and Donald Walker.
+Additional elements were taken from the formatting instructions of the \emph{International Joint Conference on Artificial Intelligence} and the \emph{Conference on Computer Vision and Pattern Recognition}.
 

### 2020-09-22 tex/sdraft.tex
chopping and tightening pass

@@
 
+\newcommand{\ehc}[1]{\textit{\textcolor{green}{[ehc]: #1}}}
+\newcommand{\mr}[1]{\textit{\textcolor{cyan}{[mr]: #1}}}
+
 %\captionsetup{belowskip=0pt}
@@
   account is framed as a formal language Epik (Epistemic KAT) for
-  defining such models.  The language is interpreted by translation
+  defining such models.  The language is implemented by translation
   into the finite state calculus, and alternatively by modeling
@@
 neither agent knows which it is.  This initial situation is possible
-world $w_1$.  Two additional worlds $w_2$ and $w_2$ are defined by
+world $w_1$.  Two additional worlds $w_2$ and $w_3$ are defined by
 sequencing events after the initial state, with events interpreted as
@@
    $w_2 = w_1 \hspace{2pt} a_1 $ \hspace{2mm}
-   $w_2 = w_1 \hspace{2pt} a_1 \hspace{2pt} b_1 $ 
+   $w_3 = w_1 \hspace{2pt} a_1 \hspace{2pt} b_1 $ 
 \end{example}
@@
 state formula as well.  For the heads-tails example, let the state
-formula and the effect formula for $a_1$ (Aly peeking at heads) be as
+formula and the effect formula for $a_1$ (Amy peeking at heads) be as
 specified in \ex{effect}. Then ${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\varphi}$
@@
 \begin{figure}
+  \footnotesize
 \begin{center}
@@
 ${\sf{B}}$ is $\{h,t\}$.
-In the elevator example,
-$\sf{P}$ is $\{u,d\}$, and $\sf{B}$ is $\{p,q\}$
-
-Assuming a trivially true state formula $\rho$, the set of atoms
-${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\rho}$ in the elevator example is
-$\{\bar q \bar p$, $\bar q p$, $q \bar p$, $q p\}$, which we write
-$\{\hat0$, $\hat1$, $\hat2$, $\hat3\}$.  In the coin example, as we
-already saw in \ex{effect}, $\mathcal{A}^\varphi_{\mathsf B}$ is
-$\{h\bar{t},\bar{h}t\}$, for which we use the shorthand $\{H,T\}$.  A
-guarded string over ${\sf{P}}$ and ${\sf{B}}$ is a strings of events
-from ${\sf{P}}$, alternating with atoms over ${\sf{B}}$, and beginning
-and ending with atoms.  \ex{gs1} gives the encoding as guarded strings
-of the worlds in \ex{elevator1} and \ex{coin1}.  The length of a
-guarded string $p$, written $|p|$ is the number of events in $p$.  An
-atom such as $H$ is a guarded string of length $0$.
-
-\begin{example}   \label{gs1}
-  \renewcommand{\arraystretch}{1.2}
-  \begin{tabular}[t]{clc}
-{\textit{World}} & {\textit{Guarded string}} & \textit{Length}\\ \hline
-%$v_1$ & $\hat 0 \,u\, \hat 1 \,u\, \hat 2 \,d\, \hat 1 \,u\, \hat 2$ & $4$ \\
-%$v_2$ & $\hat 0 \,u\, \hat 1 \,u\, \hat 2 \,u\, \hat 3$ & $3$ \\   
-$w_1$ & $H$ & $0$\\
-$w_2$ & $H \,a_1\, H$ & $1$ \\  
-$w_3$ & $H\,a_1\, H \,b_1\, H$ & $2$\\
-  \end{tabular}
-    \renewcommand{\arraystretch}{1}
-\end{example}
+
+In the coin example, as we already saw in \ex{effect},
+$\mathcal{A}^\varphi_{\mathsf B}$ is $\{h\bar{t},\bar{h}t\}$, for which we use
+the shorthand $\{H,T\}$.  A guarded string over ${\sf{P}}$ and ${\sf{B}}$ is a
+strings of events from ${\sf{P}}$, alternating with atoms over ${\sf{B}}$, and
+beginning and ending with atoms.  \ex{gs1} gives the encoding as guarded strings
+of the worlds in \ex{elevator1} and \ex{coin1}.  The length of a guarded string
+$p$, written $|p|$ is the number of events in $p$.  An atom such as $H$ is a
+guarded string of length $0$. Correspondingly $|w_1| = 0$, $|w_2| = 1$ and
+$|w_3| = 2$.
 
@@
 
-\hspace{-2mm}\(H \; \diamond \; H \,a_1\, H = H \,a_0\, H \)
+\hspace{-2mm}\(H\,b_1\,H \; \diamond \; H \,a_1\, H = H\, b_1\, H \,a_1\, H \)
 
@@
 
-Well-formed guarded strings have the interpretation of worlds in the
-application to natural-language semantics.  The set of possible worlds
-in the Kripke frame determined by an Epik specification is the set of
-well-formed guarded strings.  At this point, we could say that any set
-of worlds is a proposition, so that the set of propositions is the
-power set of the set of worlds
-\citep{montague1975formal,gallin1975intensional}.  We will instead
-define a more restrictive set of propositions corresponding to the
-regular sets of strings. This is deferred to the next section.
-Certai sets of well-formed guarded strings 
-have the addional interpretation of event types.
-An event-type is something that can ``happen'' is different worlds.
-\ex{events1} gives the event types in
-the example.
-
-\begin{example} \label{events1}
-  \renewcommand{\arraystretch}{1.1}
-\(
-\begin{array}[t]{l|cl|c}
-  &  {\mbox{Event type}} && {\mbox{Event type}} \\ \hline
-  a_1 & \{H \,a_1\, H\} &   b_1 & \{H \,b_1\, H \} \\
-  a_0 & \{T \,a_0\, T\} &   b_0 & \{T \,b_0\, T \} \\
-\end{array}
-\)
-  \renewcommand{\arraystretch}{1}
-\end{example}
+Well-formed guarded strings have the interpretation of worlds in the application
+to natural-language semantics.  The set of possible worlds in the Kripke frame
+determined by an Epik specification is the set of well-formed guarded strings.
+At this point, we could say that any set of worlds is a proposition, so that the
+set of propositions is the power set of the set of worlds
+\citep{montague1975formal,gallin1975intensional}.  We will instead define a more
+restrictive set of propositions corresponding to the regular sets of
+strings. This is deferred to the next section.  Certain sets of well-formed
+guarded strings have the addional interpretation of event types.  An event-type
+is something that can ``happen'' is different worlds. For example, $a_1$ has the
+event type $\{H\,a_1\,H\}$, and $a_0$ has the even type $\{T\,a_0\,T\}$. The
+event types for $b_0$ and $b_1$ are analogous.
 
@@
 
-Consider a scenario like the one from Figure 1, but with an additional agent Cem. The
+Consider a scenario like the one from Figure 1, but with an additional agent Cal. The
 base world $T b_0 T c_0 T$ is one where the coin is tails, and first Bob looks at
-tails, and them Cem looks at tails. The first event $b_0$ has the alternatives $b_0$
+tails, and them Cal looks at tails. The first event $b_0$ has the alternatives $b_0$
 and $b_1$ for Amy, and the second event $c_0$ has the alternatives $c_0$ and $c_1$
@@
 modal operation $\Diamond_a$ for each agent, and a unary complement operation $\neg$
-on elements of $K$.  The intended interpretation of $\Diamond_a p$ is the set of
+on elements of $K$.  Intuitively, $\Diamond_a p$ is the set of
 worlds where proposition $p$ is epistemically possible for agent $a$.  Propositional
@@
 defined in Figure \ref{fig:epik-grammar} is the disjunction of the primitive events.
-Since a world is a well-formed sequence of events, $\bullet *$ is the set of worlds.
-Multiplying by the state symbol $h$ in the term $\bullet \! * \! h$ has the effect of
-conjoining $h$ with the atom at the end of the world.  So $\bullet\!*\!h$ is the set
+Since a world is a well-formed sequence of events, $\bullet^*$ is the set of worlds.
+Multiplying by the state symbol $h$ in the term $\bullet^* h$ has the effect of
+conjoining $h$ with the atom at the end of the world.  So $\bullet^* h$ is the set
 of worlds where the coin is heads.\footnote{A mapping from English sentences to logical
@@
 \begin{tabular}[t]{ll}
-${\bullet}\!* t$ & It's tails. \\
-$\bullet \!* h$ & It's heads. \\  
-$\Box_a \bullet \!* h$  & Amy knows that it's heads. \\
+${\bullet}^*t$ & It's tails. \\
+$\bullet^*h$ & It's heads. \\  
+$\Box_a \bullet^* h$  & Amy knows that it's heads. \\
 \end{tabular}
@@
 \begin{tabular}[t]{l}
-  $\Box_b(\Box_a\!\bullet\!* t + \Box_a\!\bullet \!* \neg t)$\\
+  $\Box_b(\Box_a\bullet^*t + \Box_a\bullet^*\neg t)$\\
  Bob knows that Amy knows whether \\ \hspace{3mm} it's tails.  \\
@@
   \sem{{pq}}^{\epikspec} && \setabs{x \diamond y}{ \begin{array}{l}x \in \sem{p}^{\epikspec} \\ y \in \sem{q}^{\epikspec}\\\mbox{$x \diamond y$ is defined}\end{array}} \\
-  \sem{p\!*\!}^{\epikspec} && \displaystyle\bigcup_{n \geq 0} {\sem{ p^n }^{\epikspec}} \\
-  \sem{ \neg p }^{\epikspec} && \sem{\bullet\!*}^{\epikspec} \setminus \sem{ p }^{\epikspec} \\
+  \sem{p^*}^{\epikspec} && \displaystyle\bigcup_{n \geq 0} {\sem{ p^n }^{\epikspec}} \\
+  \sem{ \neg p }^{\epikspec} && \sem{\bullet^*}^{\epikspec} \setminus \sem{ p }^{\epikspec} \\
   \end{array}
@@
 $\Diamond_a$ is pre-image relative to $\hat{R}_a$).  This provides a
-concrete $n$-agent Kripke frame $\tuple{\sem{\bullet *}^{\epikspec},
+concrete $n$-agent Kripke frame $\tuple{\sem{\bullet^*}^{\epikspec},
   \hat{R}_1,...,\hat{R}_n}$.\footnote{ The domain of the Kripke frame
@@
 relations by contextually constrained substitutions.  In this case, is
-{\tt Squash} us a regular relation that deletes an atom (a sequence of
+{\tt Squash} is a regular relation that deletes an atom (a sequence of
 0's and 1's of a certain length) when it follows an atom.\footnote{
@@
 
-An event symbol as $a_1$ (Aly looks at heads) is in the KAT algebra a set of
+An event symbol such as $a_1$ (Amy looks at heads) is in the KAT algebra a set of
 bare events decorated with compatible tests on each side, $\{ 10 a_1 10
@@
 agent. This determines a relation in the guarded string algebra a relation between
-bare events decorated with compatible tests.  For agent Aly, this is the relation
+bare events decorated with compatible tests.  For agent Amy, this is the relation
 described in \ex{alyrel1} as a set of ordered pairs.
@@
 of $R_a$.  Where $R$ and $S$ are relations, the concatenation product
-of $R$ and $S$ is the set of pairs of the form $\pair{x_1x_2}{y_1y2}$,
+of $R$ and $S$ is the set of pairs of the form $\pair{x_1x_2}{y_1y_2}$,
 where $\pair{x_1}{y_1}$ is in relation $R$, and $\pair{x_2}{y_2}$ is
@@
 
-The bounded interpretation into lists of strings is very similar to
-the unbounded interpretation into sets of strings, except for the
-bounds checking. The full details are shown in
-Figure~\ref{fig:list-interp}. First note that when $n=0$, we simply
-return the empty list, denoted $[]$. Terms of the form $0$, $1$, $e$,
-and $\psi$ have the same denotation as before, translated into a list
-(for a set $X$, $\lfloor X \rfloor$ is a list with the same elements
-as $X$ arbitrarily ordered). The term $p+q$ denotes the list
-concatenation (written $\texttt{+\!+}$) of the denotation of $p$ and of
-$q$. The term $p;q$ denotes the fusion product (lifted to lists) of
-the denotations of $p$ and in $q$ restricted to only those strings
-shorter than $n$ (for a list $l$, $l|_n$ filters out elements longer
-than $n$). The denotation of $p^*$ uses the fact that $p^*$ and $1 +
-p;p^*$ are equivalent, and decrements the size threshold on the
-recursive denotation of $p^*$ by $i$, where $i$ is the length of the
-longest (nonzero) string in the denotation of $p$, making sure to
-filter out guarded strings that are too long. The denotation of $\neg
-p$ is the strings that occur in $\_^*$ and not in $p$ (the $\setminus$
-operator on lists is analogous to the $\setminus$ operator on
-sets). The denotation of $\Diamond_a~p$ is analogous to the set
-denotation, depicted in Figure~\ref{fig:list-interp}-comprehension
-notation\footnote{List comprehension notation is analogous to set
-  builder notation, except that it is written using square
-  brackets. The order preservation is indicated by the keyword
-  \texttt{for}}.
-
-To convert from a list of guarded strings $l$ to a set of guarded
-strings, we simply write $\lceil l\rceil$. Note that for any $p$,
-$\varphi$, $\mathcal E$, and $n$, $\left\lceil\llparenthesis p
-\rrparenthesis^{\varphi,\mathcal E}_n\right\rceil = \{g \mid g \in
-\sem{ p }^{\varphi,\mathcal E}, |g| \leq n\}$.
+The bounded interpretation into lists of strings is very similar to the
+unbounded interpretation into sets of strings, except for the bounds
+checking. The full details are shown in Figure~\ref{fig:list-interp}. First note
+that when $n=0$, we simply return the empty list, denoted $[]$. Terms of the
+form $0$, $1$, $e$, and $\psi$ have the same denotation as before, translated
+into a list (for a set $X$, $\lfloor X \rfloor$ is a list with the same elements
+as $X$ arbitrarily ordered). We compute these lists using BBDs, a standard
+technique for concisely and cannoncially representing boolean functions~\cite{BBDS}.
+
+We lift the remaining operators (except Kleene star) to their list equivalents:
+union becomes list append (written $\texttt{+\!+}$); concatenation becomes the
+fusion product (lifted to lists this time), negation is implemented using list
+difference ($\setminus$), and the modal operator lifts the alternative relation
+over lists of strings\footnote{Figure~\ref{fig:list-interp} depicts this using
+  the list comprehension notation, which is analogous to set builder notation,
+  except that it is written using square brackets. The order of elements is
+  preserved indicated by the keyword \texttt{for}}. The only caveat to these
+direct interpretations is that we restrict the operators to have size $\leq n$,
+denoted as $l|_n$ for a list of guarded strings $l$.
+
+The denotation of $p^*$ uses the fact that $p^*$ and $1 + p;p^*$ are equivalent,
+and decrements the size threshold on the recursive denotation of $p^*$ by $i$,
+where $i$ is the length of the longest (nonzero) string in the denotation of
+$p$, making sure to filter out guarded strings that are too long.
+
+%% To convert from a list of guarded strings $l$ to a set of guarded
+%% strings, we simply write $\lceil l\rceil$. Note that for any $p$,
+%% $\varphi$, $\mathcal E$, and $n$, $\left\lceil\llparenthesis p
+%% \rrparenthesis^{\varphi,\mathcal E}_n\right\rceil = \{g \mid g \in
+%% \sem{ p }^{\varphi,\mathcal E}, |g| \leq n\}$.
 
@@
 
-The lists we use here are \emph{lazy} (as opposed to \emph{strict}),
-which broadly means that computation is delayed until the value is
-needed. This allows us to avoid computing large, unnecessary
-iterations. In the elevator example, this means that $\hat 0;u;u;u^*$
-will unroll the $u^*$ 0 times, once, and then discover that it is
-impossible to unroll it a second time, because $\llparenthesis\hat
-3;u\rrparenthesis^{\varphi,\mathcal E}_n = []$, for any $n$. So
-$\llparenthesis\hat 3;u\rrparenthesis^{\varphi,\mathcal E}_n \diamond
-\llparenthesis u^*\rrparenthesis^{\varphi,\mathcal E}_{n-1}$ will
-return $[]$ without having to compute the full fixpoint of
-$\llparenthesis u\rrparenthesis^{\varphi,\mathcal E}$. Similar
-behavior occurs once $n = 0$.
-
-Conversely, if we used sets (as in the math) instead of lists we would
-need to constantly verify the set invariant (that every element is
-unique) which means processing every element in the set. Verifying
-that every element in $\llparenthesis
-u^*\rrparenthesis^{\varphi,\mathcal E}$ is unique would require the full
-computation of the fixpoint.
+The lists we use here are \emph{lazy} (as opposed to \emph{strict}), which
+broadly means that computation is delayed until the value is needed. This allows
+us to avoid computing large, unnecessary iterations.
 
@@
 
-English sentences are mapped to terms in the logical language via an
-interpreted grammar, and these terms are in turn interpreted as
-propositions (sets of possible worlds).
-The grammar is a semantically interpreted multi-modal categorial
-grammar, consisting of a lexicon of words, their categorial types, and
-interpretations in a logical lambda language.
-The grammar covers basic statives (it's heads), that- and whether-complements
-of {\em know}, sentence negation, and predicate and sentence conjunction.
-Figure \ref{lexiconfigure} gives illustrative lexical
-entries.\footnote{Category symbols use Lambek/Bar-Hillel notation for
-  slashes, so that $(d \backslash t)\slash(d \backslash_D t)$ combines
-  with $d \backslash_D t$ on the right to give a value that combines
-  with $d$ on the left to give $t$.} The grammar and semantics are
-optimized for a simple fragment of English concerned with clausal
-complementation.  The agent names {\em Amy} and {\em Bob} contribute
-the epistemic alternative relations for those agents, rather than
-individuals.  This is possible because the agents are never arguments
-of extensional predicates.  The root verb {\em know} contributes
-existential modal force.  The complementizers {\em that} and {\em
-  whether} are the heads of their dominating clauses, and assemble an
-alternative relation, modal force, and proposition contributed by the
-complement.  These complementizers introduce the dual via two
-negations, in order to express universal modal force.
+English sentences are mapped to terms in the logical language via an interpreted
+grammar, and these terms are in turn interpreted as propositions (sets of
+possible worlds).  The grammar is a semantically interpreted multi-modal
+categorial grammar, consisting of a lexicon of words, their categorial types,
+and interpretations in a logical lambda language.  The grammar covers basic
+statives (it's heads), that- and whether-complements of {\em know}, sentence
+negation, and predicate and sentence conjunction.  Figure \ref{lexiconfigure}
+gives illustrative lexical entries.\footnote{Category symbols use
+  Lambek/Bar-Hillel notation for slashes, so that $(d \backslash t)\slash(d
+  \backslash_D t)$ combines with $d \backslash_D t$ on the right to give a value
+  that combines with $d$ on the left to give $t$.}\footnote{Lambda abstractions
+  with multiple parameters are written $\lambda x~y.~e$ rather than the more verbose
+  $\lambda x.\lambda y.e$.} The grammar and semantics are optimized for a simple
+fragment of English concerned with clausal complementation.  The agent names
+{\em Amy} and {\em Bob} contribute the epistemic alternative relations for those
+agents, rather than individuals.  This is possible because the agents are never
+arguments of extensional predicates.  The root verb {\em know} contributes
+existential modal force.  The complementizers {\em that} and {\em whether} are
+the heads of their dominating clauses, and assemble an alternative relation,
+modal force, and proposition contributed by the complement.  These
+complementizers introduce the dual via two negations, in order to express
+universal modal force.
 
 \begin{figure}
-\begin{tabular}[t]{l|c|c}
+  \footnotesize
+\begin{tabular}[t]{l|c|l}
 Amy & $e$ & ${R}_a$\\
@@
 it & $d$ & $d$ \\
-heads & $d \backslash_D t$ & $\lambda x.0^c{\cdot}h$ \\
-tails & $d \backslash_D t$ & $\lambda x.0^c{\cdot}!h$ \\
-is & $(d \backslash t)\slash(d \backslash_D t)$ & $\lambda P.\lambda x.P x$ \\
-knows & $(e \backslash t)\slash_M t$ &  $\lambda p.\lambda R. {\Diamond}R p$\\
-that  & \multicolumn{2}{l}{$(((e \backslash t){\slash_M}t) \backslash (e \backslash t)) \slash t$} \\
-& \multicolumn{2}{l}{$\lambda p.\lambda m.\lambda R.{\sim}(m ({\sim}p) R)$} \\
-whether  & \multicolumn{2}{l}{$(((e \backslash t){\slash_M} t) \backslash (e \backslash t)) \slash t$} \\
-& \multicolumn{2}{l}{$\lambda p.\lambda m.\lambda R.{\sim}(m ({\sim}p) R)$} \\
-& \multicolumn{2}{r}{$+ {\sim}(m p R)$} \\
+heads & $d \backslash_D t$ & $\lambda x.\bullet^*h$ \\
+tails & $d \backslash_D t$ & $\lambda x.\bullet^*t$ \\
+is & $(d \backslash t)\slash(d \backslash_D t)$ & $\lambda P~x.~P~x$ \\
+knows & $(e \backslash t)\slash_M t$ &  $\lambda p~R.~{\Diamond}_Rp$\\
+that  & $(((e \backslash t){\slash_M}t) \backslash (e \backslash t)) \slash t$
+& $\lambda p~m~R.{\neg}(m\;({\neg}p)\;R)$ \\
+whether  & $(((e \backslash t){\slash_M} t) \backslash (e \backslash t)) \slash t$ 
+& $\lambda p~m~R.{\neg}(m\;({\neg}p)\;R)$ \\
+&& \multicolumn{1}{r}{$+ {\neg}(m\;p\;R)$} \\
 \end{tabular}
 \caption{Partial categorial grammar lexicon. The first column has a word form.
-  the second column a categorial type, and third column a semantic translation in
-a logical language that extends the Epik term language with lambda.}
+  the second column a categorial type, and third column a semantic translation
+  in a logical language that extends the Epik term language with lambda. }
 \label{lexiconfigure}
@@
 %a complement sentence on the right.
-The semantic translations in the third column of Figure
-\ref{lexiconfigure} use the Epik term language, incremented with
-lambda.  The body of $\lambda x.0^c{\cdot}h$, which is the semantic
-lexical entry for {\em heads}, is a term denoting the set of all
-worlds where the coin is heads, expressed as the set of all guarded
-strings that end with a Boolean valuation where the primitive
+The semantic translations in the third column of Figure \ref{lexiconfigure} use
+the Epik term language, incremented with lambda.  The body of $\lambda
+x.\bullet^*h$, which is the semantic lexical entry for {\em heads}, is a
+term denoting the set of all worlds where the coin is heads, expressed as the
+set of all guarded strings that end with a Boolean valuation where the primitive
 proposition $h$ (it's heads) is true.
@@
 %expletive subject.
-The body of $\lambda p.\lambda
-R. {\Diamond}R p$, which is the semantic lexical entry of {\em knows},
-is an Epik term denoting the pre-image of the world-alternative
-relation contributed by the subject.  This is not the right semantics
-for {\em Amy knows that it's heads}, because it is an existential
-modality ${\Diamond_R}p$, rather than an universal modality
-${\Box_R}p$.  This is corrected by the complementizer {\em that} or
-{\em whether}, which introduces the dual.
-
-Sentences are parsed with a chart parser for categorial grammar.  The
-semantics for complex phrases are obtained by syntactic application of
-semantic translations, accompanied by beta reduction.
-The grammar is set up so that lambda is eliminated by beta reduction
-in the semantic term corresponding to a sentence. In consequence, the
-semantic term translating a sentence is a term of the logical language
-language.  Such a term designates a set of possible words (guarded
-strings) in the possible worlds model deterimened by an Epik
-specification.  By way of example, \exi{three}{a} is an English
-sentence with predicate conjunction and three levels of clausal
-embedding.  Using the grammar and parser, the sentence is mapped to
-the term \exi{b}, which here takes the form of a Scheme
-S-expression.
+The body of $\lambda p.\lambda R. {\Diamond}_R p$, which is the semantic lexical
+entry of {\em knows}, is an Epik term denoting the pre-image of the
+world-alternative relation contributed by the subject.  This is not the right
+semantics for {\em Amy knows that it's heads}, because it is an existential
+modality ${\Diamond_R}p$, rather than an universal modality ${\Box_R}p$.  This
+is corrected by the complementizer {\em that} or {\em whether}, which introduces
+the dual.
+
+Sentences are parsed with a chart parser for categorial grammar.  The semantics
+for complex phrases are obtained by syntactic application of semantic
+translations, accompanied by beta reduction.  In consequence, the semantic term
+translating a sentence is a term of the logical language language.  Such a term
+designates a set of possible words (guarded strings) in the possible worlds
+model deterimened by an Epik specification.  By way of example, \exi{three}{a}
+is an English sentence with predicate conjunction and three levels of clausal
+embedding.  Using the grammar and parser, the sentence is mapped to the Epik
+term in \exi{three}{b}.
 
@@
 \itema Amy knows that Bob knows that Amy knows whether it is heads and knows that Bob doesnt know that Amy knows that it is tails.
-\itemb {\tt (And (Not (Diamond amy (Not (Not (Diamond bob (Not (Or (Not (Diamond amy (Not heads))) (Not (Diamond amy heads))))))))) (Not (Diamond amy (Not (Not (Not (Diamond bob (Not (Not (Diamond amy (Not tails))))))))))) }
+%% \itemb {\tt (And (Not (Diamond amy (Not (Not (Diamond bob (Not (Or (Not (Diamond amy (Not heads))) (Not (Diamond amy heads))))))))) (Not (Diamond amy (Not (Not (Not (Diamond bob (Not (Not (Diamond amy (Not tails))))))))))) }
+%% \itemc
+%% \(\neg \Diamond_{\text{Amy}} \neg \neg \Diamond_{\text{Bob}} \neg \neg (\Diamond_{\text{Amy}}\bar h + \neg \Diamond_{\text{amy}} h) \cdot
+%% \neg \Diamond_{\text{Amy}} \neg \neg \neg \Diamond_{\text{Bob}} \neg \neg \Diamond_{\text{Amy}} \bar t\)
+\itemb \(\displaystyle \Box_{\text{Amy}}\Box_{\text{Bob}}(\Box_{\text{Amy}} \bar h + \Box_{\text{Amy}}\bar t)\) \hfill \mbox{} \(\cdot~~\Box_{\text{Amy}}\neg\Box_{\text{Bob}}\Box_{\text{Amy}}\bar t\)
+
 \end{examples}
 
-\exi{three}{b} is compiled in Fst into a finite state machine with --
-nodes and -- edges that accepts a countably infinite set of strings
-(worlds).  This is a concrete computational representation of the
-propositiion denoted by \exi{three}{a}.  This contrasts with standard
-computational approaches to intensional semantics, sentences are
-translated into logical formulas that have a mathematical
-interpretation as a set of worlds, but not a computational one.  The
-finite state representations can be used to check entailment between
-sentences (via a subset check), print random worlds that satisfy the
-sentence, or the like.
+\exi{three}{b} is compiled an Fst finite state machine with -- nodes and
+-- edges that accepts a countably infinite set of strings (worlds). Or similarly
+interpreted as a lazily computation of guarded strings.  These are concrete
+computational representations of the proposition denoted by \exi{three}{a}.
+This contrasts with standard computational approaches to intensional semantics,
+sentences are translated into logical formulas that have a mathematical
+interpretation as a set of worlds, but not a computational one.  Our
+representations can be used to check entailment between sentences (via a subset
+check), print random worlds that satisfy the sentence (Fst only), print all
+worlds shorter than some bound, and the like.
 

### 2020-09-22 tex/sdraft.tex
center & shrink figures
tighten figures

@@
 \usepackage{wasysym,graphicx}
+\usepackage{tcolorbox}
 
@@
 \begin{figure}
+  \centering
   \(\begin{array}{l >{\triangleq}c l}
@@
 \begin{figure}
+\footnotesize
 \begin{hangingpar}
@@
 \begin{figure} \label{relclosure}
+  \footnotesize
 {\raggedright
   \begin{hangingpar}
-{\tt{define RelKpl(R)}}
+{\tt{define RelKpl(R)}}\\
 $\underbrace{\mbox{\tt{Squash.i\hspace{1pt}.o.}}}_c$
@@
 {\tt{define Kst(R)
-    [St\hspace{1pt}.x.\hspace{1pt}St]\hspace{1pt}|\hspace{1pt}Kpl(X);}} The Fst
+    [St\hspace{1pt}.x.\hspace{1pt}St]\hspace{1pt}|\hspace{1pt}Kpl(X);}}\\ The Fst
 operation {\tt{.x.}} is Cartesian product.  {\tt{R.i}} is the inverse of relation
@@
 also Epik, can denote infinite sets of strings, for example the term
-$\_^*$. Normally regular languages are represented using a finite
+$\bullet^*$. Normally regular languages are represented using a finite
 coalgebra~\cite{}. However the non-distributivity of $\diamond$
@@
 represents terms with the same signature as described in Section 2. We
-then paramaterize the interpretation function, $\llparenthesis p
-\rrparenthesis^{\varphi,{\mathcal{E}}}_n$ on an integer $n$ that
+then paramaterize the interpretation function on an integer $n$ that
 describes the maximum length of a string we will produce.
@@
 
-We lift the remaining operators (except Kleene star) to their list equivalents:
-union becomes list append (written $\texttt{+\!+}$); concatenation becomes the
-fusion product (lifted to lists this time), negation is implemented using list
-difference ($\setminus$), and the modal operator lifts the alternative relation
-over lists of strings\footnote{Figure~\ref{fig:list-interp} depicts this using
-  the list comprehension notation, which is analogous to set builder notation,
-  except that it is written using square brackets. The order of elements is
-  preserved indicated by the keyword \texttt{for}}. The only caveat to these
-direct interpretations is that we restrict the operators to have size $\leq n$,
-denoted as $l|_n$ for a list of guarded strings $l$.
+We lift the remaining operators (except Kleene star) to their list
+equivalents: union becomes list append (written $\texttt{+\!+}$);
+concatenation becomes the fusion product (lifted to lists this time),
+negation is implemented using list difference ($\setminus$), and the
+modal operator lifts the alternative relation over lists of
+strings\footnote{Figure~\ref{fig:list-interp} depicts this using the
+  list comprehension notation, which is analogous to set builder
+  notation, except that it is written using square brackets. Element
+  order is evoked by the keyword \texttt{for}, rather than using the
+  unordered $\forall$.}. The only caveat to these direct
+interpretations is that we restrict the operators to have size $\leq
+n$, denoted as $l|_n$ for a list of guarded strings $l$.
 
@@
 
+\newcommand{\semlst}[2]{\llparenthesis#1\rrparenthesis^{\mathsf B, \mathsf P, \phi, \zeta}_{#2}}
+\newcommand{\setToList}[1]{\lfloor#1\rfloor}
+
 \begin{figure}
-  \[\begin{array}{r >{\triangleq}c l}
-  \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_0 && [] \\
-  \llparenthesis 0 \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [] \\
-  \llparenthesis 1 \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \lfloor \mathcal A_{\mathsf B}^\varphi \rfloor \\
-  \llparenthesis e \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \lfloor \hat{\mathcal E}^\varphi(e)\rfloor \\
-  \llparenthesis \psi \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \lfloor \mathcal A_{\mathsf B}^{\varphi~\psi} \rfloor \\
-  \llparenthesis p + q \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n \mathbin{\texttt{+\!+}} \llparenthesis q \rrparenthesis^{\varphi,{\mathcal{E}}}_n \\
-  \llparenthesis p;q \rrparenthesis^{\varphi,{\mathcal{E}}}_n &&
-  \left(\llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n \diamond \llparenthesis q\rrparenthesis^{\varphi,{\mathcal{E}}}_n\right)|_n\\
-  \llparenthesis p^* \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [] + \left(\llparenthesis p\rrparenthesis^{\varphi,{\mathcal{E}}}_n \diamond  \llparenthesis p^* \rrparenthesis^{\varphi,{\mathcal{E}}}_{m}\right)|_n \\
+  \centering
+  \(\begin{array}{r >{\triangleq}c l}
+  \semlst p 0 && [] \\
+  \semlst 0 n && [] \\
+  \semlst 1 n && \setToList \AphiB \\
+  \semlst e n && [ \alpha e \beta \mid \alpha \mathrel{\zeta_e} \beta ] \\
+  \semlst b n && \setToList{ \mathcal A_{\mathsf B}^{b\psi}} \\
+  \semlst{p + q} n && \semlst p n \mathbin{\texttt{+\!+}} \semlst q n \\
+  \semlst{p;q} n &&
+  \left(\semlst p n \diamond \semlst q n \right)|_n\\
+  \semlst{p^*} n && [] + \left(\semlst p n \diamond \semlst{p^*} {n-i}\right) |_n \\
   \multicolumn 3 r {\text{where}~
     \begin{array}[t]{r >{\hspace{-.7em}}l}
-      i &= \max\{1,\min\{|g|\mid g \in \llparenthesis p \rrparenthesis_n^{\varphi,{\mathcal{E}}}\}\}\\
-      m &= \max\{0, n - i\}
+      i &= \max\{1,\min\{|g|\mid g \in \semlst p n \}\} \\
     \end{array}}\\
-  \llparenthesis \neg p \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \llparenthesis \_^* \rrparenthesis^{\varphi,{\mathcal{E}}}_n \setminus \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n \\
-  \llparenthesis \Diamond_a~p \rrparenthesis^{\varphi,{\mathcal{E}}}_n &&
-  [g' \mid g'~\hat{R_a}~g,~\texttt{for}~g~\texttt{in}~\llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n]
+  \semlst{\neg p} n && \semlst {\bullet^*} n \setminus \semlst p n \\
+  \semlst{\Diamond_a~p } n &&
+         [g' \mid g'~\hat{R_a}~g,~\texttt{for}~g~\texttt{in}~\semlst p n]
   \end{array}
-  \]
+  \)
   \caption{Bounded interpretation using lazy lists}
@@
 
+\section{Conclusion}
+
+\ehc{Insert happy words!}
+
 \pagebreak

### 2020-09-22 tex/sdraft.tex
figure hacking

@@
 \usepackage{tcolorbox}
+\usepackage{booktabs}
 
@@
 \begin{figure}
-  \[\begin{array}{lcl}
+  \fbox{
+    \begin{minipage}{\columnwidth}
+  \(\begin{array}{lcl}
   \multicolumn 3 l {\textrm{state formulas}\hfill (a \in \mathsf{B}) \hfill \mbox{}}\\
-  \rho,\sigma,\varphi &::=& a \mid0 \mid 1 \mid\rho+\sigma \mid \rho\,\sigma \mid \bar{\rho} \\ \\
+  \rho,\sigma,\varphi &::=& a \mid0 \mid 1 \mid\rho+\sigma \mid \rho\,\sigma \mid \bar{\rho} \\
   \multicolumn {3} l {\textrm{effect formulas}}\\
@@
     \mid \zeta \mathbin{\&} \eta \mid \bar{\zeta}
-  \end{array}\]
-
-\[
+  \end{array}\)
+
+  \(
   \begin{array}{lll}
@@
   \end{array}  
-\]
-
-\vspace{-2mm}
-
+  \)
+  \end{minipage}}
 \caption{Syntax of state formulas and syntax and semantics of effect formulas. Effect formulas
@@
   state} enumerates ${\sf{B}}$. The line beginning with {\tt
-  constraint} gives the state formula.  The lines beginning with {\tt
+  restrict} gives the state formula.  The lines beginning with {\tt
   event} declare events and their effect formulas.  Finally the lines
@@
 \begin{figure}
+  \begin{tcolorbox}
     \footnotesize
+    \begin{minipage}{0.45\textwidth}
       \begin{center}    
@@
             state h t \\
-constraint h!t + t!h \\
+            restrict h!t + t!h \\
             event a1 h:h \\
@@
             event b0 t:t \\
+      \end{tabular}}
+      \end{center}
+    \end{minipage} \hfill 
+    \begin{minipage}{0.45\textwidth}
+      \begin{center}
+        {\tt
+          \begin{tabular}{l}
             agent aly \\
@@
       \end{center}
+    \end{minipage}
+  \end{tcolorbox}
 
@@
 \begin{figure}
-\begin{tabular}{ll}
-events & $e \in \mathsf{P}$ \\
-states & $\sigma$ as in Figure 1 \\    
-$p,q$ & $::= e~|~\sigma~|~p+q~|~pq~|~p^*~|~ \neg p ~|~ \Diamond_a\,p$\\
-\end{tabular}
+  \fbox{
+    \begin{minipage}{\columnwidth}
+      \(\begin{array}{ll}
+      \text{events} & e \in \mathsf{P} \\
+      \text{states} & \sigma~\text{as in Figure 1} \\    
+      p,q & ::= e\,|\,\sigma\,|\,p+q\,|\,pq\,|\,p^*\,|\,\neg p\,|\, \Diamond_ap \\
+      \end{array}\)
 
@@
       \end{tabular}
+  \end{minipage}}
 
@@
 \begin{figure}
+  \fbox{\begin{minipage}{\columnwidth}
   \centering
@@
   \hspace{2mm}\(\sem{ \Diamond_a\,p } \hspace{4mm}{\triangleq} \hspace{4mm}\{x \mid \exists y . x{\hat{R}_a}y \wedge y \in \sem{ p }^{\epikspec}\} \)
+  \end{minipage}}
   \caption{Interpretation of Epik terms as sets of guarded strings}
@@
 \begin{figure}
+  \begin{tcolorbox}[size=fbox]
 \footnotesize
 \begin{hangingpar}
-{\tt St} \hspace{2mm} Tests such as 0 1 1 0. The length is the number of generators.
+  {\tt St} \hspace{2mm} \\
+  Atomic Tests such as $0110$.% The length is the number of generators.
 \end{hangingpar}
@@
 \begin{hangingpar}
-{\tt UnequalStPair} \hspace{2mm} Sequence of two unequal tests such as 0 1 1 0 0 1 1 1,
-    differing in one or more positions.
+  {\tt UnequalStPair} \hspace{2mm} \\
+  Sequence of two unequal tests such as $0110~0111$.
 \end{hangingpar}
@@
 \end{hangingpar}
-
+\end{tcolorbox}
 \caption{\raggedright Definition in Fst of KAT product and KAT Kleene star.
@@
 \begin{figure} \label{relclosure}
+  \begin{tcolorbox}[size=fbox]
   \footnotesize  
@@
   \begin{hangingpar}
+%% {\tt{define RelKpl(R)}}\\
+%% $\underbrace{\mbox{\tt{Squash.i\hspace{1pt}.o.}}}_c$
+%% $\underbrace{\mbox{\tt Wf0\hspace{1pt}.o.}}_b$
+%% $\underbrace{\mbox{\tt{[R+]}}}_a$
+%% $\underbrace{\mbox{\tt .o.\hspace{1pt}Wf0}}_b$
+%% $\underbrace{\mbox{\tt{.o.\hspace{1pt}Squash}}}_c$
 {\tt{define RelKpl(R)}}\\
-$\underbrace{\mbox{\tt{Squash.i\hspace{1pt}.o.}}}_c$
-$\underbrace{\mbox{\tt Wf0\hspace{1pt}.o.}}_b$
-$\underbrace{\mbox{\tt{[R+]}}}_a$
-$\underbrace{\mbox{\tt .o.\hspace{1pt}Wf0}}_b$
-$\underbrace{\mbox{\tt{.o.\hspace{1pt}Squash}}}_c$
+$\!\underbrace{\texttt{Squash.i}\,\texttt{.o.}}_c
+\underbrace{\mbox{\tt Wf0\,.o.}}_b
+\underbrace{\mbox{\tt{[R+]}}}_a
+\underbrace{\mbox{\tt .o.\,Wf0}}_b
+\underbrace{\mbox{\tt{.o.\,Squash}}}_c$
+
 \end{hangingpar}}
@@
 \end{hangingpar}
-
+\end{tcolorbox}
 \caption{Definition in Fst of the Kleene concatenation closure of
@@
 
-\begin{figure}
-  \centering
-  \(\begin{array}{r >{\triangleq}c l}
+\begin{figure}[!t]
+  \fbox{
+    \begin{minipage}{\columnwidth}
+      \hspace{-0.9em}\(\begin{array}{r >{\triangleq}c l}
       \semlst p 0 && [] \\
@@
       \semlst b n && \setToList{ \mathcal A_{\mathsf B}^{b\psi}} \\
-  \semlst{p + q} n && \semlst p n \mathbin{\texttt{+\!+}} \semlst q n \\
+      \semlst{p\! + \!q} n && \semlst p n \mathbin{\texttt{+\!+}} \semlst q n \\
       \semlst{p;q} n &&
-  \left(\semlst p n \diamond \semlst q n \right)|_n\\
-  \semlst{p^*} n && [] + \left(\semlst p n \diamond \semlst{p^*} {n-i}\right) |_n \\
-  \multicolumn 3 r {\text{where}~
+      (\semlst p n \diamond \semlst q n)|_n\\
+      \semlst{p^*} n && [] + (\semlst p n \diamond \semlst{p^*} {n-i}) |_n \\
+      \multicolumn 3 r {\footnotesize \text{where}~
         \begin{array}[t]{r >{\hspace{-.7em}}l}
@@
       \semlst{\neg p} n && \semlst {\bullet^*} n \setminus \semlst p n \\
-  \semlst{\Diamond_a~p } n &&
-         [g' \mid g'~\hat{R_a}~g,~\texttt{for}~g~\texttt{in}~\semlst p n]
+      \semlst{\Diamond_ap } n &&
+             [g' \mid g'\,\hat{R_a}\,g,\texttt{\footnotesize for}\,g\,\texttt{\footnotesize in}\,\semlst p n]
       \end{array}
       \)
+    \end{minipage}}
   \caption{Bounded interpretation using lazy lists}
@@
 
-\begin{figure}
+\begin{figure}[!t]
+  \vspace{1em}
+  \centering
   \footnotesize
-\begin{tabular}[t]{l|c|l}
+  \begin{tabular}[t]{lcc}
+    \toprule
+        \textsc{Item} & \textsc{Type} & \textsc{Semantics}\\ \midrule  
         Amy & $e$ & ${R}_a$\\
@@
         that  & $(((e \backslash t){\slash_M}t) \backslash (e \backslash t)) \slash t$
-& $\lambda p~m~R.{\neg}(m\;({\neg}p)\;R)$ \\
+        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
         whether  & $(((e \backslash t){\slash_M} t) \backslash (e \backslash t)) \slash t$ 
-& $\lambda p~m~R.{\neg}(m\;({\neg}p)\;R)$ \\
-&& \multicolumn{1}{r}{$+ {\neg}(m\;p\;R)$} \\
+        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
+        && \multicolumn{1}{r}{$+ {\neg}(m\,p\,R)$} \\ \bottomrule
       \end{tabular}
@@
 \bibliography{anthology,acl2020}
-\bibliographystyle{acl_natbib}
+\bibliographystyle{chicago}%{acl_natbib}
 

### 2020-09-23 tex/sdraft.tex
add change of state

@@
 \usepackage{latexsym}
+\usepackage{enumitem}
 %\usepackage{caption}
@@
   Coin}.  Amy and Bob are seated at a table.  There is a coin on the
-table under a cup, heads up.  The coin could be heads or tails, and
-neither agent knows which it is.  This initial situation is possible
-world $w_1$.  Two additional worlds $w_2$ and $w_3$ are defined by
-sequencing events after the initial state, with events interpreted as
-in \ex{coin1}. The truth values for English sentences shown in
-\ex{coin3} are observed.
-
-\begin{example}  \label{coin1}
-\hspace{-4mm} \begin{tabular}[t]{ll}
-  $a_1$  & Amy peeks at heads, by tipping the
-  \\ & cup. Bob sees she's peeking, but not\\
-  & what she sees. \\
-  $b_1$  & Bob peeks at heads. \\
-  $a_0$  & Amy peeks at tails. \\
-  $b_0$  & Bob peeks at tails. \\     
+table under a cup, heads up.  The coin could be heads-up $H$ or
+tails-up $T$, and neither agent knows which it is.  This initial
+situation is possible world $w_1$.  Two additional worlds $w_2$ and
+$w_3$ are defined by sequencing events after the initial state, with
+events interpreted as in \ex{coin1}. The truth values for English
+sentences shown in \ex{coin3} are observed, where $0$ stands for
+falsity and $1$ for truth.
+
+\newcommand{\apH}{\ensuremath{a_1}}
+\newcommand{\apT}{\ensuremath{a_0}}
+\newcommand{\bpH}{\ensuremath{b_1}}
+\newcommand{\bpT}{\ensuremath{b_0}}
+
+\newcommand{\atTH}{\ensuremath{a_{01}}}
+\newcommand{\atHT}{\ensuremath{a_{10}}}
+\newcommand{\btTH}{\ensuremath{b_{01}}}
+\newcommand{\btHT}{\ensuremath{b_{10}}}
+
+\begin{example}  \label{coin1} \begin{enumerate}[align=left,leftmargin=0.75em]
+  \item[\apH] Amy peeks at $H$, by tipping the cup. Bob sees she's peeking, but not what she sees.
+  \item[\bpH] Bob peeks at $H$.
+  \item[\apT] Amy peeks at $T$.
+  \item[\bpT] Bob peeks at $T$. 
+  \item[\atTH] Amy secretly turns the coin from $T$ to $H$.
+    She knows she turned the coin over, but not which side was face up. Bob thinks nothing happened.
+  \item[\atHT] Amy secretly turns the coin from $H$ to $T$.
+  \item[\atTH] Bob secretly turns the coin from $T$ to $H$.
+  \item[\btHT] Bob secretly turns the coin from $H$ to $T$.
+\end{enumerate}
+\end{example}
+
+\begin{example} \label{coin2}
+  \(w_2 = w_1\apH \hfill
+  w_3 = w_2\bpH \hfill
+  w_4 = w_3\atHT\btTH\bpH\)
+\end{example}
+
+\begin{example} \label{coin3}
+\hspace{-7mm} \begin{tabular}[t]{ccccl}
+  $w_1$ & $w_2$  & $w_3$ & $w_4$ & Sentence\\
+  $0$ & $1$ & $1$& $0$ & Amy knows it's heads. \\
+  $0$ & $0$ & $1$& $1$ & Bob knows it's heads. \\
+  $0$ & $0$ & $1$& $0$ & Bob knows Amy \\ &&&& knows it's heads. \\
+  $0$ & $1$ & $1$& $0$ & Bob knows Amy \\ &&&& knows  whether it's \\ &&&& heads or tails. \\
 \end{tabular}
@@
 
-\begin{example} \label{coin2}
-   $w_2 = w_1 \hspace{2pt} a_1 $ \hspace{2mm}
-   $w_3 = w_1 \hspace{2pt} a_1 \hspace{2pt} b_1 $ 
-\end{example}
-
-\begin{example} \label{coin3}
-\hspace{-9mm} \begin{tabular}[t]{cccl}
-  $w_1$ & $w_2$  & $w_3$ & Sentence\\
-  false & true & true& Amy knows that it's \\ &&& heads. \\
-  false & false & true& Bob knows thats it's \\&&&heads. \\
-  false & false & true& Bob knows Amy \\ &&& knows it's heads. \\
-  false & true & true& Bob knows Amy \\ &&& knows  whether it's \\ &&& heads or tails. \\
-\end{tabular}
-\end{example}
-
-\newcommand{\ev}[2]{#1 \!\! : \!\! #2}
-\newcommand{\evr}[2]{#2 \!\! : \!\! #1}
-
-The events come with pre-conditions.
-Amy can peek at heads only if the coin is heads up, so $a_1$ has the precondition of
-the coin being heads up.  Let $h$ be the Boolean proposition that the
-coin is heads up.
-Then preconditions
-can be described by Boolean formulas, with $h$ being the precondition
-of $a_1$
-$u$. Events come as well with a relation between
-prior and following state, for instance with $u$ incrementing the
-floor.  This is expressed using an operator ``$:$'' (read ``and
-next'') that pairs Boolean formulas. The formula $h:h$
-describes $a_1$ (Amy looking at heads) as happening only in an $h$
-state, and as not changing the state.  Symmetrically, $a_0$ (Amy
-looking at tails) can happen only in a not-$h$ state, and does not
-change the state ($\ev{\bar{h}}{\bar{h}}$).
+\ehc{I'm not sure if these sentences are the most interesting anymore!}
+
+\newcommand{\ev}[2]{#1 \mathbin{:} #2}
+\newcommand{\evr}[2]{\ev{#2}{#1}}
+
+The events come with pre-conditions and post-conditions.  Amy can turn
+the coin from heads to tails only if the coin is heads-up, so $\atHT$
+has the pre-condition of the coin being heads up. Once she turns the
+coin over, tails must be face-up, so $a_{10}$ has the post-condition
+of the coin being tails-up. Let $h$ be the Boolean proposition that
+the coin is heads up and $t$ be the Boolean proposition that the coin
+is tails-up.  Then pre- and post-conditions can be described by
+Boolean formulas, with $h$ being the pre-condition of $\atHT$ and
+$a_{01}$ being the post-condition. This is expressed using an operator
+``$:$'' (read ``and next'') that pairs Boolean formulas. The formula
+$\ev h t$ describes $\atHT$ (Amy turning the coin from heads
+to tails) as happening only in an $h$ state, and concluding in a
+not-$h$ state.  Events don't have to change state: the event $\apT$
+(Amy peeking at tails) can happen only in a $t$ state, and does
+not change the state ($\ev{t}{t}$).
 
@@
 
-\section{Epistemic guarded string models}
-
-Suppose that in the coin example, we have an additional primitive
-stative proposition (or atomic test) $t$, interpreted as tails.  A
-sequence such as $\bar{h}{t}$ can be viewed as a valuation of
-primitive propositions, which is used to describe world state. The
-primitives are listed in fixed order, and left unmarked (indicating
-true) or marked with the overbar (indicating false).  Since a coin is
-heads or tails but not both, we want to allow the valuations
-$h\bar{t}$ and $\bar{h}t$, and disallow $ht$ and $\bar{h}\bar{t}$.
-This is enforced by a {\em state formula}, which is a Boolean formula,
-in this case the one given on the second line of \ex{effect}.  Where
-{\sf{B}} is a set of atomic tests and $\phi$ is a state constraint
-over {\sf{B}}, ${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\phi}$ is the set of
-valuations of {\sf{B}} that make formula $\phi$ true. Valuations are
-called atoms, because they correspond to the atoms of a Boolean
-algebra of tests \citep{kozen2001automata}.
+However a coin cannot be showing both heads and tails! Currently the
+precondition $h$ of $\atHT$ only says that heads must be showing, and
+says nothing about the fact that tails must be face-down, indicated by
+the formula $\bar t$. We will further restrict the feasible conditions
+for our actions by restricting the space of valuations for our formulae.
+
+A sequence such as $\bar{h}{t}$ can be viewed both as a formula and as
+a valuation of primitive propositions, which we use to describe world
+state. The primitives are listed in fixed order, and left unmarked
+(indicating true) or marked with the overbar (indicating false).
+Since a coin is heads or tails but not both, we want to allow the
+valuations $h\bar{t}$ and $\bar{h}t$, and disallow $ht$ and
+$\bar{h}\bar{t}$.  This is enforced by a {\em state formula}, which is
+a Boolean formula, in this case the one given on the second line of
+\ex{effect}.  Where {\sf{B}} is a set of atomic tests and $\phi$ is a
+state constraint over {\sf{B}},
+${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\phi}$ is the set of valuations of
+{\sf{B}} that make formula $\phi$ true. Valuations are called atoms,
+because they correspond to the atoms of a Boolean algebra of tests
+\citep{kozen2001automata}.
 
@@
 
+\section{Epistemic guarded string models}
+
 Epik is a specification language for possible worlds models that
@@
 semi-privately looking at the coin. The line beginning with {\tt
-  state} enumerates ${\sf{B}}$. The line beginning with {\tt
-  restrict} gives the state formula.  The lines beginning with {\tt
-  event} declare events and their effect formulas.  Finally the lines
-beginning with {\tt agent} define {\em event alternative} relations
-for agents.  Each clause with an arrow has a single event symbol on
-the left, and a disjunction of alternative events on the right of the
-arrow. The interpretation of Amy's alternatives for $b_1$ (Bill peeks
-at heads), is that when $b_1$ happens, for Amy either $b_1$ or $b_0$
-(Bill peeks at tails) could be happening.
+  state} enumerates ${\sf{B}}$. The line beginning with {\tt restrict}
+gives the state formula.  The lines beginning with {\tt event} declare
+events and their effect formulas.  Finally the lines beginning with
+{\tt agent} define {\em event alternative} relations for agents.  Each
+clause with an arrow has a single event symbol on the left, and a
+disjunction of alternative events on the right of the arrow. The
+interpretation of Amy's alternatives for $b_1$ (Bob peeks at heads),
+is that when $b_1$ happens, for Amy either $b_1$ or $b_0$ (Bob peeks
+at tails) could be happening. Her alternatives for $a_{01}$ and
+$a_{10}$ (she turns the coin over) are $a_{10}$ and $a_{01}$,
+indicating that she doesn't know, \emph{a priori}, whether she's
+turning the coin from $H$ to $T$ or from $T$ to $H$. Similarly, Bob
+secretly turns the coin over, in $b_{10}$ or $b_{01}$, she doesn't
+know anything has happened, so her alternatives are the no-operation
+events $o_1$ and $o_0$ for heads-worlds and tails-worlds
+respectively. Bob's alternatives are the same, \emph{mutatis mutandi}.
 
@@
     \footnotesize
-    \begin{minipage}{0.45\textwidth}
+    \begin{minipage}{0.48\textwidth}
       \begin{center}    
@@
             state h t \\
-            restrict h!t + t!h \\
+            restrict  h!t \\
+            \phantom{restrict }+ t!h \\
+            event o1  h:h \\
+            event o0  t:t \\
             event a1  h:h \\
@@
             event b0  t:t \\
+            event a10 h:t \\
+            event a01 t:h \\
+            event b10 h:t \\
+            event b01 t:h \\
       \end{tabular}}
@@
     \end{minipage} \hfill 
-    \begin{minipage}{0.45\textwidth}
+    \begin{minipage}{0.51\textwidth}
       \begin{center}
@@
             agent aly \\
+            \hspace{3mm} o1  -> o1 \\
+            \hspace{3mm} o0  -> o0 \\
             \hspace{3mm} a1  -> a1 \\
@@
             \hspace{3mm} b0  -> b1 + b0\\
+            \hspace{3mm} a10 -> a10 + a01 \\
+            \hspace{3mm} a01 -> a10 + a01 \\
+            \hspace{3mm} b10 -> o0 + o1 \\
+            \hspace{3mm} b01 -> o0 + o1 \\
             agent bob \\
-            \hspace{3mm} b1 -> b1 \\
-            \hspace{3mm} b0 -> b0 \\
-            \hspace{3mm} a1 -> a1 + a0\\
-            \hspace{3mm} a0 -> a1 + a0\\
+            \hspace{3mm} <\textrm{\textit{sim.} swap }a\textrm{ and }b>
+            %% agent bob \\
+            %% \hspace{3mm} b1 -> b1 \\
+            %% \hspace{3mm} b0 -> b0 \\
+            %% \hspace{3mm} a1 -> a1 + a0\\
+            %% \hspace{3mm} a0 -> a1 + a0\\
         \end{tabular} }
@@
 
-\caption{Epik program describing a possible-worlds event sequence model for two agents
-  with information about one coin, and events of the agents semi-privately looking
-  at the coin.}
+\caption{Epik program describing a possible-worlds event sequence
+  model for two agents with information about one coin, and events of
+  the agents semi-privately looking at the coin. }
 \label{fig:epik-example}
@@
 from ${\sf{P}}$.  In the algebra described by Figure \ref{epikfigure}, 
-${\sf{P}}$ is the set of events $\{a_1,a_0,b_1,b_0\}$, and
+${\sf{P}}$ is the set of events $\{a_1,a_0,b_1,b_0,a_{10},a_{01},b_{10},b_{01}\}$, and
 ${\sf{B}}$ is $\{h,t\}$.
@@
 
-\hspace{-2mm}\(H \; \diamond \; T \,a_1\, T = \textit{undefined} \)
+\hspace{-2mm}\(T\,b_{01}\,H \; \diamond \; T \,a_1\, T = \textit{undefined} \)
 \end{example}
@@
 conjoining $h$ with the atom at the end of the world.  So $\bullet^* h$ is the set
-of worlds where the coin is heads.\footnote{A mapping from English sentences to logical
+of worlds where the coin ends heads-up\footnote{A mapping from English sentences to logical
   terms in epistemic KAT is presented in Section 6}.
@@
 \begin{tabular}[t]{l}
-  $\Box_b(\Box_a\bullet^*t + \Box_a\bullet^*\neg t)$\\
+  $\Box_b(\Box_a\bullet^*t + \Box_a\neg\bullet^*t)$\\
  Bob knows that Amy knows whether \\ \hspace{3mm} it's tails.  \\
@@
 %% \neg \Diamond_{\text{Amy}} \neg \neg \neg \Diamond_{\text{Bob}} \neg \neg \Diamond_{\text{Amy}} \bar t\)
-\itemb \(\displaystyle \Box_{\text{Amy}}\Box_{\text{Bob}}(\Box_{\text{Amy}} \bar h + \Box_{\text{Amy}}\bar t)\) \hfill \mbox{} \(\cdot~~\Box_{\text{Amy}}\neg\Box_{\text{Bob}}\Box_{\text{Amy}}\bar t\)
+\itemb \(\displaystyle \Box_{\text{Amy}}\Box_{\text{Bob}}(\Box_{\text{Amy}} \neg \bullet^* h + \Box_{\text{Amy}}\neg \bullet^* t)\) \hfill \mbox{} \(~~\cdot~~\Box_{\text{Amy}}\neg\Box_{\text{Bob}}\Box_{\text{Amy}}\neg \bullet^* t\)
 
@@
 \bibliography{anthology,acl2020}
-\bibliographystyle{chicago}%{acl_natbib}
+\bibliographystyle{acl_natbib}
 

### 2020-09-25 tex/sdraft.tex
add modal axioms

@@
 
-\newcommand{\ehc}[1]{\textit{\textcolor{green}{[ehc]: #1}}}
+\newcommand{\ehc}[1]{\textit{\textcolor{red}{[ehc]: #1}}}
 \newcommand{\mr}[1]{\textit{\textcolor{cyan}{[mr]: #1}}}
+
+\newcommand{\K}{\mathcal{K}}
+\newcommand{\B}{\mathcal{B}}
 
@@
           \begin{tabular}{l}
-            agent aly \\
+            agent amy \\
             \hspace{3mm} o1  -> o1 \\
@@
     \begin{minipage}{\columnwidth}
-      \(\begin{array}{ll}
-      \text{events} & e \in \mathsf{P} \\
-      \text{states} & \sigma~\text{as in Figure 1} \\    
-      p,q & ::= e\,|\,\sigma\,|\,p+q\,|\,pq\,|\,p^*\,|\,\neg p\,|\, \Diamond_ap \\
-      \end{array}\)
-
-      \vspace{1mm}
-      \begin{tabular}{cl}
-        $\Box_a p$ & $\triangleq \neg\Diamond_a\neg p$ \\
-        \(p \wedge q\) &  \(\triangleq \neg(\neg p + \neg q)\) \\
-        $\bullet$ & \( \triangleq \sum_{e \in \mathsf P}e\)
-      \end{tabular}
+      \(\begin{array}{llll}
+      \multicolumn{4}{l}{\text{events}~~e\in \mathsf{P}} \\
+      p,q &\multicolumn{3}{l}{::= e\,|\,\sigma\,|\,p+q\,|\,pq\,|\,p^*\,|\,\neg p\,|\, \Diamond_ap} \\
+      \Box_a p & \triangleq \neg\Diamond_a\neg p &
+      \bullet & \triangleq \sum_{e \in \mathsf P}e \\      
+      p \wedge q &  \triangleq \neg(\neg p + \neg q) &
+      p \to q & \neg p + q      
+      \end{array} \)
   \end{minipage}}
@@
 \begin{example}   \label{interpretation1}
-\begin{tabular}[t]{ll}
+\hspace{-1.5em}\begin{tabular}[t]{ll}
 ${\bullet}^*t$ & It's tails. \\
 $\bullet^*h$ & It's heads. \\  
-$\Box_a \bullet^* h$  & Amy knows that it's heads. \\
+  $ \bullet^* h \wedge \Box_a \bullet^* h$  & Amy knows it's heads. \\
 \end{tabular}
 \vspace{1mm}
-
-\begin{tabular}[t]{l}
+\hspace{-1.5em}\begin{tabular}[t]{l}
   $\Box_b(\Box_a\bullet^*t + \Box_a\neg\bullet^*t)$\\
- Bob knows that Amy knows whether \\ \hspace{3mm} it's tails.  \\
+ Bob believes Amy knows whether it's tails.  \\
 \end{tabular}
 \end{example}
-
+\ehc{I took the `that's out of these sentences. I remember from class
+  that even if they are assigned null in the PF they still exist in
+  the LF. Is it okay to leave it out here?}
+
+\subsection{Standard Epistemic Modalities}
+
+Using our existential modal primitive $\Diamond_a$, and the dual
+encodings of $\Box_a$ and $\wedge$, we can encode the standard modal
+operators expressing knowledge $(\K_a)$ and belief $(\B_a)$ into
+Epik\footnote{The choice to refer to $\Diamond_a$ as an ``epistemic''
+  primitive rather than a ``doxastic'' primitive, is that it is
+  essential in constructing both the doxastic and epistemic
+  operators. A ``doxastic'' moniker would suggest the
+  unrepresentability of epistemic modalities}. These encodings are
+shown below:
+
+\begin{example}\label{encodings}
+  \( \begin{array}[t]{ccl}
+    \textsc{Belief} & \B_a~p &\triangleq \Box_a~p \\
+    \textsc{Knowledge} & \K_a~p & \triangleq p \wedge \B_a~p
+    \end{array}\)
+\end{example}
+
+%% As a consequence of these encodings, the standard dualities
+%% hold. These are shown in \ex{modalprops}.
+%% \begin{example} \label{modalprops}
+%%   \(\K_a~p \to \B_a~p \qquad \B_a~p \to \neg\K_a\neg p \)
+%% \end{example}
+
+To construct different types of reasoners (e.g. accurate, inaccurate,
+etc), users of Epik can construct an agents alternative relation
+accordingly. Observe the discussion of modal axioms \textbf{T} and
+\textbf{D} in Section~\ref{sec:axioms}.
+
+\subsection{Guarded String Interpretation}
 \newcommand{\epikspec}{\test,\event,\varphi,\zeta}
@@
 
-% $\tuple{K,+,\cdot,*,\bar{},0,1,{}^c,\Diamond_1,...,\Diamond_n}$
+\subsection{Axiomatic Classification}
+\label{sec:axioms}
+
+To situate our logic as a modal logic, we investigate the soundness of
+the standard modal axioms~\cite{????} in Epik's semantics. Some of
+these standard axioms (\ex{valid-axioms}) hold all the time, i.e. they
+are \emph{valid}, and the remaining axioms are valid when $\hat{R}_a$
+has a certain shape. The axioms in \ex{conditionally-valid-axioms}
+have a nontrivial condition validity condition.
+
+\begin{example}\label{valid-axioms}
+  \begin{enumerate}[align=left]
+  \item[\textbf{N}] If $p$ is valid, then $\Box_ap$ is valid
+  \item[\textbf{K}] $\Box_a(p \to q) \to \Box_a~p \to \Box_a~q$ is valid.
+  \end{enumerate}
+\end{example}
+
+\begin{example}\label{conditionally-valid-axioms}
+  \(\begin{array}[t]{lll}
+  \textbf{T} & \Box_a p \to p & \text{if } g\mathrel{\hat{R}_a} g, \forall g \\
+  \textbf{D} & \Box_a p \to \Diamond_a p & \text{if } g \in \dom(\hat{R}_a), \forall g \\
+  \textbf{4} & \Box_a p \to \Box_a\Box_a p & \text{if } \hat{R}_a \text{ idempotent}
+  \end{array}\)
+\end{example}
+
+The condition on \textbf{4} is essentially just a restatement of the
+theorem in relational terms. The conditions on the remaining standard
+modal axioms, \textbf{B} and \textbf{5}, are of a similarly trivial
+flavor.
 
@@
 sequences (sequences of 0's and 1's) are used for atoms, and these alternate with
-event symbols in the encoding of a world. \ex{gs1fst} gives the encoding of worlds
+event symbols in the encoding of a world. \ex{gs1fst} gives the encoding of worlds $w_2$ and $w_3$
 from the example.  A string is a finite sequence of symbols, and {\tt 0},
@@
 \end{tcolorbox}
-\caption{\raggedright Definition in Fst of KAT product and KAT Kleene star.
-  Where {\tt X} and {\tt Y} are regular sets and
-  {\tt R} and {\tt S} are regular relatiions,
-  {\tt X{\&}Y} is the intersection of {\tt X} and {\tt Y},
-  {\tt X|Y} is the union of {\tt X} and {\tt Y},
-  {\tt \~{}X} is the complement of {\tt X},
-  {\tt R{.o.}S} is the composition of {\tt R} and {\tt S},
-  {\tt R{.}l} is the co-domain of {\tt R}, and
-  {\tt {\$}X} is the set of strings that have a substring in $X$.   
+\caption{\raggedright Definition in Fst of KAT product and star.
+  %% Where {\tt X} and {\tt Y} are regular sets and
+  %% {\tt R} and {\tt S} are regular relatiions,
+  %% {\tt X{\&}Y} is the intersection of {\tt X} and {\tt Y},
+  %% {\tt X|Y} is the union of {\tt X} and {\tt Y},
+  %% {\tt \~{}X} is the complement of {\tt X},
+  %% {\tt R{.o.}S} is the composition of {\tt R} and {\tt S},
+  %% {\tt R{.}l} is the co-domain of {\tt R}, and
+  %% {\tt {\$}X} is the set of strings that have a substring in $X$.   
 }
@@
   \renewcommand{\arraystretch}{1.2}
-Worlds coded as strings
-\vspace{1mm}
-
- \hspace{-13mm} \begin{tabular}[t]{cl}
-{\textit{World}} & {\textit{String}} \\
+\begin{tabular}[t]{cl}
+%% {\textit{World}} & {\textit{String}} \\
 % $v_1$ & $ 0 \,u\,  1 \,u\,  2 \,d\,  1 \,u\,  2$ \\
-$w_1$ & {\tt 1 0} \\
+%% $w_1$ & {\tt 1 0} \\
 $w_2$ & {\tt 1 0 a1 1 0} \\
@@
 
-English sentences are mapped to terms in the logical language via an interpreted
-grammar, and these terms are in turn interpreted as propositions (sets of
-possible worlds).  The grammar is a semantically interpreted multi-modal
-categorial grammar, consisting of a lexicon of words, their categorial types,
-and interpretations in a logical lambda language.  The grammar covers basic
-statives (it's heads), that- and whether-complements of {\em know}, sentence
-negation, and predicate and sentence conjunction.  Figure \ref{lexiconfigure}
-gives illustrative lexical entries.\footnote{Category symbols use
-  Lambek/Bar-Hillel notation for slashes, so that $(d \backslash t)\slash(d
-  \backslash_D t)$ combines with $d \backslash_D t$ on the right to give a value
-  that combines with $d$ on the left to give $t$.}\footnote{Lambda abstractions
-  with multiple parameters are written $\lambda x~y.~e$ rather than the more verbose
-  $\lambda x.\lambda y.e$.} The grammar and semantics are optimized for a simple
-fragment of English concerned with clausal complementation.  The agent names
-{\em Amy} and {\em Bob} contribute the epistemic alternative relations for those
-agents, rather than individuals.  This is possible because the agents are never
-arguments of extensional predicates.  The root verb {\em know} contributes
-existential modal force.  The complementizers {\em that} and {\em whether} are
-the heads of their dominating clauses, and assemble an alternative relation,
+English sentences are mapped to terms in the logical language via a
+semantically interpreted multi-modal categorial grammar, consisting of
+a lexicon of words, their categorial types, and interpretations in a
+logical lambda language.  The grammar covers basic statives ({\em it's
+  heads}), {\em that-} and {\em whether}-complements of {\em know}, predicate and
+sentence negation, and predicate and sentence conjunction.  Figure
+\ref{lexiconfigure} gives illustrative lexical
+entries.\footnote{Category symbols use Lambek/Bar-Hillel notation for
+  slashes, so that $(d \backslash t)\slash(d \backslash_D t)$ combines
+  with $d \backslash_D t$ on the right to give a value that combines
+  with $d$ on the left to give $t$.}\footnote{Lambda abstractions with
+  multiple parameters are written $\lambda x~y.~e$ rather than the
+  more verbose $\lambda x.\lambda y.e$.} The grammar and semantics are
+optimized for a simple fragment of English concerned with clausal
+complementation.  The agent names {\em Amy} and {\em Bob} contribute
+the epistemic alternative relations for those agents, rather than
+individuals.  The root verb {\em know} contributes existential modal
+force.  The complementizers {\em that} and {\em whether} are the heads
+of their dominating clauses, and assemble an alternative relation,
 modal force, and proposition contributed by the complement.  These
-complementizers introduce the dual via two negations, in order to express
-universal modal force.
+complementizers introduce the dual via two negations, in order to
+express universal modal force.
 
@@
         is & $(d \backslash t)\slash(d \backslash_D t)$ & $\lambda P~x.~P~x$ \\
-        knows & $(e \backslash t)\slash_M t$ &  $\lambda p~R.~{\Diamond}_Rp$\\
+        knows & $(e \backslash t)\slash_M t$ &  $\lambda p~R.~p\vee{\Diamond}_Rp$\\
+        believes & $(e \backslash t)\slash_M t$ &  $\lambda p~R.~{\Diamond}_Rp$\\
         that  & $(((e \backslash t){\slash_M}t) \backslash (e \backslash t)) \slash t$
@@
 The semantic translations in the third column of Figure \ref{lexiconfigure} use
-the Epik term language, incremented with lambda.  The body of $\lambda
+the logical language, incremented with lambda.  The body of $\lambda
 x.\bullet^*h$, which is the semantic lexical entry for {\em heads}, is a
@@
 The body of $\lambda p.\lambda R. {\Diamond}_R p$, which is the semantic lexical
-entry of {\em knows}, is an Epik term denoting the pre-image of the
+entry of {\em knows}, is an term denoting the pre-image of the
 world-alternative relation contributed by the subject.  This is not the right
@@
 
-Sentences are parsed with a chart parser for categorial grammar.  The semantics
-for complex phrases are obtained by syntactic application of semantic
-translations, accompanied by beta reduction.  In consequence, the semantic term
-translating a sentence is a term of the logical language language.  Such a term
-designates a set of possible words (guarded strings) in the possible worlds
-model deterimened by an Epik specification.  By way of example, \exi{three}{a}
-is an English sentence with predicate conjunction and three levels of clausal
-embedding.  Using the grammar and parser, the sentence is mapped to the Epik
-term in \exi{three}{b}.
+Sentences are parsed with a chart parser for categorial grammar.  The
+semantics for complex phrases are obtained by application of semantic
+translations, accompanied by beta reductions that eliminate all
+lambdas in logical forms for clauses.  In consequence, the semantic
+term translating a sentence is a term of the logical language.  Such a
+term designates a set of possible words (guarded strings) in the
+possible worlds model determined by an Epik specification.  By way of
+example, \exi{three}{a} is an English sentence with conjunction and
+several levels of clausal embedding.  Using the grammar and parser,
+the sentence is mapped to the term in \exi{three}{b}. Then
+\exi{three}{c} shows a simplified logical from constructed form
+\exi{three}{b} using simply logical equivalences. The negations are
+propositional. The double negation, while logically redundant, comes
+syntactically from the grammar. This term is compiled in an
+implementation of the finite state calculus to a finite state machine
+with six states and eight arcs, which accepts a countably infinite set
+of worlds.  In this way the methodology ``directly'' represents the
+set of worlds denoted by \exi{three}{a}.
+
+\newcommand{\HEADS}{\bullet^*h}
+\newcommand{\AKWH}{\neg (\neg \HEADS \vee \Diamond_{a} \neg\HEADS) + \neg(\HEADS \wedge \Diamond_{a}\HEADS)}
+\newcommand{\BKAKWH}{\neg(\neg(\AKWH) + \Diamond_{b}\neg(\AKWH))}
+\newcommand{\AKBKAKWH}{\neg(\neg(\BKAKWH) + \Diamond_{a}\neg(\BKAKWH))}
 
 \begin{examples} \label{three}
-\itema Amy knows that Bob knows that Amy knows whether it is heads and knows that Bob doesnt know that Amy knows that it is tails.
-%% \itemb {\tt (And (Not (Diamond amy (Not (Not (Diamond bob (Not (Or (Not (Diamond amy (Not heads))) (Not (Diamond amy heads))))))))) (Not (Diamond amy (Not (Not (Not (Diamond bob (Not (Not (Diamond amy (Not tails))))))))))) }
-%% \itemc
-%% \(\neg \Diamond_{\text{Amy}} \neg \neg \Diamond_{\text{Bob}} \neg \neg (\Diamond_{\text{Amy}}\bar h + \neg \Diamond_{\text{amy}} h) \cdot
-%% \neg \Diamond_{\text{Amy}} \neg \neg \neg \Diamond_{\text{Bob}} \neg \neg \Diamond_{\text{Amy}} \bar t\)
-\itemb \(\displaystyle \Box_{\text{Amy}}\Box_{\text{Bob}}(\Box_{\text{Amy}} \neg \bullet^* h + \Box_{\text{Amy}}\neg \bullet^* t)\) \hfill \mbox{} \(~~\cdot~~\Box_{\text{Amy}}\neg\Box_{\text{Bob}}\Box_{\text{Amy}}\neg \bullet^* t\)
+\itema It's tails and Amy knows that Bob knows that Amy knows whether it's heads.
+% (_*;t) & ~ <amy> ~ ~ <bob> ~((~ (<amy> ~ (_*;h))) + ~ (<amy> (_*;h)) )
+\itemb \( \displaystyle \bullet^*t \wedge \AKBKAKWH \)\\
+\itemc \(\bullet^*t \wedge \K_{a}(\K_{b}(\K_{a} \bullet^* h + \K_{a} \neg \bullet ^* h))\)
+\itemd Amy knows that it's tails.
+\iteme \(
+\bullet ^*t \wedge \neg \Diamond_{a} \neg \bullet^*t~~~~~~(\equiv \K_a \bullet^* t)
+\)
 
 \end{examples}
-
-\exi{three}{b} is compiled an Fst finite state machine with -- nodes and
--- edges that accepts a countably infinite set of strings (worlds). Or similarly
-interpreted as a lazily computation of guarded strings.  These are concrete
-computational representations of the proposition denoted by \exi{three}{a}.
-This contrasts with standard computational approaches to intensional semantics,
-sentences are translated into logical formulas that have a mathematical
-interpretation as a set of worlds, but not a computational one.  Our
-representations can be used to check entailment between sentences (via a subset
-check), print random worlds that satisfy the sentence (Fst only), print all
-worlds shorter than some bound, and the like.
-
-\section{Conclusion}
-
-\ehc{Insert happy words!}
-
-\pagebreak
+Sentence \exi{three}{d} is assigned the logical form \exi{three}{e} by
+the grammar.  Logical relations between propositions are checked in
+the finite state calculus by checking set-theoretic relations between
+sets of worlds.  In this case, the propositions \exi{three}{b} and
+\exi{three}{e} are equivalent, in the sense that they denote the same
+set of worlds.  This equivalence is attested computationally in the
+possible worlds model.
+
+%% \begin{examples} \label{entail}
+%% \itema Amy knows that it's tails.
+%% \itemb \(
+%% \neg \Diamond_{a} \neg \bullet^*t
+%% \)
+%% % \itemc \(\displaystyle \Box_{a}\Box_{b}(\Box_{a} \neg \bullet^* h + \Box_{a}\neg \bullet^* t)\) \hfill \mbox{} \(~~\cdot~~\Box_{a}\neg\Box_{b}\Box_{a}\neg \bullet^* t\)
+%% \end{examples}
+
+\section{Discussion}
+The methodology presented here is designed for use in research in
+linguistic semantics, and for education at the level of a second
+graduate course in formal semantics, covering intensionality.  There
+are straighforward extensions to additional semantic phenomena, such
+as tense and perfective aspect as in \exi{extension}{a}, and the combination of
+metaphysical modality and prospective aspect in \exi{extension}{b}.
+
+\begin{examples}   \label{extension}
+  \itema Amy has learned that Bob had learned that it's heads.
+  \itemb Amy might learn that it's heads.
+\end{examples}
+
+The development here was concerned with defining concrete possible
+worlds models, and applying them in natural language semantics.  An
+issue for further investigation are the mathematical characterizations
+of epistemic KATs, e.g. sound and complete axioms, coalgebra, decidability.
 

### 2020-09-25 tex/sdraft.tex
tighten section 4

@@
 \section{Translation into the finite state calculus}
-The finite state calculus is an algebra of regular sets of strings and regular
-relations between strings that was designed for use in computationonal phonology and
-morphographemics \citep{kaplan1994regular,beesley2003finite}.  Current
-implementations allow for the definition of functions that have the status of defined
-operations on regular sets and relations \citep{hulden2009foma,karttunen2010update}.
-Such definitions are used here to construct of a model for epistemic KAT inside the
-finite state calculus.  The space of worlds is a set of ordinary strings.  Bit
-sequences (sequences of 0's and 1's) are used for atoms, and these alternate with
-event symbols in the encoding of a world. \ex{gs1fst} gives the encoding of worlds $w_2$ and $w_3$
-from the example.  A string is a finite sequence of symbols, and {\tt 0},
-{\tt 1}, {\tt u}, and {\tt d}, are symbols.  {\tt a0} and {\tt b0} are
-multi-character symbols that are found in implementations of the finite state
-calculus. The multi-character symbol {\tt a0} is an element of the alphabet that has
-no connection with the element {\tt a}.
+The finite state calculus is an algebra of regular sets of strings and
+regular relations between strings that was designed for use in
+computationonal phonology and morphographemics
+\citep{kaplan1994regular,beesley2003finite}.  Current implementations
+allow for the definition of functions on regular sets and relations
+\citep{hulden2009foma,karttunen2010update}.  Such definitions are used
+here to construct of a model for epistemic KAT inside the finite state
+calculus. We describe our translation from Epik terms to Fst programs
+here.
+
+First, the space of worlds is a set of ordinary (as opposed to
+guarded) strings.  Bit sequences (sequences of $0$'s and $1$'s) encode
+atoms, as before, these alternate with event symbols to encode a
+world. \ex{gs1fst} gives the encoding of worlds $w_2$ and $w_3$ from
+the example. The valuation of $h$ preceeds the valuation of $t$, so
+$1~0$ indicates the valuation $h\bar{t}$.
+
+%% unnecessary detail
+%% A string is a finite sequence of symbols, and {\tt 0}, {\tt 1}, {\tt
+%%   u}, and {\tt d}, are symbols.  {\tt a0} and {\tt b0} are
+%% multi-character symbols that are found in implementations of the
+%% finite state calculus. The multi-character symbol {\tt a0} is an
+%% element of the alphabet that has no connection with the element {\tt
+%%   a}.
 
@@
 functions (defined as macros) mapping one or more regular sets or
-relations to a regular set or relation.  The finite state calculus has
-a product operation of string concatenation raised to sets.
-Concatenation of strings with atoms (Boolean vectors) at both ends has
-the effect of doubling atoms at the juncture, and does not enforce
-matching of atoms at the juncture.  Therefore KAT product can not be
-identified with product in the finite state calculus. Instead, KAT
-product and KAT Kleene star are defined operations, see Figure
-\ref{fstfigure1}.  The binary product operation {\tt Cn} and the unary
-Kleene star operation {\tt Kst} combine strings in the string algebra,
-remove strings with non-matching atoms, and then delete the second of
-two tests to create well-formed guarded strings.  Matching of atoms is
-enforced with the set {\tt Wf0}, which is the set of strings that do
-not contain non-matching Boolean vectors.  The containment operator
-(expressed by the dollar sign) and complement (expressed by tilde) are
-operators of the finite state calculus.  The set of non-matching
-sequences of atoms {\tt UnequalStPair} is defined by a finite
-disjunction.  Deletion is accomplished by a re-write rule in the
-finite state calculus, which is a notation for defining regular
-relations by contextually constrained substitutions.  In this case, is
-{\tt Squash} is a regular relation that deletes an atom (a sequence of
-0's and 1's of a certain length) when it follows an atom.\footnote{
-  This is a non-equal length regular relation. The finite state
-  calculus includes such relations, and they can be used with relation
-  composition and relation domain and co-domain.  They are restricted
-  in that the complement and set difference for non-equal length
-  relations is not defined.  The epistemic alterative relations that
-  are defined in Figure -- are equal-length relations.  }
+relations to a regular set or relation.
+
+Now, translating the Epik terms $0$, $1$, $b$, and $e$ are
+straightforward: we simply convert the atoms as previously described,
+ensuring to decorate the events $e$ with their compatible atoms
+first. For example $a_1$ becomes $\{01a_101\}$ Fst also has built-in notions of union (\texttt{|}),
+intersection (\texttt{\&}), and complement (\texttt{\~}). Translating
+the the remaining operators requires more work.
+
+Implementing fusion product using Fst's standard string concatenation
+(denoted by juxatposition $\texttt{X} \texttt{Y}$) requires more
+care. Naively concatenating strings with atoms (Boolean vectors) at
+both ends doubles the atoms at the juncture, and does not enforce the
+requisite atom equality.
+
+To implement the fusion product, we define the binary product
+operation \texttt{Cn}, which combines strings in the underlying string
+algebra, remove strings with non-matching atoms, and then delete the
+second of two tests to create well-formed guarded strings (see Figure
+\ref{fstfigure1}). The set {\tt Wf0} describes pairs of equal atoms as
+the complement of set of unequal atom pairs, implemented using Fst's
+containment operator \texttt{\$}. Then. to delete the second of the
+two (matching) pairs, the \texttt{Squash} operator uses Fst's rewrite
+notation to delete atoms (\texttt{St}) that are preceded by another
+atom. \footnote{ This is a non-equal length regular relation. The
+  finite state calculus includes such relations, and they can be used
+  with relation composition and relation domain and co-domain.  They
+  are restricted in that the complement and set difference for
+  non-equal length relations is not defined.  The epistemic alterative
+  relations that are defined in Figure -- are equal-length relations.
+} The squash operator is applied via the relational composition
+\texttt{.o.} and codomain \texttt{.l} operators.
+%% Deletion is accomplished by a
+%% re-write rule in the finite state calculus, which is a notation for
+%% defining regular relations by contextually constrained substitutions.
+%% In this case, is {\tt Squash} is a regular relation that deletes an
+%% atom (a sequence of 0's and 1's of a certain length) when it follows
+%% an atom.
+
+Kleene star is defined in a similar way (see Figure \ref{fstfigure1}),
+unioning the set of atoms \texttt{St} with \texttt{Kpl} which uses
+Fst's non-empty iteration operator \texttt{+} and normalizes the
+result in the same way as \texttt{Cn}.
 
@@
 
-An event symbol such as $a_1$ (Amy looks at heads) is in the KAT algebra a set of
-bare events decorated with compatible tests on each side, $\{ 10 a_1 10
-\}$ in this case.  This is a unit set rather than a guarded string, because elements
-of the KAT algebra are sets. Worlds in the KAT algebra are defined by sequencing
-events using {\em Kst}. The operation enforces compatibility of states, so that
-$(a_1 + a_0)(b_1 + b_0)$ contains two worlds rather than four.  The program in
-Figure \ref{epikfigure} as interpreted in FST defines a countably infinite set of possible worlds by
-KAT Kleene closure as $\mbox{\em Kst}(a_1 + a_0 +b_1 + b_0)$.
+%% An event symbol such as $a_1$ (Amy looks at heads) is in the KAT algebra a set of
+%% bare events decorated with compatible tests on each side, $\{ 10 a_1 10
+%% \}$ in this case.  This is a unit set rather than a guarded string, because elements
+%% of the KAT algebra are sets. Worlds in the KAT algebra are defined by sequencing
+%% events using {\em Kst}. The operation enforces compatibility of states, so that
+%% $(a_1 + a_0)(b_1 + b_0)$ contains two worlds rather than four.  The program in
+%% Figure \ref{epikfigure} as interpreted in FST defines a countably infinite set of possible worlds by
+%% KAT Kleene closure as $\mbox{\em Kst}(a_1 + a_0 +b_1 + b_0)$.
 
@@
 pre-conditions of the event alternative $d$).  This principle is found
-in earlier literature (Moore 198x, Baltag, Moss and Solecki 20xx).  In
-the construction in Fst, the definition of world alternatives takes a
-simple form. Where $R_a$ is the relation on decorated events for agent
-$a$, the the corresponding relation on worlds in is the Kleene closure
-of $R_a$.  Where $R$ and $S$ are relations, the concatenation product
-of $R$ and $S$ is the set of pairs of the form $\pair{x_1x_2}{y_1y_2}$,
-where $\pair{x_1}{y_1}$ is in relation $R$, and $\pair{x_2}{y_2}$ is
-in relation $S$.  The Kleene closure of relation $R$ is \(\cup_{n \geq
-  0}R^n\), where $R^n$ is the $n$-times concatenation product of $R$
-with itself (the $0$-times concatenation product is $\sem{ 1
-}^{\varphi,\mathcal E}$). This is an operation in the finite
-state calculus.  Figure \ref{relclosure} defines the corresponding
-operation in the guarded string algebra.  The epistemic alternative
-relation on worlds for an agent is then defined as the concatenation
-closure of the event alternative relation for the agent.
+in earlier literature (Moore 198x, Baltag, Moss and Solecki 20xx).
+
+In the construction in Fst, the definition of world alternatives takes
+a simple form. Where $R_a$ is the relation on decorated events for
+agent $a$, the corresponding relation on worlds in is the Kleene
+closure of $R_a$.  Where $R$ and $S$ are relations, the concatenation
+product of $R$ and $S$ is the set of pairs of the form
+$\pair{x_1x_2}{y_1y_2}$, where $\pair{x_1}{y_1}$ is in relation $R$,
+and $\pair{x_2}{y_2}$ is in relation $S$.  The Kleene closure of
+relation $R$ is \(\cup_{n \geq 0}R^n\), where $R^n$ is the $n$-times
+concatenation product of $R$ with itself (the $0$-times concatenation
+product is $\sem{ 1 }^{\varphi,\mathcal E}$). This is an operation in
+the finite state calculus.  Figure \ref{relclosure} defines the
+corresponding operation in the guarded string algebra.  The epistemic
+alternative relation on worlds for an agent is then defined as the
+concatenation closure of the event alternative relation for the agent.
 

### 2020-09-25 tex/sdraft.tex
prose tightening in 4 - 6

@@
 \newcommand{\mr}[1]{\textit{\textcolor{cyan}{[mr]: #1}}}
+%% \newcommand{\ehc}[1]{}
+%% \newcommand{\mr}[1]{}
 
@@
 \begin{hangingpar}
-{\tt{define Kst(R)
+{\tt{define RelKst(R)
     [St\hspace{1pt}.x.\hspace{1pt}St]\hspace{1pt}|\hspace{1pt}Kpl(X);}}\\ The Fst
@@
 
-In the construction in Fst, the definition of world alternatives takes
-a simple form. Where $R_a$ is the relation on decorated events for
-agent $a$, the corresponding relation on worlds in is the Kleene
-closure of $R_a$.  Where $R$ and $S$ are relations, the concatenation
-product of $R$ and $S$ is the set of pairs of the form
-$\pair{x_1x_2}{y_1y_2}$, where $\pair{x_1}{y_1}$ is in relation $R$,
-and $\pair{x_2}{y_2}$ is in relation $S$.  The Kleene closure of
-relation $R$ is \(\cup_{n \geq 0}R^n\), where $R^n$ is the $n$-times
-concatenation product of $R$ with itself (the $0$-times concatenation
-product is $\sem{ 1 }^{\varphi,\mathcal E}$). This is an operation in
-the finite state calculus.  Figure \ref{relclosure} defines the
+In Fst, we use the Kleene closure of the concatenation product
+operation on relations to lift a relation on events for an agent $a$,
+$R_a$, to the corresponding relation on worlds.  First define the
+concatenation product $R~S$ of two relations $R$ and $S$ to be the set
+of pairs of the form $\pair{x_1x_2}{y_1y_2}$, where $x_1 \mathrel R
+y_1$, and $x_2 \mathrel S y_2$. Then $R+$ is the closure of $R$ with
+respect to this operation. Figure \ref{relclosure} defines the
 corresponding operation in the guarded string algebra.  The epistemic
@@
 
-We also implement the semantics of Epik terms using lazy
-lists~\cite{LazyLists} in Haskell, rather than the direct
-interpretation as sets. Unfortunately, regular expressions, and hence
-also Epik, can denote infinite sets of strings, for example the term
-$\bullet^*$. Normally regular languages are represented using a finite
-coalgebra~\cite{}. However the non-distributivity of $\diamond$
-accross $;$ complicates this construction\footnote{Axiomatic and
-  Coalgebraic models for Epik are open questions}. To sidestep this,
-we parameterize the interpretation function on a positive integer $n$
-and only produce guarded strings of length $n$ or less.
-
-We translate the Epik terms into a Haskell algebraic datatype that
-represents terms with the same signature as described in Section 2. We
-then paramaterize the interpretation function on an integer $n$ that
-describes the maximum length of a string we will produce.
-
-The bounded interpretation into lists of strings is very similar to the
-unbounded interpretation into sets of strings, except for the bounds
-checking. The full details are shown in Figure~\ref{fig:list-interp}. First note
-that when $n=0$, we simply return the empty list, denoted $[]$. Terms of the
-form $0$, $1$, $e$, and $\psi$ have the same denotation as before, translated
-into a list (for a set $X$, $\lfloor X \rfloor$ is a list with the same elements
-as $X$ arbitrarily ordered). We compute these lists using BBDs, a standard
-technique for concisely and cannoncially representing boolean functions~\cite{BBDS}.
+We also implement the semantics of Epik terms using lazy lists in
+Haskell, rather than the direct interpretation as sets. Using lists
+sidesteps checking the set invariant (elements are unique) for large
+sets, such as $\bullet^*$, and laziness allows us to delay computing
+these large sets until they are actually needed. To sidestep the
+infiteness of models, we parameterize the interpretation function on a
+positive integer $n$ and only produce guarded strings of length $n$ or
+less.
+
+%% We translate the Epik terms into a Haskell algebraic datatype that
+%% represents terms with the same signature as described in Section 2. We
+%% then paramaterize the interpretation function on an integer $n$ that
+%% describes the maximum length of a string we will produce.
+
+The bounded interpretation into lists of strings is very similar to
+the unboundeds interpretation into sets of strings, except for the
+bounds checking. The full details are shown in
+Figure~\ref{fig:list-interp}. First note that when $n=0$, the
+denotation is empty, denoted $[]$. Terms of the form $0$, $1$, $e$,
+and $\psi$ have the same denotation as before, translated into a list
+(denoted $\lfloor S\rfloor$, for a set $S$). We compute atoms using
+BBDs, which concisely represent boolean functions~\cite{BBDS}.
 
@@
 equivalents: union becomes list append (written $\texttt{+\!+}$);
-concatenation becomes the fusion product (lifted to lists this time),
-negation is implemented using list difference ($\setminus$), and the
-modal operator lifts the alternative relation over lists of
+fusion product is lifted to lists instead of sets, negation is
+implemented using list difference ($\setminus$), and the modal
+operator lifts the alternative relation over lists of
 strings\footnote{Figure~\ref{fig:list-interp} depicts this using the
@@
 
-The lists we use here are \emph{lazy} (as opposed to \emph{strict}), which
-broadly means that computation is delayed until the value is needed. This allows
-us to avoid computing large, unnecessary iterations.
+%% The lists we use here are \emph{lazy} (as opposed to \emph{strict}), which
+%% broadly means that computation is delayed until the value is needed. This allows
+%% us to avoid computing large, unnecessary iterations.
 
@@
 lambdas in logical forms for clauses.  In consequence, the semantic
-term translating a sentence is a term of the logical language.  Such a
-term designates a set of possible words (guarded strings) in the
-possible worlds model determined by an Epik specification.  By way of
-example, \exi{three}{a} is an English sentence with conjunction and
-several levels of clausal embedding.  Using the grammar and parser,
-the sentence is mapped to the term in \exi{three}{b}. Then
-\exi{three}{c} shows a simplified logical from constructed form
-\exi{three}{b} using simply logical equivalences. The negations are
-propositional. The double negation, while logically redundant, comes
-syntactically from the grammar. This term is compiled in an
+term translating a sentence is an Epik term.  Such a term designates a
+set of possible words (guarded strings).  By way of example,
+\exi{three}{a} is an English sentence with conjunction and several
+levels of clausal embedding.  Using the grammar and parser, the
+sentence is mapped to the term in \exi{three}{b}. Then \exi{three}{c}
+shows a simplified logical from constructed form \exi{three}{b} using
+simply logical equivalences. This term is compiled in an
 implementation of the finite state calculus to a finite state machine
 with six states and eight arcs, which accepts a countably infinite set
-of worlds.  In this way the methodology ``directly'' represents the
-set of worlds denoted by \exi{three}{a}.
+of worlds. \ehc{Is this right anymore? the formula has changed
+  drastically}. In this way the methodology ``directly'' represents
+the set of worlds denoted by \exi{three}{a}.
 
@@
 % (_*;t) & ~ <amy> ~ ~ <bob> ~((~ (<amy> ~ (_*;h))) + ~ (<amy> (_*;h)) )
-\itemb \( \displaystyle \bullet^*t \wedge \AKBKAKWH \)\\
+\itemb \( \displaystyle \bullet^*t \wedge \AKBKAKWH \)
 \itemc \(\bullet^*t \wedge \K_{a}(\K_{b}(\K_{a} \bullet^* h + \K_{a} \neg \bullet ^* h))\)
@@
 graduate course in formal semantics, covering intensionality.  There
-are straighforward extensions to additional semantic phenomena, such
+are straightforward extensions to additional semantic phenomena, such
 as tense and perfective aspect as in \exi{extension}{a}, and the combination of

### 2020-09-26 tex/rdraft.tex
Edits all over in sdraft.

--- 2020-09-25 tex/sdraft.tex
@@
 
-\newcommand{\ehc}[1]{\textit{\textcolor{red}{[ehc]: #1}}}
+\newcommand{\ehc}[1]{\textit{\textcolor{green}{[ehc]: #1}}}
 \newcommand{\mr}[1]{\textit{\textcolor{cyan}{[mr]: #1}}}
-
-\newcommand{\K}{\mathcal{K}}
-\newcommand{\B}{\mathcal{B}}
 
@@
           \begin{tabular}{l}
-            agent amy \\
+            agent aly \\
             \hspace{3mm} o1  -> o1 \\
@@
     \begin{minipage}{\columnwidth}
-      \(\begin{array}{llll}
-      \multicolumn{4}{l}{\text{events}~~e\in \mathsf{P}} \\
-      p,q &\multicolumn{3}{l}{::= e\,|\,\sigma\,|\,p+q\,|\,pq\,|\,p^*\,|\,\neg p\,|\, \Diamond_ap} \\
-      \Box_a p & \triangleq \neg\Diamond_a\neg p &
-      \bullet & \triangleq \sum_{e \in \mathsf P}e \\      
-      p \wedge q &  \triangleq \neg(\neg p + \neg q) &
-      p \to q & \neg p + q      
-      \end{array} \)
+      \(\begin{array}{ll}
+      \text{events} & e \in \mathsf{P} \\
+      \text{states} & \sigma~\text{as in Figure 1} \\    
+      p,q & ::= e\,|\,\sigma\,|\,p+q\,|\,pq\,|\,p^*\,|\,\neg p\,|\, \Diamond_ap \\
+      \end{array}\)
+
+      \vspace{1mm}
+      \begin{tabular}{cl}
+        $\Box_a p$ & $\triangleq \neg\Diamond_a\neg p$ \\
+        \(p \wedge q\) &  \(\triangleq \neg(\neg p + \neg q)\) \\
+        $\bullet$ & \( \triangleq \sum_{e \in \mathsf P}e\)
+      \end{tabular}
   \end{minipage}}
@@
 \begin{example}   \label{interpretation1}
-\hspace{-1.5em}\begin{tabular}[t]{ll}
+\begin{tabular}[t]{ll}
 ${\bullet}^*t$ & It's tails. \\
 $\bullet^*h$ & It's heads. \\  
-  $ \bullet^* h \wedge \Box_a \bullet^* h$  & Amy knows it's heads. \\
+$\Box_a \bullet^* h$  & Amy knows that it's heads. \\
 \end{tabular}
 \vspace{1mm}
-\hspace{-1.5em}\begin{tabular}[t]{l}
+
+\begin{tabular}[t]{l}
   $\Box_b(\Box_a\bullet^*t + \Box_a\neg\bullet^*t)$\\
- Bob believes Amy knows whether it's tails.  \\
+ Bob knows that Amy knows whether \\ \hspace{3mm} it's tails.  \\
 \end{tabular}
 \end{example}
-\ehc{I took the `that's out of these sentences. I remember from class
-  that even if they are assigned null in the PF they still exist in
-  the LF. Is it okay to leave it out here?}
-
-\subsection{Standard Epistemic Modalities}
-
-Using our existential modal primitive $\Diamond_a$, and the dual
-encodings of $\Box_a$ and $\wedge$, we can encode the standard modal
-operators expressing knowledge $(\K_a)$ and belief $(\B_a)$ into
-Epik\footnote{The choice to refer to $\Diamond_a$ as an ``epistemic''
-  primitive rather than a ``doxastic'' primitive, is that it is
-  essential in constructing both the doxastic and epistemic
-  operators. A ``doxastic'' moniker would suggest the
-  unrepresentability of epistemic modalities}. These encodings are
-shown below:
-
-\begin{example}\label{encodings}
-  \( \begin{array}[t]{ccl}
-    \textsc{Belief} & \B_a~p &\triangleq \Box_a~p \\
-    \textsc{Knowledge} & \K_a~p & \triangleq p \wedge \B_a~p
-    \end{array}\)
-\end{example}
-
-%% As a consequence of these encodings, the standard dualities
-%% hold. These are shown in \ex{modalprops}.
-%% \begin{example} \label{modalprops}
-%%   \(\K_a~p \to \B_a~p \qquad \B_a~p \to \neg\K_a\neg p \)
-%% \end{example}
-
-To construct different types of reasoners (e.g. accurate, inaccurate,
-etc), users of Epik can construct an agents alternative relation
-accordingly. Observe the discussion of modal axioms \textbf{T} and
-\textbf{D} in Section~\ref{sec:axioms}.
-
-\subsection{Guarded String Interpretation}
+
 \newcommand{\epikspec}{\test,\event,\varphi,\zeta}
@@
 
-\subsection{Axiomatic Classification}
-\label{sec:axioms}
-
-To situate our logic as a modal logic, we investigate the soundness of
-the standard modal axioms~\cite{????} in Epik's semantics. Some of
-these standard axioms (\ex{valid-axioms}) hold all the time, i.e. they
-are \emph{valid}, and the remaining axioms are valid when $\hat{R}_a$
-has a certain shape. The axioms in \ex{conditionally-valid-axioms}
-have a nontrivial condition validity condition.
-
-\begin{example}\label{valid-axioms}
-  \begin{enumerate}[align=left]
-  \item[\textbf{N}] If $p$ is valid, then $\Box_ap$ is valid
-  \item[\textbf{K}] $\Box_a(p \to q) \to \Box_a~p \to \Box_a~q$ is valid.
-  \end{enumerate}
-\end{example}
-
-\begin{example}\label{conditionally-valid-axioms}
-  \(\begin{array}[t]{lll}
-  \textbf{T} & \Box_a p \to p & \text{if } g\mathrel{\hat{R}_a} g, \forall g \\
-  \textbf{D} & \Box_a p \to \Diamond_a p & \text{if } g \in \dom(\hat{R}_a), \forall g \\
-  \textbf{4} & \Box_a p \to \Box_a\Box_a p & \text{if } \hat{R}_a \text{ idempotent}
-  \end{array}\)
-\end{example}
-
-The condition on \textbf{4} is essentially just a restatement of the
-theorem in relational terms. The conditions on the remaining standard
-modal axioms, \textbf{B} and \textbf{5}, are of a similarly trivial
-flavor.
+% $\tuple{K,+,\cdot,*,\bar{},0,1,{}^c,\Diamond_1,...,\Diamond_n}$
 
@@
 sequences (sequences of 0's and 1's) are used for atoms, and these alternate with
-event symbols in the encoding of a world. \ex{gs1fst} gives the encoding of worlds $w_2$ and $w_3$
+event symbols in the encoding of a world. \ex{gs1fst} gives the encoding of worlds
 from the example.  A string is a finite sequence of symbols, and {\tt 0},
@@
 \end{tcolorbox}
-\caption{\raggedright Definition in Fst of KAT product and star.
-  %% Where {\tt X} and {\tt Y} are regular sets and
-  %% {\tt R} and {\tt S} are regular relatiions,
-  %% {\tt X{\&}Y} is the intersection of {\tt X} and {\tt Y},
-  %% {\tt X|Y} is the union of {\tt X} and {\tt Y},
-  %% {\tt \~{}X} is the complement of {\tt X},
-  %% {\tt R{.o.}S} is the composition of {\tt R} and {\tt S},
-  %% {\tt R{.}l} is the co-domain of {\tt R}, and
-  %% {\tt {\$}X} is the set of strings that have a substring in $X$.   
+\caption{\raggedright Definition in Fst of KAT product and KAT Kleene star.
+  Where {\tt X} and {\tt Y} are regular sets and
+  {\tt R} and {\tt S} are regular relatiions,
+  {\tt X{\&}Y} is the intersection of {\tt X} and {\tt Y},
+  {\tt X|Y} is the union of {\tt X} and {\tt Y},
+  {\tt \~{}X} is the complement of {\tt X},
+  {\tt R{.o.}S} is the composition of {\tt R} and {\tt S},
+  {\tt R{.}l} is the co-domain of {\tt R}, and
+  {\tt {\$}X} is the set of strings that have a substring in $X$.   
 }
@@
   \renewcommand{\arraystretch}{1.2}
-\begin{tabular}[t]{cl}
-%% {\textit{World}} & {\textit{String}} \\
+Worlds coded as strings
+\vspace{1mm}
+
+ \hspace{-13mm} \begin{tabular}[t]{cl}
+{\textit{World}} & {\textit{String}} \\
 % $v_1$ & $ 0 \,u\,  1 \,u\,  2 \,d\,  1 \,u\,  2$ \\
-%% $w_1$ & {\tt 1 0} \\
+$w_1$ & {\tt 1 0} \\
 $w_2$ & {\tt 1 0 a1 1 0} \\
@@
         is & $(d \backslash t)\slash(d \backslash_D t)$ & $\lambda P~x.~P~x$ \\
-        knows & $(e \backslash t)\slash_M t$ &  $\lambda p~R.~p\vee{\Diamond}_Rp$\\
-        believes & $(e \backslash t)\slash_M t$ &  $\lambda p~R.~{\Diamond}_Rp$\\
+        knows & $(e \backslash t)\slash_M t$ &  $\lambda p~R.~{\Diamond}_Rp$\\
         that  & $(((e \backslash t){\slash_M}t) \backslash (e \backslash t)) \slash t$
@@
 possible worlds model determined by an Epik specification.  By way of
-example, \exi{three}{a} is an English sentence with conjunction and
-several levels of clausal embedding.  Using the grammar and parser,
-the sentence is mapped to the term in \exi{three}{b}. Then
-\exi{three}{c} shows a simplified logical from constructed form
-\exi{three}{b} using simply logical equivalences. The negations are
-propositional. The double negation, while logically redundant, comes
-syntactically from the grammar. This term is compiled in an
-implementation of the finite state calculus to a finite state machine
-with six states and eight arcs, which accepts a countably infinite set
-of worlds.  In this way the methodology ``directly'' represents the
-set of worlds denoted by \exi{three}{a}.
-
-\newcommand{\HEADS}{\bullet^*h}
-\newcommand{\AKWH}{\neg (\neg \HEADS \vee \Diamond_{a} \neg\HEADS) + \neg(\HEADS \wedge \Diamond_{a}\HEADS)}
-\newcommand{\BKAKWH}{\neg(\neg(\AKWH) + \Diamond_{b}\neg(\AKWH))}
-\newcommand{\AKBKAKWH}{\neg(\neg(\BKAKWH) + \Diamond_{a}\neg(\BKAKWH))}
+example, \exi{three}{a} is an English sentence with 
+conjunction and several levels of clausal embedding.  Using the
+grammar and parser, the sentence is mapped to the term in
+\exi{three}{b}. The negations are propositional. The double
+negation, while logically redundant, comes syntactically from the grammar. This term
+is compiled in an implementation of the finite state calculus to a
+finite state machine with six states and eight arcs, which accepts a
+countably infinite set of worlds.  In this way the methodology
+``directly'' represents the set of worlds denoted by \exi{three}{a}.
 
 \begin{examples} \label{three}
-\itema It's tails and Amy knows that Bob knows that Amy knows whether it's heads.
+\itema Its tails and Amy knows that Bob knows that Amy knows whether its heads.
 % (_*;t) & ~ <amy> ~ ~ <bob> ~((~ (<amy> ~ (_*;h))) + ~ (<amy> (_*;h)) )
-\itemb \( \displaystyle \bullet^*t \wedge \AKBKAKWH \)\\
-\itemc \(\bullet^*t \wedge \K_{a}(\K_{b}(\K_{a} \bullet^* h + \K_{a} \neg \bullet ^* h))\)
-\itemd Amy knows that it's tails.
-\iteme \(
-\bullet ^*t \wedge \neg \Diamond_{a} \neg \bullet^*t~~~~~~(\equiv \K_a \bullet^* t)
+\itemb \( \displaystyle
+\bullet^*t \wedge \neg \Diamond_{\text{Amy}} \neg \neg \Diamond_{\text{Bob}} \neg(
+\neg \Diamond_{\text{Amy}} \neg \bullet^*h 
++
+\neg \Diamond_{\text{Amy}} \bullet^*h 
+)
 \)
-
 \end{examples}
-Sentence \exi{three}{d} is assigned the logical form \exi{three}{e} by
-the grammar.  Logical relations between propositions are checked in
-the finite state calculus by checking set-theoretic relations between
-sets of worlds.  In this case, the propositions \exi{three}{b} and
-\exi{three}{e} are equivalent, in the sense that they denote the same
-set of worlds.  This equivalence is attested computationally in the
-possible worlds model.
-
-%% \begin{examples} \label{entail}
-%% \itema Amy knows that it's tails.
-%% \itemb \(
-%% \neg \Diamond_{a} \neg \bullet^*t
-%% \)
-%% % \itemc \(\displaystyle \Box_{a}\Box_{b}(\Box_{a} \neg \bullet^* h + \Box_{a}\neg \bullet^* t)\) \hfill \mbox{} \(~~\cdot~~\Box_{a}\neg\Box_{b}\Box_{a}\neg \bullet^* t\)
-%% \end{examples}
+Sentence \exi{entail}{a} is assigned the logical form \exi{entail}{b} by the grammar.
+Logical relations between propositions are checked in the finite state calculus
+by checking set-theoretic relations between sets of worlds.
+In this case, the propositions $p_{11}$ (from \ex{three}) and
+$p_{12}$ (from \ex{entail}) have an empty
+symmetric difference 
+$(p_{11} - p_{12}) + (p_{12} - p_{11})$, and so sentences
+\exi{three}{a} and \exi{entail}{b} are equivalent, in the sense that they
+denote the same set of worlds.  This relation is assessed computationally
+in the possible worlds model.
+
+\begin{examples} \label{entail}
+\itema Amy knows that its tails.
+\itemb \( \displaystyle
+\neg \Diamond_{\text{Amy}} \neg \bullet^*t
+)
+\)
+% \itemc \(\displaystyle \Box_{\text{Amy}}\Box_{\text{Bob}}(\Box_{\text{Amy}} \neg \bullet^* h + \Box_{\text{Amy}}\neg \bullet^* t)\) \hfill \mbox{} \(~~\cdot~~\Box_{\text{Amy}}\neg\Box_{\text{Bob}}\Box_{\text{Amy}}\neg \bullet^* t\)
+\end{examples}
 
@@
 worlds models, and applying them in natural language semantics.  An
-issue for further investigation are the mathematical characterizations
-of epistemic KATs, e.g. sound and complete axioms, coalgebra, decidability.
+issue for further investigation is the mathematical characterizatiton
+of doxastic KATs, e.g. axioms relating product and modality.
 

### 2020-09-26 tex/sdraft.tex
Tiny edits in sdraft.

@@
 and computable modeling of epistemic alternatives for agents.  We
-describe a certain systematic way of defining such models, and
+describe a systematic way of defining such models, and
 illustrate how to apply them in natural language semantics.  The focus
@@
 
-\ehc{I'm not sure if these sentences are the most interesting anymore!}
+%\ehc{I'm not sure if these sentences are the most interesting anymore!}
 
@@
 
-Formulas like the ones in \ex{event1} that describe pre- and
+Formulas that describe pre- and
 post-conditions are {\em effect formulas}. They are interpreted as
@@
 an Epik program that describes a possible worlds model for two agents
-with information about one coin, and events of the agents
-semi-privately looking at the coin. The line beginning with {\tt
+with information about one coin, events of the agents
+semi-privately looking at the coin, and events of secretly turning the coin.
+The line beginning with {\tt
   state} enumerates ${\sf{B}}$. The line beginning with {\tt restrict}
@@
 and notation follow \cite{kozen2001automata}.  We add syntax and
-semantics is included to cover multi-agent epistemic semantics.
+semantics to cover multi-agent epistemic semantics.
 
@@
 In the coin example, as we already saw in \ex{effect},
-$\mathcal{A}^\varphi_{\mathsf B}$ is $\{h\bar{t},\bar{h}t\}$, for which we use
-the shorthand $\{H,T\}$.  A guarded string over ${\sf{P}}$ and ${\sf{B}}$ is a
-strings of events from ${\sf{P}}$, alternating with atoms over ${\sf{B}}$, and
-beginning and ending with atoms.  \ex{gs1} gives the encoding as guarded strings
-of the worlds in \ex{elevator1} and \ex{coin1}.  The length of a guarded string
-$p$, written $|p|$ is the number of events in $p$.  An atom such as $H$ is a
-guarded string of length $0$. Correspondingly $|w_1| = 0$, $|w_2| = 1$ and
-$|w_3| = 2$.
+$\mathcal{A}^\varphi_{\mathsf B}$ is $\{h\bar{t},\bar{h}t\}$, for
+which we use the shorthand $\{H,T\}$.  A guarded string over
+${\sf{P}}$ and ${\sf{B}}$ is a strings of events from ${\sf{P}}$,
+alternating with atoms over ${\sf{B}}$, and beginning and ending with
+atoms.  In this construction, $w_1=H$, $w_2=Ha_1H$, $w_3=Ha_1Hb_1H$,
+and $w_4=Ha_1Hb_1Ha_{10}Tb_{01}H$.
 
@@
 coming from the state and effect formulas.  \ex{wf1} defines the
-well-formed guarded strings determined by and Epik specification.  Condition (i)
+well-formed guarded strings determined by an Epik specification.  Condition (i)
 says that each atom is consistent with the state constraint, and condition (ii)
@@
 is something that can ``happen'' is different worlds. For example, $a_1$ has the
-event type $\{H\,a_1\,H\}$, and $a_0$ has the even type $\{T\,a_0\,T\}$. The
-event types for $b_0$ and $b_1$ are analogous.
+event type $\{H\,a_1\,H\}$, and $a_0$ has the event type $\{T\,a_0\,T\}$.
 
 The construction so far defines a set of worlds from an Epik specification.  Normally
-the set is countably infinite, though some choices of effect
-formulas can result in a finite set of worlds.  The next step is to define an
-alternative relation $R_a$ on worlds for each agent $a$. This will result in a Kripke
-frame $\tuple{W,R_1,...,R_n}$ consisting of a set of worlds, and a world-alternative
-relation for each agent \citep{kripke1963semantical}.  An Epik specification defines
-an alternative relation on bare events for each agent $a$, which we notate as
-$\hat{R}_{a}$.  This should be lifted to a relation ${R}_a$ on worlds.  The basic
-idea is that when a world $w$ is incremented with an event $e$, in the resulting
-world $w \diamond e$, epistemic alternatives for agent $a$ are of the form $w'
-\diamond e'$, where $w'$ is an alternative to for $a$ in $w$, and $e'$ is and
-event-alternative to $e$ for $a$.\footnote{In this it is important that the
-  event-alternative relation for an agent is constant across worlds. We anticipate
-  that the definition given here produces results equivalent to what is found in
-  literature on event alternatives in dynamic epistemic semantics, though we have not
-  verified this.  That literature primarily focuses on mapping an epistemic model for
-  a single time and situation to another, and uses general first-order models, rather
-  than guarded string models. See \citet{baltag1999logic}, \citet{van2007dynamic}, and
-  articles in \citet{ditmarsch2015handbook}.  Previous literature is motivated by epistemic
-  logic and AI planning, rather computable possible worlds models in natural language semantics.
-} This needs to be
-implemented in a way that takes account of pre- and post-conditions for events. For
-this, our approach is to refer the definition of well-formed guarded strings.
-\ex{raise1} defines a relation on worlds from a relation on bare events.
+the set is countably infinite, though some choices of effect formulas can result in a
+finite set of worlds.  The next step is to define an alternative relation $R_a$ on
+worlds for each agent $a$. This will result in a Kripke frame $\tuple{W,R_1,...,R_n}$
+consisting of a set of worlds, and a world-alternative relation for each agent
+\citep{kripke1963semantical}.  An Epik specification defines an alternative relation
+on bare events for each agent $a$, which we notate as ${R}_{a}$.  This should be
+lifted to a relation $\hat{R}_a$ on worlds.  The basic idea is that when a world $w$
+is incremented with an event $e$, in the resulting world $w \diamond e$, epistemic
+alternatives for agent $a$ are of the form $w' \diamond e'$, where $w'$ is an
+alternative to for $a$ in $w$, and $e'$ is and event-alternative to $e$ for
+$a$.\footnote{In this it is important that the event-alternative relation for an
+  agent is constant across worlds. We anticipate that the definition given here
+  produces results equivalent to what is found in literature on event alternatives in
+  dynamic epistemic semantics, though we have not verified this.  That literature
+  primarily focuses on mapping an epistemic model for a single time and situation to
+  another, and uses general first-order models, rather than guarded string
+  models. See \citet{baltag1999logic}, \citet{van2007dynamic}, and articles in
+  \citet{ditmarsch2015handbook}.  Previous literature is motivated by epistemic logic
+  and AI planning, rather than computable possible worlds models in natural language
+  semantics.  } This needs to be implemented in a way that takes account of pre- and
+post-conditions for events. For this, our approach is to refer the definition of
+well-formed guarded strings.  \ex{raise1} defines a relation on worlds from a
+relation on bare events.
 
@@
   Let $W$ be a set of guarded strings over events P and primitive
-  tests B, and $\hat{R}$ be a relation on $P$. The corresponding
-  relation ${R}$ on $W$ holds between a guarded string
+  tests B, and ${R}$ be a relation on $P$. The corresponding
+  relation $\hat{R}$ on $W$ holds between a guarded string
 \(
@@
 \alpha_0'e_0' ... e_n'\alpha_{n+1}', 
-\), where for $0 \leq n$, $\element{\tuple{e_i,e'_i}}{\hat{R}}$.
+\), where for $0 \leq n$, $\element{\tuple{e_i,e'_i}}{{R}}$.
 \end{example}
@@
 operation $\cdot$ is fusion product raised to sets, $*$ is Kleene star, the operation
-$\;\bar{}\;$ is complement for tests, $0$ the empty set, and $1$ is the set of
+$\;\bar{}\;$ is complement for tests, $0$ is the empty set, and $1$ is the set of
 atoms.\footnote{$0$ has the dual role the identity for $+$ (union), and as False for
@@
 are used to represent the proposositional semantic values of English sentences.
-\ex{interpretation1} gives some examples.  To explain the first one $\bullet$ as
+\ex{interpretation1} gives some examples.  To explain the first one, $\bullet$ as
 defined in Figure \ref{fig:epik-grammar} is the disjunction of the primitive events.
@@
 conjoining $h$ with the atom at the end of the world.  So $\bullet^* h$ is the set
-of worlds where the coin ends heads-up\footnote{A mapping from English sentences to logical
-  terms in epistemic KAT is presented in Section 6}.
+of worlds where the coin ends heads-up.
 
@@
 \end{example}
-\ehc{I took the `that's out of these sentences. I remember from class
-  that even if they are assigned null in the PF they still exist in
-  the LF. Is it okay to leave it out here?}
-
-\subsection{Standard Epistemic Modalities}
-
+xs
+{\em Standard Epistemic Modalities}
 Using our existential modal primitive $\Diamond_a$, and the dual
 encodings of $\Box_a$ and $\wedge$, we can encode the standard modal
-operators expressing knowledge $(\K_a)$ and belief $(\B_a)$ into
-Epik\footnote{The choice to refer to $\Diamond_a$ as an ``epistemic''
-  primitive rather than a ``doxastic'' primitive, is that it is
-  essential in constructing both the doxastic and epistemic
-  operators. A ``doxastic'' moniker would suggest the
-  unrepresentability of epistemic modalities}. These encodings are
-shown below:
+operators expressing knowledge $(\K_a)$ and belief $(\B_a)$
+as in \ex{encodings}.\footnote{Deeper analyisis of the lexical semantics
+  of {\em know} requires adding modeling of presupposition
+  \citep{collard2018finite}. The grammar fragment in Section 6 does not model
+  the presupposition of {\em know}, except (depending on the Epik model) as
+  an entailment.}
 
@@
 
-To construct different types of reasoners (e.g. accurate, inaccurate,
-etc), users of Epik can construct an agents alternative relation
-accordingly. Observe the discussion of modal axioms \textbf{T} and
-\textbf{D} in Section~\ref{sec:axioms}.
-
-\subsection{Guarded String Interpretation}
+Different types of reasoners (e.g. accurate, inaccurate,
+etc) are modeled using the event-alternatives in an
+Epik specification.\footnote{See the discussion of modal axioms \textbf{T} and
+  \textbf{D} below.}  The agents in Figure 1 do not always have reliable beliefs,
+because of the possibility of secret turning.
+
 \newcommand{\epikspec}{\test,\event,\varphi,\zeta}
@@
 
-A term $p$ of the logical language is interpreted as a set of guarded
+{\em Guarded String Interpretation.} A term $p$ of the logical language is interpreted as a set of guarded
 strings $\sem{p}^{\epikspec}$, where superscript captures dependence
@@
 The epistemic formula $\Diamond_a p$ is interpreted with Kripke
-semantics for epistemic modality, as the pre-image of the embedded
-proposition under the world-alternative relation ${R}_a$.
+semantics for epistemic modality, as the pre-image of $p$ 
+under the world-alternative relation $\hat{R}_a$.
 
@@
 
-\subsection{Axiomatic Classification}
 \label{sec:axioms}
 
-To situate our logic as a modal logic, we investigate the soundness of
-the standard modal axioms~\cite{????} in Epik's semantics. Some of
-these standard axioms (\ex{valid-axioms}) hold all the time, i.e. they
+{\em Axiomatic Classification}. To situate our logic as a modal logic, consieer the soundness of
+the standard modal axioms given our semantics \citep{hughes1996new}. Some of
+these standard axioms (see \ex{valid-axioms}) hold all the time, i.e. they
 are \emph{valid}, and the remaining axioms are valid when $\hat{R}_a$
@@
 
-First, the space of worlds is a set of ordinary (as opposed to
-guarded) strings.  Bit sequences (sequences of $0$'s and $1$'s) encode
-atoms, as before, these alternate with event symbols to encode a
-world. \ex{gs1fst} gives the encoding of worlds $w_2$ and $w_3$ from
-the example. The valuation of $h$ preceeds the valuation of $t$, so
-$1~0$ indicates the valuation $h\bar{t}$.
-
-%% unnecessary detail
-%% A string is a finite sequence of symbols, and {\tt 0}, {\tt 1}, {\tt
-%%   u}, and {\tt d}, are symbols.  {\tt a0} and {\tt b0} are
-%% multi-character symbols that are found in implementations of the
-%% finite state calculus. The multi-character symbol {\tt a0} is an
-%% element of the alphabet that has no connection with the element {\tt
-%%   a}.
+The space of worlds is a set of ordinary (as opposed to guarded) strings.  Bit
+sequences (sequences of $0$'s and $1$'s) encode atoms, and as before, these alternate
+with event symbols to encode a world.  In this construction, $w_3$ of the example is
+the string {\tt{10\hspace{2pt}a1\hspace{2pt}10\hspace{2pt}b1\hspace{2pt}10}}.
 
@@
 
-\begin{example}   \label{gs1fst}
-  \renewcommand{\arraystretch}{1.2}
-\begin{tabular}[t]{cl}
-%% {\textit{World}} & {\textit{String}} \\
-% $v_1$ & $ 0 \,u\,  1 \,u\,  2 \,d\,  1 \,u\,  2$ \\
-%% $w_1$ & {\tt 1 0} \\
-$w_2$ & {\tt 1 0 a1 1 0} \\
-$w_3$ & {\tt 1 0 a1 1 0 b1 1 0} \\  
-  \end{tabular}
-    \renewcommand{\arraystretch}{1}
-\end{example}
-
 Terms in the finite state calculus are interpreted as sets of strings,
@@
 
-Now, translating the Epik terms $0$, $1$, $b$, and $e$ are
+Translating the Epik terms $0$, $1$, $b$, and $e$ are
 straightforward: we simply convert the atoms as previously described,
-ensuring to decorate the events $e$ with their compatible atoms
-first. For example $a_1$ becomes $\{01a_101\}$ Fst also has built-in notions of union (\texttt{|}),
-intersection (\texttt{\&}), and complement (\texttt{\~}). Translating
-the the remaining operators requires more work.
-
-Implementing fusion product using Fst's standard string concatenation
-(denoted by juxatposition $\texttt{X} \texttt{Y}$) requires more
-care. Naively concatenating strings with atoms (Boolean vectors) at
-both ends doubles the atoms at the juncture, and does not enforce the
-requisite atom equality.
-
-To implement the fusion product, we define the binary product
-operation \texttt{Cn}, which combines strings in the underlying string
-algebra, remove strings with non-matching atoms, and then delete the
-second of two tests to create well-formed guarded strings (see Figure
-\ref{fstfigure1}). The set {\tt Wf0} describes pairs of equal atoms as
-the complement of set of unequal atom pairs, implemented using Fst's
-containment operator \texttt{\$}. Then. to delete the second of the
-two (matching) pairs, the \texttt{Squash} operator uses Fst's rewrite
-notation to delete atoms (\texttt{St}) that are preceded by another
-atom. \footnote{ This is a non-equal length regular relation. The
-  finite state calculus includes such relations, and they can be used
-  with relation composition and relation domain and co-domain.  They
-  are restricted in that the complement and set difference for
-  non-equal length relations is not defined.  The epistemic alterative
-  relations that are defined in Figure -- are equal-length relations.
-} The squash operator is applied via the relational composition
-\texttt{.o.} and codomain \texttt{.l} operators.
+decorate the events $e$ with their compatible atoms. For example $a_1$ becomes
+an Fst term denoting
+$\{01a_101\}$. Fst has built-in operations of union (\texttt{|}),
+and intersection (\texttt{\&}), which define
+the sum and intersection operations in the guarded string algebra.
+Fst set difference (\texttt{-}) is used to define propositional complement as the
+difference between the set of worlds and the argument.
+
+Defining KAT product using Fst's set-lifted string concatenation (denoted by
+juxatposition $\texttt{X} \texttt{Y}$) requires more care. Naively concatenating
+strings with atoms (Boolean vectors) at both ends doubles atoms at the juncture, and
+does not enforce the requisite atom equality.  To implement KAT product, we define
+the binary operation \texttt{Cn}, which concatenates strings in the string algebra,
+removes strings with non-matching atoms, and then deletes the second of two atoms to
+create a set of well-formed guarded strings.  See Figure \ref{fstfigure1}. {\tt Wf0}
+is the set of ordinary strings that does not contain unequal pairs of atoms, as
+defined using Fst's containment operator \texttt{\$}. The \texttt{Squash} relation
+uses Fst's rewrite notation to delete atoms (elements of \texttt{St}) that are
+preceded by another atom. \footnote{ This is a non-equal length regular relation. The
+  finite state calculus includes such relations, and they can be used with relation
+  composition and relation domain and co-domain.  They are restricted in that the
+  complement and set difference for non-equal length relations is not defined.
+  Epistemic alterative relations are equal-length relations.} This relation is
+applied via the relational composition (\texttt{.o.}) and codomain (\texttt{.l})
+operators.
 %% Deletion is accomplished by a
@@
 
-Kleene star is defined in a similar way (see Figure \ref{fstfigure1}),
-unioning the set of atoms \texttt{St} with \texttt{Kpl} which uses
-Fst's non-empty iteration operator \texttt{+} and normalizes the
-result in the same way as \texttt{Cn}.
+KAT Kleene plus is defined in a similar way using Kleene plus in the string algebra,
+with checks for equality of atoms and deletion of atoms. See Figure \ref{fstfigure1}.
+KAT Kleene star is definied from KAT Kleene plus and the multiplicative identity,
+which is the set of well-formed atoms \texttt{St}.
 
@@
   a relation between guarded strings.}
-\label{fstfigure2}
+\label{relclosure}
 \end{figure}
@@
 It remains to define an epistemic alternative relation on worlds for each agent.  The
-relevant information in Figure \ref{epikfigure} is a relation between bare events for each
-agent. This determines a relation in the guarded string algebra a relation between
-bare events decorated with compatible tests.  For agent Amy, this is the relation
-described in \ex{alyrel1} as a set of ordered pairs.
-
-\vspace{-2mm}
-\begin{example} \label{alyrel1}
- \hspace{-4mm} \(
- \left\{
- \begin{array}{l}
-   \pair{10a_110}{10a_110},  \pair{01a_001}{01a_001}, \\
-   \pair{10b_110}{10b_110},  \pair{10b_110}{01b_001}, \\
-   \pair{01b_001}{10b_110},  \pair{01b_001}{01b_001} \\
- \end{array}
- \right\}
- \) 
-\end{example}
-
-% Fixes junky vertical spacing of the set above.
-% Possibly remove it in the final version.
-\raggedbottom
-
-The relation on decorated events needs to be generalized to a relation
-of worlds.  The principle for this is that an epistemic alternative to
-a world of the form $we$ is a world of the form $vd$, where $v$ is a
-world-alternative to $w$, $d$ is an event-alternative to $e$, and $vd$
-is defined (i.e.  the world alternative $v$ satisfies the
-pre-conditions of the event alternative $d$).  This principle is found
-in earlier literature (Moore 198x, Baltag, Moss and Solecki 20xx).
-
-In Fst, we use the Kleene closure of the concatenation product
-operation on relations to lift a relation on events for an agent $a$,
-$R_a$, to the corresponding relation on worlds.  First define the
-concatenation product $R~S$ of two relations $R$ and $S$ to be the set
-of pairs of the form $\pair{x_1x_2}{y_1y_2}$, where $x_1 \mathrel R
-y_1$, and $x_2 \mathrel S y_2$. Then $R+$ is the closure of $R$ with
-respect to this operation. Figure \ref{relclosure} defines the
-corresponding operation in the guarded string algebra.  The epistemic
-alternative relation on worlds for an agent is then defined as the
-concatenation closure of the event alternative relation for the agent.
+relevant information in Figure \ref{epikfigure} is a relation between bare events for
+each agent. This determines a relation between bare events decorated with compatible
+atoms. 
+In Fst, we use the closure of the concatenation
+product operation on relations to lift a relation on decorated events for an agent to
+the corresponding relation on worlds.  The concatenation product $RS$ of two
+relations $R$ and $S$ is the set of pairs of the form $\pair{x_1x_2}{y_1y_2}$, where
+$x_1 \mathrel R y_1$, and $x_2 \mathrel S y_2$.  In Fst, {\tt{R+}} is the closure of relation {\tt R} with
+respect to this operation. Figure \ref{relclosure} defines the corresponding
+operation on sets of guarded strings as encoded in Fst.\footnote{Relation concatenation in Fst
+  differs from relation composition (.o.), and the closure under discussion here is
+  the closure of the former rather than the latter.}  The epistemic alternative
+relation on worlds for an agent is then defined as the KAT relational concatenation
+closure {\tt RelKst} of the decorated-event alternative relation for the agent.
 
@@
   with $d \backslash_D t$ on the right to give a value that combines
-  with $d$ on the left to give $t$.}\footnote{Lambda abstractions with
-  multiple parameters are written $\lambda x~y.~e$ rather than the
-  more verbose $\lambda x.\lambda y.e$.} The grammar and semantics are
+  with $d$ on the left to give $t$. In the semantics, lambda abstractions with
+  multiple parameters are written $\lambda x~y.~e$ rather than $\lambda x.\lambda y.e$.
+$d$ is the category of {\em it}.}
+The grammar and semantics are
 optimized for a simple fragment of English concerned with clausal
@@
         \textsc{Item} & \textsc{Type} & \textsc{Semantics}\\ \midrule  
-        Amy & $e$ & ${R}_a$\\
-        Bob & $e$ & ${R}_b$\\
+        Amy & $e$ & $\hat{R}_a$\\
+        Bob & $e$ & $\hat{R}_b$\\
         it & $d$ & $d$ \\
@@
 Multimodal categories such as $\backslash_D$ and $\backslash_M$ are
-used to control the derivation.
+used to control the derivation---phrases with these top-level slashes
+can only combine syntactically as arguments.
 %
@@
 world-alternative relation contributed by the subject.  This is not the right
-semantics for {\em Amy knows that it's heads}, because it is an existential
+semantics for {\em Amy knows that it's heads}, because it has an existential
 modality ${\Diamond_R}p$, rather than an universal modality ${\Box_R}p$.  This
@@
 translations, accompanied by beta reductions that eliminate all
-lambdas in logical forms for clauses.  In consequence, the semantic
+lambdas logical forms for clauses.  In consequence, the semantic
 term translating a sentence is an Epik term.  Such a term designates a
@@
 levels of clausal embedding.  Using the grammar and parser, the
-sentence is mapped to the term in \exi{three}{b}. Then \exi{three}{c}
+sentence is mapped to the term in \exi{three}{b}. \exi{three}{c}
 shows a simplified logical from constructed form \exi{three}{b} using
-simply logical equivalences. This term is compiled in an
+logical equivalences. The origial term is compiled in an
 implementation of the finite state calculus to a finite state machine
-with six states and eight arcs, which accepts a countably infinite set
+with ====== states and ====== arcs, which accepts a countably infinite set
 of worlds. \ehc{Is this right anymore? the formula has changed
-  drastically}. In this way the methodology ``directly'' represents
+  drastically. MR: No, we need to write the Epik program and compile it
+into Fst. The sentences might not be equivalent any more.}. In this way the methodology ``directly'' represents
 the set of worlds denoted by \exi{three}{a}.
@@
 \end{examples}
-Sentence \exi{three}{d} is assigned the logical form \exi{three}{e} by
-the grammar.  Logical relations between propositions are checked in
-the finite state calculus by checking set-theoretic relations between
-sets of worlds.  In this case, the propositions \exi{three}{b} and
-\exi{three}{e} are equivalent, in the sense that they denote the same
-set of worlds.  This equivalence is attested computationally in the
-possible worlds model.
+Sentence \exi{three}{d} is assigned the logical form \exi{three}{e} by the grammar.
+Logical relations between propositions are checked in the finite state calculus by
+checking set-theoretic relations between sets of worlds (as represented by finite
+state machines in an interpreter for the finite state calculus).  In this case, the
+propositions \exi{three}{b} and \exi{three}{e} are equivalent, in the sense that they
+denote the same set of worlds.  This equivalence is attested computationally in the
+possible worlds model.  Entailment can be tested with a subset check on sets of
+worlds, and logical compatibility with a check for non-empty intersection.
 
@@
 \section{Discussion}
-The methodology presented here is designed for use in research in
-linguistic semantics, and for education at the level of a second
-graduate course in formal semantics, covering intensionality.  There
-are straightforward extensions to additional semantic phenomena, such
-as tense and perfective aspect as in \exi{extension}{a}, and the combination of
-metaphysical modality and prospective aspect in \exi{extension}{b}.
+The methodology presented here is designed for use in research in linguistic
+semantics, and for education at the level of a second graduate course in formal
+semantics, covering intensionality.  There are straightforward extensions to
+additional linguistic phenomena, such as tense and perfective aspect as in
+\exi{extension}{a}, and the combination of metaphysical modality and prospective
+aspect in \exi{extension}{b}.
 
@@
 
-The development here was concerned with defining concrete possible
-worlds models, and applying them in natural language semantics.  An
-issue for further investigation are the mathematical characterizations
-of epistemic KATs, e.g. sound and complete axioms, coalgebra, decidability.
+The model framework is a constructive branching-time framework with metaphysical
+modality and epistemic modality, which should be applicable in linguistic semantic
+research on combinations of tense, metaphysical modality, and epistemic
+complementation \citep{thomason1984combinations, abusch1998generalizing,
+  condoravdi2002temporal}.  Connections with research on temporal constitution of
+events in a related formal setting remain to be explored
+\citep{fernando2004finite,fernando2007observing,carlson2009tense}.
+
+The development here is concerned with defining concrete possible
+worlds models, and applying them in natural language semantics.  
+Issues for further investigation are mathematical characterizations
+of epistemic KATs, e.g. sound and complete axioms, coalgebra, and decidability.
+
+Open source code will be distributed in conjunction with the conference.
 

### 2020-09-27 tex/sdraft.tex
Edits in sdraft, nearly there!

@@
 \newcommand{\mr}[1]{\textit{\textcolor{cyan}{[mr]: #1}}}
-%% \newcommand{\ehc}[1]{}
-%% \newcommand{\mr}[1]{}
+% \newcommand{\ehc}[1]{}
+% \newcommand{\mr}[1]{}
 
@@
 \usepackage{stmaryrd}
+\SetSymbolFont{stmry}{bold}{U}{stmry}{m}{n}
 \usepackage{array}
@@
   \citep{collard2018finite}. The grammar fragment in Section 6 does not model
-  the presupposition of {\em know}, except (depending on the Epik model) as
+  the presupposition of {\em know}, except as
   an entailment.}
@@
 operation on sets of guarded strings as encoded in Fst.\footnote{Relation concatenation in Fst
-  differs from relation composition (.o.), and the closure under discussion here is
+  differs from relation composition ({\tt{.o.}}), and the closure under discussion here is
   the closure of the former rather than the latter.}  The epistemic alternative
@@
 operator lifts the alternative relation over lists of
-strings\footnote{Figure~\ref{fig:list-interp} depicts this using the
+strings\footnote{Figure~7 depicts this using the
   list comprehension notation, which is analogous to set builder
@@
 
-Sentences are parsed with a chart parser for categorial grammar.  The
-semantics for complex phrases are obtained by application of semantic
-translations, accompanied by beta reductions that eliminate all
-lambdas logical forms for clauses.  In consequence, the semantic
-term translating a sentence is an Epik term.  Such a term designates a
-set of possible words (guarded strings).  By way of example,
-\exi{three}{a} is an English sentence with conjunction and several
-levels of clausal embedding.  Using the grammar and parser, the
-sentence is mapped to the term in \exi{three}{b}. \exi{three}{c}
-shows a simplified logical from constructed form \exi{three}{b} using
-logical equivalences. The origial term is compiled in an
-implementation of the finite state calculus to a finite state machine
-with ====== states and ====== arcs, which accepts a countably infinite set
-of worlds. \ehc{Is this right anymore? the formula has changed
-  drastically. MR: No, we need to write the Epik program and compile it
-into Fst. The sentences might not be equivalent any more.}. In this way the methodology ``directly'' represents
-the set of worlds denoted by \exi{three}{a}.
+Sentences are parsed with a chart parser for categorial grammar.  The semantics for
+complex phrases are obtained by application of semantic translations, accompanied by
+beta reductions that eliminate all lambdas logical forms for clauses.  In
+consequence, the semantic term translating a sentence is an Epik term.  Such a term
+designates a set of possible words (guarded strings).  By way of example,
+\exi{three}{a} is an English sentence with conjunction and several levels of clausal
+embedding.  Using the grammar and parser, the sentence is mapped to the term in
+\exi{three}{b}. \exi{three}{c} shows a simplified logical from constructed form
+\exi{three}{b} using logical equivalences. Either term is compiled in an
+implementation of the finite state calculus to a finite state machine with 10
+states and 23 edges, which accepts a countably infinite set of worlds.\footnote{
+Machine sizes need not be small, especially as the cardinality of $\sf{B}$
+increases.  A certain Epik model with fourteen primitive tests has the set of worlds
+representented by a finite state machine with 184794 states and 257881
+edges.}
+In this way the methodology ``directly'' represents the set of worlds
+denoted by \exi{three}{a}.
 
@@
 Logical relations between propositions are checked in the finite state calculus by
-checking set-theoretic relations between sets of worlds (as represented by finite
-state machines in an interpreter for the finite state calculus).  In this case, the
-propositions \exi{three}{b} and \exi{three}{e} are equivalent, in the sense that they
-denote the same set of worlds.  This equivalence is attested computationally in the
-possible worlds model.  Entailment can be tested with a subset check on sets of
-worlds, and logical compatibility with a check for non-empty intersection.
+checking set-theoretic relations between sets of worlds.
+For instance entailment $p \rightarrow q$ is decided by checking in an interpreter
+for the finite sate calculus whether $p - q$ is non-empty.
+In the model defined by Figure 2, the
+propositions \exi{three}{b} and \exi{three}{e} are
+independent. They are equivalent in a version 
+without secret flipping.
 
@@
 The model framework is a constructive branching-time framework with metaphysical
-modality and epistemic modality, which should be applicable in linguistic semantic
+modality and epistemic modality, which will be applicable in linguistic semantic
 research on combinations of tense, metaphysical modality, and epistemic
@@
 
-The development here is concerned with defining concrete possible
+The development here is concerned with defining concrete computable possible
 worlds models, and applying them in natural language semantics.  

### 2020-09-27 tex/sdraft.tex
fix typos and bib

@@
   \begin{array}{lll}
-    \sem{\rho\mathbin{:}\sigma}^\varphi & = &
+    \sem{\rho\mathbin{:}\sigma}^\varphi & \triangleq &
         {\mathcal A}^{\rho\varphi}_{\mathsf{B}}\times{\mathcal A}^{\sigma\varphi}_{\mathsf{B}}\\[.1em]
-        \sem{\zeta +\eta}^\varphi & = & \sem{\zeta}^\varphi\cup\sem{\eta}^\varphi\\
-        \sem{\zeta \mathbin{\&}  \eta}^\varphi & = & \sem{\zeta}^\varphi\cap\sem{\eta}^\varphi\\
-        \sem{\bar{\zeta}}^\varphi & = & {\mathcal A}^{\varphi}_{\mathsf{B}} \times {\mathcal A}^{\varphi}_{\mathsf{B}} \setminus \sem{\zeta}^\varphi\\
+        \sem{\zeta +\eta}^\varphi & \triangleq & \sem{\zeta}^\varphi\cup\sem{\eta}^\varphi\\
+        \sem{\zeta \mathbin{\&}  \eta}^\varphi & \triangleq & \sem{\zeta}^\varphi\cap\sem{\eta}^\varphi\\
+        \sem{\bar{\zeta}}^\varphi & \triangleq & {\mathcal A}^{\varphi}_{\mathsf{B}} \times {\mathcal A}^{\varphi}_{\mathsf{B}} \setminus \sem{\zeta}^\varphi\\
   \end{array}  
@@
 \end{example}
-xs
+
 {\em Standard Epistemic Modalities}
@@
 
-{\em Axiomatic Classification}. To situate our logic as a modal logic, consieer the soundness of
+{\em Axiomatic Classification}. To situate our logic as a modal logic, consider the soundness of
 the standard modal axioms given our semantics \citep{hughes1996new}. Some of
@@
 Defining KAT product using Fst's set-lifted string concatenation (denoted by
-juxatposition $\texttt{X} \texttt{Y}$) requires more care. Naively concatenating
+juxtaposition $\texttt{X}~\texttt{Y}$) requires more care. Naively concatenating
 strings with atoms (Boolean vectors) at both ends doubles atoms at the juncture, and
@@
 
-\begin{figure} \label{relclosure}
+\begin{figure}
   \begin{tcolorbox}[size=fbox]
@@
 product operation on relations to lift a relation on decorated events for an agent to
-the corresponding relation on worlds.  The concatenation product $RS$ of two
+the corresponding relation on worlds.  The concatenation product $R~S$ of two
 relations $R$ and $S$ is the set of pairs of the form $\pair{x_1x_2}{y_1y_2}$, where
@@
 bounds checking. The full details are shown in
-Figure~\ref{fig:list-interp}. First note that when $n=0$, the
+Figure~\ref{fig:bound-interp}. First note that when $n=0$, the
 denotation is empty, denoted $[]$. Terms of the form $0$, $1$, $e$,
@@
 (denoted $\lfloor S\rfloor$, for a set $S$). We compute atoms using
-BBDs, which concisely represent boolean functions~\cite{BBDS}.
+BDDs, which concisely represent boolean functions~\cite{lee1959representation}.
 
@@
 For instance entailment $p \rightarrow q$ is decided by checking in an interpreter
-for the finite sate calculus whether $p - q$ is non-empty.
+for the finite state calculus whether $p - q$ is non-empty.
 In the model defined by Figure 2, the
@@
 
-Open source code will be distributed in conjunction with the conference.
+We will release Epik's source code under an open source license prior
+to the conference.
 

### 2021-01-12 tex/pdraft.tex
Initiate publication draft pdraft.tex
Add page 9, add appendix, change to nop events in Figure 2.

--- 2020-09-27 tex/sdraft.tex
@@
 
-%\aclfinalcopy % Uncomment this line for the final submission
+% MR249 UNCOMMENTED FOR FINAL VERSION
+\aclfinalcopy % Uncomment this line for the final submission
 %\def\aclpaperid{***} %  Enter the acl Paper ID here
@@
 
-%\author{Eric Campbell \\
-%  Cornell University\\
-%  \texttt{ehc86@cornell.edu} \\\And
-%  Mats Rooth \\
-%  Cornell University\\
-%  \texttt{mr249@cornell.edu} \\}
+\author{Eric Campbell \\
+  Cornell University\\
+  \texttt{ehc86@cornell.edu} \\\And
+  Mats Rooth \\
+  Cornell University\\
+  \texttt{mr249@cornell.edu} \\}
 
@@
 an Epik program that describes a possible worlds model for two agents
-with information about one coin, events of the agents
-semi-privately looking at the coin, and events of secretly turning the coin.
-The line beginning with {\tt
-  state} enumerates ${\sf{B}}$. The line beginning with {\tt restrict}
-gives the state formula.  The lines beginning with {\tt event} declare
-events and their effect formulas.  Finally the lines beginning with
-{\tt agent} define {\em event alternative} relations for agents.  Each
-clause with an arrow has a single event symbol on the left, and a
-disjunction of alternative events on the right of the arrow. The
-interpretation of Amy's alternatives for $b_1$ (Bob peeks at heads),
-is that when $b_1$ happens, for Amy either $b_1$ or $b_0$ (Bob peeks
-at tails) could be happening. Her alternatives for $a_{01}$ and
-$a_{10}$ (she turns the coin over) are $a_{10}$ and $a_{01}$,
+with information about one coin, events of the agents semi-privately
+looking at the coin, and events of secretly turning the coin.  The
+line beginning with {\tt state} enumerates ${\sf{B}}$. The line
+beginning with {\tt restrict} gives the state formula.  The lines
+beginning with {\tt event} declare events and their effect formulas.
+Finally the lines beginning with {\tt agent} define {\em event
+  alternative} relations for agents.  Each clause with an arrow has a
+single event symbol on the left, and a disjunction of alternative
+events on the right of the arrow. The interpretation of Amy's
+alternatives for $b_1$ (Bob peeks at heads), is that when $b_1$
+happens, for Amy either $b_1$ or $b_0$ (Bob peeks at tails) could be
+happening. Her alternatives for $a_{01}$ and $a_{10}$ (she turns the
+coin over in one state or the other) are $a_{10}$ and $a_{01}$,
 indicating that she doesn't know, \emph{a priori}, whether she's
-turning the coin from $H$ to $T$ or from $T$ to $H$. Similarly, Bob
-secretly turns the coin over, in $b_{10}$ or $b_{01}$, she doesn't
-know anything has happened, so her alternatives are the no-operation
-events $o_1$ and $o_0$ for heads-worlds and tails-worlds
-respectively. Bob's alternatives are the same, \emph{mutatis mutandi}.
+turning the coin from $H$ to $T$ or from $T$ to $H$. When Bob secretly
+turns the coin over, in event $b_{10}$ or $b_{01}$, she doesn't know
+anything has happened, so her alternative is the no-operation,
+no-information event event $o$. Bob's event relation is symmetric.
 
@@
             \phantom{restrict }+ t!h \\
-            event o1  h:h \\
-            event o0  t:t \\
+            event o  h:h + t:t\\
             event a1  h:h \\
@@
             agent amy \\
-            \hspace{3mm} o1  -> o1 \\
-            \hspace{3mm} o0  -> o0 \\
+            \hspace{3mm} o  -> o \\
             \hspace{3mm} a1  -> a1 \\
@@
   model for two agents with information about one coin, and events of
-  the agents semi-privately looking at the coin. }
+  the agents semi-privately looking at the coin, and privately turning
+  the coin.}
 \label{fig:epik-example}
@@
 express universal modal force.
-
-\begin{figure}[!t]
-  \vspace{1em}
-  \centering
-  \footnotesize
-  \begin{tabular}[t]{lcc}
-    \toprule
-        \textsc{Item} & \textsc{Type} & \textsc{Semantics}\\ \midrule  
-        Amy & $e$ & $\hat{R}_a$\\
-        Bob & $e$ & $\hat{R}_b$\\
-        it & $d$ & $d$ \\
-        heads & $d \backslash_D t$ & $\lambda x.\bullet^*h$ \\
-        tails & $d \backslash_D t$ & $\lambda x.\bullet^*t$ \\
-        is & $(d \backslash t)\slash(d \backslash_D t)$ & $\lambda P~x.~P~x$ \\
-        knows & $(e \backslash t)\slash_M t$ &  $\lambda p~R.~p\vee{\Diamond}_Rp$\\
-        believes & $(e \backslash t)\slash_M t$ &  $\lambda p~R.~{\Diamond}_Rp$\\
-        that  & $(((e \backslash t){\slash_M}t) \backslash (e \backslash t)) \slash t$
-        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
-        whether  & $(((e \backslash t){\slash_M} t) \backslash (e \backslash t)) \slash t$ 
-        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
-        && \multicolumn{1}{r}{$+ {\neg}(m\,p\,R)$} \\ \bottomrule
-      \end{tabular}
-  \caption{Partial categorial grammar lexicon. The first column has a word form.
-    the second column a categorial type, and third column a semantic translation
-    in a logical language that extends the Epik term language with lambda. }
-  \label{lexiconfigure}
-\end{figure}
 
@@
 
+\begin{figure}[!t]
+  \vspace{1em}
+  \centering
+  \footnotesize
+  \begin{tabular}[t]{lcc}
+    \toprule
+        \textsc{Item} & \textsc{Type} & \textsc{Semantics}\\ \midrule  
+        Amy & $e$ & $\hat{R}_a$\\
+        Bob & $e$ & $\hat{R}_b$\\
+        it & $d$ & $d$ \\
+        heads & $d \backslash_D t$ & $\lambda x.\bullet^*h$ \\
+        tails & $d \backslash_D t$ & $\lambda x.\bullet^*t$ \\
+        is & $(d \backslash t)\slash(d \backslash_D t)$ & $\lambda P~x.~P~x$ \\
+        knows & $(e \backslash t)\slash_M t$ &  $\lambda p~R.~p\vee{\Diamond}_Rp$\\
+        believes & $(e \backslash t)\slash_M t$ &  $\lambda p~R.~{\Diamond}_Rp$\\
+        that  & $(((e \backslash t){\slash_M}t) \backslash (e \backslash t)) \slash t$
+        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
+        whether  & $(((e \backslash t){\slash_M} t) \backslash (e \backslash t)) \slash t$ 
+        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
+        && \multicolumn{1}{r}{$+ {\neg}(m\,p\,R)$} \\ \bottomrule
+  \end{tabular}
+    \caption{Partial categorial grammar lexicon. The first column has a word form.
+    the second column a categorial type, and third column a semantic translation
+    in a language that extends the Epik logical language with lambda. The complete grammar
+    is in the appendix.}
+  \label{lexiconfigure}
+\end{figure}
+
 Sentences are parsed with a chart parser for categorial grammar.  The semantics for
@@
 \end{examples}
-Sentence \exi{three}{d} is assigned the logical form \exi{three}{e} by the grammar.
-Logical relations between propositions are checked in the finite state calculus by
-checking set-theoretic relations between sets of worlds.
-For instance entailment $p \rightarrow q$ is decided by checking in an interpreter
-for the finite state calculus whether $p - q$ is non-empty.
-In the model defined by Figure 2, the
-propositions \exi{three}{b} and \exi{three}{e} are
-independent. They are equivalent in a version 
-without secret flipping.
+
+Sentence \exi{three}{d} is assigned the logical form \exi{three}{e} by
+the grammar.  Logical relations between propositions are checked in
+the finite state calculus by checking set-theoretic relations between
+sets of worlds.  For instance entailment $p \rightarrow q$ is decided
+by checking in an interpreter for the finite state calculus whether $p
+- q$ is non-empty.  In the model defined by Figure 2, the propositions
+\exi{three}{b} and \exi{three}{e} are independent. They are equivalent
+in a version without secret flipping.
 
@@
 The methodology presented here is designed for use in research in linguistic
-semantics, and for education at the level of a second graduate course in formal
+semantics, for research in computational syntax and semantics on model-theoretically
+grounded semantics, 
+and for education at the level of a second graduate course in formal
 semantics, covering intensionality.  There are straightforward extensions to
@@
 
-The model framework is a constructive branching-time framework with metaphysical
-modality and epistemic modality, which will be applicable in linguistic semantic
-research on combinations of tense, metaphysical modality, and epistemic
-complementation \citep{thomason1984combinations, abusch1998generalizing,
-  condoravdi2002temporal}.  Connections with research on temporal constitution of
-events in a related formal setting remain to be explored
+The model framework is a constructive branching-time framework with
+metaphysical modality and epistemic modality, which will be applicable
+in linguistic semantic research on combinations of tense, metaphysical
+modality, and epistemic complementation
+\citep{thomason1984combinations, abusch1998generalizing,
+  condoravdi2002temporal}.  Connections with research on temporal
+constitution of events in a related formal setting remain to be
+explored
 \citep{fernando2004finite,fernando2007observing,carlson2009tense}.
+
+The syntactic part of the grammar formalism uses categorial grammar to
+map between strings (or trees or derivations) and logical
+translations.  This way of using a categorial grammar is standard
+\citep{steedman2000syntactic,bozsahin2012combinatory}, and there are
+computational implementations for it
+\citep{barker2005parser,bozsahin2021ccglab}.\footnote{Our grammar uses
+  basic categorial grammar, not additional features such as
+  combinators, type raising, continuations, or multimodal slashes that
+  interact with the grammar in non-trivial ways.  These features would
+  become relevant in more sophisticated grammars.  Multimodal
+  categories are used for phrases such as a tenseless verb phrase that
+  semantically are functional, but can combine in the grammar only as
+  arguments.}  While this way of proceeding is simple and attractive
+for the applications we are interested in, it would be possible to use
+our semantics with other frameworks that have a semantics that uses a
+lambda extension of a logical language.
+
+%\newcommand{\Prime}[1]{{\rm\bf{#1}}}
+
+The semantics presented here is exclusively concerned with events and
+worlds.  Semantic type systems for natural language usually include a
+type for individuals.  In the coin example, the individuals are hidden
+in the primitive event symbols such as $a_1$ and $b_1$. This situation
+would get worse in a more elaborate model with more agents and
+multiple coins.  The solution to this should be to base the world
+construction on grounded event terms such as
+$\Prime{look}(\Prime{a},\Prime{k},1)$, for ``agent $\Prime{a}$ see
+that coin $\Prime{k}$ is heads'', rather than atomic events.  This way
+of proceeding is found in research on situation calculus
+\citep{reiter2001knowledge}.  How to incorporate it in the scheme for
+Epik specifications and into the computational parts of the proposal
+is a topic for future research.  In this it is not immediately clear
+whether it is possible to introduce quantification over individuals,
+together with individual constants, while maintaining the finite state
+semantics. \citet{rooth2017finite} develops an approach to a positive
+answer, based on introducing markers for witnesses for discourse
+referents in the construction of worlds.
 
@@
 Issues for further investigation are mathematical characterizations
-of epistemic KATs, e.g. sound and complete axioms, coalgebra, and decidability.
-
-We will release Epik's source code under an open source license prior
-to the conference.
+of epistemic KATs, e.g. sound and complete axioms and coalgebra.
+
+Source code, examples, and instructions for Epik are distributed at
+\url{https://github.com/ericthewry/epik}.
 
@@
 % \bibliographystyle{chicago}
+
+\clearpage 
 \bibliography{fs}
 
+\clearpage
+
+\onecolumn
+
+\section*{Appendix}
+Complete categorial grammar extending Figure \ref{lexiconfigure}.  The online supplement
+includes a version of this grammar that can be run using Barker and Shan's
+chart parser for categorial grammar \citep{barker2005parser}.
+
+  \begin{tabular}[t]{lcc}
+    \toprule
+    \textsc{Item} & \textsc{Type} & \textsc{Semantics}\\ \midrule
+    \multicolumn{3}{c}{{\bf Statives with expletive subject}} \\ 
+        heads & $d \backslash_D t$ & $\lambda x.\bullet^*h$ \\
+        tails & $d \backslash_D t$ & $\lambda x.\bullet^*t$ \\ \hline
+    \multicolumn{3}{c}{{\bf Agents}} \\ 
+        Amy & $e$ & $\hat{R}_a$\\
+        Bob & $e$ & $\hat{R}_b$\\ \hline
+    \multicolumn{3}{c}{{\bf Auxiliary verbs and expletive subjects}} \\        
+        it & $d$ & {\em{dummy}} \\
+        is & $(d \backslash t)\slash(d \backslash_D t)$ & $\lambda Px.Px$ \\
+        it's & $(t \slash(d \backslash_D t)$ & $\lambda Px.Px$ \\
+        isn't & $t\slash(d \backslash_D t)$ & $\lambda Px.\neg P x$ \\  
+        doesn't & $(e \backslash t)\slash(d \backslash_V t)$ & $\lambda Px.\neg Px$ \\  \hline        
+    \multicolumn{3}{c}{{\bf Tensed attitude verbs}} \\        
+        knows & $(e \backslash t)\slash_M t$ &  $\lambda pR.p\vee{\Diamond}_Rp$\\
+        believes & $(e \backslash t)\slash_M t$ &  $\lambda pR.{\Diamond}_Rp$\\ \hline
+    \multicolumn{3}{c}{{\bf Base form attitude verbs}} \\        
+        know & $(e \backslash_V t)\slash_M t$ &  $\lambda pR.p\vee{\Diamond}_Rp$\\
+        believe & $(e \backslash_V t)\slash_M t$ &  $\lambda pR.{\Diamond}_Rp$\\ \hline       
+    \multicolumn{3}{c}{{\bf Complementizers}} \\        
+        that  & $(((e \backslash t){\slash_M}t) \backslash (e \backslash t)) \slash t$
+        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
+        whether  & $(((e \backslash t){\slash_M} t) \backslash (e \backslash t)) \slash t$ 
+        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
+        && \multicolumn{1}{r}{$+ {\neg}(m\,p\,R)$} \\ \bottomrule
+    \multicolumn{3}{c}{{\bf Complementizers for base form verbs}} \\        
+        that  & $(((e \backslash_V t){\slash_M}t) \backslash (e \backslash_V t)) \slash t$
+        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
+        whether  & $(((e \backslash_V t){\slash_M} t) \backslash (e \backslash_V t)) \slash t$ 
+        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
+        && \multicolumn{1}{r}{$+ {\neg}(m\,p\,R)$} \\ % \bottomrule    
+      \end{tabular}
 \end{document}
+
+\begin{figure}[!t]
+  \vspace{1em}
+  \centering
+  \footnotesize
+  \begin{tabular}[t]{lcc}
+    \toprule
+    \textsc{Item} & \textsc{Type} & \textsc{Semantics}\\ \midrule
+    \multicolumn{3}{c}{{\bf Statives with expletive subject}} \\ 
+        heads & $d \backslash_D t$ & $\lambda x.\bullet^*h$ \\
+        tails & $d \backslash_D t$ & $\lambda x.\bullet^*t$ \\ \hline
+    \multicolumn{3}{c}{{\bf Agents}} \\ 
+        Amy & $e$ & $\hat{R}_a$\\
+        Bob & $e$ & $\hat{R}_b$\\ \hline
+    \multicolumn{3}{c}{{\bf Auxiliary verbs and expletive subjects}} \\        
+        it & $d$ & {\em{dummy}} \\
+        is & $(d \backslash t)\slash(d \backslash_D t)$ & $\lambda Px.Px$ \\
+        it's & $(t \slash(d \backslash_D t)$ & $\lambda Px.Px$ \\
+        isn't & $t\slash(d \backslash_D t)$ & $\lambda Px.\neg P x$ \\  
+        doesn't & $(e \backslash t)\slash(d \backslash_V t)$ & $\lambda Px.\neg Px$ \\  \hline        
+    \multicolumn{3}{c}{{\bf Tensed attitude verbs}} \\        
+        knows & $(e \backslash t)\slash_M t$ &  $\lambda pR.p\vee{\Diamond}_Rp$\\
+        believes & $(e \backslash t)\slash_M t$ &  $\lambda pR.{\Diamond}_Rp$\\ \hline
+    \multicolumn{3}{c}{{\bf Base form attitude verbs}} \\        
+        know & $(e \backslash_V t)\slash_M t$ &  $\lambda pR.p\vee{\Diamond}_Rp$\\
+        believe & $(e \backslash_V t)\slash_M t$ &  $\lambda pR.{\Diamond}_Rp$\\ \hline       
+    \multicolumn{3}{c}{{\bf Complementizers}} \\        
+        that  & $(((e \backslash t){\slash_M}t) \backslash (e \backslash t)) \slash t$
+        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
+        whether  & $(((e \backslash t){\slash_M} t) \backslash (e \backslash t)) \slash t$ 
+        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
+        && \multicolumn{1}{r}{$+ {\neg}(m\,p\,R)$} \\ \hline
+    \multicolumn{3}{c}{{\bf Complementizers for base form verbs}} \\        
+        that  & $(((e \backslash_V t){\slash_M}t) \backslash (e \backslash_V t)) \slash t$
+        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
+        whether  & $(((e \backslash_V t){\slash_M} t) \backslash (e \backslash_V t)) \slash t$ 
+        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
+        && \multicolumn{1}{r}{$+ {\neg}(m\,p\,R)$} \\ % \bottomrule    
+      \end{tabular}
+    \caption{Categorial grammar lexicon. The first column has a word form.
+    the second column a categorial type, and third column a semantic translation
+    in a language that extends the Epik logical language with lambda.}
+  \label{lexiconfigure}
+\end{figure}
+
+Old version of grammar figure.
+
+  \begin{tabular}[t]{lcc}
+    \toprule
+        \textsc{Item} & \textsc{Type} & \textsc{Semantics}\\ \midrule  
+        Amy & $e$ & $\hat{R}_a$\\
+        Bob & $e$ & $\hat{R}_b$\\
+        it & $d$ & $d$ \\
+        heads & $d \backslash_D t$ & $\lambda x.\bullet^*h$ \\
+        tails & $d \backslash_D t$ & $\lambda x.\bullet^*t$ \\
+        is & $(d \backslash t)\slash(d \backslash_D t)$ & $\lambda P~x.~P~x$ \\
+        knows & $(e \backslash t)\slash_M t$ &  $\lambda p~R.~p\vee{\Diamond}_Rp$\\
+        believes & $(e \backslash t)\slash_M t$ &  $\lambda p~R.~{\Diamond}_Rp$\\
+        that  & $(((e \backslash t){\slash_M}t) \backslash (e \backslash t)) \slash t$
+        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
+        whether  & $(((e \backslash t){\slash_M} t) \backslash (e \backslash t)) \slash t$ 
+        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
+        && \multicolumn{1}{r}{$+ {\neg}(m\,p\,R)$} \\ \bottomrule
+      \end{tabular}
 

### 2021-01-13 tex/pdraft.tex
fix figure 8 and write KA+H section

@@
 \usepackage{enumitem}
+\usepackage{mathtools}
 %\usepackage{caption}
@@
   \footnotesize
-  \begin{tabular}[t]{lcc}
+  %% \begin{tabular}[t]{lcc}
+  %%   \toprule
+  %%       \textsc{Item} & \textsc{Type} & \textsc{Semantics}\\ \midrule
+  %%       Amy & $e$ & $\hat{R}_a$\\
+  %%       Bob & $e$ & $\hat{R}_b$\\
+  %%       it & $d$ & $d$ \\
+  %%       heads & $d \backslash_D t$ & $\lambda x.\bullet^*h$ \\
+  %%       tails & $d \backslash_D t$ & $\lambda x.\bullet^*t$ \\
+  %%       is & $(d \backslash t)\slash(d \backslash_D t)$ & $\lambda P~x.~P~x$ \\
+  %%       knows & $(e \backslash t)\slash_M t$ &  $\lambda p~R.~p\vee{\Diamond}_Rp$\\
+  %%       believes & $(e \backslash t)\slash_M t$ &  $\lambda p~R.~{\Diamond}_Rp$\\
+  %%       that  & $(((e \backslash t){\slash_M}t) \backslash (e \backslash t)) \slash t$
+  %%       & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
+  %%       whether  & $(((e \backslash t){\slash_M} t) \backslash (e \backslash t)) \slash t$
+  %%       & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
+  %%       && \multicolumn{1}{r}{$+ {\neg}(m\,p\,R)$} \\ \bottomrule
+  %% \end{tabular}
+
+  \newcommand{\hdrspc}{0.3em}
+    \begin{tabular}[t]{l>{\hspace{-.8em}}c >{\hspace{-.6em}}c}
     \toprule
     \textsc{Item} & \textsc{Type} & \textsc{Semantics}\\ \midrule
+    \multicolumn{3}{c}{{\bf Statives with expletive subject}} \\[\hdrspc]
+        heads & $d \backslash_D t$ & $\lambda x.\,\bullet^*h$ \\
+        tails & $d \backslash_D t$ & $\lambda x.\,\bullet^*t$ \\ \midrule
+    \multicolumn{3}{c}{{\bf Agents}} \\
         Amy & $e$ & $\hat{R}_a$\\
-        Bob & $e$ & $\hat{R}_b$\\
-        it & $d$ & $d$ \\
-        heads & $d \backslash_D t$ & $\lambda x.\bullet^*h$ \\
-        tails & $d \backslash_D t$ & $\lambda x.\bullet^*t$ \\
-        is & $(d \backslash t)\slash(d \backslash_D t)$ & $\lambda P~x.~P~x$ \\
-        knows & $(e \backslash t)\slash_M t$ &  $\lambda p~R.~p\vee{\Diamond}_Rp$\\
-        believes & $(e \backslash t)\slash_M t$ &  $\lambda p~R.~{\Diamond}_Rp$\\
+        Bob & $e$ & $\hat{R}_b$\\ \midrule
+    \multicolumn{3}{c}{{\bf Auxiliary verbs and expletive subjects}} \\[\hdrspc]
+        it & $d$ & {\em{dummy}} \\
+        is & $(d \backslash t)\slash(d \backslash_D t)$ & $\lambda P\,x.\,P\,x$ \\
+        it's & $(t \slash(d \backslash_D t)$ & $\lambda P\,x.\,P\,x$ \\
+        isn't & $t\slash(d \backslash_D t)$ & $\lambda P\,x.\,\neg P\,x$ \\
+        doesn't & $(e \backslash t)\slash(d \backslash_V t)$ & $\lambda P\,x.\,\neg P\,x$ \\  \midrule
+    \multicolumn{3}{c}{{\bf Tensed attitude verbs}} \\[\hdrspc]
+        knows & $(e \backslash t)\slash_M t$ &  $\lambda p\,R.\,p + {\Diamond}_Rp$\\
+        believes & $(e \backslash t)\slash_M t$ &  $\lambda p\,R.\,{\Diamond}_Rp$\\ \midrule
+    \multicolumn{3}{c}{{\bf Base form attitude verbs}} \\[\hdrspc]
+        know & $(e \backslash_V t)\slash_M t$ &  $\lambda p\,R.\,p + {\Diamond}_Rp$\\
+        believe & $(e \backslash_V t)\slash_M t$ &  $\lambda p\,R.\,{\Diamond}_Rp$\\ \midrule
+    \multicolumn{3}{c}{{\bf Complementizers}} \\[\hdrspc]
         that  & $(((e \backslash t){\slash_M}t) \backslash (e \backslash t)) \slash t$
-        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
+        & $\begin{pmatrix*}[l]
+          \lambda p\,m\,R.\\
+          \,\,{\neg}(m\,({\neg}p)\,R)
+        \end{pmatrix*}$ \\[1.1em]
         whether  & $(((e \backslash t){\slash_M} t) \backslash (e \backslash t)) \slash t$
-        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
-        && \multicolumn{1}{r}{$+ {\neg}(m\,p\,R)$} \\ \bottomrule
-  \end{tabular}
-    \caption{Partial categorial grammar lexicon. The first column has a word form.
-    the second column a categorial type, and third column a semantic translation
-    in a language that extends the Epik logical language with lambda. The complete grammar
-    is in the appendix.}
+        & $\begin{pmatrix*}[l]
+          \lambda p\,m\,R.\\
+                  \,\, {\neg}(m\,({\neg}p)\,R) \\
+          \,\,\,\, +\,{\neg}(m\,p\,R)
+        \end{pmatrix*}$\\ \midrule
+    \multicolumn{3}{c}{{\bf Complementizers for base form verbs}} \\[\hdrspc]
+        that  & $(((e \backslash_V t){\slash_M}t) \backslash (e \backslash_V t)) \slash t$
+        & $\begin{pmatrix*}[l]
+          \lambda p\,m\,R. \\
+          \,\, {\neg}(m\,({\neg}p)\,R)
+        \end{pmatrix*}$ \\[1.1em]
+        whether & $(((e \backslash_V t){\slash_M} t) \backslash (e \backslash_V t)) \slash t$
+        & $\begin{pmatrix*}[l]
+          \lambda p\,m\,R.\\
+                  \,\, {\neg}(m\,({\neg}p)\,R) \\
+          \,\,\,\, +\,{\neg}(m\,p\,R)
+        \end{pmatrix*}$ \\
+        \bottomrule
+      \end{tabular}%%
+    \caption{Categorial grammar lexicon. The first column has a word
+      form. the second column a categorial type, and third column a
+      semantic translation in a language that extends the Epik logical
+      language with lambda. The online supplement includes a version
+      of this grammar that can be run using Barker and Shan's chart
+      parser for categorial grammar \citep{barker2005parser}.}
   \label{lexiconfigure}
@@
 
-The development here is concerned with defining concrete computable possible
-worlds models, and applying them in natural language semantics.  
-Issues for further investigation are mathematical characterizations
-of epistemic KATs, e.g. sound and complete axioms and coalgebra.
+Epik is closely related to Kleene Algebra with Hypotheses\cite{},
+\textsc{Ka+H}, which permits user-specification of futher equations
+beyond the axioms of \textsc{Ka}. In practice, these hypotheses are
+used to express domain-specific properties of the axioms, such as the
+commutativity of certain uninterpreted actions in program
+verification, such as $x := 3; y := 5 \equiv y := 5; x := 3$). We can
+model both the state formulae and the event formulae as \textsc{KA+H}
+hypotheses. Specifically, the state formula $\varphi$ can be thought
+of as a hypothesis $\varphi \equiv 1$, and an event formula $\zeta_a$
+for event $a$, is simply the hypothesis $a \equiv
+\sum_{\varphi,\varphi' \in \llbracket \zeta_a \rrbracket}
+\varphi;a;\varphi'$. Notwithstanding some early results about
+\textsc{Kat} with hypotheses of the form $cp=c$~\cite{}, using basic
+\textsc{KA} implementations would rely on using the exponential
+encoding of \textsc{Kat} into \textsc{Ka}~\cite{}.
+
+The key facet that differentiates Epik from \textsc{Ka+H} is the
+modality. Each alternative relation represents a set of
+uni-directional rewrites that must be applied for that agent, rather
+than the universal, optional rewrites represented by the \textsc{Ka+H}
+hypotheses. It may be possible to encode Epik's alternative relations
+into \textsc{Ka+H} hypotheses, but the translation would be
+non-trivial, and we leave it as future work.
+
+The development here is concerned with defining concrete computable
+possible worlds models, and applying them in natural language
+semantics. Issues for further investigation are mathematical
+characterizations of epistemic KATs, e.g. sound and complete axioms
+and coalgebrae.
 
@@
 
-\section*{Appendix}
-Complete categorial grammar extending Figure \ref{lexiconfigure}.  The online supplement
-includes a version of this grammar that can be run using Barker and Shan's
-chart parser for categorial grammar \citep{barker2005parser}.
-
-  \begin{tabular}[t]{lcc}
-    \toprule
-    \textsc{Item} & \textsc{Type} & \textsc{Semantics}\\ \midrule
-    \multicolumn{3}{c}{{\bf Statives with expletive subject}} \\ 
-        heads & $d \backslash_D t$ & $\lambda x.\bullet^*h$ \\
-        tails & $d \backslash_D t$ & $\lambda x.\bullet^*t$ \\ \hline
-    \multicolumn{3}{c}{{\bf Agents}} \\ 
-        Amy & $e$ & $\hat{R}_a$\\
-        Bob & $e$ & $\hat{R}_b$\\ \hline
-    \multicolumn{3}{c}{{\bf Auxiliary verbs and expletive subjects}} \\        
-        it & $d$ & {\em{dummy}} \\
-        is & $(d \backslash t)\slash(d \backslash_D t)$ & $\lambda Px.Px$ \\
-        it's & $(t \slash(d \backslash_D t)$ & $\lambda Px.Px$ \\
-        isn't & $t\slash(d \backslash_D t)$ & $\lambda Px.\neg P x$ \\  
-        doesn't & $(e \backslash t)\slash(d \backslash_V t)$ & $\lambda Px.\neg Px$ \\  \hline        
-    \multicolumn{3}{c}{{\bf Tensed attitude verbs}} \\        
-        knows & $(e \backslash t)\slash_M t$ &  $\lambda pR.p\vee{\Diamond}_Rp$\\
-        believes & $(e \backslash t)\slash_M t$ &  $\lambda pR.{\Diamond}_Rp$\\ \hline
-    \multicolumn{3}{c}{{\bf Base form attitude verbs}} \\        
-        know & $(e \backslash_V t)\slash_M t$ &  $\lambda pR.p\vee{\Diamond}_Rp$\\
-        believe & $(e \backslash_V t)\slash_M t$ &  $\lambda pR.{\Diamond}_Rp$\\ \hline       
-    \multicolumn{3}{c}{{\bf Complementizers}} \\        
-        that  & $(((e \backslash t){\slash_M}t) \backslash (e \backslash t)) \slash t$
-        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
-        whether  & $(((e \backslash t){\slash_M} t) \backslash (e \backslash t)) \slash t$ 
-        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
-        && \multicolumn{1}{r}{$+ {\neg}(m\,p\,R)$} \\ \bottomrule
-    \multicolumn{3}{c}{{\bf Complementizers for base form verbs}} \\        
-        that  & $(((e \backslash_V t){\slash_M}t) \backslash (e \backslash_V t)) \slash t$
-        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
-        whether  & $(((e \backslash_V t){\slash_M} t) \backslash (e \backslash_V t)) \slash t$ 
-        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
-        && \multicolumn{1}{r}{$+ {\neg}(m\,p\,R)$} \\ % \bottomrule    
-      \end{tabular}
+%% \section*{Appendix}
+%% Complete categorial grammar extending Figure \ref{lexiconfigure}.  The online supplement
+%% includes a version of this grammar that can be run using Barker and Shan's
+%% chart parser for categorial grammar \citep{barker2005parser}.
+
 \end{document}

### 2021-01-14 tex/pdraft.tex
General edits.

@@
 
-\title{Epistemic Semantics in Guarded String Models}
+\title{Epistemic Semantics in Guarded String Models\thanks{Thanks guys.}}
+\renewcommand\footnotemark{}
+%%\renewcommand\footnoterule{}
 
@@
   \item[\atHT] Amy secretly turns the coin from $H$ to $T$.
-  \item[\atTH] Bob secretly turns the coin from $T$ to $H$.
+  \item[\btTH] Bob secretly turns the coin from $T$ to $H$.
   \item[\btHT] Bob secretly turns the coin from $H$ to $T$.
@@
 Boolean formulas, with $h$ being the pre-condition of $\atHT$ and
-$a_{01}$ being the post-condition. This is expressed using an operator
+$t$ being the post-condition. This is expressed using an operator
 ``$:$'' (read ``and next'') that pairs Boolean formulas. The formula
@@
 to tails) as happening only in an $h$ state, and concluding in a
-not-$h$ state.  Events don't have to change state: the event $\apT$
+$t$ state.  Events don't have to change state: the event $\apT$
 (Amy peeking at tails) can happen only in a $t$ state, and does
@@
 the formula $\bar t$. We will further restrict the feasible conditions
-for our actions by restricting the space of valuations for our formulae.
+for our actions by restricting the space of valuations for our formulas.
 
@@
 valuations $h\bar{t}$ and $\bar{h}t$, and disallow $ht$ and
-$\bar{h}\bar{t}$.  This is enforced by a {\em state formula}, which is
+$\bar{h}\bar{t}$.\footnote{Two Booleans constrained by a state formula, rather than one Boolean, are stipulated here to
+ illustrate state formulas.
+}  This is enforced by a {\em state formula}, which is
 a Boolean formula, in this case the one given on the second line of
-\ex{effect}.  Where {\sf{B}} is a set of atomic tests and $\phi$ is a
-state constraint over {\sf{B}},
+\ex{effect}.  Where {\sf{B}} is a set of state primitives and $\phi$ is a
+ constraint over {\sf{B}},
 ${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\phi}$ is the set of valuations of
@@
 turns the coin over, in event $b_{10}$ or $b_{01}$, she doesn't know
-anything has happened, so her alternative is the no-operation,
-no-information event event $o$. Bob's event relation is symmetric.
+anything has happened, so her alternative is the ``no-operation'' or
+``no-information'' event event $o$. Bob's event relation is symmetric.
 
@@
 
-This paper focuses on defining a concrete possible worlds model from
-an Epik specification.  The models are an extension of guarded-string
-models for Kleene Algebra with Tests (KAT). This is an algebraic
-theory that has model classes including guarded string models,
-relational models, finite models, and matrix models.  Our definitions
-and notation follow \cite{kozen2001automata}.  We add syntax and
+This paper focuses on defining a concrete possible worlds model from an Epik
+specification.  The models are an extension of guarded-string models for Kleene
+Algebra with Tests (KAT). This is an algebraic theory that has model classes
+including guarded string models, relational models, finite models, and matrix models.
+Our definitions and notation follow \cite{kozen2001automata}.  We add syntax and
 semantics to cover multi-agent epistemic semantics.
 
-Guarded strings over a finite alphabet ${\sf{P}}$
-are like ordinary strings, but with atoms over a set ${\sf{B}}$  
-alternating with the symbols
-from ${\sf{P}}$.  In the algebra described by Figure \ref{epikfigure}, 
-${\sf{P}}$ is the set of events $\{a_1,a_0,b_1,b_0,a_{10},a_{01},b_{10},b_{01}\}$, and
-${\sf{B}}$ is $\{h,t\}$.
-
-In the coin example, as we already saw in \ex{effect},
-$\mathcal{A}^\varphi_{\mathsf B}$ is $\{h\bar{t},\bar{h}t\}$, for
-which we use the shorthand $\{H,T\}$.  A guarded string over
-${\sf{P}}$ and ${\sf{B}}$ is a strings of events from ${\sf{P}}$,
-alternating with atoms over ${\sf{B}}$, and beginning and ending with
-atoms.  In this construction, $w_1=H$, $w_2=Ha_1H$, $w_3=Ha_1Hb_1H$,
-and $w_4=Ha_1Hb_1Ha_{10}Tb_{01}H$.
-
-The discussion of \ex{coin2} mentioned building worlds by incrementing
-worlds with events.  This is accomplished in guarded string models
-with fusion product $\diamond$, a partial operation that combines two
-guarded strings, subject to the condition that the atom at the end of
-the the first argument is identical to the atom at the start of the
-second one.  \ex{fusion1} gives some examples.
+Guarded strings over a finite alphabet ${\sf{P}}$ are like ordinary strings, but with
+atoms over a set ${\sf{B}}$ alternating with the symbols from ${\sf{P}}$.  In the
+algebra described by Figure \ref{epikfigure}, ${\sf{P}}$ is the set of events
+$\{a_1,a_0,b_1,b_0,a_{10},a_{01},b_{10},b_{01},o\}$, and ${\sf{B}}$ is $\{h,t\}$.  As
+we already saw in \ex{effect}, $\mathcal{A}^\varphi_{\mathsf B}$ is
+$\{h\bar{t},\bar{h}t\}$, for which we use the shorthand $\{H,T\}$.  A guarded string
+over ${\sf{P}}$ and ${\sf{B}}$ is a string of events from ${\sf{P}}$, alternating
+with atoms over ${\sf{B}}$, and beginning and ending with atoms.  In this
+construction, $w_1=H$, $w_2=Ha_1H$, $w_3=Ha_1Hb_1H$, and
+$w_4=Ha_1Hb_1Ha_{10}Tb_{01}Hb_1H$.
+
+The discussion of \ex{coin2} mentioned building worlds by incrementing worlds with
+events.  This is accomplished in guarded string models with fusion product
+$\diamond$, a partial operation that combines two guarded strings, subject to the
+condition that the atom at the end of the the first argument is identical to the atom
+at the start of the second one.  \ex{fusion1} gives some examples.
 
@@
 
-\hspace{2mm}(ii) $\element{\tuple{\alpha_i,\alpha_{i+1}}}{\sem{\zeta_{ei}}^{\varphi}}$, ($0{\leq} i {\le} n$).
+\hspace{2mm}(ii) $\element{\tuple{\alpha_i,\alpha_{i+1}}}{\sem{\zeta_{e_i}}^{\varphi}}$, ($0{\leq} i {\le} n$).
 \end{example}
@@
 guarded strings have the addional interpretation of event types.  An event-type
-is something that can ``happen'' is different worlds. For example, $a_1$ has the
-event type $\{H\,a_1\,H\}$, and $a_0$ has the event type $\{T\,a_0\,T\}$.
+is something that can ``happen'' is different worlds. For example, $a_1$ corresponds to the 
+event type $\{H\,a_1\,H\}$, and $a_0$ corresponds to the event type $\{T\,a_0\,T\}$. Event
+types need not be singletons.
 
@@
 finite set of worlds.  The next step is to define an alternative relation $R_a$ on
-worlds for each agent $a$. This will result in a Kripke frame $\tuple{W,R_1,...,R_n}$
-consisting of a set of worlds, and a world-alternative relation for each agent
-\citep{kripke1963semantical}.  An Epik specification defines an alternative relation
-on bare events for each agent $a$, which we notate as ${R}_{a}$.  This should be
-lifted to a relation $\hat{R}_a$ on worlds.  The basic idea is that when a world $w$
-is incremented with an event $e$, in the resulting world $w \diamond e$, epistemic
-alternatives for agent $a$ are of the form $w' \diamond e'$, where $w'$ is an
-alternative to for $a$ in $w$, and $e'$ is and event-alternative to $e$ for
-$a$.\footnote{In this it is important that the event-alternative relation for an
-  agent is constant across worlds. We anticipate that the definition given here
-  produces results equivalent to what is found in literature on event alternatives in
-  dynamic epistemic semantics, though we have not verified this.  That literature
-  primarily focuses on mapping an epistemic model for a single time and situation to
-  another, and uses general first-order models, rather than guarded string
-  models. See \citet{baltag1999logic}, \citet{van2007dynamic}, and articles in
-  \citet{ditmarsch2015handbook}.  Previous literature is motivated by epistemic logic
-  and AI planning, rather than computable possible worlds models in natural language
-  semantics.  } This needs to be implemented in a way that takes account of pre- and
-post-conditions for events. For this, our approach is to refer the definition of
-well-formed guarded strings.  \ex{raise1} defines a relation on worlds from a
-relation on bare events.
+worlds for each agent $a$. This will result in a general modal frame
+$\tuple{W,R_1,...,R_n,K}$ consisting of a set of worlds, a world-alternative relation
+for each agent, and a set $K$ of propositions, where each proposition is a subset of
+$W$ \citep{chagrov1997modal}.\footnote{As explained in the next section, $K$ will not
+  be the power set of $W$, rather it will consists of the regular subsets of $W$.}
+An Epik specification defines an alternative relation on bare events for each agent
+$a$, which we notate as ${R}_{a}$.  This should be lifted to a relation $\hat{R}_a$
+on worlds.  The basic idea is that when a world $w$ is incremented with an event $e$,
+in the resulting world $w \diamond e$, epistemic alternatives for agent $a$ are of
+the form $w' \diamond e'$, where $w'$ is an alternative to for $a$ in $w$, and $e'$
+is and event-alternative to $e$ for $a$.\footnote{In this it is important that the
+  event-alternative relation for an agent is constant across worlds. We anticipate
+  that the definition given here produces results equivalent to what is found in
+  literature on event alternatives in dynamic epistemic semantics, though we have not
+  verified this.  That literature primarily focuses on mapping an epistemic model for
+  a single time and situation to another, and uses general first-order models, rather
+  than guarded string models. See \citet{baltag1999logic}, \citet{van2007dynamic},
+  and articles in \citet{ditmarsch2015handbook}.  This literature is motivated by
+  epistemic logic and AI planning, rather than computable possible worlds models in
+  natural language semantics.  } This needs to be implemented in a way that takes
+account of pre- and post-conditions for events. For this, our approach is to refer
+the definition of well-formed guarded strings.  \ex{raise1} defines an epistemic
+alternative relation on worlds from an alternative relation on bare events.
 
@@
 \)
-in $W$ and a guarded string  $q$ iff $q$ is an element of $W$ and is of the form
+in $W$ and a guarded string  $q$ in $W$ iff $q$ is of the form
 \(
 \alpha_0'e_0' ... e_n'\alpha_{n+1}', 
-\), where for $0 \leq n$, $\element{\tuple{e_i,e'_i}}{{R}}$.
+\), where for $0 \leq n$, $e_iRe'_i$.
 \end{example}
@@
 This requires that in an alternative world, each constituent event $e'_i$ is an
-alternative to the event $e_i$ in the base world.
+alternative to the corresponding event $e_i$ in the base world.
 Compatibilities between events in the alternative world are enforced by the
@@
 With modalities and propositional negation added, the signature of $n$-agent
-epistemic KAT is $\tuple{K,+,\cdot,*,\bar{},0,1,\neg,\Diamond_1
-  ... \Diamond_n}$. Figure \ref{termfigure} defines the syntax of the language.
+epistemic KAT is $\tuple{K,+,\cdot,*,\bar{},0,1,\neg,\Diamond_1,
+  ..., \Diamond_n}$. Figure \ref{termfigure} defines the syntax of the language.
 Juxtaposition is used for product.  Terms in this language
@@
 \hspace{-1.5em}\begin{tabular}[t]{l}
-  $\Box_b(\Box_a\bullet^*t + \Box_a\neg\bullet^*t)$\\
+  $\Box_b(\bullet^*t \wedge \Box_a\bullet^*t + \neg\bullet^*t \wedge \Box_a\neg\bullet^*t)$\\
  Bob believes Amy knows whether it's tails.  \\
@@
 
-Summing up, given an Epik specification $\epikspec$, term $p$ (as
-defined syntactically in Figure \ref{termfigure}) is interpreted as a
-set of guarded strings $\sem{p}^{\epikspec}$.  Let $K^{\epikspec}$ be
-the sets that are interpretations of terms.  Then
+Summing up, given an Epik specification $\epikspec$, term $p$ (as defined
+syntactically in Figure \ref{termfigure}) is interpreted as a set of guarded strings
+$\sem{p}^{\epikspec}$.  Let $K^{\epikspec}$ be the sets that are interpretations of
+terms.  Then
 $\tuple{K^{\epikspec},+,\cdot,*,\bar{},0,1,\neg,\Diamond_{a_1},...,\Diamond_{a_n}}$
-is a concrete guarded string interpretation for the signature of
-epistemic KAT, with operations as in Figure \ref{fig:interp2}
-(e.g. the binary operation $+$ is union, and the unary operation
-$\Diamond_a$ is pre-image relative to $\hat{R}_a$).  This provides a
-concrete $n$-agent Kripke frame $\tuple{\sem{\bullet^*}^{\epikspec},
-  \hat{R}_1,...,\hat{R}_n}$.\footnote{ The domain of the Kripke frame
-  differs from the domain of the guarded string model, because the
-  former is the set of worlds, while the latter is the set of
-  propositions.  } The frame consists of a set of worlds, and an
-epistemic-alternative relation for each agent.  It is used as a target
+is a concrete guarded string interpretation for the signature of epistemic KAT, with
+operations as in Figure \ref{fig:interp2} (e.g. the binary operation $+$ is union,
+and the unary operation $\Diamond_a$ is pre-image relative to $\hat{R}_a$).  This
+provides a concrete $n$-agent general modal frame
+$\tuple{\sem{\bullet^*}^{\epikspec}, \hat{R}_1,...,\hat{R}_n,K^{\epikspec}}$.  The
+frame consists of a set of worlds, an epistemic-alternative relation for each agent,
+and a set of propositions, each of which is a set of worlds.  It is used as a target
 for natural-language interpretation in Section 6.
@@
 allow for the definition of functions on regular sets and relations
-\citep{hulden2009foma,karttunen2010update}.  Such definitions are used
+\citep{hulden2009foma,linden2009hfst,karttunen2010update}.  Such definitions are used
 here to construct of a model for epistemic KAT inside the finite state
-calculus. We describe our translation from Epik terms to Fst programs
+calculus. We describe our translation from Epik logical terms to Fst terms
 here.
@@
   {\tt St} \hspace{2mm} \\
-  Atomic Tests such as $0110$.% The length is the number of generators.
+  Atoms such as $0110$.% The length is the number of generators.
 \end{hangingpar}
@@
   {\tt UnequalStPair} \hspace{2mm} \\
-  Sequence of two unequal tests such as $0110~0111$.
+  Sequence of two unequal atoms such as $0110~0111$.
 \end{hangingpar}
@@
 \end{verbatim}
-String that doesn't contain a non-matching test pair.
+String that doesn't contain a non-matching pair of atoms.
 \end{hangingpar}
@@
 \end{verbatim}
-Rewrite relation deleting the second of two tests.
+Rewrite relation deleting the second of two atoms.
 \end{hangingpar}
@@
 straightforward: we simply convert the atoms as previously described,
-decorate the events $e$ with their compatible atoms. For example $a_1$ becomes
+and decorate the events $e$ with their compatible atoms. For example $a_1$ becomes
 an Fst term denoting
 $\{01a_101\}$. Fst has built-in operations of union (\texttt{|}),
-and intersection (\texttt{\&}), which define
-the sum and intersection operations in the guarded string algebra.
+and intersection (\texttt{\&}), which are
+the sum and intersection operations in the guarded string algebra as represented
+in Fst.
 Fst set difference (\texttt{-}) is used to define propositional complement as the
@@
 KAT Kleene plus is defined in a similar way using Kleene plus in the string algebra,
-with checks for equality of atoms and deletion of atoms. See Figure \ref{fstfigure1}.
+with checks for equality of atoms and deletion of atoms. 
 KAT Kleene star is definied from KAT Kleene plus and the multiplicative identity,
@@
   $b$ & Constrain domain and co-domain to contain \\
-      & no unmatched tests. \\
-  $c$ & Reduce doubled tests to a single \\
-      & test in the domain and co-domain. \\
+      & no unmatched atoms. \\
+  $c$ & Reduce doubled atoms to a single \\
+      & atom in the domain and co-domain. \\
 \end{tabular}
@@
 
-%% An event symbol such as $a_1$ (Amy looks at heads) is in the KAT algebra a set of
-%% bare events decorated with compatible tests on each side, $\{ 10 a_1 10
-%% \}$ in this case.  This is a unit set rather than a guarded string, because elements
-%% of the KAT algebra are sets. Worlds in the KAT algebra are defined by sequencing
-%% events using {\em Kst}. The operation enforces compatibility of states, so that
-%% $(a_1 + a_0)(b_1 + b_0)$ contains two worlds rather than four.  The program in
-%% Figure \ref{epikfigure} as interpreted in FST defines a countably infinite set of possible worlds by
-%% KAT Kleene closure as $\mbox{\em Kst}(a_1 + a_0 +b_1 + b_0)$.
-
 It remains to define an epistemic alternative relation on worlds for each agent.  The
@@
 each agent. This determines a relation between bare events decorated with compatible
-atoms. 
-In Fst, we use the closure of the concatenation
-product operation on relations to lift a relation on decorated events for an agent to
-the corresponding relation on worlds.  The concatenation product $R~S$ of two
-relations $R$ and $S$ is the set of pairs of the form $\pair{x_1x_2}{y_1y_2}$, where
-$x_1 \mathrel R y_1$, and $x_2 \mathrel S y_2$.  In Fst, {\tt{R+}} is the closure of relation {\tt R} with
+atoms.  In Fst, we use the closure of the concatenation product operation on
+relations to lift a relation on decorated events for an agent to the corresponding
+relation on worlds.  The concatenation product $R~S$ of two relations $R$ and $S$ is
+the set of pairs of the form $\pair{x_1x_2}{y_1y_2}$, where $x_1 \mathrel R y_1$, and
+$x_2 \mathrel S y_2$.  In Fst, {\tt{R+}} is the closure of relation {\tt R} with
 respect to this operation. Figure \ref{relclosure} defines the corresponding
-operation on sets of guarded strings as encoded in Fst.\footnote{Relation concatenation in Fst
-  differs from relation composition ({\tt{.o.}}), and the closure under discussion here is
-  the closure of the former rather than the latter.}  The epistemic alternative
-relation on worlds for an agent is then defined as the KAT relational concatenation
-closure {\tt RelKst} of the decorated-event alternative relation for the agent.
+operation on sets of guarded strings as encoded in Fst.\footnote{Relation
+  concatenation in Fst differs from relation composition ({\tt{.o.}}), and the
+  closure under discussion here is the closure of the former rather than the latter.}
+The epistemic alternative relation on worlds for an agent is then defined as the KAT
+relation concatenation closure {\tt RelKst} of the decorated-event alternative
+relation for the agent.
 
@@
 
-English sentences are mapped to terms in the logical language via a
-semantically interpreted multi-modal categorial grammar, consisting of
-a lexicon of words, their categorial types, and interpretations in a
-logical lambda language.  The grammar covers basic statives ({\em it's
-  heads}), {\em that-} and {\em whether}-complements of {\em know}, predicate and
-sentence negation, and predicate and sentence conjunction.  Figure
-\ref{lexiconfigure} gives illustrative lexical
-entries.\footnote{Category symbols use Lambek/Bar-Hillel notation for
-  slashes, so that $(d \backslash t)\slash(d \backslash_D t)$ combines
-  with $d \backslash_D t$ on the right to give a value that combines
-  with $d$ on the left to give $t$. In the semantics, lambda abstractions with
-  multiple parameters are written $\lambda x~y.~e$ rather than $\lambda x.\lambda y.e$.
-$d$ is the category of {\em it}.}
-The grammar and semantics are
-optimized for a simple fragment of English concerned with clausal
-complementation.  The agent names {\em Amy} and {\em Bob} contribute
-the epistemic alternative relations for those agents, rather than
-individuals.  The root verb {\em know} contributes existential modal
-force.  The complementizers {\em that} and {\em whether} are the heads
-of their dominating clauses, and assemble an alternative relation,
-modal force, and proposition contributed by the complement.  These
-complementizers introduce the dual via two negations, in order to
-express universal modal force.
+English sentences are mapped to terms in the logical language via a semantically
+interpreted multi-modal categorial grammar, consisting of a lexicon of words, their
+categorial types, and interpretations in a logical lambda language.  The grammar
+covers basic statives ({\em it's heads}), {\em that-} and {\em whether}-complements,
+negation, and predicate and sentence
+conjunction.\marginpar{Conjunction missing!}  Figure \ref{lexiconfigure} gives the
+lexicon.\footnote{Category symbols use Lambek/Bar-Hillel notation for slashes, so
+  that $(d \backslash t)\slash(d \backslash_D t)$ combines with $d \backslash_D t$ on
+  the right to give a value that combines with $d$ on the left to give $t$. In the
+  semantics, lambda abstractions with multiple parameters are written $\lambda
+  x~y.~e$ rather than $\lambda x.\lambda y.e$.  $d$ is the category of expletive {\em it}.}
+The grammar and semantics are optimized for a simple fragment of English concerned
+with clausal complementation.  The agent names {\em Amy} and {\em Bob} contribute the
+epistemic alternative relations for those agents, rather than individuals.  The root
+verb {\em believe} contributes existential modal force.  The complementizers {\em that}
+and {\em whether} are the heads of their dominating clauses, and assemble an
+alternative relation, modal force, and proposition contributed by the complement.
+These complementizers introduce the dual via two negations, in order to express
+universal modal force.
 
@@
 The body of $\lambda p.\lambda R. {\Diamond}_R p$, which is the semantic lexical
-entry of {\em knows}, is an term denoting the pre-image of the
-world-alternative relation contributed by the subject.  This is not the right
-semantics for {\em Amy knows that it's heads}, because it has an existential
-modality ${\Diamond_R}p$, rather than an universal modality ${\Box_R}p$.  This
-is corrected by the complementizer {\em that} or {\em whether}, which introduces
-the dual.
+entry of {\em believes}, is an term denoting the pre-image of the world-alternative
+relation contributed by the subject.  This is not the right semantics for {\em Amy
+  believes that it's heads}, because it has an existential modality ${\Diamond_R}p$,
+rather than an universal modality ${\Box_R}p$.  This is corrected by the
+complementizer {\em that}, which introduces the dual.
 
@@
 complex phrases are obtained by application of semantic translations, accompanied by
-beta reductions that eliminate all lambdas logical forms for clauses.  In
-consequence, the semantic term translating a sentence is an Epik term.  Such a term
-designates a set of possible words (guarded strings).  By way of example,
+beta reductions that eliminate all lambdas in logical forms for clauses.  In
+consequence, the semantic term translating a sentence is an Epik logical term.  Such
+a term designates a set of possible words (guarded strings).  By way of example,
 \exi{three}{a} is an English sentence with conjunction and several levels of clausal
@@
 \exi{three}{b} using logical equivalences. Either term is compiled in an
-implementation of the finite state calculus to a finite state machine with 10
-states and 23 edges, which accepts a countably infinite set of worlds.\footnote{
-Machine sizes need not be small, especially as the cardinality of $\sf{B}$
-increases.  A certain Epik model with fourteen primitive tests has the set of worlds
-representented by a finite state machine with 184794 states and 257881
-edges.}
-In this way the methodology ``directly'' represents the set of worlds
-denoted by \exi{three}{a}.
+implementation of the finite state calculus to a finite state machine with 10 states
+and 23 edges, which accepts a countably infinite set of worlds.\footnote{ Machine
+  sizes need not be small, especially as the cardinality of $\sf{B}$ increases.  A
+  certain Epik model with fourteen primitive Booleans has the set of worlds
+  representented by a finite state machine with 184,794 states and 257,881 edges.}  In
+this way the methodology ``directly'' represents the set of worlds denoted by
+\exi{three}{a}.
 
@@
 The methodology presented here is designed for use in research in linguistic
-semantics, for research in computational syntax and semantics on model-theoretically
+semantics, for research in computational linguistics on model-theoretically
 grounded semantics, 
@@
   combinators, type raising, continuations, or multimodal slashes that
-  interact with the grammar in non-trivial ways.  These features would
+  interact with the grammar in non-trivial ways.  These features 
   become relevant in more sophisticated grammars.  Multimodal
   categories are used for phrases such as a tenseless verb phrase that
-  semantically are functional, but can combine in the grammar only as
+  semantically are functional, but combine in the grammar only as
   arguments.}  While this way of proceeding is simple and attractive
 for the applications we are interested in, it would be possible to use
-our semantics with other frameworks that have a semantics that uses a
+our semantics with other frameworks that have a semantics with a
 lambda extension of a logical language.
@@
 worlds.  Semantic type systems for natural language usually include a
-type for individuals.  In the coin example, the individuals are hidden
+type for individuals \citep{gallin1975intensional}.  In the coin example, the individuals are hidden
 in the primitive event symbols such as $a_1$ and $b_1$. This situation
@@
 construction on grounded event terms such as
-$\Prime{look}(\Prime{a},\Prime{k},1)$, for ``agent $\Prime{a}$ see
-that coin $\Prime{k}$ is heads'', rather than atomic events.  This way
+$\Prime{look}(\Prime{a},\!\Prime{k},\!\Prime{1})$, for ``agent $\Prime{a}$ see
+that coin $\Prime{k}$ is heads'', rather than atomic event symbols.  This way
 of proceeding is found in research on situation calculus
@@
 Epik specifications and into the computational parts of the proposal
-is a topic for future research.  In this it is not immediately clear
+is a topic for future research.  In this it is not clear
 whether it is possible to introduce quantification over individuals,
-together with individual constants, while maintaining the finite state
-semantics. \citet{rooth2017finite} develops an approach to a positive
+together with individual constants. \citet{rooth2017finite} develops an approach to a positive
 answer, based on introducing markers for witnesses for discourse
@@
 
-\clearpage 
 \bibliography{fs}
-
-\clearpage
-
-\onecolumn
 

### 2021-01-14 tex/pdraft.tex
add citations

--- 2021-01-13 tex/pdraft.tex
@@
 
-Epik is closely related to Kleene Algebra with Hypotheses\cite{},
-\textsc{Ka+H}, which permits user-specification of futher equations
-beyond the axioms of \textsc{Ka}. In practice, these hypotheses are
-used to express domain-specific properties of the axioms, such as the
-commutativity of certain uninterpreted actions in program
-verification, such as $x := 3; y := 5 \equiv y := 5; x := 3$). We can
-model both the state formulae and the event formulae as \textsc{KA+H}
-hypotheses. Specifically, the state formula $\varphi$ can be thought
-of as a hypothesis $\varphi \equiv 1$, and an event formula $\zeta_a$
-for event $a$, is simply the hypothesis $a \equiv
-\sum_{\varphi,\varphi' \in \llbracket \zeta_a \rrbracket}
-\varphi;a;\varphi'$. Notwithstanding some early results about
-\textsc{Kat} with hypotheses of the form $cp=c$~\cite{}, using basic
-\textsc{KA} implementations would rely on using the exponential
-encoding of \textsc{Kat} into \textsc{Ka}~\cite{}.
+Epik is closely related to Kleene Algebra with Hypotheses
+\textsc{Ka+H}~\cite{cohen1994hypotheses}, which permits
+user-specification of futher equations beyond the axioms of
+\textsc{Ka}. In practice, these hypotheses are used to express
+domain-specific properties of the axioms, such as the commutativity of
+certain uninterpreted actions in program verification, such as $x :=
+3; y := 5 \equiv y := 5; x := 3$). We can model both the state
+formulae and the event formulae as \textsc{KA+H} hypotheses.
+Specifically, the state formula $\varphi$ can be thought of as a
+hypothesis $\varphi \equiv 1$, and an event formula $\zeta_a$ for
+event $a$, is simply the hypothesis $a \equiv \sum_{\varphi,\varphi'
+  \in \llbracket \zeta_a \rrbracket} \varphi;a;\varphi'$. In general,
+the Horn theory of \textsc{KA+H} is undecidable, however many special
+cases for \textsc{Ka} and for \textsc{Kat} are
+decidable~\cite{doumane2019kleene, hardin2002elimination,
+  kozen1997kleenecompleteness}.
 

### 2021-01-14 tex/pdraft.tex

--- 2021-01-14 tex/pdraft.tex
@@
 
-\author{Eric Campbell \\
+\author{Eric Hayden Campbell \\
   Cornell University\\
@@
   models in an epistemic extension of Kleene Algebra with Tests.  The
-  account is framed as a formal language Epik (Epistemic KAT) for
+  account is framed as a formal language Epik (Epistemic \textsc{Kat}) for
   defining such models.  The language is implemented by translation
@@
 
-The approach taken here synthesizes five antecedents in a certain way.
+The approach taken here synthesizes five antecedents.
 John McCarthy's {\em Situation Calculus} is the source of the idea of
@@
 to express intensions \cite{fernando2017intensions}.  Work on {\em
-  finite state intensiononal semantics} has investigated how to do the
+  finite state intensional semantics} has investigated how to do the
 semantics of intensional complementation in a setting where
@@
 
-A running example of an event-sequence model is {\em The Concealed
-  Coin}.  Amy and Bob are seated at a table.  There is a coin on the
-table under a cup, heads up.  The coin could be heads-up $H$ or
-tails-up $T$, and neither agent knows which it is.  This initial
-situation is possible world $w_1$.  Two additional worlds $w_2$ and
-$w_3$ are defined by sequencing events after the initial state, with
-events interpreted as in \ex{coin1}. The truth values for English
-sentences shown in \ex{coin3} are observed, where $0$ stands for
-falsity and $1$ for truth.
+A running example of an event-sequence model is {\em The Concealed Coin}. Amy
+and Bob are seated at a table. There is a coin on the table under a cup, heads
+up. The coin could be heads-up $H$ or tails-up $T$, and neither agent knows
+which it is. This initial situation is possible world $w_1$. We can construct
+additional worlds by sequencing events like those shown in \ex{coin1}.
 
@@
 
+The worlds $w_1,\ldots,w_4$ in \ex{coin2} are examples constructed from the
+events in \ex{coin1}. Juxtaposition indicates the order and occurence of events,
+for example $w_3$ indicates that after the events of $w_2$, event $b_1$ occurs,
+fully, this indicates beginning in world $w_1$, Amy surreptitiously peeking at
+a heads-up coin, and then Bob doing the same.
+
 \begin{example} \label{coin2}
-  \(w_2 = w_1\apH \hfill
-  w_3 = w_2\bpH \hfill
-  w_4 = w_3\atHT\btTH\bpH\)
+  \(w_2 \triangleq w_1\apH \hfill
+  w_3 \triangleq w_2\bpH \hfill
+  w_4 \triangleq w_3\atHT\btTH\bpH\)
 \end{example}
+
+The truth values for English sentences are observed at the end of the event
+sequences specified by $w_1,\ldots, w_4$, where $0$ stands for falsity and $1$
+for truth.
 
@@
 a Boolean formula, in this case the one given on the second line of
-\ex{effect}.  Where {\sf{B}} is a set of state primitives and $\phi$ is a
+\ex{effect}.  Where {\sf{B}} is a set of state primitives and $\varphi$ is a
  constraint over {\sf{B}},
-${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\phi}$ is the set of valuations of
-{\sf{B}} that make formula $\phi$ true. Valuations are called atoms,
+${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\varphi}$ is the set of valuations of
+{\sf{B}} that make formula $\varphi$ true. Valuations are called atoms,
 because they correspond to the atoms of a Boolean algebra of tests
@@
 beginning with {\tt event} declare events and their effect formulas.
-Finally the lines beginning with {\tt agent} define {\em event
-  alternative} relations for agents.  Each clause with an arrow has a
-single event symbol on the left, and a disjunction of alternative
-events on the right of the arrow. The interpretation of Amy's
-alternatives for $b_1$ (Bob peeks at heads), is that when $b_1$
-happens, for Amy either $b_1$ or $b_0$ (Bob peeks at tails) could be
-happening. Her alternatives for $a_{01}$ and $a_{10}$ (she turns the
-coin over in one state or the other) are $a_{10}$ and $a_{01}$,
-indicating that she doesn't know, \emph{a priori}, whether she's
-turning the coin from $H$ to $T$ or from $T$ to $H$. When Bob secretly
-turns the coin over, in event $b_{10}$ or $b_{01}$, she doesn't know
+
+Finally, the lines beginning with {\tt agent} define {\em event alternative}
+relations for the epistemic agents in the model. Each clause with an arrow has a
+single event symbol on the left, and a disjunction of alternative events on the
+right of the arrow. The interpretation of Amy's alternatives for $\bpH$ (Bob
+peeks at heads), is that when $\bpH$ happens, for Amy either $\bpH$ or $\bpT$ (Bob
+peeks at tails) could be happening, indicating that she doesn't know whether Bob saw heads or tails, only that he peeked at the coin.
+
+Similarly, Her alternatives for $\atTH$ (she turns the coin over from heads to
+tails) are $\atTH$ and $\atHT$, indicating that she doesn't know, \emph{a
+  priori}, whether she's turning the coin from $H$ to $T$ or from $T$ to $H$.
+She has the same alternatives for the event $\atHT$. Conversely, When Bob
+secretly turns the coin over, in event $\btTH$ or $\btHT$, she doesn't know
 anything has happened, so her alternative is the ``no-operation'' or
@@
 specification. The models are an extension of guarded-string models for Kleene
-Algebra with Tests (KAT). This is an algebraic theory that has model classes
-including guarded string models, relational models, finite models, and matrix models.
-Our definitions and notation follow \cite{kozen2001automata}.  We add syntax and
+Algebra with Tests (\textsc{Kat}). This algebraic theory has model classes including
+guarded string models, relational models, finite models, and matrix models. Our
+definitions and notation follow \cite{kozen2001automata}. We add syntax and
 semantics to cover multi-agent epistemic semantics.
 
-Guarded strings over a finite alphabet ${\sf{P}}$ are like ordinary strings, but with
-atoms over a set ${\sf{B}}$ alternating with the symbols from ${\sf{P}}$.  In the
-algebra described by Figure \ref{epikfigure}, ${\sf{P}}$ is the set of events
-$\{a_1,a_0,b_1,b_0,a_{10},a_{01},b_{10},b_{01},o\}$, and ${\sf{B}}$ is $\{h,t\}$.  As
-we already saw in \ex{effect}, $\mathcal{A}^\varphi_{\mathsf B}$ is
-$\{h\bar{t},\bar{h}t\}$, for which we use the shorthand $\{H,T\}$.  A guarded string
-over ${\sf{P}}$ and ${\sf{B}}$ is a string of events from ${\sf{P}}$, alternating
-with atoms over ${\sf{B}}$, and beginning and ending with atoms.  In this
-construction, $w_1=H$, $w_2=Ha_1H$, $w_3=Ha_1Hb_1H$, and
+Guarded strings over a finite alphabet ${\sf{P}}$ are like ordinary strings, but
+with atoms over a set ${\sf{B}}$ alternating with the symbols from ${\sf{P}}$.
+In the algebra described by Figure \ref{epikfigure}, ${\sf{P}}$ is the set of
+events $\{\apH,\apT,\bpH,\bpT,\atTH,\atHT,\btTH,\btHT,o\}$, and ${\sf{B}}$ is
+$\{h,t\}$. As we already saw in \ex{effect}, $\mathcal{A}^\varphi_{\mathsf B}$
+is $\{h\bar{t},\bar{h}t\}$, for which we use the shorthand $\{H,T\}$. A guarded
+string over ${\sf{P}}$ and ${\sf{B}}$ is a string of events from ${\sf{P}}$,
+alternating with atoms over ${\sf{B}}$, and beginning and ending with atoms. In
+this construction, $w_1=H$, $w_2=Ha_1H$, $w_3=Ha_1Hb_1H$, and
 $w_4=Ha_1Hb_1Ha_{10}Tb_{01}Hb_1H$.
@@
 Rather than individual guarded strings, elements of a guarded string
-model for KAT are sets of guarded strings.  In our application, these
+model for \textsc{Kat} are sets of guarded strings.  In our application, these
 elements have the interpretation of propositions, which are sets of possible
-worlds.  In a free guarded string model for KAT, any event can be
+worlds.  In a free guarded string model for \textsc{Kat}, any event can be
 adjacent to any atom in a guarded string that is an element of the
@@
   be the power set of $W$, rather it will consists of the regular subsets of $W$.}
-An Epik specification defines an alternative relation on bare events for each agent
-$a$, which we notate as ${R}_{a}$.  This should be lifted to a relation $\hat{R}_a$
-on worlds.  The basic idea is that when a world $w$ is incremented with an event $e$,
-in the resulting world $w \diamond e$, epistemic alternatives for agent $a$ are of
-the form $w' \diamond e'$, where $w'$ is an alternative to for $a$ in $w$, and $e'$
-is and event-alternative to $e$ for $a$.\footnote{In this it is important that the
-  event-alternative relation for an agent is constant across worlds. We anticipate
-  that the definition given here produces results equivalent to what is found in
-  literature on event alternatives in dynamic epistemic semantics, though we have not
-  verified this.  That literature primarily focuses on mapping an epistemic model for
-  a single time and situation to another, and uses general first-order models, rather
-  than guarded string models. See \citet{baltag1999logic}, \citet{van2007dynamic},
-  and articles in \citet{ditmarsch2015handbook}.  This literature is motivated by
-  epistemic logic and AI planning, rather than computable possible worlds models in
-  natural language semantics.  } This needs to be implemented in a way that takes
-account of pre- and post-conditions for events. For this, our approach is to refer
-the definition of well-formed guarded strings.  \ex{raise1} defines an epistemic
-alternative relation on worlds from an alternative relation on bare events.
+
+An Epik specification defines an alternative relation on bare events for each
+agent $a$, which is notated ${R}_{a}$, and lifted to a relation $\hat{R}_a$ on
+worlds. The basic idea is that when a world $w$ is incremented with an event
+$e$, in the resulting world $w \diamond e$, epistemic alternatives for agent $a$
+are of the form $w' \diamond e'$, where $w'$ is an alternative to for $a$ in
+$w$, and $e'$ is and event-alternative to $e$ for $a$.\footnote{In this it is
+  important that the event-alternative relation for an agent is constant across
+  worlds. We anticipate that the definition given here produces results
+  equivalent to what is found in literature on event alternatives in dynamic
+  epistemic semantics, though we have not verified this. That literature
+  primarily focuses on mapping an epistemic model for a single time and
+  situation to another, and uses general first-order models, rather than guarded
+  string models. See \citet{baltag1999logic}, \citet{van2007dynamic}, and
+  articles in \citet{ditmarsch2015handbook}. This literature is motivated by
+  epistemic logic and AI planning, rather than computable possible worlds models
+  in natural language semantics. } This needs to be implemented in a way that
+takes account of pre- and post-conditions for events. For this, our approach is
+to refer the definition of well-formed guarded strings. \ex{raise1} defines an
+epistemic alternative relation on worlds from an alternative relation on bare
+events.
 
@@
 
-Consider a scenario like the one from Figure 1, but with an additional agent Cal. The
-base world $T b_0 T c_0 T$ is one where the coin is tails, and first Bob looks at
-tails, and them Cal looks at tails. The first event $b_0$ has the alternatives $b_0$
-and $b_1$ for Amy, and the second event $c_0$ has the alternatives $c_0$ and $c_1$
-for Amy.  This results in four combinations $b_0c_0$, $b_0c_1$, $b_1c_0$, and
-$b_1c_1$. But these are filtered by post- and pre-conditions of events in the
-alternative world, so that the set of alernatives for Amy in $T b_0 T c_0 T$ is $\{T
-b_0 T c_0 T,H b_1 H c_1 H\}$, with two world-alternatives instead of four.
+Consider a scenario like the one from Figure 1, but with an additional agent
+Cal. The base world $T b_0 T c_0 T$ is one where the coin is tails, and first
+Bob looks at tails, and them Cal looks at tails. The first event $b_0$ has the
+alternatives $b_0$ and $b_1$ for Amy, and the second event $c_0$ has the
+alternatives $c_0$ and $c_1$ for Amy. This results in four combinations
+$b_0c_0$, $b_0c_1$, $b_1c_0$, and $b_1c_1$. But some of these worlds are bogus,
+such as $b_0c_1$, which says that Bob peeked at the coin and saw tails, and Cal
+peeked at the same coin and impossibly saw heads. These contradictions are
+filtered by post- and pre-conditions of events in the alternative world, so that
+the set of alernatives for Amy in $T b_0 T c_0 T$ is $\{T b_0 T c_0 T,H b_1 H
+c_1 H\}$, with two world-alternatives instead of four.
 
@@
     \begin{minipage}{\columnwidth}
-      \(\begin{array}{llll}
+      \(\begin{array}{rlrl}
       \multicolumn{4}{l}{\text{events}~~e\in \mathsf{P}} \\
@@
       p \wedge q &  \triangleq \neg(\neg p + \neg q) &
-      p \to q & \neg p + q      
+      p \to q & \triangleq \neg p + q
       \end{array} \)
@@
 
-\section{The logical language of Epistemic KAT}
+\section{The logical language of Epistemic \textsc{Kat}}
 The standard language for Kleene algebra with tests has the signature
 $\tuple{K,+,\cdot,*,\bar{},0,1}$ \citep{kozen2001automata}.  In a guarded string
-model for KAT, $K$ is a set of sets of guarded strings, $+$ is set union, the
+model for \textsc{Kat}, $K$ is a set of sets of guarded strings, $+$ is set union, the
 operation $\cdot$ is fusion product raised to sets, $*$ is Kleene star, the operation
@@
 With modalities and propositional negation added, the signature of $n$-agent
-epistemic KAT is $\tuple{K,+,\cdot,*,\bar{},0,1,\neg,\Diamond_1,
+epistemic \textsc{Kat} is $\tuple{K,+,\cdot,*,\bar{},0,1,\neg,\Diamond_1,
   ..., \Diamond_n}$. Figure \ref{termfigure} defines the syntax of the language.
@@
 
-{\em Standard Epistemic Modalities}
+{\em Standard Epistemic Modalities.}
 Using our existential modal primitive $\Diamond_a$, and the dual
@@
 $\tuple{K^{\epikspec},+,\cdot,*,\bar{},0,1,\neg,\Diamond_{a_1},...,\Diamond_{a_n}}$
-is a concrete guarded string interpretation for the signature of epistemic KAT, with
+is a concrete guarded string interpretation for the signature of epistemic \textsc{Kat}, with
 operations as in Figure \ref{fig:interp2} (e.g. the binary operation $+$ is union,
@@
 and a set of propositions, each of which is a set of worlds.  It is used as a target
-for natural-language interpretation in Section 6.
+for natural-language interpretation in Section~\ref{sec:categorial-grammar}.
 
@@
 \citep{hulden2009foma,linden2009hfst,karttunen2010update}.  Such definitions are used
-here to construct of a model for epistemic KAT inside the finite state
+here to construct of a model for epistemic \textsc{Kat} inside the finite state
 calculus. We describe our translation from Epik logical terms to Fst terms
@@
 \end{verbatim}
-KAT product.
+\textsc{Kat} product.
 \end{hangingpar}
@@
 \end{verbatim}
-KAT Kleene plus and Kleene star. The Fst operation {\tt |} is union.
+\textsc{Kat} Kleene plus and Kleene star. The Fst operation {\tt |} is union.
 \end{hangingpar}
 \end{tcolorbox}
-\caption{\raggedright Definition in Fst of KAT product and star.
+\caption{\raggedright Definition in Fst of \textsc{Kat} product and star.
   %% Where {\tt X} and {\tt Y} are regular sets and
@@
 Fst set difference (\texttt{-}) is used to define propositional complement as the
-difference between the set of worlds and the argument.
-
-Defining KAT product using Fst's set-lifted string concatenation (denoted by
+difference between the set of all worlds and the argument.
+
+Defining \textsc{Kat} product using Fst's set-lifted string concatenation (denoted by
 juxtaposition $\texttt{X}~\texttt{Y}$) requires more care. Naively concatenating
 strings with atoms (Boolean vectors) at both ends doubles atoms at the juncture, and
-does not enforce the requisite atom equality.  To implement KAT product, we define
+does not enforce the requisite atom equality.  To implement \textsc{Kat} product, we define
 the binary operation \texttt{Cn}, which concatenates strings in the string algebra,
@@
 
-KAT Kleene plus is defined in a similar way using Kleene plus in the string algebra,
+\textsc{Kat} Kleene plus is defined in a similar way using Kleene plus in the string algebra,
 with checks for equality of atoms and deletion of atoms. 
-KAT Kleene star is definied from KAT Kleene plus and the multiplicative identity,
+\textsc{Kat} Kleene star is defined from \textsc{Kat} Kleene plus and the multiplicative identity,
 which is the set of well-formed atoms \texttt{St}.
@@
   closure under discussion here is the closure of the former rather than the latter.}
-The epistemic alternative relation on worlds for an agent is then defined as the KAT
+The epistemic alternative relation on worlds for an agent is then defined as the \textsc{Kat}
 relation concatenation closure {\tt RelKst} of the decorated-event alternative
@@
 The bounded interpretation into lists of strings is very similar to
-the unboundeds interpretation into sets of strings, except for the
+the unbounded interpretation into sets of strings, except for the
 bounds checking. The full details are shown in
@@
 \section{Syntax-semantics interface}
+\label{sec:categorial-grammar}
 
 English sentences are mapped to terms in the logical language via a semantically
-interpreted multi-modal categorial grammar, consisting of a lexicon of words, their
-categorial types, and interpretations in a logical lambda language.  The grammar
-covers basic statives ({\em it's heads}), {\em that-} and {\em whether}-complements,
-negation, and predicate and sentence
-conjunction.\marginpar{Conjunction missing!}  Figure \ref{lexiconfigure} gives the
-lexicon.\footnote{Category symbols use Lambek/Bar-Hillel notation for slashes, so
-  that $(d \backslash t)\slash(d \backslash_D t)$ combines with $d \backslash_D t$ on
-  the right to give a value that combines with $d$ on the left to give $t$. In the
-  semantics, lambda abstractions with multiple parameters are written $\lambda
-  x~y.~e$ rather than $\lambda x.\lambda y.e$.  $d$ is the category of expletive {\em it}.}
-The grammar and semantics are optimized for a simple fragment of English concerned
-with clausal complementation.  The agent names {\em Amy} and {\em Bob} contribute the
-epistemic alternative relations for those agents, rather than individuals.  The root
-verb {\em believe} contributes existential modal force.  The complementizers {\em that}
-and {\em whether} are the heads of their dominating clauses, and assemble an
-alternative relation, modal force, and proposition contributed by the complement.
-These complementizers introduce the dual via two negations, in order to express
-universal modal force.
+interpreted multi-modal categorial grammar, consisting of a lexicon of words,
+their categorial types, and interpretations in a logical lambda language. The
+grammar covers basic statives ({\em it's heads}), {\em that-} and {\em
+  whether}-complements, negation, and predicate and sentence
+conjunction.\marginpar{Conjunction missing!} Figure \ref{lexiconfigure} gives
+the lexicon.\footnote{Category symbols use Lambek/Bar-Hillel notation for
+  slashes, so that $(d \backslash t)\slash(d \backslash_D t)$ combines with $d
+  \backslash_D t$ on the right to give a value that combines with $d$ on the
+  left to give $t$. In the semantics, lambda abstractions with multiple
+  parameters are written $\lambda x~y.~e$ rather than $\lambda x.\lambda y.e$.
+  $d$ is the category of expletive {\em it}.} The online supplement includes a
+version of this grammar that can be run using Barker and Shan's chart parser for
+categorial grammar \citep{barker2005parser}. The grammar and semantics are
+optimized for a simple fragment of English concerned with clausal
+complementation. The agent names {\em Amy} and {\em Bob} contribute the
+epistemic alternative relations for those agents, rather than individuals. The
+root verb {\em believe} contributes existential modal force. The complementizers
+{\em that} and {\em whether} are the heads of their dominating clauses, and
+assemble an alternative relation, modal force, and proposition contributed by
+the complement. These complementizers introduce the dual via two negations, in
+order to express universal modal force.
 
@@
       \end{tabular}%%
-    \caption{Categorial grammar lexicon. The first column has a word
-      form. the second column a categorial type, and third column a
-      semantic translation in a language that extends the Epik logical
-      language with lambda. The online supplement includes a version
-      of this grammar that can be run using Barker and Shan's chart
-      parser for categorial grammar \citep{barker2005parser}.}
+    \caption{Categorial grammar lexicon, showing word form (column 1), a
+      categorial type (column 2), and a semantic translation in Epik extended
+      with lambda abstraction (column 3).}
   \label{lexiconfigure}
@@
 \section{Discussion}
-The methodology presented here is designed for use in research in linguistic
-semantics, for research in computational linguistics on model-theoretically
-grounded semantics, 
-and for education at the level of a second graduate course in formal
+Epistemic \textsc{Kat} is designed for use in research in linguistic semantics,
+for research in computational linguistics on model-theoretically grounded
+semantics, and for education at the level of a second graduate course in formal
 semantics, covering intensionality. There are straightforward extensions to
-additional linguistic phenomena, such as tense and perfective aspect as in
+additional linguistic phenomena, such as tense and perfective aspect
 \exi{extension}{a}, and the combination of metaphysical modality and prospective
-aspect in \exi{extension}{b}.
+aspect \exi{extension}{b}.
 
@@
 
-Epik is closely related to Kleene Algebra with Hypotheses\cite{},
-\textsc{Ka+H}, which permits user-specification of futher equations
-beyond the axioms of \textsc{Ka}. In practice, these hypotheses are
-used to express domain-specific properties of the axioms, such as the
-commutativity of certain uninterpreted actions in program
-verification, such as $x := 3; y := 5 \equiv y := 5; x := 3$). We can
-model both the state formulae and the event formulae as \textsc{KA+H}
-hypotheses. Specifically, the state formula $\varphi$ can be thought
-of as a hypothesis $\varphi \equiv 1$, and an event formula $\zeta_a$
-for event $a$, is simply the hypothesis $a \equiv
-\sum_{\varphi,\varphi' \in \llbracket \zeta_a \rrbracket}
-\varphi;a;\varphi'$. Notwithstanding some early results about
-\textsc{Kat} with hypotheses of the form $cp=c$~\cite{}, using basic
-\textsc{KA} implementations would rely on using the exponential
-encoding of \textsc{Kat} into \textsc{Ka}~\cite{}.
-
-The key facet that differentiates Epik from \textsc{Ka+H} is the
-modality. Each alternative relation represents a set of
-uni-directional rewrites that must be applied for that agent, rather
-than the universal, optional rewrites represented by the \textsc{Ka+H}
-hypotheses. It may be possible to encode Epik's alternative relations
-into \textsc{Ka+H} hypotheses, but the translation would be
-non-trivial, and we leave it as future work.
+Epik is closely related to Kleene Algebra with Hypotheses
+\textsc{Ka+H}~\cite{cohen1994hypotheses}, which permits user-specification of
+futher equations beyond the axioms of \textsc{Ka}. In practice, these hypotheses
+are used to express domain-specific properties of the events, such as the
+commutativity. We can model both the state formulae and the event formulae as
+\textsc{KA+H} hypotheses. Specifically, the state formula $\varphi$ can be
+thought of as a hypothesis $\varphi \equiv 1$, and an event formula $\zeta_a$
+for event $a$, is simply the hypothesis $a \equiv \sum_{\varphi,\varphi' \in
+  \llbracket \zeta_a \rrbracket} \varphi;a;\varphi'$. In general, the Horn
+theory of \textsc{KA+H} is undecidable, however many special cases
+are~\cite{doumane2019kleene, hardin2002elimination,
+  kozen1997kleenecompleteness}.
+
+The key facet that differentiates Epik from \textsc{Ka+H} is the modality. Each
+alternative relation represents a set of uni-directional rewrites that must be
+applied for that agent, rather than the universal, optional rewrites represented
+by the \textsc{Ka+H} hypotheses. It may be possible to encode Epik's alternative
+relations into \textsc{Ka+H} hypotheses, but such an encoding is future work.
 
@@
 semantics. Issues for further investigation are mathematical
-characterizations of epistemic KATs, e.g. sound and complete axioms
+characterizations of epistemic \textsc{Kat}s, e.g. sound and complete axioms
 and coalgebrae.
 
-Source code, examples, and instructions for Epik are distributed at
+Source code, examples, and instructions for Epik are available at
 \url{https://github.com/ericthewry/epik}.

### 2021-01-14 tex/pdraft.tex
Additions to grammar figure, and thanks.

--- 2021-01-14 tex/pdraft.tex
@@
 
-\title{Epistemic Semantics in Guarded String Models\thanks{Thanks guys.}}
+\title{Epistemic Semantics in Guarded String Models\thanks{Thanks to Tobias Kappe for insightful reactions to the paper.
+Thanks also to three SCiL reviewers for helpful comments.
+Earlier versions of the work were presented at in the LUSH talk series at Utrecht in December 2019, and in the Workshop on formal/philosophical/computational approaches to natural language meaning at Rochester in December 2020.
+Thanks to the audiences for their reactions.
+    .}}
 \renewcommand\footnotemark{}
@@
 
+\newcommand{\ebackt}{(e \backslash t)}
 \begin{figure}[!t]
@@
           \,\,\,\, +\,{\neg}(m\,p\,R)
-        \end{pmatrix*}$ \\
+        \end{pmatrix*}$ \\ \midrule
+        \multicolumn{3}{c}{{\bf Conjunction}} \\[\hdrspc]
+        and & $(t \backslash t)\slash t$ &  $\lambda p\,q.\,p \wedge q$\\
+        or & $(t \backslash t)\slash t$ &  $\lambda p\,q.\,p + q$\\
+        and & $((\ebackt \backslash \ebackt)\slash \ebackt$ &  $\lambda p\,q\,x.\,p(x) \wedge q(x)$\\
+        or & $((\ebackt \backslash \ebackt)\slash \ebackt$ &  $\lambda p\,q\,x.\,p(x) + q(x)$\\        
         \bottomrule
@@
       semantic translation in a language that extends the Epik logical
-      language with lambda. The online supplement includes a version
-      of this grammar that can be run using Barker and Shan's chart
-      parser for categorial grammar \citep{barker2005parser}.}
+      language with lambda. }
   \label{lexiconfigure}
@@
 \url{https://github.com/ericthewry/epik}.
+
+\clearpage
 

### 2021-01-14 tex/pdraft.tex
typo

@@
 Thanks also to three SCiL reviewers for helpful comments.
-Earlier versions of the work were presented at in the LUSH talk series at Utrecht in December 2019, and in the Workshop on formal/philosophical/computational approaches to natural language meaning at Rochester in December 2020.
+Earlier versions of the work were presented in the LUSH talk series at Utrecht in December 2019, and in the Workshop on formal/philosophical/computational approaches to natural language meaning at Rochester in December 2020.
 Thanks to the audiences for their reactions.

### 2021-01-14 tex/pdraft.tex

--- 2021-01-14 tex/pdraft.tex
@@
 
-\title{Epistemic Semantics in Guarded String Models\thanks{Thanks guys.}}
+\title{Epistemic Semantics in Guarded String Models\thanks{Thanks to Tobias Kappe for insightful reactions to the paper.
+Thanks also to three SCiL reviewers for helpful comments.
+Earlier versions of the work were presented in the LUSH talk series at Utrecht in December 2019, and in the Workshop on formal/philosophical/computational approaches to natural language meaning at Rochester in December 2020.
+Thanks to the audiences for their reactions.
+    .}}
 \renewcommand\footnotemark{}

### 2021-01-14 tex/pdraft.tex
editing pass

@@
 
-\title{Epistemic Semantics in Guarded String Models\thanks{Thanks to Tobias Kappe for insightful reactions to the paper.
+\title{Epistemic Semantics in Guarded String Models\thanks{Thanks to Tobias Kapp\'{e} for insightful reactions to the paper.
 Thanks also to three SCiL reviewers for helpful comments.
 Earlier versions of the work were presented in the LUSH talk series at Utrecht in December 2019, and in the Workshop on formal/philosophical/computational approaches to natural language meaning at Rochester in December 2020.
-Thanks to the audiences for their reactions.
-    .}}
+Thanks to the audiences for their reactions.}}
 \renewcommand\footnotemark{}
@@
 
-{\em Axiomatic Classification}. To situate our logic as a modal logic, consider the soundness of
-the standard modal axioms given our semantics \citep{hughes1996new}. Some of
-these standard axioms (see \ex{valid-axioms}) hold all the time, i.e. they
-are \emph{valid}, and the remaining axioms are valid when $\hat{R}_a$
-has a certain shape. The axioms in \ex{conditionally-valid-axioms}
-have a nontrivial condition validity condition.
+{\em Axiomatic Classification}. To situate our logic as a modal logic, consider
+the soundness of the standard modal axioms given our semantics
+\citep{hughes1996new}. The axioms in \ex{valid-axioms}
 
@@
 
+The axioms in \ex{conditionally-valid-axioms} are not valid, but do hold when
+when the relation $R_a$ has a certain shape.
+
 \begin{example}\label{conditionally-valid-axioms}
   \(\begin{array}[t]{lll}
-  \textbf{T} & \Box_a p \to p & \text{if } g\mathrel{\hat{R}_a} g, \forall g \\
-  \textbf{D} & \Box_a p \to \Diamond_a p & \text{if } g \in \dom(\hat{R}_a), \forall g \\
-  \textbf{4} & \Box_a p \to \Box_a\Box_a p & \text{if } \hat{R}_a \text{ idempotent}
+  \textbf{T} & \Box_a p \to p & \text{if } g \mathrel{\hat{R}_a} g, \forall g \\
+  \textbf{D} & \Box_a p \to \Diamond_a p & \text{if } \hat{R}_a \text{ is a function} \\
+  \textbf{4} & \Box_a p \to \Box_a\Box_a p & \text{if } \hat{R}_a \text{ is idempotent}
   \end{array}\)
@@
 theorem in relational terms. The conditions on the remaining standard
-modal axioms, \textbf{B} and \textbf{5}, are of a similarly trivial
-flavor.
+modal axioms, \textbf{B} and \textbf{5}, are similar.
 
@@
 here to construct of a model for epistemic \textsc{Kat} inside the finite state
-calculus. We describe our translation from Epik logical terms to Fst terms
-here.
+calculus (Figures~\ref{fstfigure1}\&\ref{relclosure}).
 
@@
 removes strings with non-matching atoms, and then deletes the second of two atoms to
-create a set of well-formed guarded strings.  See Figure \ref{fstfigure1}. {\tt Wf0}
-is the set of ordinary strings that does not contain unequal pairs of atoms, as
+create a set of well-formed guarded strings (Figure \ref{fstfigure1}). {\tt Wf0}
+is the set of ordinary strings with only equal pairs of atoms, as
 defined using Fst's containment operator \texttt{\$}. The \texttt{Squash} relation
 uses Fst's rewrite notation to delete atoms (elements of \texttt{St}) that are
-preceded by another atom. \footnote{ This is a non-equal length regular relation. The
+preceded by another atom\footnote{ This is a non-equal length regular relation. The
   finite state calculus includes such relations, and they can be used with relation
@@
   complement and set difference for non-equal length relations is not defined.
-  Epistemic alterative relations are equal-length relations.} This relation is
-applied via the relational composition (\texttt{.o.}) and codomain (\texttt{.l})
+  Epistemic alterative relations are equal-length relations.}. This relation is
+applied via the composition (\texttt{.o.}) and codomain (\texttt{.l})
 operators.
@@
 
-\textsc{Kat} Kleene plus is defined in a similar way using Kleene plus in the string algebra,
-with checks for equality of atoms and deletion of atoms. 
-\textsc{Kat} Kleene star is defined from \textsc{Kat} Kleene plus and the multiplicative identity,
-which is the set of well-formed atoms \texttt{St}.
+Kleene plus is defined in a similar way, using Kleene plus in the string
+algebra, with checks for equality of atoms and deletion of atoms. Kleene star is
+defined from Kleene plus the well-formed atoms \texttt{St}, which implements $1$.
 
@@
 
-It remains to define an epistemic alternative relation on worlds for each agent.  The
-relevant information in Figure \ref{epikfigure} is a relation between bare events for
-each agent. This determines a relation between bare events decorated with compatible
-atoms.  In Fst, we use the closure of the concatenation product operation on
-relations to lift a relation on decorated events for an agent to the corresponding
-relation on worlds.  The concatenation product $R~S$ of two relations $R$ and $S$ is
-the set of pairs of the form $\pair{x_1x_2}{y_1y_2}$, where $x_1 \mathrel R y_1$, and
-$x_2 \mathrel S y_2$.  In Fst, {\tt{R+}} is the closure of relation {\tt R} with
-respect to this operation. Figure \ref{relclosure} defines the corresponding
-operation on sets of guarded strings as encoded in Fst.\footnote{Relation
-  concatenation in Fst differs from relation composition ({\tt{.o.}}), and the
-  closure under discussion here is the closure of the former rather than the latter.}
-The epistemic alternative relation on worlds for an agent is then defined as the \textsc{Kat}
-relation concatenation closure {\tt RelKst} of the decorated-event alternative
-relation for the agent.
+Finally, the epistemic alternative relation for each agent, i.e. the
+\texttt{agent} relation from Figure~\ref{epikfigure} is implemented in Fst.
+Combining the agent relation with the effect formula gives a relation on events
+decorated with compatible atoms. Then the relation on worlds is constructed
+using the closure of the concatenation product operation. The concatenation
+product $R~S$ of two relations $R$ and $S$ is the set of pairs of the form
+$\pair{x_1x_2}{y_1y_2}$, where $x_1 \mathrel R y_1$, and $x_2 \mathrel S y_2$.
+In Fst, {\tt{R+}} is the closure of relation {\tt R} with respect to this
+operation. Figure \ref{relclosure} defines the corresponding operation on sets
+of guarded strings as encoded in Fst.\footnote{Relation concatenation in Fst
+  differs from relation composition ({\tt{.o.}}), and the closure under
+  discussion here is the closure of the former rather than the latter.} The
+epistemic alternative relation on worlds for an agent is then defined as the
+\textsc{Kat} relation concatenation closure {\tt RelKst} of the decorated-event
+alternative relation for the agent.
 
@@
 \label{sec:categorial-grammar}
-
-English sentences are mapped to terms in the logical language via a semantically
-interpreted multi-modal categorial grammar, consisting of a lexicon of words,
-their categorial types, and interpretations in a logical lambda language. The
-grammar covers basic statives ({\em it's heads}), {\em that-} and {\em
-  whether}-complements, negation, and predicate and sentence
-conjunction.\marginpar{Conjunction missing!} Figure \ref{lexiconfigure} gives
-the lexicon.\footnote{Category symbols use Lambek/Bar-Hillel notation for
-  slashes, so that $(d \backslash t)\slash(d \backslash_D t)$ combines with $d
-  \backslash_D t$ on the right to give a value that combines with $d$ on the
-  left to give $t$. In the semantics, lambda abstractions with multiple
-  parameters are written $\lambda x~y.~e$ rather than $\lambda x.\lambda y.e$.
-  $d$ is the category of expletive {\em it}.} The online supplement includes a
-version of this grammar that can be run using Barker and Shan's chart parser for
-categorial grammar \citep{barker2005parser}. The grammar and semantics are
-optimized for a simple fragment of English concerned with clausal
-complementation. The agent names {\em Amy} and {\em Bob} contribute the
-epistemic alternative relations for those agents, rather than individuals. The
-root verb {\em believe} contributes existential modal force. The complementizers
-{\em that} and {\em whether} are the heads of their dominating clauses, and
-assemble an alternative relation, modal force, and proposition contributed by
-the complement. These complementizers introduce the dual via two negations, in
-order to express universal modal force.
-
-Multimodal categories such as $\backslash_D$ and $\backslash_M$ are
-used to control the derivation---phrases with these top-level slashes
-can only combine syntactically as arguments.
-%
-%For instance the category of {\em
-%  heads} is $d \backslash_D t$.  The dummy expletive subject {\em it}
-%has category $d$.  However the phrase {\em it heads} of category $t$
-%can not be formed, because $\backslash_D$ is not syntactically active
-%as a function.  Instead {\em it is heads} is formed with a predicator
-%{\em is} of category $(d \backslash t)\slash(d \backslash_D t)$.
-%Similarly {\em knows} has a category with the top-level slash
-%$\slash\!_M$, and combines to form a sentence as an argument of {\em
-%  that} or {\em whether}. These complementizers have a category that
-%looks for the category of {\em know} on the left, after combining with
-%a complement sentence on the right.
-The semantic translations in the third column of Figure \ref{lexiconfigure} use
-the logical language, incremented with lambda.  The body of $\lambda
-x.\bullet^*h$, which is the semantic lexical entry for {\em heads}, is a
-term denoting the set of all worlds where the coin is heads, expressed as the
-set of all guarded strings that end with a Boolean valuation where the primitive
-proposition $h$ (it's heads) is true.
-%There is $\lambda x$ at the
-%front because the grammar formalism enforces a strict correspondence
-%between syntactic and semantic types. However, the lambda does not
-%bind anything, because sentences such as {\em it isn't heads} have an
-%expletive subject.
-The body of $\lambda p.\lambda R. {\Diamond}_R p$, which is the semantic lexical
-entry of {\em believes}, is an term denoting the pre-image of the world-alternative
-relation contributed by the subject.  This is not the right semantics for {\em Amy
-  believes that it's heads}, because it has an existential modality ${\Diamond_R}p$,
-rather than an universal modality ${\Box_R}p$.  This is corrected by the
-complementizer {\em that}, which introduces the dual.
 
@@
 
+English sentences are mapped to terms in the logical language via a semantically
+interpreted multi-modal categorial grammar, consisting of a lexicon of words,
+their categorial types, and interpretations in a logical lambda language. The
+grammar covers basic statives ({\em it's heads}), {\em that-} and {\em
+  whether}-complements, negation, and predicate and sentence
+conjunction.\marginpar{Conjunction missing!} Figure \ref{lexiconfigure} gives
+the lexicon.\footnote{Category symbols use Lambek/Bar-Hillel notation for
+  slashes, so that $(d \backslash t)\slash(d \backslash_D t)$ combines with $d
+  \backslash_D t$ on the right to give a value that combines with $d$ on the
+  left to give $t$. In the semantics, lambda abstractions with multiple
+  parameters are written $\lambda x~y.~e$ rather than $\lambda x.\lambda y.e$.
+  $d$ is the category of expletive {\em it}.} The online supplement includes a
+version of this grammar that can be run using Barker and Shan's chart parser for
+categorial grammar \citep{barker2005parser}. The grammar and semantics are
+optimized for a simple fragment of English concerned with clausal
+complementation. The agent names {\em Amy} and {\em Bob} contribute the
+epistemic alternative relations for those agents, rather than individuals. The
+root verb {\em believe} contributes existential modal force. The complementizers
+{\em that} and {\em whether} are the heads of their dominating clauses, and
+assemble an alternative relation, modal force, and proposition contributed by
+the complement. These complementizers introduce the dual via two negations, in
+order to express universal modal force.
+
+Multimodal categories such as $\backslash_D$ and $\backslash_M$ are
+used to control the derivation---phrases with these top-level slashes
+can only combine syntactically as arguments.
+%
+%For instance the category of {\em
+%  heads} is $d \backslash_D t$.  The dummy expletive subject {\em it}
+%has category $d$.  However the phrase {\em it heads} of category $t$
+%can not be formed, because $\backslash_D$ is not syntactically active
+%as a function.  Instead {\em it is heads} is formed with a predicator
+%{\em is} of category $(d \backslash t)\slash(d \backslash_D t)$.
+%Similarly {\em knows} has a category with the top-level slash
+%$\slash\!_M$, and combines to form a sentence as an argument of {\em
+%  that} or {\em whether}. These complementizers have a category that
+%looks for the category of {\em know} on the left, after combining with
+%a complement sentence on the right.
+The semantic translations in the third column of Figure \ref{lexiconfigure} use
+the logical language, incremented with lambda.  The body of $\lambda
+x.\bullet^*h$, which is the semantic lexical entry for {\em heads}, is a
+term denoting the set of all worlds where the coin is heads, expressed as the
+set of all guarded strings that end with a Boolean valuation where the primitive
+proposition $h$ (it's heads) is true.
+%There is $\lambda x$ at the
+%front because the grammar formalism enforces a strict correspondence
+%between syntactic and semantic types. However, the lambda does not
+%bind anything, because sentences such as {\em it isn't heads} have an
+%expletive subject.
+The body of $\lambda p.\lambda R. {\Diamond}_R p$, which is the semantic lexical
+entry of {\em believes}, is an term denoting the pre-image of the world-alternative
+relation contributed by the subject.  This is not the right semantics for {\em Amy
+  believes that it's heads}, because it has an existential modality ${\Diamond_R}p$,
+rather than an universal modality ${\Box_R}p$.  This is corrected by the
+complementizer {\em that}, which introduces the dual.
+
 Sentences are parsed with a chart parser for categorial grammar.  The semantics for
@@
 \section{Discussion}
-Epistemic \textsc{Kat} is designed for use in research in linguistic semantics,
-for research in computational linguistics on model-theoretically grounded
-semantics, and for education at the level of a second graduate course in formal
-semantics, covering intensionality. There are straightforward extensions to
-additional linguistic phenomena, such as tense and perfective aspect
-\exi{extension}{a}, and the combination of metaphysical modality and prospective
-aspect \exi{extension}{b}.
+Epistemic \textsc{Kat} is designed for use in linguistic semantics research, for
+compultational linguistics research on model-theoretically grounded semantics,
+and for students in graduate course modules on intensional formal semantics.
+
+There are straightforward extensions to additional linguistic phenomena, such as
+tense and perfective aspect \exi{extension}{a}, and the combination of
+metaphysical modality and prospective aspect \exi{extension}{b}.
 
@@
 
-The model framework is a constructive branching-time framework with
-metaphysical modality and epistemic modality, which will be applicable
-in linguistic semantic research on combinations of tense, metaphysical
-modality, and epistemic complementation
-\citep{thomason1984combinations, abusch1998generalizing,
-  condoravdi2002temporal}.  Connections with research on temporal
-constitution of events in a related formal setting remain to be
-explored
+The model framework is a constructive branching-time framework with metaphysical
+and epistemic modalities, which will be applicable in linguistic semantic
+research on combinations of tense, metaphysical modality, and epistemic
+complementation \citep{thomason1984combinations, abusch1998generalizing,
+  condoravdi2002temporal}. Connections with research on temporal constitution of
+events remain to be explored
 \citep{fernando2004finite,fernando2007observing,carlson2009tense}.
 
-The syntactic part of the grammar formalism uses categorial grammar to
-map between strings (or trees or derivations) and logical
-translations.  This way of using a categorial grammar is standard
-\citep{steedman2000syntactic,bozsahin2012combinatory}, and there are
-computational implementations for it
-\citep{barker2005parser,bozsahin2021ccglab}.\footnote{Our grammar uses
-  basic categorial grammar, not additional features such as
-  combinators, type raising, continuations, or multimodal slashes that
-  interact with the grammar in non-trivial ways.  These features 
-  become relevant in more sophisticated grammars.  Multimodal
-  categories are used for phrases such as a tenseless verb phrase that
-  semantically are functional, but combine in the grammar only as
-  arguments.}  While this way of proceeding is simple and attractive
-for the applications we are interested in, it would be possible to use
-our semantics with other frameworks that have a semantics with a
-lambda extension of a logical language.
+The syntactic part of the grammar formalism uses a standard approach to
+categorial grammar \citep{steedman2000syntactic,bozsahin2012combinatory} to map
+between strings (or trees or derivations) and logical terms, and is rigorous
+enough to admit computational implementations
+\citep{barker2005parser,bozsahin2021ccglab}.\footnote{Our grammar uses basic
+  categorial grammar, not additional features such as combinators, type raising,
+  continuations, or multimodal slashes that interact with the grammar in
+  non-trivial ways. These features become relevant in more sophisticated
+  grammars. Multimodal categories are used for phrases such as a tenseless verb
+  phrase that semantically are functional, but combine in the grammar only as
+  arguments.} While this approach is simple and attractive for our applications,
+it would be possible to use Epik with other frameworks that support lambda
+extensions of logical languages.
 

### 2021-01-15 tex/pdraft.tex
rewrite introduction

@@
 \begin{abstract}
-  Constructive and computable multi-agent epistemic possible worlds
-  models are defined, where possible worlds models are guarded string
-  models in an epistemic extension of Kleene Algebra with Tests.  The
-  account is framed as a formal language Epik (Epistemic \textsc{Kat}) for
-  defining such models.  The language is implemented by translation
-  into the finite state calculus, and alternatively by modeling
-  propositions as lazy lists in Haskell. The syntax-semantics
-  interface for a fragment of English is defined by a categorial
-  grammar.
+  Constructive and computable multi-agent epistemic possible worlds models are
+  as sets of guarded string models in an epistemic extension of Kleene Algebra
+  with Tests (\textsc{Kat}). The account is framed as a formal language Epik
+  (Epistemic \textsc{Kat}) for defining such models. The language is implemented
+  by translation into the finite state calculus, and alternatively by modeling
+  propositions as lazy lists in Haskell. The syntax-semantics interface for a
+  fragment of English is defined by a categorial grammar.
 \end{abstract}
@@
 Linguistic semantics in the Montague tradition proceeds by assigning
-propositional {\em semantic values} to disambiguated sentences of a
-natural language.  A proposition is a set or class of {\em possible
-  worlds}.  These are often assumed to be things with the same nature
-and complexity as the world we occupy \citep{lewis1986plurality}.  But
-alternatively, one can work with small idealized models, in order to
-illustrate and test ideas.  The point of this paper is to scale up toy
-or idealized models to countable sets of worlds, and to constructive
-and computable modeling of epistemic alternatives for agents.  We
-describe a systematic way of defining such models, and
-illustrate how to apply them in natural language semantics.  The focus
-is on epistemic semantics and clausal embedding. The fundamental move
-is to identify possible worlds with strings of primitive events, so
-that propositions are sets of strings. An advantage in this is that it
-allows for a mathematical description of an algebra of propositions,
-coupled with a computational representation using either lazy lists of
-strings, or finite state machines that describe sets of strings.
-
-The approach taken here synthesizes five antecedents.
+propositional {\em semantic values} to disambiguated sentences of a natural
+language. A proposition is a set or class of {\em possible worlds}, which are
+often assumed to have the same nature and complexity as the world we occupy
+\citep{lewis1986plurality}. But alternatively, one can work with small idealized
+models, to illustrate and test ideas. The point of this paper is to extend
+standard idealized models to countable sets of worlds, and to constructively and
+computably model alternatives for epistemic agents.
+
+Enter Epik, which is a systematic way of defining such models, and can be
+applied to natural language semantics, specifically, epistemic semantics and
+clausal embedding. The fundamental insight is to identify possible worlds with
+strings of primitive events, so that propositions are sets of (guarded) strings,
+whose regular sets have a rich algebraic characterization~\cite{kozen1997kleene}
+and computational model~\cite{kozen1997kleenecompleteness,kozen2001automata}.
+Epik leverages these algebrae to define a mathematical model
+(Sections~\ref{sec:model} \&~\ref{sec:logic}) and computationally interpret
+using finite state machines (Section~\ref{sec:fst}) and lazy lists of strings
+(Section~\ref{sec:lazy})).
+
+{\em Related Work.} Epik synthesizes five antecedents.
 John McCarthy's {\em Situation Calculus} is the source of the idea of
@@
 
-A running example of an event-sequence model is {\em The Concealed Coin}. Amy
-and Bob are seated at a table. There is a coin on the table under a cup, heads
-up. The coin could be heads-up $H$ or tails-up $T$, and neither agent knows
-which it is. This initial situation is possible world $w_1$. We can construct
-additional worlds by sequencing events like those shown in \ex{coin1}.
+{\em Example.} Consider an example event-sequence model called {\em The
+  Concealed Coin}. Amy and Bob are seated at a table. There is a coin on the
+table under a cup, heads up. The coin could be heads-up $H$ or tails-up $T$, and
+neither agent knows which it is. Call this initial possible world $w_1$.
+Additional worlds can be constructed via the events in \ex{coin1}.
 
@@
 
-The worlds $w_1,\ldots,w_4$ in \ex{coin2} are examples constructed from the
-events in \ex{coin1}. Juxtaposition indicates the order and occurence of events,
-for example $w_3$ indicates that after the events of $w_2$, event $b_1$ occurs,
-fully, this indicates beginning in world $w_1$, Amy surreptitiously peeking at
-a heads-up coin, and then Bob doing the same.
+The worlds in \ex{coin2} are examples constructed from the events in \ex{coin1}.
+Juxtaposition indicates the order of events; for example, $w_3$ indicates that
+after the events of $w_2$, event $b_1$ occurs, which corresponds to, in the
+initial world, Amy surreptitiously peeking at a heads-up coin, and then Bob
+doing the same.
 
@@
 
-The truth values for English sentences are observed at the end of the event
-sequences specified by $w_1,\ldots, w_4$, where $0$ stands for falsity and $1$
-for truth.
+Now, the truth ($1$) or falsity ($0$) of English sentences can be evaluated in
+each world $w_i$, i.e. \emph{after} the events have happened. See \ex{coin3} for
+examples, such as in $w_4$, where Bob knows it's heads (because he peeked after
+turning the coin), but Amy does not (because Bob secretly turned the coin over).
 
@@
 
-The events come with pre-conditions and post-conditions.  Amy can turn
-the coin from heads to tails only if the coin is heads-up, so $\atHT$
-has the pre-condition of the coin being heads up. Once she turns the
-coin over, tails must be face-up, so $a_{10}$ has the post-condition
-of the coin being tails-up. Let $h$ be the Boolean proposition that
-the coin is heads up and $t$ be the Boolean proposition that the coin
-is tails-up.  Then pre- and post-conditions can be described by
-Boolean formulas, with $h$ being the pre-condition of $\atHT$ and
-$t$ being the post-condition. This is expressed using an operator
-``$:$'' (read ``and next'') that pairs Boolean formulas. The formula
-$\ev h t$ describes $\atHT$ (Amy turning the coin from heads
-to tails) as happening only in an $h$ state, and concluding in a
-$t$ state.  Events don't have to change state: the event $\apT$
-(Amy peeking at tails) can happen only in a $t$ state, and does
-not change the state ($\ev{t}{t}$).
+To create realistic models, contradictory event sequences must be prohibited.
+For example, $\apH\atTH$ would require the coin to simultaneously heads-up and
+tails-up. To mechanize this, events in Epik come with Boolean pre-conditions and
+post-conditions. In the running example, the Boolean variables $h$ (the coin is
+heads-up) and $t$ (the coin is tails-up) represent the coin's state. Then, since
+Amy can only observe heads when the coin is heads-up, event $\apH$ has the
+invariant condition $h$. Similarly, Amy can only turn the coin from tails to
+heads if it shows tails, so $\atTH$ has the tails-up precondition $t$ and the
+heads-up postcondition $h$.
+
+Pre- and post- conditions are expressed using an operator ``$:$'' (read ``and
+next'') that pairs Boolean formulas into an \emph{effect formula}. Specifically,
+the effects for $\apH$ are written $\ev h h$, and for $\atTH$ are written $\ev t
+h$. Effect formulae are interpreted relations on boolean valuations, as defined
+in Figure~\ref{effectfigure}.
+
+However, a coin cannot simultanously show both heads and tails! Currently the
+precondition $h$ of $\atHT$ only says that heads must be showing, and says
+nothing about the fact that tails must be face-down, indicated by the formula
+$\bar t$. The effect valuations are restricted to only the feasible ones via a
+\emph{state formula}, shown in \ref{effect}, and demonstrated for $\apH$ in
+\ref{effect}. The state formula, also shown in \ref{effect}, says that the coin
+is either heads-up or tails-up, but not both\footnote{Two Booleans constrained
+  by a state formula, is a more illustrative example than using a single boolean
+  and modelling tails as $\bar h$.}. Here juxtaposition represents conjuction,
+$+$ is disjunction, and overbar is negation.
+
+With the state formula in hand, $\apH\atTH$ is evidently contradictory, the
+post-condition $h$ of $\apH$ is incompatible with the pre-condition $t$ of
+$\atTH$. See Figure~\ref{effectfigure} for more details.
 
@@
 
-However a coin cannot be showing both heads and tails! Currently the
-precondition $h$ of $\atHT$ only says that heads must be showing, and
-says nothing about the fact that tails must be face-down, indicated by
-the formula $\bar t$. We will further restrict the feasible conditions
-for our actions by restricting the space of valuations for our formulas.
-
-A sequence such as $\bar{h}{t}$ can be viewed both as a formula and as
-a valuation of primitive propositions, which we use to describe world
-state. The primitives are listed in fixed order, and left unmarked
-(indicating true) or marked with the overbar (indicating false).
-Since a coin is heads or tails but not both, we want to allow the
-valuations $h\bar{t}$ and $\bar{h}t$, and disallow $ht$ and
-$\bar{h}\bar{t}$.\footnote{Two Booleans constrained by a state formula, rather than one Boolean, are stipulated here to
- illustrate state formulas.
-}  This is enforced by a {\em state formula}, which is
-a Boolean formula, in this case the one given on the second line of
-\ex{effect}.  Where {\sf{B}} is a set of state primitives and $\varphi$ is a
- constraint over {\sf{B}},
-${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\varphi}$ is the set of valuations of
-{\sf{B}} that make formula $\varphi$ true. Valuations are called atoms,
-because they correspond to the atoms of a Boolean algebra of tests
-\citep{kozen2001automata}.
-
-Formulas that describe pre- and
-post-conditions are {\em effect formulas}. They are interpreted as
-defining relations between atoms, as defined in Figure
-\ref{effectfigure}.  The atoms they relate are constrained by the
-state formula as well.  For the heads-tails example, let the state
-formula and the effect formula for $a_1$ (Amy peeking at heads) be as
-specified in \ex{effect}. Then ${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\varphi}$
-and the relation on atoms for the event $a_1$ are as given at
-the bottom in \ex{effect}.
+Given a set {\sf{B}} of state primitives and $\varphi$ a state formula over
+{\sf{B}}, define ${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\varphi}$ to be the set of
+valuations of {\sf{B}} that make formula $\varphi$ true. The valuations for the
+coin example are shown in \ref{effect}, using the sequence notation for
+valuations, e.g. $\bar{h}{t}$, wherein every state primitive is listed in fixed
+order, and left unmarked (indicating true) or marked with the overbar
+(indicating false). Valuations are called atoms, because they correspond to the
+atoms of a Boolean algebra of tests \citep{kozen2001automata}.
 
@@
 \mbox{state formula $\varphi$} & h\bar{t} + \bar{h}t \\
-\mbox{effect formula $\zeta$ for $a_1$} &  h \mathbin{:} h \\
+\mbox{effect formula $\zeta_{a_1}$ for $a_1$} &  h \mathbin{:} h \\
  \mathcal{A}^\varphi_{\mathsf B} & \{h\bar{t},\bar{h}t\}   \\
-\sem{\zeta}^\varphi & \{\pair{h\bar{t}}{h\bar{t}}\}   \\
+\sem{\zeta_{a_1}}^\varphi & \{\pair{h\bar{t}}{h\bar{t}}\}   \\
 \end{array}\)
 \end{example}
+
+Thus far, the presentation is closely related to Kleene Algebra with Hypotheses
+(\textsc{Ka+H})~\cite{cohen1994hypotheses}, which permits user-specification of
+futher equations beyond the axioms of \textsc{Ka}. Specifically, the state
+formula $\varphi$ can be thought of as a hypothesis $\varphi \equiv 1$, and an
+event formula $\zeta_a$ for event $a$, is simply the hypothesis $a \equiv
+\sum_{\varphi,\varphi' \in \llbracket \zeta_a \rrbracket} \varphi;a;\varphi'$.
+Rather than rely on general techniques ~\cite{doumane2019kleene,
+  hardin2002elimination, kozen1997kleenecompleteness}, we relalize our specific
+hypotheses directly.
+
+The key facet that differentiates Epik from \textsc{Ka+H} is the epistemic
+modality, which cannot be naively represented using universal
+hypotheses\footnote{The existence of this encoding is an interesting future
+  topic}. Sections~\ref{sec:model} and~\ref{sec:logic} discuss the
+model-theoretic and logical development of the epistemic operators.
 
@@
 \section{Epistemic guarded string models}
+\label{sec:model}
 
@@
 \section{The logical language of Epistemic \textsc{Kat}}
+\label{sec:logic}
+
 The standard language for Kleene algebra with tests has the signature
@@
 \section{Translation into the finite state calculus}
+\label{sec:fst}
+
 The finite state calculus is an algebra of regular sets of strings and
@@
 \section{Bounded Lazy Interpretation}
+\label{sec:lazy}
 
@@
 
-Epik is closely related to Kleene Algebra with Hypotheses
-\textsc{Ka+H}~\cite{cohen1994hypotheses}, which permits user-specification of
-futher equations beyond the axioms of \textsc{Ka}. In practice, these hypotheses
-are used to express domain-specific properties of the events, such as the
-commutativity. We can model both the state formulae and the event formulae as
-\textsc{KA+H} hypotheses. Specifically, the state formula $\varphi$ can be
-thought of as a hypothesis $\varphi \equiv 1$, and an event formula $\zeta_a$
-for event $a$, is simply the hypothesis $a \equiv \sum_{\varphi,\varphi' \in
-  \llbracket \zeta_a \rrbracket} \varphi;a;\varphi'$. In general, the Horn
-theory of \textsc{KA+H} is undecidable, however many special cases
-are~\cite{doumane2019kleene, hardin2002elimination,
-  kozen1997kleenecompleteness}.
-
-The key facet that differentiates Epik from \textsc{Ka+H} is the modality. Each
-alternative relation represents a set of uni-directional rewrites that must be
-applied for that agent, rather than the universal, optional rewrites represented
-by the \textsc{Ka+H} hypotheses. It may be possible to encode Epik's alternative
-relations into \textsc{Ka+H} hypotheses, but such an encoding is future work.
-
 The development here is concerned with defining concrete computable

### 2021-01-15 tex/pdraft.tex
a_1 -> a_h, epik -> epikat and prose cleaning

@@
   as sets of guarded string models in an epistemic extension of Kleene Algebra
-  with Tests (\textsc{Kat}). The account is framed as a formal language Epik
-  (Epistemic \textsc{Kat}) for defining such models. The language is implemented
-  by translation into the finite state calculus, and alternatively by modeling
-  propositions as lazy lists in Haskell. The syntax-semantics interface for a
-  fragment of English is defined by a categorial grammar.
+  with Tests (\textsc{Kat}). The account is framed as a formal language
+  Epi\textsc{Kat} (Epistemic \textsc{Kat}) for defining such models. The
+  language is implemented by translation into the finite state calculus, and
+  alternatively by modeling propositions as lazy lists in Haskell. The
+  syntax-semantics interface for a fragment of English is defined by a
+  categorial grammar.
 \end{abstract}
@@
 
-Enter Epik, which is a systematic way of defining such models, and can be
+Enter Epi\textsc{Kat}, which is a systematic way of defining such models, and can be
 applied to natural language semantics, specifically, epistemic semantics and
@@
 and computational model~\cite{kozen1997kleenecompleteness,kozen2001automata}.
-Epik leverages these algebrae to define a mathematical model
+Epi\textsc{Kat} leverages these algebrae to define a mathematical model
 (Sections~\ref{sec:model} \&~\ref{sec:logic}) and computationally interpret
@@
 
-{\em Related Work.} Epik synthesizes five antecedents.
+{\em Related Work.} Epi\textsc{Kat} synthesizes five antecedents.
 John McCarthy's {\em Situation Calculus} is the source of the idea of
@@
 
-\newcommand{\apH}{\ensuremath{a_1}}
-\newcommand{\apT}{\ensuremath{a_0}}
-\newcommand{\bpH}{\ensuremath{b_1}}
-\newcommand{\bpT}{\ensuremath{b_0}}
-
-\newcommand{\atTH}{\ensuremath{a_{01}}}
-\newcommand{\atHT}{\ensuremath{a_{10}}}
-\newcommand{\btTH}{\ensuremath{b_{01}}}
-\newcommand{\btHT}{\ensuremath{b_{10}}}
+\newcommand{\nop}{\ensuremath{o}}
+
+\newcommand{\apH}{\ensuremath{a_h}}
+\newcommand{\apT}{\ensuremath{a_t}}
+\newcommand{\bpH}{\ensuremath{b_h}}
+\newcommand{\bpT}{\ensuremath{b_t}}
+\newcommand{\cpH}{\ensuremath{c_h}}
+\newcommand{\cpT}{\ensuremath{c_t}}
+
+\newcommand{\atTH}{\ensuremath{a_{th}}}
+\newcommand{\atHT}{\ensuremath{a_{ht}}}
+\newcommand{\btTH}{\ensuremath{b_{th}}}
+\newcommand{\btHT}{\ensuremath{b_{ht}}}
 
@@
 For example, $\apH\atTH$ would require the coin to simultaneously heads-up and
-tails-up. To mechanize this, events in Epik come with Boolean pre-conditions and
+tails-up. To mechanize this, events in Epi\textsc{Kat} come with Boolean pre-conditions and
 post-conditions. In the running example, the Boolean variables $h$ (the coin is
@@
 
-Thus far, the presentation is closely related to Kleene Algebra with Hypotheses
-(\textsc{Ka+H})~\cite{cohen1994hypotheses}, which permits user-specification of
+Thus far, the presentation is closely related to Kleene Algebra with
+Hypotheses~\cite{cohen1994hypotheses}, which permits user-specification of
 futher equations beyond the axioms of \textsc{Ka}. Specifically, the state
@@
 Rather than rely on general techniques ~\cite{doumane2019kleene,
-  hardin2002elimination, kozen1997kleenecompleteness}, we relalize our specific
+  hardin2002elimination, kozen1997kleenecompleteness}, we realize our specific
 hypotheses directly.
 
-The key facet that differentiates Epik from \textsc{Ka+H} is the epistemic
-modality, which cannot be naively represented using universal
-hypotheses\footnote{The existence of this encoding is an interesting future
-  topic}. Sections~\ref{sec:model} and~\ref{sec:logic} discuss the
-model-theoretic and logical development of the epistemic operators.
+Epi\textsc{Kat} is first and foremost a multi-agent epistemic logic, with epistemic
+operators that cannot naively be represented as
+hypothesis\footnote{Characterizing this relationship precisely is a future
+  research question.}. Aside from formalizing the metaphysical modality seen so
+far, Sections~\ref{sec:model} and~\ref{sec:logic} develop the epistemic
+modality.
 
@@
   \fbox{
-    \begin{minipage}{\columnwidth}
+    \begin{minipage}{0.96\columnwidth}
   \(\begin{array}{lcl}
@@
 
-Epik is a specification language for possible worlds models that
+Epi\textsc{Kat} is a specification language for possible worlds models that
 includes declarations of events and states, state formulas, effect
 formulas, and additional information.  Figure \ref{epikfigure} shows
-an Epik program that describes a possible worlds model for two agents
+an Epi\textsc{Kat} program that describes a possible worlds model for two agents
 with information about one coin, events of the agents semi-privately
@@
 anything has happened, so her alternative is the ``no-operation'' or
-``no-information'' event event $o$. Bob's event relation is symmetric.
+``no-information'' event event $\nop$. Bob's event relation is symmetric.
 
@@
             \hspace{3mm} a01 -> a10 + a01 \\
-            \hspace{3mm} b10 -> o0 + o1 \\
-            \hspace{3mm} b01 -> o0 + o1 \\
+            \hspace{3mm} b10 -> o \\
+            \hspace{3mm} b01 -> o \\
             agent bob \\
@@
 
-\caption{Epik program describing a possible-worlds event sequence
+\caption{Epi\textsc{Kat} program describing a possible-worlds event sequence
   model for two agents with information about one coin, and events of
@@
 \label{fig:epik-example}
-\label{epikfigure}
 \end{figure}
 
-This paper focuses on defining a concrete possible worlds model from an Epik
+The sequel focuses on defining a concrete possible worlds model from an Epi\textsc{Kat}
 specification. The models are an extension of guarded-string models for Kleene
@@
 with atoms over a set ${\sf{B}}$ alternating with the symbols from ${\sf{P}}$.
-In the algebra described by Figure \ref{epikfigure}, ${\sf{P}}$ is the set of
-events $\{\apH,\apT,\bpH,\bpT,\atTH,\atHT,\btTH,\btHT,o\}$, and ${\sf{B}}$ is
+In the model described by Figure~\ref{fig:epik-example}, ${\sf{P}}$ is the set of
+events $\{\apH,\apT,\bpH,\bpT,\atTH,\atHT,\btTH,\btHT,\nop\}$, and ${\sf{B}}$ is
 $\{h,t\}$. As we already saw in \ex{effect}, $\mathcal{A}^\varphi_{\mathsf B}$
@@
 alternating with atoms over ${\sf{B}}$, and beginning and ending with atoms. In
-this construction, $w_1=H$, $w_2=Ha_1H$, $w_3=Ha_1Hb_1H$, and
-$w_4=Ha_1Hb_1Ha_{10}Tb_{01}Hb_1H$.
-
-The discussion of \ex{coin2} mentioned building worlds by incrementing worlds with
-events.  This is accomplished in guarded string models with fusion product
-$\diamond$, a partial operation that combines two guarded strings, subject to the
-condition that the atom at the end of the the first argument is identical to the atom
-at the start of the second one.  \ex{fusion1} gives some examples.
+this construction, $w_1=H$, $w_2=H \apH H$, $w_3=H \apH H \bpH H$, and
+$w_4=H \apH H \bpH H \atHT T \btTH H \bpH H$.
+
+The discussion of \ex{coin2} mentioned building worlds by incrementing worlds
+with events and maintaining the pre-and-post conditions. This is accomplished in
+guarded string models with the fusion product ($x \diamond y$), a partial
+operation that combines two guarded strings $x$ and $y$, subject to the
+condition that the atom at the end of the $x$ is identical to the atom at the
+start of $y$. \ex{fusion1} gives some examples.
 
@@
 
-\hspace{-2mm}\(H\,b_1\,H \; \diamond \; H \,a_1\, H = H\, b_1\, H \,a_1\, H \)
-
-\hspace{-2mm}\(T\,b_{01}\,H \; \diamond \; T \,a_1\, T = \textit{undefined} \)
+\hspace{-2mm}\(H\,\bpH\,H \; \diamond \; H \,\apH\, H = H\, \bpH\, H \,\apH\, H \)
+
+\hspace{-2mm}\(T\,\btTH\,H \; \diamond \; T \,\apH\, T = \textit{undefined} \)
 \end{example}
 
-Rather than individual guarded strings, elements of a guarded string
-model for \textsc{Kat} are sets of guarded strings.  In our application, these
-elements have the interpretation of propositions, which are sets of possible
+Rather than individual guarded strings, elements of a guarded string model for
+\textsc{Kat} are sets of guarded strings. In Epi\textsc{Kat}, these elements
+have the interpretation of propositions, which are (regular) sets of possible
 worlds. In a free guarded string model for \textsc{Kat}, any event can be
-adjacent to any atom in a guarded string that is an element of the
-underlying set for the algebra.  We instead impose the constraints
-coming from the state and effect formulas.  \ex{wf1} defines the
-well-formed guarded strings determined by an Epik specification.  Condition (i)
-says that each atom is consistent with the state constraint, and condition (ii)
-says that each constituent token event $\alpha_ie_i\alpha_{i+1}$ is
-consistent with the effect constraint on $e_i$.\footnote{
-An alternative is to define equations such as
-$\bar{\phi}=0$ (from the state formula $\phi$)
-and $a_1\,{=}\,ha_{1}h$ (from the effect formula $h\mathbin{:}h$ for event $a_1$),
-and construct a quotient algebra from the equivalence relation generated
-by these equations.  This results in equating sets of guarded strings in the free algebra that
-differ by guarded strings that are ill-formed according to the state and
-effect formulas.  In the development in the text, we instead use
-a set of guarded strings that are well-formed according to the state
-and effect formulas as the representative of the equivalence class.
-}
+adjacent to any atom in a guarded string that is an element of the underlying
+set for the algebra. However, this subsumes the valid worlds defined by the
+state and effect formulae. \ex{wf1} defines the \emph{well-formed} guarded
+strings (valid worlds) determined by an Epi\textsc{Kat} specification. Condition
+(i) says that each atom is consistent with the state constraint, and condition
+(ii) says that each constituent token event $\alpha_ie_i\alpha_{i+1}$ is
+consistent with its effects\footnote{ An alternative is to
+  define equations such as $\bar{\phi}=0$ (from the state formula $\phi$) and
+  $\apH\,{=}\,h \apH h$ (from the effect formula $h\mathbin{:}h$ for event
+  $\apH$), and construct a quotient algebra from the equivalence relation
+  generated by these equations. This results in equating sets of guarded strings
+  in the free algebra that differ by guarded strings that are ill-formed
+  according to the state and effect formulas. In the development in the text, we
+  instead use a set of guarded strings that are well-formed according to the
+  state and effect formulas as the representative of the equivalence class.}.
 
@@
 Well-formed guarded strings have the interpretation of worlds in the application
-to natural-language semantics.  The set of possible worlds in the Kripke frame
-determined by an Epik specification is the set of well-formed guarded strings.
-At this point, we could say that any set of worlds is a proposition, so that the
-set of propositions is the power set of the set of worlds
-\citep{montague1975formal,gallin1975intensional}.  We will instead define a more
-restrictive set of propositions corresponding to the regular sets of
-strings. This is deferred to the next section.  Certain sets of well-formed
-guarded strings have the addional interpretation of event types.  An event-type
-is something that can ``happen'' is different worlds. For example, $a_1$ corresponds to the 
-event type $\{H\,a_1\,H\}$, and $a_0$ corresponds to the event type $\{T\,a_0\,T\}$. Event
-types need not be singletons.
-
-The construction so far defines a set of worlds from an Epik specification.  Normally
+to natural-language semantics. The set of possible worlds in the modal frame
+determined by an Epi\textsc{Kat} specification is the set of well-formed guarded
+strings, and propositions are regular sets of guarded strings, rather than the
+power set \citep{montague1975formal,gallin1975intensional}. Certain sets of
+well-formed guarded strings have the addional interpretation of event types. An
+event-type is something that can ``happen'' in different worlds. For example,
+$\apH$ corresponds to the event type $\{H\,\apH\,H\}$, and $\nop$ to the event
+type $\{T\,\nop\,T, H\,\nop\,H\}$.
+
+The construction so far defines a set of worlds from an Epi\textsc{Kat} specification.  Normally
 the set is countably infinite, though some choices of effect formulas can result in a
@@
 
-An Epik specification defines an alternative relation on bare events for each
+An Epi\textsc{Kat} specification defines an alternative relation on bare events for each
 agent $a$, which is notated ${R}_{a}$, and lifted to a relation $\hat{R}_a$ on
@@
 
-Consider a scenario like the one from Figure 1, but with an additional agent
-Cal. The base world $T b_0 T c_0 T$ is one where the coin is tails, and first
-Bob looks at tails, and them Cal looks at tails. The first event $b_0$ has the
-alternatives $b_0$ and $b_1$ for Amy, and the second event $c_0$ has the
-alternatives $c_0$ and $c_1$ for Amy. This results in four combinations
-$b_0c_0$, $b_0c_1$, $b_1c_0$, and $b_1c_1$. But some of these worlds are bogus,
-such as $b_0c_1$, which says that Bob peeked at the coin and saw tails, and Cal
-peeked at the same coin and impossibly saw heads. These contradictions are
-filtered by post- and pre-conditions of events in the alternative world, so that
-the set of alernatives for Amy in $T b_0 T c_0 T$ is $\{T b_0 T c_0 T,H b_1 H
-c_1 H\}$, with two world-alternatives instead of four.
+Consider a scenario like the one from Figure~\ref{fig:epik-example}, augmented
+with an additional agent Cal. The base world $T\,\bpT\,T\,\cpT\,T$ is one where
+the coin is tails, and first Bob looks at tails, and them Cal looks at tails.
+The first event $\bpT$ has the alternatives $\bpT$ and $\bpH$ for Amy, and the
+second event $\bpT$ has the alternatives $\cpT$ and $\cpH$ for Amy. This results
+in four combinations $\bpT\cpT$, $\bpT\cpH$, $\bpH\cpT$, and $\bpH\cpH$, some of
+which, like $\bpT\cpH$, are bogus. These contradictions are filtered by post-
+and pre-conditions of events in the alternative world, so that the set of
+alternatives for Amy in $T\,\bpT\,T\,\cpT\,T$ is $\{T\,\bpT\,T\,\cpT\,T,
+H\,\bpH\,H\,\cpH\,H\}$, with two world-alternatives instead of four.
 
@@
   \fbox{
-    \begin{minipage}{\columnwidth}
+    \begin{minipage}{0.96\columnwidth}
       \(\begin{array}{rlrl}
@@
 
-\caption{The language of Epik terms and key derived
+\caption{The language of Epi\textsc{Kat} terms and key derived
   operators.}
@@
 With modalities and propositional negation added, the signature of $n$-agent
-epistemic \textsc{Kat} is $\tuple{K,+,\cdot,*,\bar{},0,1,\neg,\Diamond_1,
+Epi\textsc{Kat} is $\tuple{K,+,\cdot,*,\bar{},0,1,\neg,\Diamond_1,
   ..., \Diamond_n}$. Figure \ref{termfigure} defines the syntax of the language.
 Juxtaposition is used for product.  Terms in this language
-are used to represent the proposositional semantic values of English sentences.
-\ex{interpretation1} gives some examples.  To explain the first one, $\bullet$ as
-defined in Figure \ref{fig:epik-grammar} is the disjunction of the primitive events.
-Since a world is a well-formed sequence of events, $\bullet^*$ is the set of worlds.
-Multiplying by the state symbol $h$ in the term $\bullet^* h$ has the effect of
-conjoining $h$ with the atom at the end of the world.  So $\bullet^* h$ is the set
-of worlds where the coin ends heads-up.
+are used to represent the propositional semantic values of English sentences.
+
+Consider the examples in \ex{interpretation1}. The first term, $\bullet^*\,h$ uses
+the syntactic sugar $\bullet$, defined in Figure \ref{fig:epik-grammar} to be
+the disjunction of the primitive events. Since a world is a well-formed sequence
+of events, $\bullet^*$ is the set of worlds. Multiplying by the state symbol $h$
+in the term $\bullet^* h$ has the effect of conjoining $h$ with the atom at the
+end of the world. So $\bullet^* h$ is the set of worlds where the coin ends
+heads-up.
 
@@
 
-{\em Standard Epistemic Modalities.}
-Using our existential modal primitive $\Diamond_a$, and the dual
-encodings of $\Box_a$ and $\wedge$, we can encode the standard modal
-operators expressing knowledge $(\K_a)$ and belief $(\B_a)$
-as in \ex{encodings}.\footnote{Deeper analyisis of the lexical semantics
-  of {\em know} requires adding modeling of presupposition
-  \citep{collard2018finite}. The grammar fragment in Section 6 does not model
-  the presupposition of {\em know}, except as
-  an entailment.}
+{\em Standard Epistemic Modalities.} Using the modal primitive $\Diamond_a$, and the
+dual encodings of $\Box_a$ and $\wedge$, we can encode the standard modal
+operators expressing knowledge $(\K_a)$ and belief $(\B_a)$ as
+\ex{encodings}.\footnote{Deeper analysis of the lexical semantics of {\em know}
+  requires adding modeling of presupposition \citep{collard2018finite}. The
+  grammar fragment in Section 6 does not model the presupposition of {\em know},
+  except as an entailment.}
 
@@
 etc) are modeled using the event-alternatives in an
-Epik specification.\footnote{See the discussion of modal axioms \textbf{T} and
+Epi\textsc{Kat} specification.\footnote{See the discussion of modal axioms \textbf{T} and
   \textbf{D} below.}  The agents in Figure 1 do not always have reliable beliefs,
@@
 
-\newcommand{\epikspec}{\test,\event,\varphi,\zeta}
+\newcommand{\epikspec}{\test,\varphi,\zeta}
 \newcommand{\AphiB}{\mathcal A^\varphi_{\mathsf B}}
@@
 strings $\sem{p}^{\epikspec}$, where superscript captures dependence
-on an Epik specification.  Figure \ref{fig:interp2} defines the
+on an Epi\textsc{Kat} specification.  Figure \ref{fig:interp2} defines the
 interpretation.  The interpretation $\sem{ 1
@@
 \begin{figure}
-  \fbox{\begin{minipage}{\columnwidth}
+  \fbox{\begin{minipage}{0.97\columnwidth}
   \centering
@@
   \sem{ \bar{\sigma} }^{\epikspec} && \AphiB \setminus \sem{\sigma}^{\epikspec}\\
-  \sem{ e }^{\epikspec}&&  \setabs{\alpha e \beta}{\alpha \zeta_e \beta} \\
+  \sem{ e }^{\epikspec}&&  \setabs{\alpha e \beta}{\alpha \mathrel{\sem{\zeta_e}^\varphi} \beta} \\
   \sem{{p+q}}^{\epikspec} && \sem{{p}}^{\epikspec} \cup \sem{ q }^{\epikspec}\\
@@
   \end{minipage}}
-  \caption{Interpretation of Epik terms as sets of guarded strings}
+  \caption{Interpretation of Epi\textsc{Kat} terms as sets of guarded strings}
   \label{fig:interp2}
@@
 
-Summing up, given an Epik specification $\epikspec$, term $p$ (as defined
-syntactically in Figure \ref{termfigure}) is interpreted as a set of guarded strings
-$\sem{p}^{\epikspec}$.  Let $K^{\epikspec}$ be the sets that are interpretations of
-terms.  Then
+Summing up, given an Epi\textsc{Kat} specification $\epikspec$, term $p$ (as
+defined syntactically in Figure \ref{termfigure}) is interpreted as a set of
+guarded strings $\sem{p}^{\epikspec}$. Let $K^{\epikspec}$ be the sets that are
+interpretations of terms. Then
 $\tuple{K^{\epikspec},+,\cdot,*,\bar{},0,1,\neg,\Diamond_{a_1},...,\Diamond_{a_n}}$
-is a concrete guarded string interpretation for the signature of epistemic \textsc{Kat}, with
-operations as in Figure \ref{fig:interp2} (e.g. the binary operation $+$ is union,
-and the unary operation $\Diamond_a$ is pre-image relative to $\hat{R}_a$).  This
-provides a concrete $n$-agent general modal frame
-$\tuple{\sem{\bullet^*}^{\epikspec}, \hat{R}_1,...,\hat{R}_n,K^{\epikspec}}$.  The
-frame consists of a set of worlds, an epistemic-alternative relation for each agent,
-and a set of propositions, each of which is a set of worlds.  It is used as a target
-for natural-language interpretation in Section~\ref{sec:categorial-grammar}.
+is a concrete guarded string interpretation for the signature of epistemic
+\textsc{Kat}, with operations as in Figure \ref{fig:interp2}. This provides a
+concrete $n$-agent general modal frame $\tuple{\sem{\bullet^*}^{\epikspec},
+  \hat{R}_1,...,\hat{R}_n,K^{\epikspec}}$. The frame consists of a set of
+worlds, an epistemic-alternative relation for each agent, and a set of
+propositions, each of which is a set of worlds. It is used as a target for
+natural-language interpretation in Section~\ref{sec:categorial-grammar}.
 
@@
 the soundness of the standard modal axioms given our semantics
-\citep{hughes1996new}. The axioms in \ex{valid-axioms}
+\citep{hughes1996new}. The axioms in \ex{valid-axioms} are valid.
 
@@
 regular relations between strings that was designed for use in
-computationonal phonology and morphographemics
+computational phonology and morphographemics
 \citep{kaplan1994regular,beesley2003finite}.  Current implementations
@@
 \citep{hulden2009foma,linden2009hfst,karttunen2010update}.  Such definitions are used
-here to construct of a model for epistemic \textsc{Kat} inside the finite state
+here to construct a model for Epi\textsc{Kat} inside the finite state
 calculus (Figures~\ref{fstfigure1}\&\ref{relclosure}).
@@
 with event symbols to encode a world.  In this construction, $w_3$ of the example is
-the string {\tt{10\hspace{2pt}a1\hspace{2pt}10\hspace{2pt}b1\hspace{2pt}10}}.
+the string {\tt{1\!0\;ah\;1\!0\;bh\;1\!0}}.
 
@@
   {\tt UnequalStPair} \hspace{2mm} \\
-  Sequence of two unequal atoms such as $0110~0111$.
+  Sequence of two unequal atoms such as $0110~~0111$.
 \end{hangingpar}
@@
 \end{tcolorbox}
-\caption{\raggedright Definition in Fst of \textsc{Kat} product and star.
-  %% Where {\tt X} and {\tt Y} are regular sets and
-  %% {\tt R} and {\tt S} are regular relatiions,
-  %% {\tt X{\&}Y} is the intersection of {\tt X} and {\tt Y},
-  %% {\tt X|Y} is the union of {\tt X} and {\tt Y},
-  %% {\tt \~{}X} is the complement of {\tt X},
-  %% {\tt R{.o.}S} is the composition of {\tt R} and {\tt S},
-  %% {\tt R{.}l} is the co-domain of {\tt R}, and
-  %% {\tt {\$}X} is the set of strings that have a substring in $X$.   
-}
+\caption{\raggedright Epi\textsc{Kat} product and star defined in Fst.}
 \label{fstfigure1}
@@
 
-Translating the Epik terms $0$, $1$, $b$, and $e$ are
-straightforward: we simply convert the atoms as previously described,
-and decorate the events $e$ with their compatible atoms. For example $a_1$ becomes
+Translating the Epi\textsc{Kat} terms $0$, $1$, $b$, and $e$ is
+straightforward: convert the atoms as previously described,
+and decorate the events $e$ with their compatible atoms. For example, $a_1$ becomes
 an Fst term denoting
-$\{01a_101\}$. Fst has built-in operations of union (\texttt{|}),
-and intersection (\texttt{\&}), which are
-the sum and intersection operations in the guarded string algebra as represented
-in Fst.
-Fst set difference (\texttt{-}) is used to define propositional complement as the
-difference between the set of all worlds and the argument.
+$\{01\apH01\}$.
+
+Further, Epi\textsc{Kat}'s union $(+)$ and intersection $(\wedge)$ operators can
+be directly translated to Fst's built-in analogous operators ((\texttt{|}) and
+(\texttt{\&}). Propositional complement in Epi\textsc{Kat} becomes Fst set
+difference (\texttt{-}) from the set of all worlds.
 
@@
 Finally, the epistemic alternative relation for each agent, i.e. the
-\texttt{agent} relation from Figure~\ref{epikfigure} is implemented in Fst.
+\texttt{agent} relation from Figure~\ref{fig:epik-example} is implemented in Fst.
 Combining the agent relation with the effect formula gives a relation on events
@@
 
-We also implement the semantics of Epik terms using lazy lists in
-Haskell, rather than the direct interpretation as sets. Using lists
-sidesteps checking the set invariant (elements are unique) for large
-sets, such as $\bullet^*$, and laziness allows us to delay computing
-these large sets until they are actually needed. To sidestep the
-infiteness of models, we parameterize the interpretation function on a
-positive integer $n$ and only produce guarded strings of length $n$ or
-less.
-
-%% We translate the Epik terms into a Haskell algebraic datatype that
+Epi\textsc{Kat} specifications also target lazy lists in Haskell, rather than the direct
+interpretation as sets. Using lists sidesteps checking the set invariant (that
+elements are unique) for large sets, such as $\bullet^*$, and laziness allows us
+to delay computing these large sets until they are actually needed. To sidestep
+the infiteness of models, we parameterize the interpretation function on a
+positive integer $n$ and only produce guarded strings of length $n$ or less.
+
+%% We translate the Epi\textsc{Kat} terms into a Haskell algebraic datatype that
 %% represents terms with the same signature as described in Section 2. We
@@
 the unbounded interpretation into sets of strings, except for the
-bounds checking. The full details are shown in
+(lazy) bounds checking. The full details are shown in
 Figure~\ref{fig:bound-interp}. First note that when $n=0$, the
-denotation is empty, denoted $[]$. Terms of the form $0$, $1$, $e$,
+denotation is empty, written $[]$. Terms of the form $0$, $1$, $e$,
 and $\psi$ have the same denotation as before, translated into a list
-(denoted $\lfloor S\rfloor$, for a set $S$). We compute atoms using
+(written $\lfloor S\rfloor$, for a set $S$). We compute atoms using
 BDDs, which concisely represent boolean functions~\cite{lee1959representation}.
 
-We lift the remaining operators (except Kleene star) to their list
-equivalents: union becomes list append (written $\texttt{+\!+}$);
-fusion product is lifted to lists instead of sets, negation is
-implemented using list difference ($\setminus$), and the modal
-operator lifts the alternative relation over lists of
-strings\footnote{Figure~7 depicts this using the
-  list comprehension notation, which is analogous to set builder
-  notation, except that it is written using square brackets. Element
-  order is evoked by the keyword \texttt{for}, rather than using the
-  unordered $\forall$.}. The only caveat to these direct
-interpretations is that we restrict the operators to have size $\leq
-n$, denoted as $l|_n$ for a list of guarded strings $l$.
+We lift the remaining operators (except Kleene star) to their list equivalents:
+union becomes list append (written $\texttt{+\!+}$); fusion product is lifted to
+lists instead of sets, negation is implemented using list difference
+($\setminus$), and the modal operator lifts the alternative relation over lists
+of strings\footnote{Figure~7 depicts this using the list comprehension notation,
+  which is analogous to set builder notation, except that it is written using
+  square brackets. Element order is evoked by the keyword \texttt{for}, rather
+  than using the unordered $\forall$.}. The only caveat to these direct
+interpretations is that we lazily restrict the strings to have size $\leq n$,
+written as $l|_n$ for a list of guarded strings $l$.
 
@@
 
-\newcommand{\semlst}[2]{\llparenthesis#1\rrparenthesis^{\mathsf B, \mathsf P, \phi, \zeta}_{#2}}
+\newcommand{\semlst}[2]{\llparenthesis#1\rrparenthesis^{\mathsf B, \varphi, \zeta}_{#2}}
 \newcommand{\setToList}[1]{\lfloor#1\rfloor}
@@
   \fbox{
-    \begin{minipage}{\columnwidth}
-      \hspace{-0.9em}\(\begin{array}{r >{\triangleq}c l}
+    \begin{minipage}{0.95\columnwidth}
+      \hspace{-0.9em}\(\begin{array}{r >{\triangleq}c >{}l}
       \semlst p 0 && [] \\
@@
       \semlst 1 n && \setToList \AphiB \\
-      \semlst e n && [ \alpha e \beta \mid \alpha \mathrel{\zeta_e} \beta ] \\
+      \semlst e n && [ \alpha e \beta \mid \alpha \mathrel{\sem{\zeta_e}^\varphi} \beta ] \\
       \semlst b n && \setToList{ \mathcal A_{\mathsf B}^{b\psi}} \\
@@
       \semlst{p^*} n && [] + (\semlst p n \diamond \semlst{p^*} {n-i}) |_n \\
-      \multicolumn 3 r {\footnotesize \text{where}~
-        \begin{array}[t]{r >{\hspace{-.7em}}l}
-          i &= \max\{1,\min\{|g|\mid g \in \semlst p n \}\} \\
-      \end{array}}\\
+      \multicolumn 3 r {\scriptstyle \text{where~}
+          i \mathbin{=} \max\{1,\min\{|g|\mid g \in \semlst p n \}\}} \\
       \semlst{\neg p} n && \semlst {\bullet^*} n \setminus \semlst p n \\
@@
     \caption{Categorial grammar lexicon, showing word form (column 1), a
-      categorial type (column 2), and a semantic translation in Epik extended
+      categorial type (column 2), and a semantic translation in Epi\textsc{Kat} extended
       with lambda abstraction (column 3).}
@@
   whether}-complements, negation, and predicate and sentence
-conjunction.\marginpar{Conjunction missing!} Figure \ref{lexiconfigure} gives
+conjunction. Figure \ref{lexiconfigure} gives
 the lexicon.\footnote{Category symbols use Lambek/Bar-Hillel notation for
@@
 beta reductions that eliminate all lambdas in logical forms for clauses.  In
-consequence, the semantic term translating a sentence is an Epik logical term.  Such
+consequence, the semantic term translating a sentence is an Epi\textsc{Kat} logical term.  Such
 a term designates a set of possible words (guarded strings).  By way of example,
@@
   sizes need not be small, especially as the cardinality of $\sf{B}$ increases.  A
-  certain Epik model with fourteen primitive Booleans has the set of worlds
+  certain Epi\textsc{Kat} model with fourteen primitive Booleans has the set of worlds
   representented by a finite state machine with 184,794 states and 257,881 edges.}  In
@@
 \section{Discussion}
-Epistemic \textsc{Kat} is designed for use in linguistic semantics research, for
-compultational linguistics research on model-theoretically grounded semantics,
+Epi\textsc{Kat} is designed for use in linguistic semantics research, for
+computational linguistics research on model-theoretically grounded semantics,
 and for students in graduate course modules on intensional formal semantics.
+Currently, the grammar covers statives, {\em that-} and {\em
+  whether}-complements, negation, and predicate and sentence conjunction.
 
@@
   arguments.} While this approach is simple and attractive for our applications,
-it would be possible to use Epik with other frameworks that support lambda
+it would be possible to use Epi\textsc{Kat} with other frameworks that support lambda
 extensions of logical languages.
@@
 
-The semantics presented here is exclusively concerned with events and
-worlds.  Semantic type systems for natural language usually include a
-type for individuals \citep{gallin1975intensional}.  In the coin example, the individuals are hidden
-in the primitive event symbols such as $a_1$ and $b_1$. This situation
-would get worse in a more elaborate model with more agents and
-multiple coins.  The solution to this should be to base the world
-construction on grounded event terms such as
-$\Prime{look}(\Prime{a},\!\Prime{k},\!\Prime{1})$, for ``agent $\Prime{a}$ see
-that coin $\Prime{k}$ is heads'', rather than atomic event symbols.  This way
-of proceeding is found in research on situation calculus
-\citep{reiter2001knowledge}.  How to incorporate it in the scheme for
-Epik specifications and into the computational parts of the proposal
-is a topic for future research.  In this it is not clear
-whether it is possible to introduce quantification over individuals,
-together with individual constants. \citet{rooth2017finite} develops an approach to a positive
-answer, based on introducing markers for witnesses for discourse
-referents in the construction of worlds.
-
-The development here is concerned with defining concrete computable
-possible worlds models, and applying them in natural language
-semantics. Issues for further investigation are mathematical
-characterizations of epistemic \textsc{Kat}s, e.g. sound and complete axioms
-and coalgebrae.
-
-Source code, examples, and instructions for Epik are available at
-\url{https://github.com/ericthewry/epik}.
+The semantics presented here is exclusively concerned with events and worlds.
+Semantic type systems for natural language usually include a type for
+individuals \citep{gallin1975intensional}. In the coin example, the individuals
+are hidden in the primitive event symbols such as $a_1$ and $b_1$. This
+situation would get worse in a more elaborate model with more agents and
+multiple coins. The solution to this should be to base the world construction on
+grounded or parametric event terms such as
+$\Prime{peek}(\Prime{a},\!\Prime{k},\!\Prime{h})$, for ``agent $\Prime{a}$ peeks
+at heads-up coin $\Prime{k}$'', rather than atomic event symbols. This approach
+is found in research on situation calculus \citep{reiter2001knowledge}. How to
+incorporate it in the scheme for Epi\textsc{Kat} specifications and into the
+computational parts of the proposal is a topic for future research. The
+ramifications of quantification or abstraction over individuals is unknown.
+\citet{rooth2017finite} develops an approach based on introducing markers for
+witnesses for discourse referents in the construction of worlds.
+
+The development here is concerned with defining concrete computable possible
+worlds models, and applying them in natural language semantics. Logical and
+computational characterizations of Epi\textsc{Kat}, such as sound and complete
+axioms, coalgebrae, and relations to other logics, warrant further investigation.
+
+Source code, examples, and instructions for building and running Epi\textsc{Kat}
+are available under an open-source license at \url{https://github.com/ericthewry/epikat}.
 
@@
     the second column a categorial type, and third column a semantic translation
-    in a language that extends the Epik logical language with lambda.}
+    in a language that extends the Epi\textsc{Kat} logical language with lambda.}
   \label{lexiconfigure}

### 2021-01-15 tex/pdraft.tex
fix typo
Substitute for (12) plus changes to end of Section 6.

@@
 \newcommand{\HEADS}{\bullet^*h}
-\newcommand{\AKWH}{\neg (\neg \HEADS \vee \Diamond_{a} \neg\HEADS) + \neg(\HEADS \wedge \Diamond_{a}\HEADS)}
-\newcommand{\BKAKWH}{\neg(\neg(\AKWH) + \Diamond_{b}\neg(\AKWH))}
-\newcommand{\AKBKAKWH}{\neg(\neg(\BKAKWH) + \Diamond_{a}\neg(\BKAKWH))}
+\newcommand{\TAILS}{\bullet^*t}
 
 \begin{examples} \label{three}
-\itema It's tails and Amy knows that Bob knows that Amy knows whether it's heads.
-% (_*;t) & ~ <amy> ~ ~ <bob> ~((~ (<amy> ~ (_*;h))) + ~ (<amy> (_*;h)) )
-\itemb \( \displaystyle \bullet^*t \wedge \AKBKAKWH \)
-\itemc \(\bullet^*t \wedge \K_{a}(\K_{b}(\K_{a} \bullet^* h + \K_{a} \neg \bullet ^* h))\)
-\itemd Amy knows that it's tails.
-\iteme \(
-\bullet ^*t \wedge \neg \Diamond_{a} \neg \bullet^*t~~~~~~(\equiv \K_a \bullet^* t)
+  \itema Amy knows that it's tails and doesn't believe that it's heads and
+  believes that Bob believes that it's heads.
+\itemb \(
+\begin{array}[t]{ll}
+  \neg(\neg \TAILS + \Diamond_a \neg \TAILS) & \wedge \\
+  \neg \neg \Diamond_a \neg \HEADS & \wedge \\
+  \neg \Diamond_a \neg\neg\Diamond_b \neg \HEADS
+\end{array}  
 \)
-
+\itemc \(\K_{a}\!\TAILS \; \wedge \, \neg\B_{a}\!\HEADS \; \wedge \; \B_{a}\B_{b}\HEADS\)
+\itemd Bob believes that it's tails.
+\iteme \(\B_{b} \TAILS \)
 \end{examples}
 
-Sentence \exi{three}{d} is assigned the logical form \exi{three}{e} by
-the grammar.  Logical relations between propositions are checked in
+% This conversion was done by hand. It would be good to put it into the scheme program.
+% edge      : 695 Amy knows that its tails and doesnt believe that its heads and believes that Bob believes that its heads        (0 19)  t
+%semantics : (((Not ((Not tails) + (Diamond amy (Not tails)))) & (Not (Not (Diamond amy (Not heads))))) & (Not (Diamond amy (Not (Not (Diamond bob (Not heads)))))))
+% (((Not ((Not tails) + (Diamond amy (Not tails))))
+% & (Not (Not (Diamond amy (Not heads)))))
+% & (Not (Diamond amy (Not (Not (Diamond bob (Not heads)))))))
+%\[
+%\begin{array}{ll}
+%  \neg(\neg \TAILS + \Diamond_a \neg \TAILS) & \wedge \\
+%  \neg \neg \Diamond_a \neg \HEADS & \wedge \\
+%  \neg \Diamond_a \neg\neg\Diamond_b \neg \HEADS
+%\end{array}  
+%\]
+
+Sentence \exi{three}{d} is assigned a logical form that is syntactically
+equivalent to \exi{three}{e}. Logical relations between propositions are checked in
 the finite state calculus by checking set-theoretic relations between
-sets of worlds.  For instance entailment $p \rightarrow q$ is decided
+sets of worlds.  Propositional entailment from $p$ to $q$ is decided
 by checking in an interpreter for the finite state calculus whether $p
-- q$ is non-empty.  In the model defined by Figure 2, the propositions
-\exi{three}{b} and \exi{three}{e} are independent. They are equivalent
-in a version without secret flipping.
-
-%% \begin{examples} \label{entail}
-%% \itema Amy knows that it's tails.
-%% \itemb \(
-%% \neg \Diamond_{a} \neg \bullet^*t
-%% \)
-%% % \itemc \(\displaystyle \Box_{a}\Box_{b}(\Box_{a} \neg \bullet^* h + \Box_{a}\neg \bullet^* t)\) \hfill \mbox{} \(~~\cdot~~\Box_{a}\neg\Box_{b}\Box_{a}\neg \bullet^* t\)
-%% \end{examples}
+- q$ is non-empty.  In the model defined by Figure 2, proposition
+\exi{three}{c} entails proposition \exi{three}{e}.
+
+Another way of using the Fst interpreter is to display worlds in a given proposition.
+\ex{p4worlds} shows the three words of length three in proposition \exi{three}{c}.
+In the first one, Bob looks at heads, then Amy looks at heads, then Amy secretly
+turns the coin, believing that she is turning it from heads to tails because she
+believed after two steps that the coin was heads. Intuitively sentence
+\exi{three}{a} is true in this scenario.
+
+\begin{example} \label{p4worlds}
+\(
+\begin{array}[t]{l}  
+1 0 \; bh \; 1 0 \; ah \; 1 0 \; aht \;  0 1 \\
+1 0 \; bh \; 1 0 \; aht \; 0 1 \; at \; 0 1 \\
+1 0 \; ah \; 1 0 \; bh \; 1 0 \; aht \; 0 1 \\
+\end{array}
+\)
+\end{example}
 
 \section{Discussion}
-Epi\textsc{Kat} is designed for use in linguistic semantics research, for
+Epi\textsc{Kat} is designed for research in linguistic semantics, for
 computational linguistics research on model-theoretically grounded semantics,
-and for students in graduate course modules on intensional formal semantics.
+and for course modules on intensional formal semantics.
 Currently, the grammar covers statives, {\em that-} and {\em
-  whether}-complements, negation, and predicate and sentence conjunction.
-
+  whether}-complements, negation, and conjunction.
 There are straightforward extensions to additional linguistic phenomena, such as
@@
 
-\clearpage
+%\clearpage
 

### 2021-01-15 tex/pdraft.tex
49 states and 110 edges

@@
 \exi{three}{b} using logical equivalences. Either term is compiled in an
-implementation of the finite state calculus to a finite state machine with 10 states
-and 23 edges, which accepts a countably infinite set of worlds.\footnote{ Machine
+implementation of the finite state calculus to a finite state machine with 49 states
+and 110 edges, which accepts a countably infinite set of worlds.\footnote{ Machine
   sizes need not be small, especially as the cardinality of $\sf{B}$ increases.  A

### 2021-01-15 tex/pdraft.tex

--- 2021-01-15 tex/pdraft.tex
@@
   Constructive and computable multi-agent epistemic possible worlds models are
-  as sets of guarded string models in an epistemic extension of Kleene Algebra
-  with Tests (\textsc{Kat}). The account is framed as a formal language
-  Epi\textsc{Kat} (Epistemic \textsc{Kat}) for defining such models. The
-  language is implemented by translation into the finite state calculus, and
+  interpreted as sets of guarded string models in an epistemic extension of
+  Kleene Algebra with Tests (\textsc{Kat}). The account is framed as a formal
+  language Epi\textsc{Kat} (Epistemic \textsc{Kat}) for defining such models.
+  The language is implemented by translation into the finite state calculus, and
   alternatively by modeling propositions as lazy lists in Haskell. The
@@
 and computational model~\cite{kozen1997kleenecompleteness,kozen2001automata}.
-Epi\textsc{Kat} leverages these algebrae to define a mathematical model
+Epi\textsc{Kat} leverages these algebras to define a mathematical model
 (Sections~\ref{sec:model} \&~\ref{sec:logic}) and computationally interpret
@@
 
-{\em Related Work.} Epi\textsc{Kat} synthesizes five antecedents.
-John McCarthy's {\em Situation Calculus} is the source of the idea of
-constructing possible worlds as event sequences
-\citep{mccarthy1963situations,reiter2001knowledge}.  The algebraic
-theory of {\em Kleene Algebra with Tests} characterizes algebras with
-elements corresponding to propositions and event types in our
-application \citep{kozen2001automata}.  {\em Action models} in dynamic
-epistemic semantics introduce the technique of constructing epistemic
-models from primitive alternative relations on events, in order to
-capture the epistemic consequences of perceptual and communicative
-events \citep{baltag1999logic}.  Literature on {\em finite state
-  methods in linguistic semantics} has used event strings and sets of
-event strings to theorize about tense and aspect in natural language
-semantics
-\citep{fernando2004finite,fernando2007observing,carlson2009tense} and
-to express intensions \cite{fernando2017intensions}.  Work on {\em
-  finite state intensional semantics} has investigated how to do the
-semantics of intensional complementation in a setting where
-compositional semantics is expressed in a finite state calculus
-\citep{rooth2017finite,collard2018finite}.
+{\em Related Work.} Epi\textsc{Kat} synthesizes five antecedents. John
+McCarthy's {\em Situation Calculus} is the source of the idea of constructing
+possible worlds as event sequences
+\citep{mccarthy1963situations,reiter2001knowledge}. The algebraic theory of {\em
+  Kleene Algebra with Tests} (\textsc{Kat}) characterizes algebras of regular
+sets of guarded strings \citep{kozen2001automata}, which form the basis for
+Epi\textsc{Kat}'s propositions and event types. {\em Action models} in dynamic
+epistemic semantics introduce the technique of constructing epistemic models
+from primitive alternative relations on events, in order to capture the
+epistemic consequences of perceptual and communicative events
+\citep{baltag1999logic}. Literature on {\em finite state methods in linguistic
+  semantics} has used event strings and sets of event strings to theorize about
+tense and aspect in natural language semantics
+\citep{fernando2004finite,fernando2007observing,carlson2009tense} and to express
+intensions \cite{fernando2017intensions}. Work on {\em finite state intensional
+  semantics} has investigated how to do the semantics of intensional
+complementation in a setting where compositional semantics is expressed in a
+finite state calculus \citep{rooth2017finite,collard2018finite}.
 
@@
   Concealed Coin}. Amy and Bob are seated at a table. There is a coin on the
-table under a cup, heads up. The coin could be heads-up $H$ or tails-up $T$, and
+table under a cup. The coin could be heads-up ($H$) or tails-up ($T$), and
 neither agent knows which it is. Call this initial possible world $w_1$.
@@
 Juxtaposition indicates the order of events; for example, $w_3$ indicates that
-after the events of $w_2$, event $b_1$ occurs, which corresponds to, in the
-initial world, Amy surreptitiously peeking at a heads-up coin, and then Bob
-doing the same.
+after the events of $w_2$, event $b_1$ occurs, which, starting in the initial
+world $w_1$ corresponds to Amy peeking at a heads-up coin, and then Bob doing
+the same.
 
 \begin{example} \label{coin2}
+  \hspace{-1.5em}
   \(w_2 \triangleq w_1\apH \hfill
@@
 each world $w_i$, i.e. \emph{after} the events have happened. See \ex{coin3} for
-examples, such as in $w_4$, where Bob knows it's heads (because he peeked after
+examples such as in $w_4$, where Bob knows it's heads (because he peeked after
 turning the coin), but Amy does not (because Bob secretly turned the coin over).
@@
 To create realistic models, contradictory event sequences must be prohibited.
-For example, $\apH\atTH$ would require the coin to simultaneously heads-up and
-tails-up. To mechanize this, events in Epi\textsc{Kat} come with Boolean pre-conditions and
-post-conditions. In the running example, the Boolean variables $h$ (the coin is
-heads-up) and $t$ (the coin is tails-up) represent the coin's state. Then, since
-Amy can only observe heads when the coin is heads-up, event $\apH$ has the
-invariant condition $h$. Similarly, Amy can only turn the coin from tails to
-heads if it shows tails, so $\atTH$ has the tails-up precondition $t$ and the
-heads-up postcondition $h$.
+For example, $\apH\atTH$ would require the coin to simultaneously be heads-up
+and tails-up. To mechanize such reasoning, events in Epi\textsc{Kat} come with
+Boolean pre-conditions and post-conditions \emph{\`a la} Hoare
+Logic~\cite{hoare1969axiomatic}. In the running example, the Boolean variables
+$h$ (the coin is heads-up) and $t$ (the coin is tails-up) represent the coin's
+state. Then, since Amy can only observe heads when the coin is heads-up, event
+$\apH$ has the invariant condition $h$. Similarly, Amy can only turn the coin
+from tails to heads if it shows tails, so $\atTH$ has the tails-up precondition
+$t$ and the heads-up postcondition $h$.
 
@@
 the effects for $\apH$ are written $\ev h h$, and for $\atTH$ are written $\ev t
-h$. Effect formulae are interpreted relations on boolean valuations, as defined
+h$. Effect formulas are interpreted as relations on boolean valuations, as defined
 in Figure~\ref{effectfigure}.
 
-However, a coin cannot simultanously show both heads and tails! Currently the
+However, a coin cannot simultanously show both heads and tails! Currently, the
 precondition $h$ of $\atHT$ only says that heads must be showing, and says
@@
 $\bar t$. The effect valuations are restricted to only the feasible ones via a
-\emph{state formula}, shown in \ref{effect}, and demonstrated for $\apH$ in
-\ref{effect}. The state formula, also shown in \ref{effect}, says that the coin
-is either heads-up or tails-up, but not both\footnote{Two Booleans constrained
-  by a state formula, is a more illustrative example than using a single boolean
-  and modelling tails as $\bar h$.}. Here juxtaposition represents conjuction,
-$+$ is disjunction, and overbar is negation.
-
-With the state formula in hand, $\apH\atTH$ is evidently contradictory, the
-post-condition $h$ of $\apH$ is incompatible with the pre-condition $t$ of
+\emph{state formula}, shown, and demonstrated for $\apH$ in \ref{effect}. The
+state formula says that the coin is either heads-up or tails-up, but not
+both\footnote{Two Booleans constrained by a state formula, is more illustrative
+  a single boolean $h$ where tails is just $\bar h$.}. Here juxtaposition
+represents conjuction, $+$ is disjunction, and overbar is negation.
+
+With the state formula in hand, $\apH\atTH$ is evidently contradictory, since
+the post-condition $h$ of $\apH$ is incompatible with the pre-condition $t$ of
 $\atTH$. See Figure~\ref{effectfigure} for more details.
@@
 \mbox{state formula $\varphi$} & h\bar{t} + \bar{h}t \\
-\mbox{effect formula $\zeta_{a_1}$ for $a_1$} &  h \mathbin{:} h \\
+\mbox{effect formula $\zeta_{\apH}$ for $\apH$} &  h \mathbin{:} h \\
  \mathcal{A}^\varphi_{\mathsf B} & \{h\bar{t},\bar{h}t\}   \\
-\sem{\zeta_{a_1}}^\varphi & \{\pair{h\bar{t}}{h\bar{t}}\}   \\
+\sem{\zeta_{\apH}}^\varphi & \{\pair{h\bar{t}}{h\bar{t}}\}   \\
 \end{array}\)
@@
 
-Epi\textsc{Kat} is first and foremost a multi-agent epistemic logic, with epistemic
-operators that cannot naively be represented as
-hypothesis\footnote{Characterizing this relationship precisely is a future
-  research question.}. Aside from formalizing the metaphysical modality seen so
-far, Sections~\ref{sec:model} and~\ref{sec:logic} develop the epistemic
-modality.
+Epi\textsc{Kat} is first and foremost a multi-agent epistemic logic, with
+epistemic operators that cannot naively be represented as
+hypotheses\footnote{Characterizing this relationship precisely is future work.}.
+Aside from formalizing the metaphysical modality seen so far,
+Sections~\ref{sec:model} and~\ref{sec:logic} develop the epistemic modality.
 
@@
 includes declarations of events and states, state formulas, effect
-formulas, and additional information.  Figure \ref{epikfigure} shows
+formulas, and additional information.  Figure \ref{fig:epik-example} shows
 an Epi\textsc{Kat} program that describes a possible worlds model for two agents
@@
 
-Similarly, Her alternatives for $\atTH$ (she turns the coin over from heads to
+Similarly, her alternatives for $\atTH$ (she turns the coin over from heads to
 tails) are $\atTH$ and $\atHT$, indicating that she doesn't know, \emph{a
   priori}, whether she's turning the coin from $H$ to $T$ or from $T$ to $H$.
-She has the same alternatives for the event $\atHT$. Conversely, When Bob
+She has the same alternatives for the event $\atHT$. Conversely, when Bob
 secretly turns the coin over, in event $\btTH$ or $\btHT$, she doesn't know
 anything has happened, so her alternative is the ``no-operation'' or
-``no-information'' event event $\nop$. Bob's event relation is symmetric.
+``no-information'' event $\nop$. Bob's event relation is symmetric.
 
@@
 
-The discussion of \ex{coin2} mentioned building worlds by incrementing worlds
-with events and maintaining the pre-and-post conditions. This is accomplished in
-guarded string models with the fusion product ($x \diamond y$), a partial
-operation that combines two guarded strings $x$ and $y$, subject to the
-condition that the atom at the end of the $x$ is identical to the atom at the
-start of $y$. \ex{fusion1} gives some examples.
+The discussion of \ex{coin2} mentioned building worlds by incrementing smaller
+worlds with events and maintaining the pre- and post- conditions. This is
+accomplished in guarded string models with the fusion product ($x \diamond y$),
+a partial operation that combines two guarded strings $x$ and $y$, subject to
+the condition that the atom at the end of the $x$ is identical to the atom at
+the start of $y$. \ex{fusion1} gives some examples.
 
@@
 set for the algebra. However, this subsumes the valid worlds defined by the
-state and effect formulae. \ex{wf1} defines the \emph{well-formed} guarded
+state and effect formulas. \ex{wf1} defines the \emph{well-formed} guarded
 strings (valid worlds) determined by an Epi\textsc{Kat} specification. Condition
@@
 in four combinations $\bpT\cpT$, $\bpT\cpH$, $\bpH\cpT$, and $\bpH\cpH$, some of
-which, like $\bpT\cpH$, are bogus. These contradictions are filtered by post-
-and pre-conditions of events in the alternative world, so that the set of
+which, like $\bpT\cpH$, are bogus. These contradictions are filtered by pre-
+and post-conditions of events in the alternative world, so that the set of
 alternatives for Amy in $T\,\bpT\,T\,\cpT\,T$ is $\{T\,\bpT\,T\,\cpT\,T,
@@
 $\tuple{K^{\epikspec},+,\cdot,*,\bar{},0,1,\neg,\Diamond_{a_1},...,\Diamond_{a_n}}$
-is a concrete guarded string interpretation for the signature of epistemic
-\textsc{Kat}, with operations as in Figure \ref{fig:interp2}. This provides a
+is a concrete guarded string interpretation for the signature of
+Epi\textsc{Kat}, with operations as in Figure \ref{fig:interp2}. This provides a
 concrete $n$-agent general modal frame $\tuple{\sem{\bullet^*}^{\epikspec},
@@
   \textbf{D} & \Box_a p \to \Diamond_a p & \text{if } \hat{R}_a \text{ is a function} \\
-  \textbf{4} & \Box_a p \to \Box_a\Box_a p & \text{if } \hat{R}_a \text{ is idempotent}
   \end{array}\)
@@
 
-The condition on \textbf{4} is essentially just a restatement of the
-theorem in relational terms. The conditions on the remaining standard
-modal axioms, \textbf{B} and \textbf{5}, are similar.
+The conditions on axioms \textbf{4}, \textbf{5} and \textbf{B} are just a
+restatements of the theorem in relational terms.
 
@@
 embedding.  Using the grammar and parser, the sentence is mapped to the term in
-\exi{three}{b}. \exi{three}{c} shows a simplified logical from constructed form
-\exi{three}{b} using logical equivalences. Either term is compiled in an
+\exi{three}{b}. \exi{three}{c} shows a simplified logical form constructed from
+\exi{three}{b} using syntactic equivalences. Either term is compiled in an
 implementation of the finite state calculus to a finite state machine with 10 states
@@
 \begin{array}[t]{l}  
-1 0 \; bh \; 1 0 \; ah \; 1 0 \; aht \;  0 1 \\
-1 0 \; bh \; 1 0 \; aht \; 0 1 \; at \; 0 1 \\
-1 0 \; ah \; 1 0 \; bh \; 1 0 \; aht \; 0 1 \\
+1 0 \; \bpH \; 1 0 \; \apH \; 1 0 \; \atHT \;  0 1 \\
+1 0 \; \bpH \; 1 0 \; \atHT \; 0 1 \; \apT \; 0 1 \\
+1 0 \; \apH \; 1 0 \; \bpH \; 1 0 \; \atHT \; 0 1 \\
 \end{array}
@@
 individuals \citep{gallin1975intensional}. In the coin example, the individuals
-are hidden in the primitive event symbols such as $a_1$ and $b_1$. This
+are hidden in the primitive event symbols such as $\apH$ and $\bpH$. This
 situation would get worse in a more elaborate model with more agents and
@@
 computational characterizations of Epi\textsc{Kat}, such as sound and complete
-axioms, coalgebrae, and relations to other logics, warrant further investigation.
+axioms, coalgebras, and relations to other logics, warrant further investigation.
 

### 2021-01-15 tex/pdraft.tex

@@
 
-Sentences are parsed with a chart parser for categorial grammar.  The semantics for
-complex phrases are obtained by application of semantic translations, accompanied by
-beta reductions that eliminate all lambdas in logical forms for clauses.  In
-consequence, the semantic term translating a sentence is an Epi\textsc{Kat} logical term.  Such
-a term designates a set of possible words (guarded strings).  By way of example,
-\exi{three}{a} is an English sentence with conjunction and several levels of clausal
-embedding.  Using the grammar and parser, the sentence is mapped to the term in
-\exi{three}{b}. \exi{three}{c} shows a simplified logical form constructed from
-\exi{three}{b} using syntactic equivalences. Either term is compiled in an
-implementation of the finite state calculus to a finite state machine with 10 states
-and 23 edges, which accepts a countably infinite set of worlds.\footnote{ Machine
-  sizes need not be small, especially as the cardinality of $\sf{B}$ increases.  A
-  certain Epi\textsc{Kat} model with fourteen primitive Booleans has the set of worlds
-  representented by a finite state machine with 184,794 states and 257,881 edges.}  In
-this way the methodology ``directly'' represents the set of worlds denoted by
-\exi{three}{a}.
+Sentences are parsed with a chart parser for categorial grammar. The semantics
+for complex phrases are obtained by application of semantic translations,
+accompanied by beta reductions that eliminate all lambdas in logical forms for
+clauses. In consequence, the semantic term translating a sentence is an
+Epi\textsc{Kat} logical term. Such a term designates a set of possible words
+(guarded strings). By way of example, \exi{three}{a} is an English sentence with
+conjunction and several levels of clausal embedding. Using the grammar and
+parser, the sentence is mapped to the term in \exi{three}{b}. \exi{three}{c}
+shows a simplified logical form constructed from \exi{three}{b} using syntactic
+equivalences. Either term is compiled in an implementation of the finite state
+calculus to a finite state machine with 10 states and 23 edges, which accepts a
+countably infinite set of worlds.\footnote{ Machine sizes need not be small,
+  especially as the cardinality of $\sf{B}$ increases. A certain Epi\textsc{Kat}
+  model with fourteen primitive Booleans has the set of worlds representented by
+  a finite state machine with 184,794 states and 257,881 edges.} In this way the
+methodology ``directly'' represents the set of worlds denoted by \exi{three}{a}.
 

### 2021-01-16 tex/pdraft.tex
Random edits everywhere.

@@
 models, to illustrate and test ideas. The point of this paper is to extend
-standard idealized models to countable sets of worlds, and to constructively and
+idealized models to countable sets of worlds, and to constructively and
 computably model alternatives for epistemic agents.
@@
 $\bar t$. The effect valuations are restricted to only the feasible ones via a
-\emph{state formula}, shown, and demonstrated for $\apH$ in \ref{effect}. The
+\emph{state formula}, shown and demonstrated for $\apH$ in \ex{effect}. The
 state formula says that the coin is either heads-up or tails-up, but not
-both\footnote{Two Booleans constrained by a state formula, is more illustrative
+both\footnote{Two Booleans constrained by a state formula, is more illustrative than
   a single boolean $h$ where tails is just $\bar h$.}. Here juxtaposition
@@
 
-With the state formula in hand, $\apH\atTH$ is evidently contradictory, since
+With the state formula in hand, $\apH\atTH$ is contradictory, since
 the post-condition $h$ of $\apH$ is incompatible with the pre-condition $t$ of
@@
 epistemic operators that cannot naively be represented as
-hypotheses\footnote{Characterizing this relationship precisely is future work.}.
+hypotheses.\footnote{Characterizing this relationship precisely is future work.}
 Aside from formalizing the metaphysical modality seen so far,
@@
 (ii) says that each constituent token event $\alpha_ie_i\alpha_{i+1}$ is
-consistent with its effects\footnote{ An alternative is to
+consistent with its effects.\footnote{ An alternative is to
   define equations such as $\bar{\phi}=0$ (from the state formula $\phi$) and
@@
   instead use a set of guarded strings that are well-formed according to the
-  state and effect formulas as the representative of the equivalence class.}.
+  state and effect formulas as the representative of the equivalence class.}
 
@@
 determined by an Epi\textsc{Kat} specification is the set of well-formed guarded
-strings, and propositions are regular sets of guarded strings, rather than the
-power set \citep{montague1975formal,gallin1975intensional}. Certain sets of
+strings, and propositions are sets of guarded strings. Certain sets of
 well-formed guarded strings have the addional interpretation of event types. An
@@
 worlds for each agent $a$. This will result in a general modal frame
-$\tuple{W,R_1,...,R_n,K}$ consisting of a set of worlds, a world-alternative relation
+$\tuple{W,\hat{R}_1,...,\hat{R}_n,K}$ consisting of a set of worlds, a world-alternative relation
 for each agent, and a set $K$ of propositions, where each proposition is a subset of
@@
 
-Consider a scenario like the one from Figure~\ref{fig:epik-example}, augmented
-with an additional agent Cal. The base world $T\,\bpT\,T\,\cpT\,T$ is one where
-the coin is tails, and first Bob looks at tails, and them Cal looks at tails.
-The first event $\bpT$ has the alternatives $\bpT$ and $\bpH$ for Amy, and the
-second event $\bpT$ has the alternatives $\cpT$ and $\cpH$ for Amy. This results
-in four combinations $\bpT\cpT$, $\bpT\cpH$, $\bpH\cpT$, and $\bpH\cpH$, some of
-which, like $\bpT\cpH$, are bogus. These contradictions are filtered by pre-
+Consider a scenario like the one from Figure~\ref{fig:epik-example}, augmented with
+an additional agent Cal. The base world $T\,\bpT\,T\,\cpT\,T$ is one where the coin
+is tails, and first Bob looks at tails, and them Cal looks at tails.  The first event
+$\bpT$ has the alternatives $\bpT$ and $\bpH$ for Amy, and the second event $\bpT$
+has the alternatives $\cpT$ and $\cpH$ for Amy. This results in four combinations
+$\bpT,\cpT$, $\bpT,\cpH$, $\bpH,\cpT$, and $\bpH,\cpH$.  These are filtered by pre-
 and post-conditions of events in the alternative world, so that the set of
@@
 With modalities and propositional negation added, the signature of $n$-agent
-Epi\textsc{Kat} is $\tuple{K,+,\cdot,*,\bar{},0,1,\neg,\Diamond_1,
-  ..., \Diamond_n}$. Figure \ref{termfigure} defines the syntax of the language.
-Juxtaposition is used for product.  Terms in this language
-are used to represent the propositional semantic values of English sentences.
-
-Consider the examples in \ex{interpretation1}. The first term, $\bullet^*\,h$ uses
-the syntactic sugar $\bullet$, defined in Figure \ref{fig:epik-grammar} to be
-the disjunction of the primitive events. Since a world is a well-formed sequence
-of events, $\bullet^*$ is the set of worlds. Multiplying by the state symbol $h$
-in the term $\bullet^* h$ has the effect of conjoining $h$ with the atom at the
-end of the world. So $\bullet^* h$ is the set of worlds where the coin ends
-heads-up.
+epistemic \textsc{Kat} is $\tuple{K,+,\cdot,*,\bar{},0,1,\neg,\Diamond_1, ...,
+  \Diamond_n}$. Figure \ref{termfigure} defines the syntax of the language.
+Juxtaposition is used for product.  Terms in this language are used to represent the
+propositional semantic values of English sentences.  \ex{interpretation1} gives
+examples. The abbreviation $\bullet$ as defined in Figure \ref{fig:epik-grammar} is
+the disjunction of the primitive events. Since a world is a well-formed sequence of
+events, $\bullet^*$ is the set of worlds. Multiplying by the state symbol $h$ in the
+term $\bullet^* h$ has the effect of conjoining $h$ with the atom at the end of the
+world. So $\bullet^* h$ is the set of worlds where the coin ends heads-up.
 
@@
 \hspace{-1.5em}\begin{tabular}[t]{l}
-  $\Box_b(\bullet^*t \wedge \Box_a\bullet^*t + \neg\bullet^*t \wedge \Box_a\neg\bullet^*t)$\\
+  $\Box_b(\bullet^*t \wedge \Box_a\bullet^*t + (\neg\bullet^*t) \wedge \Box_a\neg\bullet^*t)$\\
  Bob believes Amy knows whether it's tails.  \\
@@
 the form of $e$ flanked by compatible atoms, as determined by the
-event formula $\zeta_e$. The product $pq$ is
+effect formula $\zeta_e$. The product $pq$ is
 interpreted with fusion product raised to sets of guarded strings.
@@
 
-Summing up, given an Epi\textsc{Kat} specification $\epikspec$, term $p$ (as
-defined syntactically in Figure \ref{termfigure}) is interpreted as a set of
-guarded strings $\sem{p}^{\epikspec}$. Let $K^{\epikspec}$ be the sets that are
+Summing up, given an Epi\textsc{Kat} specification $\tuple{\epikspec}$, term $p$ (as
+defined syntactically in Figure \ref{termfigure}) is interpreted as a set of guarded
+strings $\sem{p}^{\epikspec}$. Let $K^{\epikspec}$ be the sets that are
 interpretations of terms. Then
 $\tuple{K^{\epikspec},+,\cdot,*,\bar{},0,1,\neg,\Diamond_{a_1},...,\Diamond_{a_n}}$
-is a concrete guarded string interpretation for the signature of
-Epi\textsc{Kat}, with operations as in Figure \ref{fig:interp2}. This provides a
-concrete $n$-agent general modal frame $\tuple{\sem{\bullet^*}^{\epikspec},
-  \hat{R}_1,...,\hat{R}_n,K^{\epikspec}}$. The frame consists of a set of
-worlds, an epistemic-alternative relation for each agent, and a set of
-propositions, each of which is a set of worlds. It is used as a target for
-natural-language interpretation in Section~\ref{sec:categorial-grammar}.
+is a concrete guarded string interpretation for the signature of Epi\textsc{Kat},
+with operations as in Figure \ref{fig:interp2}. This provides a concrete $n$-agent
+general modal frame $\tuple{\sem{\bullet^*}^{\epikspec},
+  \hat{R}_1,...,\hat{R}_n,K^{\epikspec}}$. The frame consists of a set of worlds, an
+epistemic-alternative relation for each agent, and a set of propositions, each of
+which is a set of worlds. It is used as a target for natural-language interpretation
+in Section~\ref{sec:categorial-grammar}.  Since $K^{\epikspec}$ is the set of sets of
+worlds that are interpretations of terms, and there are only countably many terms,
+not all sets of worlds are propositions in the general modal frame.  Rather,
+propositions are regular sets of worlds.
 
@@
 with event symbols to encode a world.  In this construction, $w_3$ of the example is
-the string {\tt{1\!0\;ah\;1\!0\;bh\;1\!0}}.
+the string ``$1 \, 0 \,ah \,1 \,0 \,bh \,1 \,0$''.\footnote{The Fst programming language of the finite
+state calculus allows for multicharacter symbols, so that $ah$ is a single symbol in the string.}
 
@@
 
-Translating the Epi\textsc{Kat} terms $0$, $1$, $b$, and $e$ is
-straightforward: convert the atoms as previously described,
-and decorate the events $e$ with their compatible atoms. For example, $a_1$ becomes
-an Fst term denoting
-$\{01\apH01\}$.
-
-Further, Epi\textsc{Kat}'s union $(+)$ and intersection $(\wedge)$ operators can
-be directly translated to Fst's built-in analogous operators ((\texttt{|}) and
-(\texttt{\&}). Propositional complement in Epi\textsc{Kat} becomes Fst set
-difference (\texttt{-}) from the set of all worlds.
+Translating the Epi\textsc{Kat} terms $0$, $1$, $b$, and $e$ is straightforward:
+convert the atoms as previously described, and decorate the events $e$ with their
+compatible atoms. For example, $a_1$ becomes an Fst term denoting $\{01\apH01\}$.
+Union $(+)$ and intersection $(\wedge)$ operators are directly translated to Fst's
+built-in analogous operators ((\texttt{|}) and (\texttt{\&})). Propositional
+complement in Epi\textsc{Kat} becomes Fst set difference (\texttt{-}) from the set of
+all worlds.
 
@@
 uses Fst's rewrite notation to delete atoms (elements of \texttt{St}) that are
-preceded by another atom\footnote{ This is a non-equal length regular relation. The
+preceded by another atom.\footnote{ This is a non-equal length regular relation. The
   finite state calculus includes such relations, and they can be used with relation
@@
   complement and set difference for non-equal length relations is not defined.
-  Epistemic alterative relations are equal-length relations.}. This relation is
+  Epistemic alterative relations are equal-length relations.} This relation is
 applied via the composition (\texttt{.o.}) and codomain (\texttt{.l})
@@
 
-Finally, the epistemic alternative relation for each agent, i.e. the
-\texttt{agent} relation from Figure~\ref{fig:epik-example} is implemented in Fst.
-Combining the agent relation with the effect formula gives a relation on events
-decorated with compatible atoms. Then the relation on worlds is constructed
-using the closure of the concatenation product operation. The concatenation
-product $R~S$ of two relations $R$ and $S$ is the set of pairs of the form
-$\pair{x_1x_2}{y_1y_2}$, where $x_1 \mathrel R y_1$, and $x_2 \mathrel S y_2$.
-In Fst, {\tt{R+}} is the closure of relation {\tt R} with respect to this
-operation. Figure \ref{relclosure} defines the corresponding operation on sets
-of guarded strings as encoded in Fst.\footnote{Relation concatenation in Fst
-  differs from relation composition ({\tt{.o.}}), and the closure under
-  discussion here is the closure of the former rather than the latter.} The
-epistemic alternative relation on worlds for an agent is then defined as the
-\textsc{Kat} relation concatenation closure {\tt RelKst} of the decorated-event
-alternative relation for the agent.
+Finally, world alternative relations are constructed in Fst.  Combining the agent
+relation with the effect formula gives a relation on events decorated with compatible
+atoms. Then the corresponding relation on worlds is constructed using the closure of
+the relational concatenation product operation. The concatenation product $R~S$ of
+two relations $R$ and $S$ is the set of pairs of the form $\pair{x_1x_2}{y_1y_2}$,
+where $x_1 \mathrel R y_1$, and $x_2 \mathrel S y_2$.  In Fst, {\tt{R+}} is the
+closure of relation {\tt R} with respect to this operation. Figure \ref{relclosure}
+defines the corresponding operation on sets of guarded strings as encoded in
+Fst.\footnote{Relation concatenation in Fst differs from relation composition
+  ({\tt{.o.}}), and the closure under discussion here is the closure of the former
+  rather than the latter.} The epistemic alternative relation on worlds for an agent
+is then defined as the \textsc{Kat} relation concatenation closure {\tt RelKst} of
+the decorated-event alternative relation for the agent.
 
@@
 English sentences are mapped to terms in the logical language via a semantically
-interpreted multi-modal categorial grammar, consisting of a lexicon of words,
-their categorial types, and interpretations in a logical lambda language. The
-grammar covers basic statives ({\em it's heads}), {\em that-} and {\em
-  whether}-complements, negation, and predicate and sentence
-conjunction. Figure \ref{lexiconfigure} gives
-the lexicon.\footnote{Category symbols use Lambek/Bar-Hillel notation for
-  slashes, so that $(d \backslash t)\slash(d \backslash_D t)$ combines with $d
-  \backslash_D t$ on the right to give a value that combines with $d$ on the
-  left to give $t$. In the semantics, lambda abstractions with multiple
-  parameters are written $\lambda x~y.~e$ rather than $\lambda x.\lambda y.e$.
-  $d$ is the category of expletive {\em it}.} The online supplement includes a
-version of this grammar that can be run using Barker and Shan's chart parser for
-categorial grammar \citep{barker2005parser}. The grammar and semantics are
-optimized for a simple fragment of English concerned with clausal
-complementation. The agent names {\em Amy} and {\em Bob} contribute the
-epistemic alternative relations for those agents, rather than individuals. The
-root verb {\em believe} contributes existential modal force. The complementizers
-{\em that} and {\em whether} are the heads of their dominating clauses, and
-assemble an alternative relation, modal force, and proposition contributed by
-the complement. These complementizers introduce the dual via two negations, in
-order to express universal modal force.
+interpreted multi-modal categorial grammar, consisting of a lexicon of words, their
+categorial types, and interpretations in a logical lambda language. The grammar
+covers basic statives ({\em it's heads}), {\em that-} and {\em whether}-complements,
+negation, and predicate and sentence conjunction. Figure \ref{lexiconfigure} gives
+the lexicon.\footnote{Category symbols use Lambek/Bar-Hillel notation for slashes, so
+  that $(d \backslash t)\slash(d \backslash_D t)$ combines with $d \backslash_D t$ on
+  the right to give a value that combines with $d$ on the left to give $t$. In the
+  semantics, lambda abstractions with multiple parameters are written $\lambda
+  x~y.~e$ rather than $\lambda x.\lambda y.e$.  $d$ is the category of expletive {\em
+    it}.}  The grammar and semantics are optimized for a simple fragment of English
+concerned with clausal complementation. The agent names {\em Amy} and {\em Bob}
+contribute the epistemic alternative relations for those agents, rather than
+individuals. The root verb {\em believe} contributes existential modal force. The
+complementizers {\em that} and {\em whether} are the heads of their dominating
+clauses, and assemble an alternative relation, modal force, and proposition
+contributed by the complement. These complementizers introduce the dual via two
+negations, in order to express universal modal force.
 
@@
 the logical language, incremented with lambda.  The body of $\lambda
-x.\bullet^*h$, which is the semantic lexical entry for {\em heads}, is a
+x.\bullet^*\!h$, which is the semantic lexical entry for {\em heads}, is a
 term denoting the set of all worlds where the coin is heads, expressed as the
@@
 equivalences. Either term is compiled in an implementation of the finite state
-calculus to a finite state machine with 10 states and 23 edges, which accepts a
+calculus to a finite state machine with 49 states and 110 edges, which accepts a
 countably infinite set of worlds.\footnote{ Machine sizes need not be small,
@@
 
-\newcommand{\HEADS}{\bullet^*h}
-\newcommand{\TAILS}{\bullet^*t}
+\newcommand{\HEADS}{\bullet^*\!h}
+\newcommand{\TAILS}{\bullet^*\!t}
 
@@
 \itemc \(\K_{a}\!\TAILS \; \wedge \, \neg\B_{a}\!\HEADS \; \wedge \; \B_{a}\B_{b}\HEADS\)
-\itemd Bob believes that it's tails.
-\iteme \(\B_{b} \TAILS \)
+\itemd Bob believes that it's heads.
+\iteme \(\B_{b} \HEADS \)
 \end{examples}
@@
 Epi\textsc{Kat} is designed for research in linguistic semantics, for
-computational linguistics research on model-theoretically grounded semantics,
+computational linguistic research on model-theoretically grounded semantics,
 and for course modules on intensional formal semantics.
@@
 complementation \citep{thomason1984combinations, abusch1998generalizing,
-  condoravdi2002temporal}. Connections with research on temporal constitution of
+  condoravdi2002temporal}. Connections with research in a finite state framework on temporal constitution of
 events remain to be explored
@@
 
-The syntactic part of the grammar formalism uses a standard approach to
-categorial grammar \citep{steedman2000syntactic,bozsahin2012combinatory} to map
-between strings (or trees or derivations) and logical terms, and is rigorous
-enough to admit computational implementations
+The grammar formalism uses the standard approach in
+categorial grammar to map
+between strings (or trees or derivations) and logical terms
+\citep{steedman2000syntactic,bozsahin2012combinatory}, which comes
+with practically useful computational implementations
 \citep{barker2005parser,bozsahin2021ccglab}.\footnote{Our grammar uses basic
@@
   arguments.} While this approach is simple and attractive for our applications,
-it would be possible to use Epi\textsc{Kat} with other frameworks that support lambda
+it would be possible to use Epi\textsc{Kat} with other grammar formalisms that support lambda
 extensions of logical languages.
@@
 Semantic type systems for natural language usually include a type for
-individuals \citep{gallin1975intensional}. In the coin example, the individuals
+individuals \citep{montague1975formal,gallin1975intensional}. In the coin example, the individuals
 are hidden in the primitive event symbols such as $\apH$ and $\bpH$. This
